---
title: "Optimal control of fractional diffusion equations on metric graphs"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
  - \newcommand{\almosteverywhere}{\mathrm{a.e.}\;}
---

Go back to the [Contents](about.html) page.

<div style="color: #2c3e50; text-align: right;">
********  
<strong>Press Show to reveal the code chunks.</strong>  

********
</div>


```{r}
# Create a clipboard button on the rendered HTML page
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```

```{r}
library(MetricGraph)
library(ggplot2)
library(reshape2)
library(plotly)
```


```{r}
capture.output(
  knitr::purl(here::here("functionality.Rmd"), output = here::here("functionality.R")),
  file = here::here("purl_log.txt")
)
source(here::here("functionality.R"))
```

# Optimal control of fractional diffusion equations on metric graphs

## Problem statement {#optimal_control_problem}

Let $\Gamma = (\mathcal{V},\mathcal{E})$ be a metric graph. Let $u_d: \Gamma \times(0, T) \rightarrow \mathbb{R}$ be the desired state and  $\mu>0$ a regularization parameter. We define the cost functional

\begin{equation}
\label{eq:costfun}
\tag{1}
    J(u, z)=\frac{1}{2} \int_0^T\left(\left\|u-u_d\right\|_{L_2(\Gamma)}^2+\mu\|z\|_{L_2(\Gamma)}^2\right) dt
\end{equation}


Let $f:\Gamma\times (0,T)\rightarrow\ar$ and $u_0: \Gamma \rightarrow \mathbb{R}$ be fixed functions. We will call them right-hand side and initial datum, respectively. Let $\alpha\in(0,2]$ and $z: \Gamma \times(0, T) \rightarrow \mathbb{R}$ denote the control variable. We shall be concerned with the following PDE-constrained optimization problem: Find

\begin{equation}
\label{eq:min_pro}
\tag{2}
    \min\; J(u, z)
\end{equation}
subject to the fractional diffusion equation
\begin{equation}
\label{eq:maineq}
\tag{3}
\left\{
\begin{aligned}
    \partial_t u(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s,t) &= f(s,t)+z(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    u(s,0) &= u_0(s), && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
with $u(\cdot,t)$ satisfying the Kirchhoff vertex conditions
\begin{equation}
\label{eq:Kcond}
\tag{4}
   \mathcal{K} =  \left\{\phi\in C(\Gamma)\;\middle|\; \forall v\in \mathcal{V}:\; \sum_{e\in\mathcal{E}_v}\partial_e \phi(v)=0 \right\}
\end{equation}
and the control constraints 
\begin{align}
\label{control_constraints}
\tag{5}
    a(s,t)\leq z(s,t)\leq b(s,t)\;\almosteverywhere (s,t)\in\Gamma \times(0, T).
\end{align}


## Numerical scheme {#num_scheme_optim_control}

The optimal variables $(\bar{u}, \bar{p}, \bar{z})$ satisfy

\begin{equation}
\label{eq:maineqoptimal}
\tag{6}
\left\{
\begin{aligned}
    \partial_t \bar{u}(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{u}(s,t) &= f(s,t)+\bar{z}(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    \bar{u}(s,0) &= u_0(s), && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
and
\begin{equation}
\label{eq:adjointeq}
\tag{7}
\left\{
\begin{aligned}
    -\partial_t \bar{p}(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{p}(s,t) &= \bar{u}(s,t)-u_d(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    \bar{p}(s,T) &= 0, && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
with 
\begin{align}
\label{zz}
\tag{8}
        \bar{z}(s,t) = \max\left\{a(s,t),\min\left\{b(s,t),-\dfrac{1}{\mu}\bar{p}(s,t)\right\}\right\}.
\end{align}
By considering the change of variable $\bar{q}(s,t) = \bar{p}(s,T-t)$, the fractional adjoint problem \eqref{eq:adjointeq} becomes a forward-in-time problem where the transformed adjoint state $\bar{q}$ solves
\begin{equation}
\label{transformed_adjoint_state}
\tag{9}
\left\{
\begin{aligned}
    \partial_t \bar{q}(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{q}(s,t) &= \bar{v}(s,t)-v_d(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    \bar{q}(s,0) &= 0, && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
since $\partial_t\bar{q}(s,t) = -\partial_t\bar{p}(s,T-t)$ and $\bar{q}(s, 0) =\bar{p}(s,T-0)= \bar{p}(s,T)=0$. Here, $\bar{v}(s,t) = \bar{u}(s,T-t)$ and $v_d(s,t) = u_d(s,T-t)$. 


Given $\bar{u}$ and $u_d$, we can time-reverse them to obtain $\bar{v}$ and $v_d$ and then use the same numerical scheme we use for the forward problem \eqref{eq:maineqoptimal} to solve the adjoint problem \eqref{transformed_adjoint_state}. The control variable $\bar{z}$ is then computed using \eqref{zz}.

The numerical scheme for \eqref{eq:maineqoptimal} and \eqref{transformed_adjoint_state} are given by (see the [Functionality](functionality.html#num_scheme) page)
\begin{align}
\label{numericalscheme1}
\tag{10}
\bar{\boldsymbol{U}}^{k+1} = \boldsymbol{C}^{-1}\left(\sum_{k=1}^{m+1} a_k\left(\dfrac{\boldsymbol{L}\boldsymbol{C}^{-1}}{\kappa^2}-p_k\boldsymbol{I}\right)^{-1} + r\boldsymbol{I}\right) (\boldsymbol{C}\bar{\boldsymbol{U}}^k+\tau (\boldsymbol{F}^{k+1}+\bar{\boldsymbol{Z}}^{k+1})),
\end{align}
and
\begin{align}
\label{thenumericalscheme2}
\tag{11}
\bar{\boldsymbol{Q}}^{k+1} = \boldsymbol{C}^{-1}\left(\sum_{k=1}^{m+1} a_k\left(\dfrac{\boldsymbol{L}\boldsymbol{C}^{-1}}{\kappa^2}-p_k\boldsymbol{I}\right)^{-1} + r\boldsymbol{I}\right) (\boldsymbol{C}\bar{\boldsymbol{Q}}^k+\tau (\bar{\boldsymbol{V}}^{k+1}-\boldsymbol{V_d}^{k+1})),
\end{align}
with initial conditions $\bar{\boldsymbol{U}}^0 = \boldsymbol{U}_0$ and $\bar{\boldsymbol{Q}}^0 = \boldsymbol{0}$. Observe that \eqref{numericalscheme1}-\eqref{thenumericalscheme2} is a coupled problem.

For notation convenience, let

 - $\bar{\boldsymbol{U}}$  be the matrix with columns $\bar{\boldsymbol{U}}^k$ obtained from the numerical scheme \eqref{numericalscheme1} 
 - $\boldsymbol{F}$ be the matrix with columns $\boldsymbol{F}^k$ and entries $\boldsymbol{F}_{ik}=(\psi^i_h,f(\cdot,t_k))_{L_2(\Gamma)}$
 - $\bar{\boldsymbol{Z}}$ be the matrix with columns $\bar{\boldsymbol{Z}}^k$ and entries $\bar{\boldsymbol{Z}}_{ik}=(\psi^i_h,\bar{z}(\cdot,t_k))_{L_2(\Gamma)}$
 - $\bar{\boldsymbol{Q}}$ be the matrix with columns $\bar{\boldsymbol{Q}}^k$ obtained from the numerical scheme \eqref{thenumericalscheme2}
 - $\bar{\boldsymbol{V}}$ be the matrix with columns $\bar{\boldsymbol{V}}^k$ and entries $\bar{\boldsymbol{V}}^{ik}=(\psi^i_h,\bar{v}(\cdot,t_k))_{L_2(\Gamma)} =(\psi^i_h,\bar{u}(\cdot,T-t_k))_{L_2(\Gamma)}$
 - $\boldsymbol{V_d}$ be the matrix with columns $\boldsymbol{V_d}^k$ and entries $\boldsymbol{V_d}^{ik}=(\psi^i_h,v_d(\cdot,t_k))_{L_2(\Gamma)} =(\psi^i_h,u_d(\cdot,T-t_k))_{L_2(\Gamma)}$
 - $\boldsymbol{A}$ be the matrices with entries $\boldsymbol{A}_{ik} = a(s_i,t_k)$
 - $\boldsymbol{B}$ be the matrices with entries $\boldsymbol{B}_{ik} = b(s_i,t_k)$
 - $\bar{\boldsymbol{P}} = \text{reversecolumns}(\bar{\boldsymbol{Q}})$


Observe that the entries of $\boldsymbol{F}$ are given by $\boldsymbol{F}_{ik}=(\psi^i_h, f(\cdot,t_k))_{L_2(\Gamma)}$ where $\psi^i_h$ is the $i$-th basis function on a mesh with $N_h$ nodes. By introducing a finer integration mesh with $N_{h_{\text{ok}}}$ nodes, we can approximate $\boldsymbol{F}_{ik}$ as $\boldsymbol{\psi}_i^\top\boldsymbol{C}^{\text{ok}}\boldsymbol{f}_k$, where $\boldsymbol{\psi}_i=[\psi^i_h(s_1)\dots \psi^i_h(s_{N_{h_{\text{ok}}}})]^\top$ for $i=1\dots, N_h$, $\boldsymbol{f}_k = [f(s_1,t_k)\dots f(s_{N_{h_{\text{ok}}}},t_k)]\top$ for $k= 0,\dots,N$, and matrix $\boldsymbol{C}^{\text{ok}}$ has entries $\boldsymbol{C}^{\text{ok}}_{nm} = (\psi_{h_\text{ok}}^n,\psi_{h_\text{ok}}^m)_{L_2(\Gamma)}$ for $n,m = 1,\dots, N_{h_\text{ok}}$. Let $\boldsymbol{f}$ denote the matrix with columns $\boldsymbol{f}_k$ and $\boldsymbol{\Psi}$ denote a matrix with entries $\boldsymbol{\Psi}_{ji}=\psi^i_h(s_j)$ for $i=1\dots, N_h$ and $j = 1,\dots N_{h_{\text{ok}}}$. Then 
\begin{equation}
\boldsymbol{F} \approx \boldsymbol{\Psi}^\top \boldsymbol{C}^{\text{ok}} \boldsymbol{f}.
\end{equation}
Similarly, if $\boldsymbol{U_d}$ is the matrix with entries $\boldsymbol{U_d}^{jk} = u_d(s_j,t_k)$ for $j = 1,\dots, N_{h_{\text{ok}}}$ and $k = 0,\dots, N$, then the matrix $\boldsymbol{V_d}$ can be approximated as
\begin{equation}
\boldsymbol{V_d} \approx \text{reversecolumns}(\boldsymbol{\Psi}^\top \boldsymbol{C}^{\text{ok}} \boldsymbol{U_d}).
\end{equation}


To approximate $\boldsymbol{F}$ and $\boldsymbol{V_d}$, we use the values of $f$ and $u_d$ on the integration mesh with $N_{h_{\text{ok}}}$ nodes. However, to approximate $\bar{\boldsymbol{Z}}$ and $\bar{\boldsymbol{V}}$, we do not even have access to the values of $\bar{z}$ and $\bar{v}$ on the coarse mesh, let alone on the fine mesh. Instead, we use their approximations on the coarse mesh, given by the matrices $\bar{\boldsymbol{z}} = \max\{\boldsymbol{A},\min\{\boldsymbol{B},-\bar{\boldsymbol{P}}/\mu\}\}$ and $\text{reversecolumns}(\bar{\boldsymbol{U}})$, respectively. With these at hand, 
we can approximate $\bar{\boldsymbol{Z}}$ and $\bar{\boldsymbol{V}}$ by first projecting $\bar{\boldsymbol{z}} = \max\{\boldsymbol{A},\min\{\boldsymbol{B},-\bar{\boldsymbol{P}}/\mu\}\}$ and $\text{reversecolumns}(\bar{\boldsymbol{U}})$ onto the fine mesh and then approximating the inner products $(\psi^i_h,\bar{z}(\cdot,t_k))_{L_2(\Gamma)}$ and $(\psi^i_h,\bar{u}(\cdot,T-t_k))_{L_2(\Gamma)}$ as
\begin{equation}
\bar{\boldsymbol{Z}} \approx \boldsymbol{\Psi}^\top \boldsymbol{C}^{\text{ok}} \boldsymbol{\Psi} \bar{\boldsymbol{z}}\quad \text{ and }\quad \bar{\boldsymbol{V}} \approx \text{reversecolumns}(\boldsymbol{\Psi}^\top \boldsymbol{C}^{\text{ok}} \boldsymbol{\Psi} \bar{\boldsymbol{U}}).
\end{equation}


## Construction of an exact solution to the optimal control problem {#exact_sol_for_optim_control_prob}

We follow the approach in @Glusa2021errorestimates to construct an exact solution to the optimal control problem \eqref{eq:costfun}-\eqref{control_constraints}. The solution to the elliptic problem
\begin{align}
    \label{eq:elliptic_problem}
    \tag{12}
   (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s)=f(s),\quad s\in\Gamma,
\end{align}
where $u(\cdot)$ satisfies the Kirchhoff vertex conditions \eqref{eq:Kcond}, is given by
\begin{align*}
    u(s) = \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}(f,e_j)_{L_2(\Gamma)}e_j(s),\quad s\in\Gamma.
\end{align*}
We choose coefficients $\{x_j\}_{j=1}^{\infty}$ and $\{y_j\}_{j=1}^{\infty}$ and set
\begin{align}
    \label{eq:rhsfunctions}
    \tag{13}
   f(s) = \sum_{j=1}^{\infty}x_je_j(s)\quad \text{ and } \quad g(s) = \sum_{j=1}^{\infty}y_je_j(s).
\end{align}
Hence the corresponding solutions are given by
\begin{align}
    \label{eq:sols}
    \tag{14}
   u(s) = \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)\quad \text{ and } \quad v(s) = \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s).
\end{align}
 We also define $\psi(t) = \cos(t)$ and $\phi(t) =\sin(T-t)$. Observe that $\psi(0)=1$ and $\phi(T)=0$. We also choose $\mu =0.1$, $a(s,t) = -0.5$ and $b(s,t)=0.5$. Now we set
\begin{align*}
    f(s,t) &= \psi'(t)u(s)+\psi(t)f(s) - \bar{z}(s,t) = -\sin(t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)+ \cos(t)\sum_{j=1}^{\infty}x_je_j(s)- \max\{-0.5,\min\{0.5, \sin(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)\}\}\\ 
    u_d(s,t) & = \psi(t)u(s)-\mu\phi'(t)v(s)+\mu\phi(t)g(s) = \cos(t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)+\mu\cos(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)+\mu\sin(T-t)\sum_{j=1}^{\infty}y_je_j(s)\\
    u_0(s)  & = u(s)= \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)
\end{align*}
The exact solution to the optimal control problem is given by
\begin{align}
    \bar{u}(s,t)&=\psi(t)u(s) = \cos(t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)\\
    \bar{p}(s,t)&=-\mu\phi(t)v(s)= - \mu\sin(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)\\
    \bar{z}(s,t)&=\max\{a(s,t),\min\{b(s,t), \phi(t)v(s)\}\}= \max\{-0.5,\min\{0.5, \sin(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)\}\} 
\end{align}

Observe that the choice of $f$ decouples the systems \eqref{numericalscheme1}-\eqref{thenumericalscheme2}.

### Checking that $\bar{u}$ and $\bar{p}$ satisfy \eqref{eq:maineqoptimal} and \eqref{eq:adjointeq}

At $t=0$, we get
\begin{align*}
    \bar{u}(s,0) = \psi(0)u(s) = u(s) = u_0(s),
\end{align*}
while for $(s,t) \in \Gamma \times (0, T)$, we obtain
\begin{align*}
    \partial_t \bar{u}(s,t) +(\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{u}(s,t) = \psi'(t)u(s) + \psi(t) (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s) = \psi'(t)u(s) + \psi(t)f(s) = f(s,t) + \bar{z}(s,t).
\end{align*}
Similarly, at $t=T$, we get
\begin{align*}
    \bar{p}(s,T) = -\mu\phi(T)v(s) = 0,
\end{align*}
while for $(s,t) \in \Gamma \times (0, T)$, we obtain
\begin{align*}
    -\partial_t \bar{p}(s,t) +(\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{p}(s,t) = \mu\phi'(t)v(s) -\mu \phi(t) (\kappa^2 - \Delta_\Gamma)^{\alpha/2} v(s) = \mu\phi'(t)v(s) -\mu \phi(t)g(s)
\end{align*}
and
\begin{align*}
    \bar{u}(s,t) - u_d(s,t) = \psi(t)u(s) - \psi(t)u(s) + \mu\phi'(t)v(s) - \mu\phi(t)g(s) = \mu\phi'(t)v(s) -\mu \phi(t)g(s).
\end{align*}
This shows that $\bar{u}$ and $\bar{p}$ satisfy \eqref{eq:maineqoptimal} and \eqref{eq:adjointeq}.

## Numerical implementation {#num_implementation_optimal_control_problem}

In this section, we implement the numerical scheme for the optimal control problem \eqref{eq:costfun}-\eqref{control_constraints} on the tadpole graph. We start by defining the parameters and constructing the exact solution. We them use \eqref{numericalscheme1}-\eqref{thenumericalscheme2} to solve the forward and adjoint problems. Finally, we compute the control variable $\bar{z}$ using \eqref{zz}.


```{r}
# Parameters
T_final <- 2
time_step <- 0.02
h <- 0.02
kappa <- 15
alpha <- 1.8
m = 4
beta <- alpha/2
mu <- 0.1
a <- - 0.5
b <- 0.5
N_finite = 4 # choose even
adjusted_N_finite <- N_finite + N_finite/2 + 1
# Coefficients for f and g
coeff_elliptic_g <- rep(0, adjusted_N_finite)
coeff_elliptic_g[7] <- 100
coeff_elliptic_f <- rep(0, adjusted_N_finite)
coeff_elliptic_f[7] <- -10
# time and spatial mesh 
time_seq <- seq(0, T_final, length.out = ((T_final - 0) / time_step + 1))
graph <- gets.graph.tadpole(h = h)
graph$compute_fem()
G <- graph$mesh$G
C <- graph$mesh$C
L <- kappa^2*C + G
I <- Matrix::Diagonal(nrow(C))
# Compute the eigenvalues and eigenfunctions
eigen_params <- gets.eigen.params(N_finite = N_finite, 
                                  kappa = kappa, 
                                  alpha = alpha, 
                                  graph = graph)
EIGENVAL_MINUS_ALPHA <- eigen_params$EIGENVAL_MINUS_ALPHA
EIGENFUN <- eigen_params$EIGENFUN
# Construct the right hand side functions f and g for the elliptic problem
elliptic_f <- as.vector(EIGENFUN %*% coeff_elliptic_f)
elliptic_g <- as.vector(EIGENFUN %*% coeff_elliptic_g)
# Construct the corresponding elliptic solution u and v
elliptic_u <- as.vector(EIGENFUN %*% (coeff_elliptic_f * EIGENVAL_MINUS_ALPHA))
elliptic_v <- as.vector(EIGENFUN %*% (coeff_elliptic_g * EIGENVAL_MINUS_ALPHA))

A <- matrix(a, nrow = nrow(C), ncol = length(time_seq))
B <- matrix(b, nrow = nrow(C), ncol = length(time_seq))
# Construct the optimal variables
psi <- cos(time_seq)
psi_prime <- - sin(time_seq)
phi <- sin(T_final - time_seq)
phi_prime <- - cos(T_final - time_seq)
# psi_rate <- 1
# phi_rate <- 1
# phi_coefficient <- 1
# psi <- exp(psi_rate*time_seq)
# psi_prime <- psi_rate * exp(psi_rate*time_seq)
# phi <- phi_coefficient * exp(phi_rate*time_seq) - phi_coefficient * exp(phi_rate*T_final)
# phi_prime <- phi_coefficient * phi_rate * exp(phi_rate*time_seq)
u_bar <- elliptic_u %*% t(psi)
p_bar <- - mu * (elliptic_v %*% t(phi))
z_bar <- pmax(A, pmin(B, - p_bar / mu))
# Construct the data for the problem
f <- elliptic_u %*% t(psi_prime) + elliptic_f %*% t(psi) - z_bar
f_plus_z_bar <- elliptic_u %*% t(psi_prime) + elliptic_f %*% t(psi)
u_d <- elliptic_u %*% t(psi) -
  mu * (elliptic_v %*% t(phi_prime)) +
  mu * (elliptic_g %*% t(phi))
u_0 <- elliptic_u
# Construct the fractional operator, which is shared for the forward and adjoint problems
my_op_frac <- my.fractional.operators.frac(L, beta, C, scale.factor = kappa^2, m = m, time_step)
# Construct the integration mesh
overkill_graph <- gets.graph.tadpole(h = 0.0001)
overkill_graph$compute_fem()
overkill_EIGENFUN <- gets.eigen.params(N_finite = N_finite, 
                                       kappa = kappa, 
                                       alpha = alpha, 
                                       graph = overkill_graph)$EIGENFUN
# Construct the right hand side functions f and g for the elliptic problem on the integration mesh
overkill_elliptic_f <- as.vector(overkill_EIGENFUN %*% coeff_elliptic_f)
overkill_elliptic_g <- as.vector(overkill_EIGENFUN %*% coeff_elliptic_g)
# Construct the corresponding elliptic solution u and v on the integration mesh
overkill_elliptic_u <- as.vector(overkill_EIGENFUN %*% (coeff_elliptic_f * EIGENVAL_MINUS_ALPHA))
overkill_elliptic_v <- as.vector(overkill_EIGENFUN %*% (coeff_elliptic_g * EIGENVAL_MINUS_ALPHA))
# Construct the projection matrix
Psi <- graph$fem_basis(overkill_graph$get_mesh_locations())
R <- t(Psi) %*% overkill_graph$mesh$C
# Construct the right hand side term for the forward problem
F_plus_Z_bar <- overkill_elliptic_u %*% t(psi_prime) + overkill_elliptic_f %*% t(psi) 
F_plus_Z_bar_innerproduct <- R %*% F_plus_Z_bar
# Solve the forward problem
U_bar <- solve_fractional_evolution(my_op_frac, 
                                     time_step, 
                                     time_seq, 
                                     val_at_0 = u_0, 
                                     RHST = F_plus_Z_bar_innerproduct)
# Construct the right hand side term for the adjoint problem
U_d <- overkill_elliptic_u %*% t(psi) -
  mu * (overkill_elliptic_v %*% t(phi_prime)) + 
  mu * (overkill_elliptic_g %*% t(phi))
V_d <- reversecolumns(R %*% U_d)
V_bar <- reversecolumns(R %*% Psi %*% U_bar)
# Solve the adjoint problem
Q_bar <- solve_fractional_evolution(my_op_frac, 
                                     time_step, 
                                     time_seq, 
                                     val_at_0 = u_0*0, 
                                     RHST = V_bar - V_d)
P_bar <- reversecolumns(Q_bar)
# Compute the control variable
Z_bar <- pmax(A, pmin(B, - P_bar / mu))
```


```{r}
start_idx <- which.min(abs(time_seq - 0))
end_idx <- which.min(abs(time_seq - T_final))
idx <- seq(start_idx, end_idx, by = 1)
```

```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Right-hand side functions $f$ and $g$ for the elliptic problem.")}
graph.plotter.3d.static(graph, list(f = elliptic_f, g = elliptic_g))

```


```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Solutions $u$ and $v$ to the elliptic problem.")}
graph.plotter.3d.static(graph, list(u = elliptic_u, v = elliptic_v))
```

```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Time evolution of the right-hand side function $f+\\bar{z}$ of the state equation.")}
graph.plotter.3d.single(graph, f_plus_z_bar[, idx], time_seq[idx])
```

```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Time evolution of the right-hand side function $\\bar{v}-v_d$ of the adjoint equation.")}
v_bar_minus_v_d <- reversecolumns(u_bar - u_d)
graph.plotter.3d.single(graph, v_bar_minus_v_d[, idx], time_seq[idx])
```

## Visualization of the results {#visualization_results}

```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Exact (in red) and numerical (in blue) optimal state variable.")}
graph.plotter.3d(graph, time_seq[idx], time_seq[idx], list(exact_state = u_bar[, idx], numerical_state = U_bar[, idx]))
```



```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Exact (in red) and numerical (in blue) optimal adjoint variable.")}
graph.plotter.3d(graph, time_seq[idx], time_seq[idx], list(exact_adjoint = p_bar[, idx], numerical_adjoint = P_bar[, idx]))
```


```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Exact (in red) and numerical (in blue) optimal control variable.")}
graph.plotter.3d(graph, time_seq[idx], time_seq[idx], list(exact_control = z_bar[, idx], numerical_control = Z_bar[, idx]))
```


```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Desired state (in red) and optimal state (in blue).")}
graph.plotter.3d(graph, time_seq[idx], time_seq[idx], list(desired_state = u_d[, idx], numerical_state = u_bar[, idx]))
```


# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```