---
title: "Control Functionality"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

Go back to the [Contents](about.html) page.

<div style="color: #2c3e50; text-align: right;">
********  
<strong>Press Show to reveal the code chunks.</strong>  

********
</div>


```{r, purl = FALSE, echo = FALSE}
# Create a clipboard button on the rendered HTML page
source(here::here("clipboard.R")); clipboard
```


```{r, purl = FALSE, class.source = "fold-hide"}
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```



```{r}
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")
library(rSPDE)
library(MetricGraph)
library(grateful)

library(ggplot2)
library(reshape2)
library(plotly)
```


# Optimal control of fractional diffusion equations on metric graphs

## Problem statement {#optimal_control_problem}

Let $\Gamma = (\mathcal{V},\mathcal{E})$ be a metric graph. Let $u_d: \Gamma \times(0, T) \rightarrow \mathbb{R}$ be the desired state and  $\mu>0$ a regularization parameter. We define the cost functional

\begin{equation}
\label{eq:costfun}
\tag{1}
    J(u, z)=\frac{1}{2} \int_0^T\left(\left\|u-u_d\right\|_{L_2(\Gamma)}^2+\mu\|z\|_{L_2(\Gamma)}^2\right) dt
\end{equation}


Let $f:\Gamma\times (0,T)\rightarrow\ar$ and $u_0: \Gamma \rightarrow \mathbb{R}$ be fixed functions. We will call them right-hand side and initial datum, respectively. Let $\alpha\in(0,2]$ and $z: \Gamma \times(0, T) \rightarrow \mathbb{R}$ denote the control variable. We shall be concerned with the following PDE-constrained optimization problem: Find

\begin{equation}
\label{eq:min_pro}
\tag{2}
    \min\; J(u, z)
\end{equation}
subject to the fractional diffusion equation
\begin{equation}
\label{eq:maineq}
\tag{3}
\left\{
\begin{aligned}
    \partial_t u(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s,t) &= f(s,t)+z(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    u(s,0) &= u_0(s), && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
with $u(\cdot,t)$ satisfying the Kirchhoff vertex conditions
\begin{equation}
\label{eq:Kcond}
\tag{4}
   \mathcal{K} =  \left\{\phi\in C(\Gamma)\;\middle|\; \forall v\in \mathcal{V}:\; \sum_{e\in\mathcal{E}_v}\partial_e \phi(v)=0 \right\}
\end{equation}
and the control constraints 
\begin{align}
\label{control_constraints}
\tag{5}
    a(s,t)\leq z(s,t)\leq b(s,t)\;\text{a.e.} (s,t)\in\Gamma \times(0, T).
\end{align}


## Optimal solution 

The optimal variables $(\bar{u}, \bar{p}, \bar{z})$ satisfy

\begin{equation}
\label{eq:maineqoptimal}
\tag{6}
\left\{
\begin{aligned}
    \partial_t \bar{u}(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{u}(s,t) &= f(s,t)+\bar{z}(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    \bar{u}(s,0) &= u_0(s), && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
and
\begin{equation}
\label{eq:adjointeq}
\tag{7}
\left\{
\begin{aligned}
    -\partial_t \bar{p}(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{p}(s,t) &= \bar{u}(s,t)-u_d(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    \bar{p}(s,T) &= 0, && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
with 
\begin{align}
\label{zz}
\tag{8}
        \bar{z}(s,t) = \max\left\{a(s,t),\min\left\{b(s,t),-\dfrac{1}{\mu}\bar{p}(s,t)\right\}\right\}.
\end{align}

## Numerical Scheme {#num_scheme_optim_control}

### Discretization by time reversal strategy

By considering the change of variable $t^* = T-t$ and defining $\bar{q}(s,t^*): = \bar{p}(s,T-t^*)$, the fractional adjoint problem \eqref{eq:adjointeq} becomes a forward-in-time problem where the transformed adjoint state $\bar{q}$ satisfies the Kirchhoff vertex conditions \eqref{eq:Kcond} and solves
    \begin{equation}
    \label{transformed_adjoint_state}
    \tag{9}
    \left\{
    \begin{aligned}
        \partial_{t^*} \bar{q}(s,t^*) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{q}(s,t^*) &= \bar{v}(s,t^*)-v_d(s,t^*), && \quad (s,t^*) \in \Gamma \times (0, T), \\
        \bar{q}(s,0) &= 0, && \quad s \in \Gamma,
    \end{aligned}
    \right.
    \end{equation}
    since $\partial_t\bar{p}(s,t) = -\partial_{t^*}\bar{q}(s,t^*)$ and $\bar{q}(s, 0)= \bar{p}(s,T)=0$. Here, $\bar{v}(s,t^*) = \bar{u}(s,T-t^*)$ and $v_d(s,t^*) = u_d(s,T-t^*)$.
    
    
Given $\bar{u}$ and $u_d$, we can time-reverse them to obtain $\bar{v}$ and $v_d$ and then use the same numerical scheme we use for the forward problem \eqref{eq:maineqoptimal} to solve the adjoint problem \eqref{transformed_adjoint_state}. The control variable $\bar{z}$ is then computed using \eqref{zz}.



The numerical scheme for \eqref{eq:maineqoptimal} and \eqref{transformed_adjoint_state} are given by (see the [Functionality](functionality.html#num_scheme) page)

\begin{align}
\tag{10}
    \label{numericalscheme1}
    \begin{cases}
    &\mathbf{\bar{U}}_{k+1}  = \mathbf{R}\mathbf{C}\mathbf{\bar{U}}_k+\tau \mathbf{R}(\mathbf{F}_{k+1}+\mathbf{C}\mathbf{\bar{Z}}_{k+1}),\quad k = 0,\dots, N-1,\\
    &\mathbf{\bar{U}}_{0}  = [u_0(s_1), \dots, u_0(s_{N_h})]^\top,
    \end{cases}
\end{align}
and
\begin{align}
\tag{11}
    \label{thenumericalscheme2}
    \begin{cases}
    &\mathbf{\bar{Q}}_{k+1}  = \mathbf{R}\mathbf{C}\mathbf{\bar{Q}}_k+\tau\mathbf{R} ((\mathbf{C}\mathbf{\bar{U}}\mathbf{J}_{N+1})_{k+1}-(\mathbf{D}\mathbf{J}_{N+1})_{k+1}),\quad k = 0,\dots, N-1,\\
    &\mathbf{\bar{Q}}_{0}  = \mathbf{0},
    \end{cases}
\end{align}
where $\mathbf{R} = \sum_{k=1}^{m+1} a_k\left(\mathbf{L}/\kappa^2-p_k\mathbf{C}\right)^{-1}$. Observe that \eqref{numericalscheme1}-\eqref{thenumericalscheme2} is a coupled problem.

Here

- $\mathbf{\bar{U}}$ has entries $\mathbf{U}_{j,k}  = \bar{u}(s_j,t_k)$,
- $\mathbf{F}$ has entries $\mathbf{F}_{j,k} =(f^{k},\psi^j_h)_{L_2(\Gamma)}$,
- $\mathbf{\bar{Z}}$ has entries $\mathbf{\bar{Z}}_{j,k} = \bar{z}(s_j,t_k)$,
- $\mathbf{D}$ has entries $\mathbf{D}_{j,k} =(u_d^{k},\psi^j_h)_{L_2(\Gamma)}$,
- $\mathbf{\bar{P}} = \mathbf{\bar{Q}}\mathbf{J}_{N+1}$ and has entries $\mathbf{\bar{P}}_{j,k} = \bar{p}(s_j,t_k)$.


If we change \eqref{thenumericalscheme2} to $\mathbf{\bar{P}}$, then we obtain 
\begin{align}
\tag{12}
    \label{thenumericalscheme3}
    \begin{cases}
    &\mathbf{\bar{P}}_{k}  = \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau \mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{k}-\mathbf{D}_{k}),\quad k = N-1,\dots, 0,\\
    &\mathbf{\bar{P}}_{N}  = \mathbf{0}.
    \end{cases}
\end{align}

### Discretization by directly solving the adjoint problem

The discrete version $\bar{P}_h^\tau\subset V_h$ of the adjoint optimal state $\bar{p}$ in problem \eqref{eq:adjointeq} solves 

\begin{equation}
\label{discrete_adjoint}
    \left\{
\begin{aligned}
    \langle\bar{\delta} \bar{P}_h^{k},\phi\rangle + \mathfrak{a}(\bar{P}_h^{k},\phi) &= \langle \bar{U}_h^{k+1}-u_d^{k+1},\phi\rangle, \quad\forall\phi\in V_h,\quad k=N-1,\dots, 0, \\
    \bar{P}^N_h &= 0
\end{aligned}
\right.
\end{equation}
The above expression can be equivalently written as
\begin{equation}
    \left\{
\begin{aligned}
    \langle\dfrac{\bar{P}_h^{k} - \bar{P}_h^{k+1}}{\tau},\phi\rangle + \mathfrak{a}(
        \bar{P}_h^{k},\phi) & = \langle \bar{U}_h^{k+1} - u_d^{k+1},\phi\rangle, \quad\forall\phi\in V_h,\quad k=N-1,\dots, 0, \\
    \bar{P}^N_h &= 0
\end{aligned}
\right.
\end{equation}
or
\begin{equation}
\label{disoptcons}
\tag{13}
    \left\{
\begin{aligned}
    \langle \bar{P}_h^{k},\phi\rangle + \tau\mathfrak{a}(
        \bar{P}_h^{k},\phi) & = \langle \bar{P}_h^{k+1},\phi\rangle + \tau\langle \bar{U}_h^{k+1} - u_d^{k+1},\phi\rangle, \quad\forall\phi\in V_h,\quad k=N-1,\dots, 0, \\
    \bar{P}^N_h &= 0
\end{aligned}
\right.
\end{equation}

At each time step $t_k$, the finite element solution $\bar{P}_h^k\in V_h$ to \eqref{disoptcons} can be expressed as a linear combination of the basis functions  $\{\psi^i_h\}_{i=1}^{N_h}$ introduced in the [Preliminaries](preliminaries.html#fem-basis) page, namely, 
\begin{align}
\label{num_sol2}
\tag{14}
    \bar{P}_h^k(s) =  \sum_{i=1}^{N_h}p_i^k\psi^i_h(s), \;s\in\Gamma.
\end{align}
Replacing \eqref{num_sol2} into \eqref{disoptcons} yields the following system
\begin{align*}
    \sum_{j=1}^{N_h}p_j^{k}[(\psi_h^j,\psi_h^i)_{L_2(\Gamma)}+ \tau\mathfrak{a}(\psi_h^j,\psi_h^i)] = \sum_{j=1}^{N_h}p_j^{k+1}(\psi_h^j,\psi_h^i)_{L_2(\Gamma)}+\tau(\bar{U}_h^{k+1} - u_d^{k+1},\psi_h^i)_{L_2(\Gamma)},\quad i = 1,\dots, N_h.
\end{align*}
In matrix notation,
\begin{align}
\label{diff_eq_discrete_adjoint}
    (\mathbf{C}+\tau \mathbf{L}^{\alpha/2})\mathbf{\bar{P}}_{k} = \mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau( \boldsymbol{\bar{\mathfrak{U}}}_{k+1} - \mathbf{D}_{k+1}),
\end{align}
or by introducing the scaling parameter $\kappa^2>0$,
\begin{align}
    (\mathbf{C}+\tau (\kappa^2)^{\alpha/2}(\mathbf{L}/\kappa^2)^{\alpha/2})\mathbf{\bar{P}}_{k} = \mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau ( \boldsymbol{\bar{\mathfrak{U}}}_{k+1} - \mathbf{D}_{k+1}),
\end{align}
where $\mathbf{C}$ has entries $\mathbf{C}_{i,j} = (\psi_h^j,\psi_h^i)_{L_2(\Gamma)}$, $\mathbf{L}^{\alpha/2}$ has entries $\mathfrak{a}(\psi_h^j,\psi_h^i)$, $\mathbf{\bar{P}}^k$ has components $p_j^k$, and $\boldsymbol{\bar{\mathfrak{U}}}^k$ has components $( \bar{u}^{k},\psi_h^i)_{L_2(\Gamma)}$. Applying $(\mathbf{L}/\kappa^2)^{-\alpha/2}$ to both sides yields
\begin{equation}
((\mathbf{L}/\kappa^2)^{-\alpha/2}\mathbf{C}+\tau (\kappa^2)^{\alpha/2}\mathbf{I})\mathbf{\bar{P}}_{k} = (\mathbf{L}/\kappa^2)^{-\alpha/2}(\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau ( \boldsymbol{\bar{\mathfrak{U}}}_{k+1} - \mathbf{D}_{k+1})).
\end{equation}
Following @rSPDE2020, we approximate $(\mathbf{L}/\kappa^2)^{-\alpha/2}$ by $\mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top$ to arrive at
\begin{equation}
\label{eq:scheme2adjoint}
\tag{15}
(\mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top \mathbf{C}+\tau(\kappa^2)^{\alpha/2} \mathbf{I})\mathbf{\bar{P}}_{k} = \mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top(\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau ( \boldsymbol{\bar{\mathfrak{U}}}_{k+1} - \mathbf{D}_{k+1})).
\end{equation}
where
\begin{equation}
\label{eq:PLPRbolinadjoint}
\tag{16}
\mathbf{P}_r = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right)\quad\text{and}\quad \mathbf{P}_\ell = \dfrac{\kappa^{2\beta}}{\texttt{factor}}\mathbf{C}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right),
\end{equation}
and $\texttt{factor} = \dfrac{c_m}{b_{m+1}}$, and
$\{r_{1i}\}_{i=1}^m$ and $\{r_{2j}\}_{j=1}^{m+1}$ are the roots of $q_1(x) =\sum_{i=0}^mc_ix^{i}$ and  $q_2(x)=\sum_{j=0}^{m+1}b_jx^{j}$, respectively. The coefficients  $\{c_i\}_{i=0}^m$ and  $\{b_j\}_{j=0}^{m+1}$ are determined as the best rational approximation $q_1/q_2$ of the function $x^{\alpha/2-1}$ over the interval $J_h: = [\kappa^{2}\lambda_{N_h,h}^{-1}, \kappa^{2}\lambda_{1,h}^{-1}]$, where $\lambda_{1,h}, \lambda_{N_h,h}>0$ are the smallest and the largest eigenvalue of $L_h$, respectively.


For the sake of clarity, we note that the numerical implementation of @rSPDE2020 actually defines $\mathbf{P}_r$ and $\mathbf{P}_\ell$ as
\begin{equation}
\label{eq:PLPRbolin}
\tag{17}
\mathbf{P}_r = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right)\quad\text{and}\quad \mathbf{P}_\ell = \dfrac{\kappa^{2\beta}}{\texttt{factor}}\mathbf{C}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right),
\end{equation}
where $\beta = \alpha/2$ and the scaling factor $(\kappa^2)^{\alpha/2}$ or $\kappa^{2\beta}$ is already incorporated in $\mathbf{P}_\ell$, a convention we adopt in the following. With this under consideration, we can rewrite \eqref{eq:scheme2adjoint} as
\begin{equation}
\tag{18}
(\mathbf{P}_r^\top \mathbf{C}+\tau \mathbf{P}_\ell^\top)\mathbf{\bar{P}}_{k} = \mathbf{P}_r^\top(\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau ( \boldsymbol{\bar{\mathfrak{U}}}_{k+1} - \mathbf{D}_{k+1})),
\label{eq:schemeadjoint}
\end{equation}
where  
\begin{equation}
\mathbf{P}_r^\top = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\quad\text{and}\quad \mathbf{P}_\ell^\top = \dfrac{\kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot \mathbf{C}
\end{equation}
since $\mathbf{L}$ and $\mathbf{C}^{-1}$ are symmetric and the factors in the product commute. Replacing these two into \eqref{eq:schemeadjoint} yields
\begin{equation}
\left(\prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\right)\mathbf{C}\mathbf{\bar{P}}_{k} = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot (\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau ( \boldsymbol{\bar{\mathfrak{U}}}_{k+1} - \mathbf{D}_{k+1})),
\end{equation}
that is,
\begin{equation}
\label{eq:final_schemeadjoint}
\tag{19}
\mathbf{\bar{P}}_{k} = \mathbf{C}^{-1}\left(\prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\right)^{-1} \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot (\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau ( \boldsymbol{\bar{\mathfrak{U}}}_{k+1} - \mathbf{D}_{k+1})).
\end{equation}
Considering the partial fraction decomposition
\begin{equation}
\label{eq:partial_fractionadjoint}
\tag{20}
\dfrac{\prod_{i=1}^m (1-r_{1i}x)}{\prod_{i=1}^m (1-r_{1i}x)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}} \prod_{j=1}^{m+1} (1-r_{2j}x)}=\sum_{k=1}^{m+1} a_k(x-p_k)^{-1} + r,
\end{equation}
scheme \eqref{eq:final_schemeadjoint} can be expressed as
\begin{equation}
\label{eq:final_scheme3adjoint}
\tag{21}
\mathbf{\bar{P}}_{k} = \mathbf{C}^{-1}\left(\sum_{k=1}^{m+1} a_k\left( \dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}-p_k\mathbf{I}\right)^{-1} + r\mathbf{I}\right) (\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau ( \boldsymbol{\bar{\mathfrak{U}}}_{k+1} - \mathbf{D}_{k+1})).
\end{equation}
In practice, since the rational function in \eqref{eq:partial_fractionadjoint} is proper, there is no remainder $r$. Moreover, since $\left( \dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}-p_k\mathbf{I}\right)^{-1}  = \mathbf{C}\left( \dfrac{\mathbf{L}}{\kappa^2}-p_k\mathbf{C}\right)^{-1}$, we have that \eqref{eq:final_scheme3adjoint} can be rewritten as

\begin{equation}
\label{eq:final_schemefinaladjoint}
\tag{21}
\mathbf{\bar{P}}_{k} = \mathbf{R} (\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau ( \boldsymbol{\bar{\mathfrak{U}}}_{k+1} - \mathbf{D}_{k+1})),\quad \mathbf{R} = \left(\sum_{k=1}^{m+1} a_k\left( \dfrac{\mathbf{L}}{\kappa^2}-p_k\mathbf{C}\right)^{-1}\right).
\end{equation}

That is,
\begin{align}
\tag{22}
\label{thenumericalscheme4}
    \begin{cases}
    &\mathbf{\bar{P}}_{k}  = \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau \mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{k+1}-\mathbf{D}_{k+1}),\quad k = N-1,\dots, 0,\\
    &\mathbf{\bar{P}}_{N}  = \mathbf{0},
    \end{cases}
\end{align}

Observe that the only difference between \eqref{thenumericalscheme3} and \eqref{thenumericalscheme4} is in the right-hand side, where in \eqref{thenumericalscheme4} we have $\mathbf{C}\mathbf{\bar{U}}_{k+1}-\mathbf{D}_{k+1}$ instead of $\mathbf{C}\mathbf{\bar{U}}_{k}-\mathbf{D}_{k}$.


## Fixed point iteration algorithm for the optimal control problem

Let $(h_\star, \tau_\star)$ denote the integration space–time mesh parameters, with $\tau_\star = T/N_\star$ and nodes $(s_i^\star, t_\ell^\star)$ for $i = 1,\dots,N_{h_\star}$ and $\ell = 0,\dots,N_\star$. The computation space-time mesh is defined analogously by $(h, \tau)$, with $\tau = T/N$ and nodes $(s_j, t_k)$ for $j = 1,\dots,N_{h}$ and $k = 0,\dots,N$. Let $\boldsymbol{\mathfrak{F}},\boldsymbol{\mathfrak{D}}\in \mathbb{R}^{N_{h_\star}\times (N+1)}$ denote the evaluations of $f$ and $u_d$ on a mixed mesh (that uses the spatial nodes of the integration mesh and the temporal nodes of the computation mesh), respectively, i.e., $\boldsymbol{\mathfrak{F}}_{i,k} = f(s_i^\star, t_k)$ and $\boldsymbol{\mathfrak{D}}_{i,k} = u_d(s_i^\star, t_k)$. With this notation at hand, we now introduce the next algorithm.

- **1. Initialization:** Initialize $\mathbf{\bar{Z}}\in \mathbb{R}^{N_{h}\times (N+1)}$ on the computation mesh and approximate $\boldsymbol{\bar{\mathcal{Z}}}\in \mathbb{R}^{N_{h}\times (N+1)}$, with entries $\boldsymbol{\bar{\mathcal{Z}}}_{j,k} =(\bar{z}^{k},\psi^j_h)_{L_2(\Gamma)}$, by $\boldsymbol{\bar{\mathcal{Z}}}\approx\boldsymbol{\Psi}^\top \mathbf{C}^{\star} \boldsymbol{\Psi} \mathbf{\bar{Z}} = \mathbf{C} \mathbf{\bar{Z}}$, where the last equality is thanks to the nestedness of the spatial meshes. Similarly, approximate $\mathbf{F}\in \mathbb{R}^{N_{h}\times (N+1)}$, with entries $\mathbf{F}_{j,k} =(f^{k},\psi^j_h)_{L_2(\Gamma)}$, by $\mathbf{F} \approx \boldsymbol{\Psi}^\top \mathbf{C}^{\star} \boldsymbol{\mathfrak{F}}$.

- **2. State solve:** Given $\boldsymbol{\bar{\mathcal{Z}}}$, $\mathbf{F}$, and $\mathbf{\bar{U}}_{0}\in \mathbb{R}^{N_{h}}$ with components $u_0(s_j)$, compute $\mathbf{\bar{U}}\in \mathbb{R}^{N_{h}\times (N+1)}$, the numerical solution corresponding to $\bar{u}$ in \eqref{eq:maineqoptimal}, with the scheme

\begin{align}
\label{statesolve}
\tag{FS}
    \begin{cases}
    \mathbf{\bar{U}}_{k+1}  &= \mathbf{R}\mathbf{C}\mathbf{\bar{U}}_k+\tau \mathbf{R}(\mathbf{F}_{k+1}+\mathbf{C}\mathbf{\bar{Z}}_{k+1})\\
    & = (\mathbf{R}\mathbf{C})^{k+1}\mathbf{\bar{U}}_0
+ \tau \sum_{i=0}^{k} (\mathbf{R}\mathbf{C})^{k-i}\mathbf{R}(\mathbf{F}_{i+1}+\mathbf{C} \mathbf{\bar{Z}}_{i+1}), \quad k = 0,\dots, N-1,\\
    \mathbf{\bar{U}}_{0}  &= [u_0(s_1), \dots, u_0(s_{N_h})]^\top,
    \end{cases}
\end{align}

- **3. Adjoint solve:** Given $\mathbf{\bar{U}}$ from before, approximate $\boldsymbol{\bar{\mathfrak{U}}}\in \mathbb{R}^{N_{h}\times (N+1)}$, with entries $\boldsymbol{\bar{\mathfrak{U}}}_{j,k} =(\bar{u}^{k},\psi^j_h)_{L_2(\Gamma)}$, by $\boldsymbol{\bar{\mathfrak{U}}} \approx \boldsymbol{\Psi}^\top \mathbf{C}^{\star} \boldsymbol{\Psi} \mathbf{\bar{U}}=\mathbf{C}\mathbf{\bar{U}}$. Similarly, approximate $\mathbf{D}\in \mathbb{R}^{N_{h}\times (N+1)}$, with entries $\mathbf{D}_{j,k} =(u_d^{k},\psi^j_h)_{L_2(\Gamma)}$, by $\mathbf{D} \approx \boldsymbol{\Psi}^\top \mathbf{C}^{\star} \boldsymbol{\mathfrak{D}}$. Compute $\mathbf{\bar{P}}\in \mathbb{R}^{N_{h}\times (N+1)}$, the numerical solution corresponding to $\bar{p}$ in \eqref{eq:adjointeq}, with the scheme

\begin{align}
\label{adjointsolve}
\tag{BS}
    \begin{cases}
    \mathbf{\bar{P}}_{k}  & = \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau \mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{k+1}-\mathbf{D}_{k+1})\\
    &=  \tau \sum_{i=0}^{N-k-1} (\mathbf{R}\mathbf{C})^{i}\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{k+i+1}-\mathbf{D}_{k+i+1}), \quad k = N-1,\dots, 0,\\
    \mathbf{\bar{P}}_{N} &  = \mathbf{0},
    \end{cases}
\end{align}

- **4. Control update:** Given $\mathbf{\bar{P}}$ from before and matrices $\mathbf{A},\mathbf{B}\in \mathbb{R}^{N_{h}\times (N+1)}$ with constant entries  $\mathbf{A}_{j,k} = a$ and $\mathbf{B}_{j,k} = b$, respectively, update 
    \begin{align}
    \label{entrywisez}
        \mathbf{\bar{Z}} = \max\{\mathbf{A},\min\{\mathbf{B},-\mathbf{\bar{P}}/\mu\}\}\quad \text{(entry-wise)}.
    \end{align}

- **5. Iteration:** Repeat steps 2--4 until convergence.
    
## Convergence analysis of the FBSM iteration {#conv_FBSM}

For illustration purposes, along with the general case $N$, we consider the particular case $N=3$. From \eqref{statesolve} and \eqref{adjointsolve}, we have

\begin{align}
\mathbf{\bar{U}}_{0} & = \mathbf{\bar{U}}_{0} \\
& \\
\mathbf{\bar{U}}_{1} & =  (\mathbf{R}\mathbf{C})^1\mathbf{\bar{U}}_0+\tau \mathbf{R}(\mathbf{F}_{1}+\mathbf{C}\mathbf{\bar{Z}}_{1})\\
& \\
\mathbf{\bar{U}}_{2} & =  \mathbf{R}\mathbf{C}\mathbf{\bar{U}}_1+\tau \mathbf{R}(\mathbf{F}_{2}+\mathbf{C}\mathbf{\bar{Z}}_{2})\\
& = \mathbf{R}\mathbf{C}((\mathbf{R}\mathbf{C})^1\mathbf{\bar{U}}_0+\tau \mathbf{R}(\mathbf{F}_{1}+\mathbf{C}\mathbf{\bar{Z}}_{1}))+\tau \mathbf{R}(\mathbf{F}_{2}+\mathbf{C}\mathbf{\bar{Z}}_{2})\\
& = (\mathbf{R}\mathbf{C})^2\mathbf{\bar{U}}_0+\tau (\mathbf{R}\mathbf{C})^1\mathbf{R}(\mathbf{F}_{1}+\mathbf{C}\mathbf{\bar{Z}}_{1})+\tau \mathbf{R}(\mathbf{F}_{2}+\mathbf{C}\mathbf{\bar{Z}}_{2})\\
& \\
\mathbf{\bar{U}}_{3} & =  \mathbf{R}\mathbf{C}\mathbf{\bar{U}}_2+\tau \mathbf{R}(\mathbf{F}_{3}+\mathbf{C}\mathbf{\bar{Z}}_{3})\\
& =  \mathbf{R}\mathbf{C}((\mathbf{R}\mathbf{C})^2\mathbf{\bar{U}}_0+\tau (\mathbf{R}\mathbf{C})^1\mathbf{R}(\mathbf{F}_{1}+\mathbf{C}\mathbf{\bar{Z}}_{1})+\tau \mathbf{R}(\mathbf{F}_{2}+\mathbf{C}\mathbf{\bar{Z}}_{2}))+\tau \mathbf{R}(\mathbf{F}_{3}+\mathbf{C}\mathbf{\bar{Z}}_{3})\\
& =  (\mathbf{R}\mathbf{C})^3\mathbf{\bar{U}}_0+\tau (\mathbf{R}\mathbf{C})^2\mathbf{R}(\mathbf{F}_{1}+\mathbf{C}\mathbf{\bar{Z}}_{1})+\tau (\mathbf{R}\mathbf{C})^1\mathbf{R}(\mathbf{F}_{2}+\mathbf{C}\mathbf{\bar{Z}}_{2})+\tau \mathbf{R}(\mathbf{F}_{3}+\mathbf{C}\mathbf{\bar{Z}}_{3})
\end{align}
and
\begin{align}
\mathbf{\bar{P}}_{3} &= \mathbf{0}, \\
&\\
\mathbf{\bar{P}}_{2} &= \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{3} + \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{3}-\mathbf{D}_{3})\\
&= \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{3}-\mathbf{D}_{3}), \\
&\\
\mathbf{\bar{P}}_{1} &= \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{2} + \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{2}-\mathbf{D}_{2}) \\
&= \tau(\mathbf{R}\mathbf{C})^1\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{3}-\mathbf{D}_{3})
+ \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{2}-\mathbf{D}_{2}), \\
&\\
\mathbf{\bar{P}}_{0} &= \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{1} + \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{1}-\mathbf{D}_{1}) \\
&= \tau(\mathbf{R}\mathbf{C})^{2}\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{3}-\mathbf{D}_{3})
+ \tau(\mathbf{R}\mathbf{C})^1\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{2}-\mathbf{D}_{2})
+ \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{1}-\mathbf{D}_{1}).
\end{align}



In block-matrix notation
\begin{align}
\begin{bmatrix}
\mathbf{\bar{U}}_{0} \\[1mm]
\mathbf{\bar{U}}_{1} \\[1mm]
\mathbf{\bar{U}}_{2} \\[1mm]
\mathbf{\bar{U}}_{3}
\end{bmatrix}
= 
\begin{bmatrix}
\mathbf{\bar{U}}_{0} \\[1mm]
(\mathbf{R}\mathbf{C})^1\mathbf{\bar{U}}_{0} \\[1mm]
(\mathbf{R}\mathbf{C})^2\mathbf{\bar{U}}_{0} \\[1mm]
(\mathbf{R}\mathbf{C})^3\mathbf{\bar{U}}_{0}
\end{bmatrix}
+\tau
\begin{bmatrix}
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{0}  \\[1mm]
\mathbf{0} & \mathbf{I} & \mathbf{0} & \mathbf{0}  \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^1 & \mathbf{I} & \mathbf{0}  \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^2 & (\mathbf{R}\mathbf{C})^1 & \mathbf{I} 
\end{bmatrix}
\begin{bmatrix}
\mathbf{R}(\mathbf{\bar{F}}_{0}+\mathbf{C}\mathbf{\bar{Z}}_{0}) \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{1}+\mathbf{C}\mathbf{\bar{Z}}_{1}) \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{2}+\mathbf{C}\mathbf{\bar{Z}}_{2}) \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{3}+\mathbf{C}\mathbf{\bar{Z}}_{3})
\end{bmatrix}
\end{align}
and
\begin{align}
\begin{bmatrix}
\mathbf{\bar{P}}_{0} \\[1mm]
\mathbf{\bar{P}}_{1} \\[1mm]
\mathbf{\bar{P}}_{2} \\[1mm]
\mathbf{\bar{P}}_{3}
\end{bmatrix}
=
\tau
\begin{bmatrix}
\mathbf{0} & \mathbf{I} & (\mathbf{R}\mathbf{C}) & (\mathbf{R}\mathbf{C})^{2} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{I} & (\mathbf{R}\mathbf{C}) \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{I} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{0}
\end{bmatrix}
\begin{bmatrix}
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{0}-\mathbf{D}_{0}) \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{1}-\mathbf{D}_{1}) \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{2}-\mathbf{D}_{2}) \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{3}-\mathbf{D}_{3})
\end{bmatrix}.
\end{align}



In general,
\begin{align}
\begin{bmatrix}
\mathbf{\bar{U}}_{0} \\[1mm]
\mathbf{\bar{U}}_{1} \\[1mm]
\mathbf{\bar{U}}_{2} \\[1mm]
\vdots \\[1mm]
\mathbf{\bar{U}}_{N}
\end{bmatrix}
&=
\begin{bmatrix}
\mathbf{\bar{U}}_{0} \\[1mm]
(\mathbf{R}\mathbf{C})^1\mathbf{\bar{U}}_{0} \\[1mm]
(\mathbf{R}\mathbf{C})^2\mathbf{\bar{U}}_{0} \\[1mm]
\vdots \\[1mm]
(\mathbf{R}\mathbf{C})^N\mathbf{\bar{U}}_{0}
\end{bmatrix}

+\tau
\begin{bmatrix}
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & \mathbf{I} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^1 & \mathbf{I} & \cdots & \mathbf{0} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^{N-1} & (\mathbf{R}\mathbf{C})^{N-2} & \cdots & \mathbf{I}
\end{bmatrix}
\begin{bmatrix}
\mathbf{R}(\mathbf{\bar{F}}_{0}+\mathbf{C}\mathbf{\bar{Z}}_{0}) \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{1}+\mathbf{C}\mathbf{\bar{Z}}_{1}) \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{2}+\mathbf{C}\mathbf{\bar{Z}}_{2}) \\[1mm]
\vdots \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{N}+\mathbf{C}\mathbf{\bar{Z}}_{N})
\end{bmatrix}.
\end{align}
and \begin{align}
\begin{bmatrix}
\mathbf{\bar{P}}_{0} \\[1mm]
\mathbf{\bar{P}}_{1} \\[1mm]
\vdots \\[1mm]
\mathbf{\bar{P}}_{N-1} \\[1mm]
\mathbf{\bar{P}}_{N}
\end{bmatrix}
=
\tau
\begin{bmatrix}
\mathbf{0} & \mathbf{I} & (\mathbf{R}\mathbf{C}) & \cdots & (\mathbf{R}\mathbf{C})^{N-1} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{I} & \cdots & (\mathbf{R}\mathbf{C})^{N-2} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{I} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0}
\end{bmatrix}
\begin{bmatrix}
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{0}-\mathbf{D}_{0}) \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{1}-\mathbf{D}_{1}) \\[1mm]
\vdots \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{N-1}-\mathbf{D}_{N-1}) \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{N}-\mathbf{D}_{N})
\end{bmatrix}.
\end{align}
By defining stacked vectors of length $(N+1)N_h$,
\begin{align*}
 \mathbf{\bar{u}} = \text{vec}(\mathbf{\bar{U}}),\quad 
\mathbf{\bar{p}} = \text{vec}(\mathbf{\bar{P}}),\quad
\mathbf{\bar{z}} = \text{vec}(\mathbf{\bar{Z}}),\quad \mathbf{f} = \text{vec}(\mathbf{F}),\quad
\mathbf{d} =\text{vec}(\mathbf{D}),
\end{align*}
\begin{align*}
    \mathbf{a} =\text{vec}(\mathbf{A}),\quad
\mathbf{b} = \text{vec}(\mathbf{B}),\quad\mathbf{h}= [\mathbf{\bar{U}}_{0}^\top ((\mathbf{R}\mathbf{C})^{1}\mathbf{\bar{U}}_{0})^\top ((\mathbf{R}\mathbf{C})^{2} \mathbf{\bar{U}}_{0})^\top\dots ((\mathbf{R}\mathbf{C})^{N}\mathbf{\bar{U}}_{0})^\top]^\top,
\end{align*}
and block matrices of dimension $(N+1)N_h\times (N+1)N_h$,
\begin{align*}
\mathbf{M} = \mathbf{I}_{N+1}\otimes  \mathbf{C},\quad \mathbf{\hat{J}}_{N+1} = \mathbf{J}_{N+1}\otimes \mathbf{I}_{N_h},\quad \mathbf{\hat{R}} = \mathbf{I}_{N+1}\otimes  \mathbf{R} ,
\end{align*}
\begin{align}
\mathbf{S}_{\mathrm{F}}
= \tau
\begin{bmatrix}
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & \mathbf{I} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^1 & \mathbf{I} & \cdots & \mathbf{0} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^{N-1} & (\mathbf{R}\mathbf{C})^{N-2} & \cdots & \mathbf{I}
\end{bmatrix}\quad \text{and}\quad
\mathbf{S}_{\mathrm{B}}
= \tau
\begin{bmatrix}
\mathbf{0} & \mathbf{I} & (\mathbf{R}\mathbf{C}) & \cdots & (\mathbf{R}\mathbf{C})^{N-1} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{I} & \cdots & (\mathbf{R}\mathbf{C})^{N-2} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{I} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0}
\end{bmatrix},
\end{align}
then the vectorized form 
\begin{align}
\label{iterationvectorized}
    \begin{cases}
    \mathbf{\bar{u}} &= \mathbf{h} + \mathbf{S}_{\mathrm{F}}\mathbf{\hat{R}}(\mathbf{f}+\mathbf{M}\mathbf{\bar{z}})\\
    \mathbf{\bar{p}} & = \mathbf{S}_{\mathrm{B}}\mathbf{\hat{R}}(\mathbf{M}\mathbf{\bar{u}} - \mathbf{d})\\
    \mathbf{\bar{z}} & = \max\{\mathbf{a}, \min\{\mathbf{b}, -\mathbf{\bar{p}}/\mu\}\} \quad \text{(componentwise)}.
     \end{cases}
\end{align}
We need to estimate $\gamma = (1/\mu)\|\boldsymbol{\mathfrak{L}}\|_{\mathbf{C}_{N+1}}$ where $\boldsymbol{\mathfrak{L}} = \mathbf{S}_{\mathrm{B}}\mathbf{\hat{R}}\mathbf{M}\mathbf{S}_{\mathrm{F}}\mathbf{\hat{R}}\mathbf{M}$. We have
\begin{align}
\mathbf{S}_{\mathrm{F}}\mathbf{\hat{R}}\mathbf{M}
= \tau
\begin{bmatrix}
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & \mathbf{R}\mathbf{C} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^2 & \mathbf{R}\mathbf{C} & \cdots & \mathbf{0} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^{N} & (\mathbf{R}\mathbf{C})^{N-1} & \cdots & \mathbf{R}\mathbf{C}
\end{bmatrix}\quad \text{and}\quad
\mathbf{S}_{\mathrm{B}}\mathbf{\hat{R}}\mathbf{M}
= \tau
\begin{bmatrix}
\mathbf{0} & \mathbf{R}\mathbf{C} & (\mathbf{R}\mathbf{C})^2 & \cdots & (\mathbf{R}\mathbf{C})^{N} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{R}\mathbf{C} & \cdots & (\mathbf{R}\mathbf{C})^{N-1} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{R}\mathbf{C} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0}
\end{bmatrix}
\end{align}
If $N=3$, then
\begin{align}
\boldsymbol{\mathfrak{L}} = \mathbf{S}_{\mathrm{B}}\mathbf{\hat{R}}\mathbf{M}\mathbf{S}_{\mathrm{F}}\mathbf{\hat{R}}\mathbf{M} 
&= \tau^2
\begin{bmatrix}
\mathbf{0} & \mathbf{R}\mathbf{C} & (\mathbf{R}\mathbf{C})^2 & (\mathbf{R}\mathbf{C})^{3} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{R}\mathbf{C} & (\mathbf{R}\mathbf{C})^2 \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{R}\mathbf{C} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{0}
\end{bmatrix}
\begin{bmatrix}
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{0}  \\[1mm]
\mathbf{0} & \mathbf{R}\mathbf{C} & \mathbf{0} & \mathbf{0}  \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^2 & \mathbf{R}\mathbf{C} & \mathbf{0}  \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^3 & (\mathbf{R}\mathbf{C})^2 & \mathbf{R}\mathbf{C}
\end{bmatrix}\\
&\\
& = \tau^2
\begin{bmatrix}
\mathbf{0} & (\mathbf{R}\mathbf{C})^2 + (\mathbf{R}\mathbf{C})^4 + (\mathbf{R}\mathbf{C})^6 
           & (\mathbf{R}\mathbf{C})^3 + (\mathbf{R}\mathbf{C})^5 
           & (\mathbf{R}\mathbf{C})^4 \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^3 + (\mathbf{R}\mathbf{C})^5 
           & (\mathbf{R}\mathbf{C})^2 + (\mathbf{R}\mathbf{C})^4 
           & (\mathbf{R}\mathbf{C})^3 \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^4 
           & (\mathbf{R}\mathbf{C})^3 
           & (\mathbf{R}\mathbf{C})^2 \\[1mm]
\mathbf{0} & \mathbf{0} 
           & \mathbf{0} 
           & \mathbf{0}
\end{bmatrix}
\end{align}

Let 
\begin{align}
\boldsymbol{\hat{\mathfrak{L}}} =  
\begin{bmatrix}
\sum_{k=1}^{N}(\mathbf{R}\mathbf{C})^{2k+(N-N)}   & \cdots  & \sum_{k=1}^{3}(\mathbf{R}\mathbf{C})^{2k+(N-3)} & \sum_{k=1}^{2}(\mathbf{R}\mathbf{C})^{2k+(N-2)}  & \sum_{k=1}^{1}(\mathbf{R}\mathbf{C})^{2k+(N-1)}  \\[1mm]
\vdots & \ddots & \vdots & \vdots & \vdots  \\[1mm]
\sum_{k=1}^{3}(\mathbf{R}\mathbf{C})^{2k+(N-3)} & \cdots & \sum_{k=1}^{3}(\mathbf{R}\mathbf{C})^{2k}  & \sum_{k=1}^{2}(\mathbf{R}\mathbf{C})^{2k+1}  & \sum_{k=1}^{1}(\mathbf{R}\mathbf{C})^{2k+2} \\[1mm]
\sum_{k=1}^{2}(\mathbf{R}\mathbf{C})^{2k+(N-2)}    & \cdots & \sum_{k=1}^{2}(\mathbf{R}\mathbf{C})^{2k+1}  & \sum_{k=1}^{2}(\mathbf{R}\mathbf{C})^{2k}  & \sum_{k=1}^{1}(\mathbf{R}\mathbf{C})^{2k+1} \\[1mm]
\sum_{k=1}^{1}(\mathbf{R}\mathbf{C})^{2k+(N-1)}   & \cdots  & \sum_{k=1}^{1}(\mathbf{R}\mathbf{C})^{2k+2}   & \sum_{k=1}^{1}(\mathbf{R}\mathbf{C})^{2k+1}  & \sum_{k=1}^{1}(\mathbf{R}\mathbf{C})^{2k} 
\end{bmatrix}
\end{align}

Then we can write

\begin{align}
\boldsymbol{\mathfrak{L}} =  \tau^2
\begin{bmatrix}
 \mathbf{0} & \boldsymbol{\hat{\mathfrak{L}}} \\[1mm]
 \mathbf{0} & \mathbf{0} 
\end{bmatrix}
\end{align}




Let $\boldsymbol{\mathfrak{B}} = \mathbf{C}_{N+1}^{\frac{1}{2}}\boldsymbol{\mathfrak{L}}\mathbf{C}_{N+1}^{-\frac{1}{2}}$. Note for example that
\begin{align*}
 \mathbf{C}^{\frac{1}{2}}(\mathbf{R}\mathbf{C})^{3}\mathbf{C}^{-\frac{1}{2}}  =  \mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}\mathbf{R}\mathbf{C}\mathbf{R}\mathbf{C}\mathbf{C}^{-\frac{1}{2}} =  (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^3.
\end{align*}
Therefore
\begin{align}
\boldsymbol{\mathfrak{B}}  = \tau^2
\begin{bmatrix}
\mathbf{0} & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^2 + (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^4 + (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^6 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^3 + (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^5 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^4 \\[1mm]
\mathbf{0} & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^3 + (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^5 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^2 + (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^4 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^3 \\[1mm]
\mathbf{0} & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^4 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^3 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^2 \\[1mm]
\mathbf{0} & \mathbf{0} 
           & \mathbf{0} 
           & \mathbf{0}
\end{bmatrix}
\end{align}
In general,
\begin{align}
\boldsymbol{\mathfrak{B}} =  \tau^2
\begin{bmatrix}
 \mathbf{0} & \boldsymbol{\hat{\mathfrak{B}}} \\[1mm]
 \mathbf{0} & \mathbf{0} 
\end{bmatrix},\quad 
\boldsymbol{\hat{\mathfrak{B}}} = \mathbf{C}_{N}^{\frac{1}{2}}\boldsymbol{\hat{\mathfrak{L}}}\mathbf{C}_{N}^{-\frac{1}{2}} = 
 \begin{bmatrix}
\sum_{k=1}^{N}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+(N-N)}   & \cdots  & \sum_{k=1}^{3}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+(N-3)} & \sum_{k=1}^{2}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+(N-2)}  & \sum_{k=1}^{1}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+(N-1)}  \\[1mm]
\vdots & \ddots & \vdots & \vdots & \vdots  \\[1mm]
\sum_{k=1}^{3}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+(N-3)} & \cdots & \sum_{k=1}^{3}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k}  & \sum_{k=1}^{2}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+1}  & \sum_{k=1}^{1}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+2} \\[1mm]
\sum_{k=1}^{2}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+(N-2)}    & \cdots & \sum_{k=1}^{2}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+1}  & \sum_{k=1}^{2}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k}  & \sum_{k=1}^{1}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+1} \\[1mm]
\sum_{k=1}^{1}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+(N-1)}   & \cdots  & \sum_{k=1}^{1}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+2}   & \sum_{k=1}^{1}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k+1}  & \sum_{k=1}^{1}(\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^{2k} 
\end{bmatrix}
\end{align}



By letting $\mathbf{\Omega} = \mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}}$, we can write
\begin{align}
\label{matrixB}
\tag{23}
\boldsymbol{\mathfrak{B}}  = \tau^2
\begin{bmatrix}
\mathbf{0} & \mathbf{\Omega}^2 + \mathbf{\Omega}^4 + \mathbf{\Omega}^6 
           & \mathbf{\Omega}^3 + \mathbf{\Omega}^5 
           & \mathbf{\Omega}^4 \\[1mm]
\mathbf{0} & \mathbf{\Omega}^3 + \mathbf{\Omega}^5 
           & \mathbf{\Omega}^2 + \mathbf{\Omega}^4 
           & \mathbf{\Omega}^3 \\[1mm]
\mathbf{0} & \mathbf{\Omega}^4 
           & \mathbf{\Omega}^3 
           & \mathbf{\Omega}^2 \\[1mm]
\mathbf{0} & \mathbf{0} 
           & \mathbf{0} 
           & \mathbf{0}
\end{bmatrix} = 
\tau^2
\begin{bmatrix}
 \mathbf{0} & \boldsymbol{\hat{\mathfrak{B}}} \\[1mm]
 \mathbf{0} & \mathbf{0} 
\end{bmatrix},\quad
\boldsymbol{\hat{\mathfrak{B}}} = \begin{bmatrix}
 \mathbf{\Omega}^2 + \mathbf{\Omega}^4 + \mathbf{\Omega}^6 
           & \mathbf{\Omega}^3 + \mathbf{\Omega}^5 
           & \mathbf{\Omega}^4 \\[1mm]
 \mathbf{\Omega}^3 + \mathbf{\Omega}^5 
           & \mathbf{\Omega}^2 + \mathbf{\Omega}^4 
           & \mathbf{\Omega}^3 \\[1mm]
 \mathbf{\Omega}^4 
           & \mathbf{\Omega}^3 
           & \mathbf{\Omega}^2 
\end{bmatrix}
\end{align}
In general,
\begin{align}
\boldsymbol{\mathfrak{B}} =  \tau^2
\begin{bmatrix}
 \mathbf{0} & \boldsymbol{\hat{\mathfrak{B}}} \\[1mm]
 \mathbf{0} & \mathbf{0} 
\end{bmatrix},\quad 
\boldsymbol{\hat{\mathfrak{B}}} = \mathbf{C}_{N}^{\frac{1}{2}}\boldsymbol{\hat{\mathfrak{L}}}\mathbf{C}_{N}^{-\frac{1}{2}} = 
 \begin{bmatrix}
\sum_{k=1}^{N}\mathbf{\Omega}^{2k+(N-N)}   & \cdots  & \sum_{k=1}^{3}\mathbf{\Omega}^{2k+(N-3)} & \sum_{k=1}^{2}\mathbf{\Omega}^{2k+(N-2)}  & \sum_{k=1}^{1}\mathbf{\Omega}^{2k+(N-1)}  \\[1mm]
\vdots & \ddots & \vdots & \vdots & \vdots  \\[1mm]
\sum_{k=1}^{3}\mathbf{\Omega}^{2k+(N-3)} & \cdots & \sum_{k=1}^{3}\mathbf{\Omega}^{2k}  & \sum_{k=1}^{2}\mathbf{\Omega}^{2k+1}  & \sum_{k=1}^{1}\mathbf{\Omega}^{2k+2} \\[1mm]
\sum_{k=1}^{2}\mathbf{\Omega}^{2k+(N-2)}    & \cdots & \sum_{k=1}^{2}\mathbf{\Omega}^{2k+1}  & \sum_{k=1}^{2}\mathbf{\Omega}^{2k}  & \sum_{k=1}^{1}\mathbf{\Omega}^{2k+1} \\[1mm]
\sum_{k=1}^{1}\mathbf{\Omega}^{2k+(N-1)}   & \cdots  & \sum_{k=1}^{1}\mathbf{\Omega}^{2k+2}   & \sum_{k=1}^{1}\mathbf{\Omega}^{2k+1}  & \sum_{k=1}^{1}\mathbf{\Omega}^{2k} 
\end{bmatrix}
\end{align}

Note that
\begin{align}
\boldsymbol{\mathfrak{B}} \mathbf{\hat{J}}_{N+1} = \begin{bmatrix}
 \boldsymbol{\hat{\mathfrak{B}}}\mathbf{\hat{J}}_{N} & \mathbf{0} \\[1mm]
 \mathbf{0} & \mathbf{0} 
\end{bmatrix}
\end{align}

Hence

\begin{align}
\label{chain1}
    \|\boldsymbol{\mathfrak{L}}\|_{\mathbf{C}_{N+1}}  = \|\boldsymbol{\mathfrak{B}}\|_2= \|\boldsymbol{\mathfrak{B}} \mathbf{\hat{J}}_{N+1} \|_2= \tau^2  \| \begin{bmatrix}
 \boldsymbol{\hat{\mathfrak{B}}}\mathbf{\hat{J}}_{N} & \mathbf{0} \\[1mm]
 \mathbf{0} & \mathbf{0} 
\end{bmatrix}\|_2  = \tau^2  \|\boldsymbol{\hat{\mathfrak{B}}}\mathbf{\hat{J}}_{N} \|_2 = \tau^2  \|\boldsymbol{\hat{\mathfrak{B}}}\|_2
\end{align}


We have that $\mathbf{Q}^\top\boldsymbol{\Omega}^k\mathbf{Q}  =  \boldsymbol{\Delta}^k$. Let $\mathbf{V} = \mathbf{I}_{N}\otimes  \mathbf{Q}$. By ,  $\mathbf{V}$ is orthogonal and
\begin{align*}
\mathbf{V}^\top \boldsymbol{\hat{\mathfrak{B}}} \mathbf{V} = 
\begin{bmatrix}
 \mathbf{\Delta}^2 + \mathbf{\Delta}^4 + \mathbf{\Delta}^6 
           & \mathbf{\Delta}^3 + \mathbf{\Delta}^5 
           & \mathbf{\Delta}^4 \\[1mm]
 \mathbf{\Delta}^3 + \mathbf{\Delta}^5 
           & \mathbf{\Delta}^2 + \mathbf{\Delta}^4 
           & \mathbf{\Delta}^3 \\[1mm]
 \mathbf{\Delta}^4 
           & \mathbf{\Delta}^3 
           & \mathbf{\Delta}^2 
\end{bmatrix}
\end{align*}
In general,
\begin{align*}
\mathbf{V}^\top \boldsymbol{\hat{\mathfrak{B}}} \mathbf{V} = 
 \begin{bmatrix}
\sum_{k=1}^{N}\mathbf{\Delta}^{2k+(N-N)}   & \cdots  & \sum_{k=1}^{3}\mathbf{\Delta}^{2k+(N-3)} & \sum_{k=1}^{2}\mathbf{\Delta}^{2k+(N-2)}  & \sum_{k=1}^{1}\mathbf{\Delta}^{2k+(N-1)}  \\[1mm]
\vdots & \ddots & \vdots & \vdots & \vdots  \\[1mm]
\sum_{k=1}^{3}\mathbf{\Delta}^{2k+(N-3)} & \cdots & \sum_{k=1}^{3}\mathbf{\Delta}^{2k}  & \sum_{k=1}^{2}\mathbf{\Delta}^{2k+1}  & \sum_{k=1}^{1}\mathbf{\Delta}^{2k+2} \\[1mm]
\sum_{k=1}^{2}\mathbf{\Delta}^{2k+(N-2)}    & \cdots & \sum_{k=1}^{2}\mathbf{\Delta}^{2k+1}  & \sum_{k=1}^{2}\mathbf{\Delta}^{2k}  & \sum_{k=1}^{1}\mathbf{\Delta}^{2k+1} \\[1mm]
\sum_{k=1}^{1}\mathbf{\Delta}^{2k+(N-1)}   & \cdots  & \sum_{k=1}^{1}\mathbf{\Delta}^{2k+2}   & \sum_{k=1}^{1}\mathbf{\Delta}^{2k+1}  & \sum_{k=1}^{1}\mathbf{\Delta}^{2k} 
\end{bmatrix}
\end{align*}
For example, if $\mathbf{\Delta} = \text{diag}(\mu_1, \mu_2,\mu_3)$, then
\begin{align}
\mathbf{V}^\top \boldsymbol{\hat{\mathfrak{B}}} \mathbf{V} =
\begin{bmatrix}
\mu_1^2+\mu_1^4+\mu_1^6 & 0 & 0 & \mu_1^3+\mu_1^5 & 0 & 0 & \mu_1^4 & 0 & 0 \\[1mm]
0 & \mu_2^2+\mu_2^4+\mu_2^6 & 0 & 0 & \mu_2^3+\mu_2^5 & 0 & 0 & \mu_2^4 & 0 \\[1mm]
0 & 0 & \mu_3^2+\mu_3^4+\mu_3^6 & 0 & 0 & \mu_3^3+\mu_3^5 & 0 & 0 & \mu_3^4 \\[2mm]
\mu_1^3+\mu_1^5 & 0 & 0 & \mu_1^2+\mu_1^4 & 0 & 0 & \mu_1^3 & 0 & 0 \\[1mm]
0 & \mu_2^3+\mu_2^5 & 0 & 0 & \mu_2^2+\mu_2^4 & 0 & 0 & \mu_2^3 & 0 \\[1mm]
0 & 0 & \mu_3^3+\mu_3^5 & 0 & 0 & \mu_3^2+\mu_3^4 & 0 & 0 & \mu_3^3 \\[2mm]
\mu_1^4 & 0 & 0 & \mu_1^3 & 0 & 0 & \mu_1^2 & 0 & 0 \\[1mm]
0 & \mu_2^4 & 0 & 0 & \mu_2^3 & 0 & 0 & \mu_2^2 & 0 \\[1mm]
0 & 0 & \mu_3^4 & 0 & 0 & \mu_3^3 & 0 & 0 & \mu_3^2
\end{bmatrix}
\end{align}
If 
\begin{align}
\mathbf{P} =
\begin{bmatrix}
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\[1mm]
0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\[1mm]
0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
\end{bmatrix}
\end{align}
then
\begin{align}
\mathbf{P}\mathbf{V}^\top \boldsymbol{\hat{\mathfrak{B}}} \mathbf{V}\mathbf{P}^\top = 
\begin{bmatrix}
\mu_1^2+\mu_1^4+\mu_1^6 & \mu_1^3+\mu_1^5 & \mu_1^4 & 0 & 0 & 0 & 0 & 0 & 0 \\[1mm]
\mu_1^3+\mu_1^5 & \mu_1^2+\mu_1^4 & \mu_1^3 & 0 & 0 & 0 & 0 & 0 & 0 \\[1mm]
\mu_1^4 & \mu_1^3 & \mu_1^2 & 0 & 0 & 0 & 0 & 0 & 0 \\[2mm]
0 & 0 & 0 & \mu_2^2+\mu_2^4+\mu_2^6 & \mu_2^3+\mu_2^5 & \mu_2^4 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & \mu_2^3+\mu_2^5 & \mu_2^2+\mu_2^4 & \mu_2^3 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & \mu_2^4 & \mu_2^3 & \mu_2^2 & 0 & 0 & 0 \\[2mm]
0 & 0 & 0 & 0 & 0 & 0 & \mu_3^2+\mu_3^4+\mu_3^6 & \mu_3^3+\mu_3^5 & \mu_3^4 \\[1mm]
0 & 0 & 0 & 0 & 0 & 0 & \mu_3^3+\mu_3^5 & \mu_3^2+\mu_3^4 & \mu_3^3 \\[1mm]
0 & 0 & 0 & 0 & 0 & 0 & \mu_3^4 & \mu_3^3 & \mu_3^2
\end{bmatrix}
\end{align}

The code below shows matrix how to build matrix $\mathbf{V}^\top \boldsymbol{\hat{\mathfrak{B}}} \mathbf{V}$ and $\mathbf{P}$ and verifies that $\mathbf{P}\mathbf{V}^\top \boldsymbol{\hat{\mathfrak{B}}} \mathbf{V}\mathbf{P}^\top$ is block diagonal with blocks $\mathbf{T}(\mu_i)$.

```{r, purl = FALSE}
make_matrix <- function(a, N) {
  v <- a^(0:(N - 1))
  toeplitz(v)
}
build_T_fast <- function(a, N) {
  M <- make_matrix(a, N)
  R <- matrix(0, N, N)
  coef <- (a^2)^(1:N)  # correct power order
  for (k in N:1) {
    R[1:k, 1:k] <- R[1:k, 1:k] + coef[N - k + 1] * M[1:k, 1:k]
  }
  R
}
build_block_matrix <- function(egs, Nh) {
  N <- length(egs)
  # total size
  m <- N * Nh
  AA <- matrix(0, nrow = m, ncol = m)
  for (j in 1:N) {
    rows <- ((j - 1) * Nh + 1):(j * Nh)
    cols <- ((j - 1) * Nh + 1):(j * Nh)
    AA[rows, cols] <- build_T_fast(egs[j], Nh)
  }
  return(AA)
}
build_perm_matrix_general <- function(N, Nh) {
  m <- N * Nh
  # target indices
  perm <- as.vector(sapply(1:Nh, function(i) i + Nh*(0:(N-1))))
  P <- diag(m)[perm, ]
  return(P)
}
Nh <- 3
egs <- c(1,3,5)
P <- build_perm_matrix_general(length(egs), Nh)
P
aux <- build_block_matrix(egs, Nh)
tVhatBV <- t(P) %*% aux %*% P
tVhatBV
PtVhatBVtP <- P %*% tVhatBV %*% t(P)
PtVhatBVtP
```

That is, in general,
\begin{align*}
    \mathbf{P}\mathbf{V}^\top \boldsymbol{\hat{\mathfrak{B}}}\mathbf{V}\mathbf{P}^\top = \text{blkdiag}(\mathbf{T}(\mu_1), \mathbf{T}(\mu_2), \dots, \mathbf{T}(\mu_{N_h})),
\end{align*}
where
\begin{align}
\mathbf{T}(\mu_i) = 
 \begin{bmatrix}
\sum_{k=1}^{N}\mu_i^{2k+(N-N)}   & \cdots  & \sum_{k=1}^{3}\mu_i^{2k+(N-3)} & \sum_{k=1}^{2}\mu_i^{2k+(N-2)}  & \sum_{k=1}^{1}\mu_i^{2k+(N-1)}  \\[1mm]
\vdots & \ddots & \vdots & \vdots & \vdots  \\[1mm]
\sum_{k=1}^{3}\mu_i^{2k+(N-3)} & \cdots & \sum_{k=1}^{3}\mu_i^{2k}  & \sum_{k=1}^{2}\mu_i^{2k+1}  & \sum_{k=1}^{1}\mu_i^{2k+2} \\[1mm]
\sum_{k=1}^{2}\mu_i^{2k+(N-2)}    & \cdots & \sum_{k=1}^{2}\mu_i^{2k+1}  & \sum_{k=1}^{2}\mu_i^{2k}  & \sum_{k=1}^{1}\mu_i^{2k+1} \\[1mm]
\sum_{k=1}^{1}\mu_i^{2k+(N-1)}   & \cdots  & \sum_{k=1}^{1}\mu_i^{2k+2}   & \sum_{k=1}^{1}\mu_i^{2k+1}  & \sum_{k=1}^{1}\mu_i^{2k} 
\end{bmatrix}
\end{align}
Let $\mathbf{T} = \mathbf{T}(\omega)$. That is,
\begin{align}
\label{matrixT}
\tag{24}
\mathbf{T} = \mathbf{T}(\omega) = 
 \begin{bmatrix}
\sum_{k=1}^{N}\omega^{2k+(N-N)}   & \cdots  & \sum_{k=1}^{3}\omega^{2k+(N-3)} & \sum_{k=1}^{2}\omega^{2k+(N-2)}  & \sum_{k=1}^{1}\omega^{2k+(N-1)}  \\[1mm]
\vdots & \ddots & \vdots & \vdots & \vdots  \\[1mm]
\sum_{k=1}^{3}\omega^{2k+(N-3)} & \cdots & \sum_{k=1}^{3}\omega^{2k}  & \sum_{k=1}^{2}\omega^{2k+1}  & \sum_{k=1}^{1}\omega^{2k+2} \\[1mm]
\sum_{k=1}^{2}\omega^{2k+(N-2)}    & \cdots & \sum_{k=1}^{2}\omega^{2k+1}  & \sum_{k=1}^{2}\omega^{2k}  & \sum_{k=1}^{1}\omega^{2k+1} \\[1mm]
\sum_{k=1}^{1}\omega^{2k+(N-1)}   & \cdots  & \sum_{k=1}^{1}\omega^{2k+2}   & \sum_{k=1}^{1}\omega^{2k+1}  & \sum_{k=1}^{1}\omega^{2k} 
\end{bmatrix}
\end{align}
Then
\begin{align}
\label{chain2}
    \|\boldsymbol{\mathfrak{L}}\|_{\mathbf{C}_{N+1}}   = \tau^2
    \|\boldsymbol{\hat{\mathfrak{B}}}\|_2 = \tau^2
    \|\mathbf{P}\mathbf{V}^\top \boldsymbol{\hat{\mathfrak{B}}} \mathbf{V}\mathbf{P}^\top\|_2 = \tau^2 \max_{i = 1,\dots, N_h}\|\mathbf{T}(\mu_i)\|_2= \tau^2\|\mathbf{T}(\omega)\|_2= \tau^2\|\mathbf{T}\|_2.
\end{align}

To see an example where we show that $\|\boldsymbol{\mathfrak{B}}\|_2 = \tau^2 \|\mathbf{T}\|_2$, go to the [this](control_functionality.html#norm_identity) section.

Following with our example, we have that 
\begin{align*}
    \mathbf{T} & =  
\begin{bmatrix}
\omega^2 & \omega^3 & \omega^4  \\[1mm]
\omega^3& \omega^2 & \omega^3  \\[1mm]
\omega^4 & \omega^3 & \omega^2  \\[2mm]
\end{bmatrix}
+
\begin{bmatrix}
\omega^4 & \omega^5 & 0  \\[1mm]
\omega^5 & \omega^4 & 0  \\[1mm]
0 & 0 & 0  \\[2mm]
\end{bmatrix}
+ 
\begin{bmatrix}
\omega^6 & 0 & 0  \\[1mm]
0 & 0 & 0  \\[1mm]
0 & 0 & 0  \\[2mm]
\end{bmatrix}
\end{align*}
or equivalently,
\begin{align*}
    \mathbf{T}& = 
    \omega^2 \begin{bmatrix}
1 & \omega & \omega^2  \\[1mm]
\omega& 1 & \omega  \\[1mm]
\omega^2 & \omega & 1  \\[2mm]
\end{bmatrix}
+
\omega^4\begin{bmatrix}
1 & \omega & 0  \\[1mm]
\omega & 1 & 0  \\[1mm]
0 & 0 & 0  \\[2mm]
\end{bmatrix}
+ 
\omega^6\begin{bmatrix}
1 & 0 & 0  \\[1mm]
0 & 0 & 0  \\[1mm]
0 & 0 & 0  \\[2mm]
\end{bmatrix}
\\
& = \omega^2 
\begin{bmatrix}
1 & 0 & 0  \\[1mm]
0 & 1 & 0  \\[1mm]
0 & 0 & 1  \\[2mm]
\end{bmatrix}
\begin{bmatrix}
1 & \omega & \omega^2  \\[1mm]
\omega& 1 & \omega  \\[1mm]
\omega^2 & \omega & 1  \\[2mm]
\end{bmatrix}
\begin{bmatrix}
1 & 0 & 0  \\[1mm]
0 & 1 & 0  \\[1mm]
0 & 0 & 1  \\[2mm]
\end{bmatrix}
\\
&+
\omega^4
\begin{bmatrix}
1 & 0 & 0  \\[1mm]
0 & 1 & 0  \\[1mm]
0 & 0 & 0  \\[2mm]
\end{bmatrix}
\begin{bmatrix}
1 & \omega & \omega^2  \\[1mm]
\omega& 1 & \omega  \\[1mm]
\omega^2 & \omega & 1  \\[2mm]
\end{bmatrix}
\begin{bmatrix}
1 & 0 & 0  \\[1mm]
0 & 1 & 0  \\[1mm]
0 & 0 & 0  \\[2mm]
\end{bmatrix}\\
&+ 
\omega^6
\begin{bmatrix}
1 & 0 & 0  \\[1mm]
0 & 0 & 0  \\[1mm]
0 & 0 & 0  \\[2mm]
\end{bmatrix}
\begin{bmatrix}
1 & \omega & \omega^2  \\[1mm]
\omega& 1 & \omega  \\[1mm]
\omega^2 & \omega & 1  \\[2mm]
\end{bmatrix}
\begin{bmatrix}
1 & 0 & 0  \\[1mm]
0 & 0 & 0  \\[1mm]
0 & 0 & 0  \\[2mm]
\end{bmatrix}
\end{align*}

The code below builds matrix $\mathbf{T}$ and shows the intermediate steps.

```{r, purl = FALSE}
build_T <- function(a, N){
  H <- make_matrix(a, N) 
  R <- H*0
  for (i in N:1) {
    K_i <- diag(c(rep(1, i), rep(0, N - i)))
    temp <- a^(2*(N - i + 1)) * K_i %*% H %*% K_i
    print(a^(2*(N - i + 1)))
    print(cbind(rep("|", N), K_i, rep("|", N), H, rep("|", N), K_i, rep("|", N)), quote = FALSE)
    R <- R + temp
  }
  return(R)
}

T_aux <- build_T(2, 3)
```


That is, in general,
\begin{align*}
    \mathbf{T} = \sum_{i=1}^N \omega^{2(N-i+1)} \mathbf{K}_i\mathbf{H}\mathbf{K}_i,
\end{align*}
where 
\begin{align*}
\mathbf{H} = \begin{bmatrix}
1 & \omega & \omega^2 & \cdots & \omega^{N-1} \\
\omega & 1 & \omega & \cdots & \omega^{N-2} \\
\omega^2 & \omega & 1 & \cdots & \omega^{N-3} \\
\vdots & \vdots & \vdots & \ddots & \vdots \\
\omega^{N-1} & \omega^{N-2} & \omega^{N-3} & \cdots & 1
\end{bmatrix},\quad 
\mathbf{K}_i = 
\mathrm{diag}(\underbrace{1, \dots, 1}_{i}, \underbrace{0, \dots, 0}_{N-i}).
\end{align*}
and $\mathbf{K}_i$ satisfies $\|\mathbf{K}_i\|_2 = 1$ for all $i=1,\dots, N$.




## Numerical implementation {#num_implementation}

### Function `my.get.roots()`

For each rational order $m$ (1,2,3,4,5,6,7,8) and smoothness parameter $\beta$ (= $\alpha/2$ with $\alpha$ between 0.5 and 2), function `my.get.roots()` (adapted from the `rSPDE` package) returns $\texttt{factor} = \dfrac{c_m}{b_{m+1}}$, and the roots $\{r_{1i}\}_{i=1}^m$ and $\{r_{2j}\}_{j=1}^{m+1}$.

The file [`data_files/chebfun_tables.RDS`](https://github.com/leninrafaelrierasegura/NAFDEMG/blob/main/data_files/chebfun_tables.RDS) contains the precomputed tables for the roots and factors for rational orders 1 to 8. These tables were generated using the [`matlab/chebfun.m`](https://github.com/leninrafaelrierasegura/NAFDEMG/blob/main/matlab/chebfun.m) and [`matlab/chebfun_tables.R`](https://github.com/leninrafaelrierasegura/NAFDEMG/blob/main/matlab/chebfun_tables.R) scripts.

```{r}
# Function to compute the roots and factor for the rational approximation
my.get.roots <- function(m, # rational order, m = 1, 2, 3, 4, 5, 6, 7, or 8
                         beta # smoothness parameter, beta = alpha/2 with alpha between 0.5 and 2
                         ) {
  # m1table <- rSPDE:::m1table
  # m2table <- rSPDE:::m2table
  # m3table <- rSPDE:::m3table
  # m4table <- rSPDE:::m4table
  # mt <- get(paste0("m", m, "table"))
  mt <- readRDS("data_files/chebfun_tables.RDS")[[m]]
  rb <- rep(0, m + 1)
  rc <- rep(0, m)
  if(m == 1) {
    rc = approx(mt$beta, mt[[paste0("rc")]], beta)$y
  } else {
    rc = sapply(1:m, function(i) {
      approx(mt$beta, mt[[paste0("rc.", i)]], beta)$y
    })
  }
  rb = sapply(1:(m+1), function(i) {
    approx(mt$beta, mt[[paste0("rb.", i)]], xout = beta)$y
  })
  factor = approx(mt$beta, mt$factor, xout = beta)$y
  return(list(pl_roots = rb, # roots \{r_{2j}\}_{j=1}^{m+1}
              pr_roots = rc, # roots \{r_{1i}\}_{i=1}^m
              factor = factor # this is c_m/b_{m+1}
              ))
}
```

### Function `poly.from.roots()`

Function `poly.from.roots()` computes the coefficients of a polynomial from its roots.

```{r}
# Function to compute polynomial coefficients from roots
poly.from.roots <- function(roots) {
  coef <- 1
  for (r in roots) {coef <- convolve(coef, c(1, -r), type = "open")}
  return(coef) # returned in increasing order like a+bx+cx^2+...
}
```

### Function `compute.partial.fraction.param()`

Given `factor`$=\texttt{factor} = \dfrac{c_m}{b_{m+1}}$, `pr_roots`$=\{r_{1i}\}_{i=1}^m$, `pl_roots`$=\{r_{2j}\}_{j=1}^{m+1}$, `time_step`$=\tau$, and `scaling`$=\kappa^{2\beta}$, function `compute.partial.fraction.param()` computes the parameters for the partial fraction decomposition \eqref{eq:partial_fractionadjoint}.

```{r}
# Function to compute the parameters for the partial fraction decomposition
compute.partial.fraction.param <- function(factor, # c_m/b_{m+1}
                                           pr_roots, # roots \{r_{1i}\}_{i=1}^m
                                           pl_roots, # roots \{r_{2j}\}_{j=1}^{m+1}
                                           time_step, # \tau
                                           scaling # \kappa^{2\beta}
                                           ) {
  pr_coef <- poly.from.roots(pr_roots)
  pl_coef <- poly.from.roots(pl_roots)
  pr_plus_pl_coef <- c(0, pr_coef) + ((scaling * time_step)/factor) * pl_coef
  poles <- Re(polyroot(rev(pr_plus_pl_coef)))
  num_vals <- pracma::polyval(pr_coef, poles)
  den_deriv <- Re(pracma::polyval(pracma::polyder(pr_plus_pl_coef), poles))
  residues <- Re(num_vals / den_deriv)
  return(list(r = residues, # residues \{a_k\}_{k=1}^{m+1}
              p = poles, # poles \{p_k\}_{k=1}^{m+1}
              k = 0 # remainder r
              )) 
}
```

### Function `my.fractional.operators.frac()`

Given the Laplacian matrix `L`, the smoothness parameter `beta`, the mass matrix `C` (not lumped), the scaling factor `scale.factor`$=\kappa^2$, the rational order `m`, and the time step `time_step`$=\tau$, function `my.fractional.operators.frac()` computes the fractional operator and returns a list containing the necessary matrices and parameters for the fractional diffusion equation.

```{r}
# Function to compute the fractional operator
my.fractional.operators.frac <- function(L, # Laplacian matrix
                                         beta, # smoothness parameter beta
                                         C, # mass matrix (not lumped)
                                         scale.factor, # scaling parameter = kappa^2
                                         m = 1, # rational order, m = 1, 2, 3, or 4
                                         time_step # time step = tau
                                         ) {
  I <- Matrix::Diagonal(dim(C)[1])
  L <- L / scale.factor 
  if(beta == 1){
    L <- L * scale.factor^beta
    return(list(C = C, # mass matrix
                L = L, # Laplacian matrix scaled
                m = m, # rational order
                beta = beta, # smoothness parameter
                LHS = C + time_step * L # left-hand side of the linear system
                ))
  } else {
    scaling <- scale.factor^beta
    roots <- my.get.roots(m, beta)
    poles_rs_k <- compute.partial.fraction.param(roots$factor, roots$pr_roots, roots$pl_roots, time_step, scaling)

    partial_fraction_terms <- list()
    for (i in 1:(m+1)) {
      # Here is where the terms in the sum in eq 12 are computed
      partial_fraction_terms[[i]] <- (L - poles_rs_k$p[i] * C)#/poles_rs_k$r[i]
      }
    return(list(C = C, # mass matrix
                L = L, # Laplacian matrix scaled
                m = m, # rational order
                beta = beta, # smoothness parameter
                partial_fraction_terms = partial_fraction_terms, # partial fraction terms
                residues = poles_rs_k$r # residues \{a_k\}_{k=1}^{m+1}
                ))
  }
}
```

### Function `my.solver.frac()`

Given the object returned by `my.fractional.operators.frac()` and a vector `v`, function `my.solver.frac()` solves the system \eqref{thenumericalscheme4} for the vector `v`. If `beta = 1`, it solves the system directly; otherwise, it uses the partial fraction decomposition.

```{r}
# Function to solve the iteration
my.solver.frac <- function(obj, # object returned by my.fractional.operators.frac()
                           v # vector to be solved for
                           ){
  beta <- obj$beta
  m <- obj$m
  if (beta == 1){
    return(solve(obj$LHS, v) # solve the linear system directly for beta = 1
           )
  } else {
    partial_fraction_terms <- obj$partial_fraction_terms
    residues <- obj$residues
    output <- v*0
    for (i in 1:(m+1)) {output <- output + residues[i] * solve(partial_fraction_terms[[i]], v)}
    return(output # solve the linear system using the partial fraction decomposition
           )
  }
}
```


### Function `solve_forward_evolution()` 

Given the object returned by `my.fractional.operators.frac()`, the time step `time_step`$=\tau$, the time sequence `time_seq`, the right-hand side term `RHST`, and the initial value `val_at_0`, function `solve_forward_evolution()` solves the forward evolution problem \eqref{statesolve}.


```{r}
solve_forward_evolution <- function(my_op_frac, time_step, time_seq, RHST, val_at_0) {
  CC <- my_op_frac$C
  N <- length(time_seq)
  SOL <- matrix(NA, nrow = nrow(CC), ncol = N)
  SOL[, 1] <- val_at_0
  for (k in 1:(N - 1)) {
    rhs <- CC %*% SOL[, k] + time_step * RHST[, k + 1]
    SOL[, k + 1] <- as.matrix(my.solver.frac(my_op_frac, rhs))
  }
  return(SOL)
}
```


### Function `solve_backward_evolution()`

Given the object returned by `my.fractional.operators.frac()`, the time step `time_step`$=\tau$, the time sequence `time_seq`, and the right-hand side term `RHST`, function `solve_backward_evolution()` solves the backward evolution problem \eqref{adjointsolve}.


```{r}
solve_backward_evolution <- function(my_op_frac, time_step, time_seq, RHST) {
  CC <- my_op_frac$C
  N <- length(time_seq)
  SOL <- matrix(NA, nrow = nrow(CC), ncol = N)
  SOL[, N] <- 0
  for (k in (N - 1):1) {
    rhs <- CC %*% SOL[, k + 1] + time_step * RHST[, k + 1] #this is how it should be in theory
    #rhs <- CC %*% SOL[, k + 1] + time_step * RHST[, k]
    SOL[, k] <- as.matrix(my.solver.frac(my_op_frac, rhs))
  }
  return(SOL)
}
```


## Auxiliary functions {#auxiliary_functions}

### Function `gets.graph.tadpole()`

Given a mesh size `h`, function `gets.graph.tadpole()` builds a tadpole graph and creates a mesh.


```{r}
# Function to build a tadpole graph and create a mesh
gets.graph.tadpole <- function(h){
  edge1 <- rbind(c(0,0),c(1,0))
  theta <- seq(from=-pi,to=pi,length.out = 10000)
  edge2 <- cbind(1+1/pi+cos(theta)/pi,sin(theta)/pi)
  edges <- list(edge1, edge2)
  graph <- metric_graph$new(edges = edges, verbose = 0)
  graph$set_manual_edge_lengths(edge_lengths = c(1,2))
  graph$build_mesh(h = h)
  return(graph)
}
```


### Function `tadpole.eig()`

Given a mode number `k` and a tadpole graph `graph`, function `tadpole.eig()` computes the eigenpairs of the tadpole graph.


```{r}
# Function to compute the eigenfunctions of the tadpole graph
tadpole.eig <- function(k,graph){
x1 <- c(0,graph$get_edge_lengths()[1]*graph$mesh$PtE[graph$mesh$PtE[,1]==1,2]) 
x2 <- c(0,graph$get_edge_lengths()[2]*graph$mesh$PtE[graph$mesh$PtE[,1]==2,2]) 

if(k==0){ 
  f.e1 <- rep(1,length(x1)) 
  f.e2 <- rep(1,length(x2)) 
  f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
  f = list(phi=f1/sqrt(3)) 
  
} else {
  f.e1 <- -2*sin(pi*k*1/2)*cos(pi*k*x1/2) 
  f.e2 <- sin(pi*k*x2/2)                  
  
  f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
  
  if((k %% 2)==1){ 
    f = list(phi=f1/sqrt(3)) 
  } else { 
    f.e1 <- (-1)^{k/2}*cos(pi*k*x1/2)
    f.e2 <- cos(pi*k*x2/2)
    f2 = c(f.e1[1],f.e2[1],f.e1[-1],f.e2[-1]) 
    f <- list(phi=f1,psi=f2/sqrt(3/2))
  }
}
return(f)
}
```

### Function `gets.eigen.params()`

Given a finite number of modes `N_finite`, a scaling parameter `kappa`, a smoothness parameter `alpha`, and a tadpole graph `graph`, function `gets.eigen.params()` computes `EIGENVAL_ALPHA` (a vector with entries $\lambda_j^{\alpha/2}$), `EIGENVAL_MINUS_ALPHA` (a vector with entries $\lambda_j^{-\alpha/2}$), and `EIGENFUN` (a matrix with columns $e_j$ on the mesh of `graph`).


```{r}
# Function to compute the eigenpairs of the tadpole graph
gets.eigen.params <- function(N_finite = 4, kappa = 1, alpha = 0.5, graph){
  EIGENVAL <- NULL
  EIGENVAL_ALPHA <- NULL
  EIGENVAL_MINUS_ALPHA <- NULL
  EIGENFUN <- NULL
  INDEX <- NULL
  for (j in 0:N_finite) {
    lambda_j <- kappa^2 + (j*pi/2)^2
    lambda_j_alpha_half <- lambda_j^(alpha/2)
    lambda_j_minus_alpha_half <- lambda_j^(-alpha/2)
    e_j <- tadpole.eig(j,graph)$phi
    EIGENVAL <- c(EIGENVAL, lambda_j)
    EIGENVAL_ALPHA <- c(EIGENVAL_ALPHA, lambda_j_alpha_half)  
    EIGENVAL_MINUS_ALPHA <- c(EIGENVAL_MINUS_ALPHA, lambda_j_minus_alpha_half)
    EIGENFUN <- cbind(EIGENFUN, e_j)
    INDEX <- c(INDEX, j)
    if (j>0 && (j %% 2 == 0)) {
      lambda_j <- kappa^2 + (j*pi/2)^2
      lambda_j_alpha_half <- lambda_j^(alpha/2)
      lambda_j_minus_alpha_half <- lambda_j^(-alpha/2)
      e_j <- tadpole.eig(j,graph)$psi
      EIGENVAL <- c(EIGENVAL, lambda_j)
      EIGENVAL_ALPHA <- c(EIGENVAL_ALPHA, lambda_j_alpha_half)    
      EIGENVAL_MINUS_ALPHA <- c(EIGENVAL_MINUS_ALPHA, lambda_j_minus_alpha_half)
      EIGENFUN <- cbind(EIGENFUN, e_j)
      INDEX <- c(INDEX, j+0.1)
      }
    }
  return(list(EIGENVAL = EIGENVAL,
              EIGENVAL_ALPHA = EIGENVAL_ALPHA, 
              EIGENVAL_MINUS_ALPHA = EIGENVAL_MINUS_ALPHA,
              EIGENFUN = EIGENFUN,
              INDEX = INDEX))
}
```


### Function `construct_piecewise_projection()` {#construct_piecewise_projection}

Given a matrix `projected_U_approx` with approximated values at discrete time points, a sequence of time points `time_seq`, and an extended sequence of time points `overkill_time_seq`, function `construct_piecewise_projection()` constructs a piecewise constant projection of the approximated values over the extended time sequence.

```{r}
# Function to construct a piecewise constant projection of approximated values
construct_piecewise_projection <- function(projected_U_approx, time_seq, overkill_time_seq) {
  projected_U_piecewise <- matrix(NA, nrow = nrow(projected_U_approx), ncol = length(overkill_time_seq))
  
  # Assign value at t = 0
  projected_U_piecewise[, which(overkill_time_seq == 0)] <- projected_U_approx[, 1]
  
  # Assign values for intervals (t_{k-1}, t_k]
  for (k in 2:length(time_seq)) {
    idxs <- which(overkill_time_seq > time_seq[k - 1] & overkill_time_seq <= time_seq[k])
    projected_U_piecewise[, idxs] <- projected_U_approx[, k]
  }
  
  return(projected_U_piecewise)
}
```

### Functions for computing the true line rates

```{r}
loglog_line_equation <- function(x1, y1, slope) {
  b <- log10(y1 / (x1 ^ slope))
  
  function(x) {
    (x ^ slope) * (10 ^ b)
  }
}
exp_line_equation <- function(x1, y1, slope) {
  lnC <- log(y1) - slope * x1
  
  function(x) {
    exp(lnC + slope * x)
  }
}
compute_guiding_lines <- function(x_axis_vector, errors, theoretical_rates, line_equation_fun) {
  guiding_lines <- matrix(NA, nrow = length(x_axis_vector), ncol = length(theoretical_rates))
  
  for (j in seq_along(theoretical_rates)) {
    guiding_lines_aux <- matrix(NA, nrow = length(x_axis_vector), ncol = length(x_axis_vector))
    
    for (k in seq_along(x_axis_vector)) {
      point_x1 <- x_axis_vector[k]
      point_y1 <- errors[k, j]
      slope <- theoretical_rates[j]
      
      line <- line_equation_fun(x1 = point_x1, y1 = point_y1, slope = slope)
      guiding_lines_aux[, k] <- line(x_axis_vector)
    }
    
    guiding_lines[, j] <- rowMeans(guiding_lines_aux)
  }
  
  return(guiding_lines)
}
```


```{r}
# Functions to compute the exact solution to the fractional diffusion equation
g_linear <- function(r, A, lambda_j_alpha_half) {
  return(A * exp(-lambda_j_alpha_half * r))
  }
G_linear <- function(t, A) {
  return(A * t)
  }
g_exp <- function(r, A, mu) {
  return(A * exp(mu * r))
  }
G_exp <- function(t, A, lambda_j_alpha_half, mu) {
  exponent <- lambda_j_alpha_half + mu
  return(A * (exp(exponent * t) - 1) / exponent)
  }
g_poly <- function(r, A, n) {
  return(A * r^n)
}
G_poly <- function(t, A, lambda_j_alpha_half, n) {
  t <- as.vector(t)
  k_vals <- 0:n
  sum_term <- sapply(t, function(tt) {
    sum(((-lambda_j_alpha_half * tt)^k_vals) / factorial(k_vals))
  })
  coeff <- ((-1)^(n + 1)) * factorial(n) / (lambda_j_alpha_half^(n + 1))
  return(A * coeff * (1 - exp(lambda_j_alpha_half * t) * sum_term))
}
g_sin <- function(r, A, omega) {
  return(A * sin(omega * r))
}
G_sin <- function(t, A, lambda_j_alpha_half, omega) {
  denom <- lambda_j_alpha_half^2 + omega^2
  numerator <- exp(lambda_j_alpha_half * t) * (lambda_j_alpha_half * sin(omega * t) - omega * cos(omega * t)) + omega
  return(A * numerator / denom)
}
g_cos <- function(r, A, theta) {
  return(A * cos(theta * r)) 
}
G_cos <- function(t, A, lambda_j_alpha_half, theta) {
  denom <- lambda_j_alpha_half^2 + theta^2
  numerator <- exp(lambda_j_alpha_half * t) * (lambda_j_alpha_half * cos(theta * t) + theta * sin(theta * t)) - lambda_j_alpha_half
  return(A * numerator / denom)
}
```

### Function `reversecolumns()`

Given a matrix `mat`, function `reversecolumns()` reverses the order of its columns.


```{r}
reversecolumns <- function(mat) {
  return(mat[, rev(seq_len(ncol(mat)))])
}
```


```{r}
# helper: measure change relative to the size of the previous iterate 
change_comparer <- function(X_new, X_old, time_step, C, relative = TRUE) {
  XX <- X_new - X_old
  num <- sqrt(as.double(time_step * sum(XX * (C %*% XX))))
  if (!relative) {
    return(num)
    }
  den <- sqrt(as.double(time_step * sum(X_new * (C %*% X_new))))
  if (den < .Machine$double.eps) {
    return(ifelse(num < .Machine$double.eps, 0, num))
  } else {
    return(num / den)
  }
}
```


```{r}
# Coupled solver with multi-criteria convergence
solve_coupled_system_multi_tol <- function(
  my_op_frac,           # operator
  time_step,            # tau
  time_seq,             # vector of times 
  u_0,                  # initial state U^0 
  F_proj,               # matrix of F 
  Z_ini,
  V_d,                  # matrix of 
  u_d,
  Psi,                  # Psi matrix
  R,                    # R matrix
  a, b, C,                # lower/upper bounds (vector or matrix broadcastable to time grid)
  mu,                   # positive scalar
  tol = 1e-8,           # scalar or named list: list(Z=..., U=..., P=...)
  maxit = 200,
  verbose = FALSE,
  nested_spatial_mesh = FALSE,
  true_sol
) {

  if (is.numeric(tol) && length(tol) == 1) {
    tol_list <- list(Z = tol, U = tol, P = tol)
  } else if (is.list(tol)) {
    tol_list <- modifyList(list(Z = 1e-8, U = 1e-8, P = 1e-8), tol)
  } else stop("tol must be scalar or list(Z=...,U=...,P=...)")

  it <- 0
  converged <- FALSE
  
  rel_history <- data.frame(iter = integer(0), variable = character(0), value = numeric(0))
  abs_history <- data.frame(iter = integer(0), variable = character(0), value = numeric(0))
  min_history <- data.frame(iter = integer(0), variable = character(0), value = numeric(0))

  Z_list <- list()
  U_list <- list()
  P_list <- list()
  
  z_prev <- Z_ini
  if(nested_spatial_mesh == TRUE){Z_mat <- C %*% z_prev}else{Z_mat <- R %*% Psi %*% z_prev}
  U_prev <- F_proj*0
  P_prev <- F_proj*0

  repeat {
    it <- it + 1

    U_mat <- solve_forward_evolution(my_op_frac, time_step, time_seq, RHST = F_proj + Z_mat, val_at_0 = u_0)
    if(nested_spatial_mesh == TRUE){V_mat <- C %*% U_mat}else{V_mat <- R %*% Psi %*% U_mat}
    P_mat <- solve_backward_evolution(my_op_frac, time_step, time_seq, RHST = V_mat - V_d)
    z_new <- matrix(pmax(a, pmin(b, - P_mat / mu)), dim(P_mat))
    if(nested_spatial_mesh == TRUE){Z_mat <- C %*% z_new}else{Z_mat <- R %*% Psi %*% z_new}
    
    # relative changes
    rel_changes_Z <- change_comparer(z_new, z_prev, time_step, C, relative = TRUE)  
    rel_changes_U <- change_comparer(U_mat, U_prev, time_step, C, relative = TRUE)
    rel_changes_P <- change_comparer(P_mat, P_prev, time_step, C, relative = TRUE)
    abs_changes_Z <- change_comparer(z_new, true_sol$z_bar, time_step, C, relative = FALSE)
    abs_changes_U <- change_comparer(U_mat, true_sol$u_bar, time_step, C, relative = FALSE)
    abs_changes_P <- change_comparer(P_mat, true_sol$p_bar, time_step, C, relative = FALSE)
    XX <- U_mat - u_d
    min_change <- 0.5 * as.double(time_step * sum(XX * (C %*% XX)))  + 0.5 * mu * as.double(time_step * sum(z_new * (C %*% z_new)))
    rel_history <- rbind(rel_history,
      data.frame(iter = it, variable = "Z", value = rel_changes_Z),
      data.frame(iter = it, variable = "U", value = rel_changes_U),
      data.frame(iter = it, variable = "P", value = rel_changes_P))
    abs_history <- rbind(abs_history,
      data.frame(iter = it, variable = "Z", value = abs_changes_Z),
      data.frame(iter = it, variable = "U", value = abs_changes_U),
      data.frame(iter = it, variable = "P", value = abs_changes_P))
    min_history <- rbind(min_history,
      data.frame(iter = it, variable = "min", value = min_change))
    
    if (verbose) {message(sprintf("iter %3d: rel(Z) = %.3e, rel(U) = %.3e, rel(P) = %.3e", it, rel_changes_Z, rel_changes_U, rel_changes_P))}

    # update stored previous iterates
    z_prev <- z_new
    U_prev <- U_mat
    P_prev <- P_mat
    
    Z_list[[paste0("iteration ", it)]] <- z_new
    U_list[[paste0("iteration ",it)]] <- U_mat
    P_list[[paste0("iteration ",it)]] <- P_mat

    # convergence check: require all rel_changes <= respective tol
    cond_Z <- rel_changes_Z <= tol_list$Z
    cond_U <- rel_changes_U <= tol_list$U
    cond_P <- rel_changes_P <= tol_list$P

    if ((cond_Z && cond_U && cond_P) || it >= maxit) {
      converged <- (cond_Z && cond_U && cond_P)
      break
    }
  }

  if (verbose && !converged) {
    message(sprintf(
      "Stopped at maxit=%d; rel_changes: Z = %.3e (tol %.3e), U = %.3e (tol %.3e), P = %.3e (tol %.3e)",
      it, rel_changes_Z, tol_list$Z, rel_changes_U, tol_list$U, rel_changes_P, tol_list$P
    ))
  }

  return(list(U = U_mat,  # solution U
              Z = z_new,  # solution z
              P = P_mat, # solution P
              iterations = it,
              converged = converged,
              tol_list = tol_list,
              rel_history = rel_history,
              abs_history = abs_history,
              min_history = min_history,
              Z_list = Z_list,
              U_list = U_list,
              P_list = P_list))
}
```


```{r}
plot_convergence_history <- function(history_df, tol_list = NULL, type = "relative") {
  if (type == "relative"){
    text_title <- "|X_{iter} - X_{iter-1}| / |X_{iter}|"
  } else if (type == "absolute") {
    text_title <- "|X_{exact} - X_{iter}|"
  } else if (type == "minimum") {
    text_title <- "J(U_{iter},z_{iter})"
  }

  p <- ggplot(history_df, aes(x = iter, y = value, color = variable)) +
    geom_line() +
    geom_point(size = 1.5) +
    scale_y_log10() +
    labs(
      title = text_title,
      x = "Iteration",
      y = "Error",
      color = "Quantity"
    ) +
    theme_minimal()
  
  # Add tolerance lines if provided
  if (!is.null(tol_list)) {
    tol_df <- data.frame(
      variable = names(tol_list),
      tol = unlist(tol_list)
    )
    p <- p + geom_hline(
      data = tol_df,
      aes(yintercept = tol, color = variable),
      linetype = "dashed"
    )
  }
  
  return(plotly::ggplotly(p))
}
```

```{r}
largest_nested_h <- function(h_fine, h_candidate) {
  Nfine <- round(1 / h_fine)       # number of intervals in fine mesh
  m0 <- floor(h_candidate / h_fine)
  
  best <- 0
  r <- floor(sqrt(Nfine))
  
  for (a in 1:r) {
    if (Nfine %% a == 0) {
      b <- Nfine / a
      if (a <= m0 && a > best) best <- a
      if (b <= m0 && b > best) best <- b
    }
  }
  
  # if no divisor found, default to h_fine
  if (best == 0) best <- 1
  
  h_coarse <- best * h_fine
  return(h_coarse)
}
```

```{r}
trunc_first_signi_digit <- function(x){
  aux <- floor(log10(x))
  return(floor(x / 10^aux) * 10^aux)
}
```

## Plotting functions {#plotting_functions}

### Function `plotting.order()`

Given a vector `v` and a graph object `graph`, function `plotting.order()` orders the mesh values for plotting.

```{r}
# Function to order the vertices for plotting
plotting.order <- function(v, graph){
  edge_number <- graph$mesh$VtE[, 1]
  pos <- sum(edge_number == 1)+1
  return(c(v[1], v[3:pos], v[2], v[(pos+1):length(v)], v[2]))
}
```

### Function `global.scene.setter()`

Given ranges for the `x`, `y`, and `z` axes, and an optional aspect ratio for the `z` axis, function `global.scene.setter()` sets the scene for 3D plots so that all plots have the same aspect ratio and camera position.

```{r}
# Function to set the scene for 3D plots
global.scene.setter <- function(x_range, y_range, z_range, z_aspectratio = 4) {
  
  return(list(xaxis = list(title = "x", range = x_range),
              yaxis = list(title = "y", range = y_range),
              zaxis = list(title = "z", range = z_range),
              aspectratio = list(x = 2*(1+2/pi), 
                                 y = 2*(2/pi), 
                                 z = z_aspectratio*(2/pi)),
              camera = list(eye = list(x = (1+2/pi)/2, 
                                       y = 4, 
                                       z = 2),
                            center = list(x = (1+2/pi)/2, 
                                          y = 0, 
                                          z = 0))))
}
```

### Function `graph.plotter.3d()`

Given a graph object `graph`, a sequence of time points `time_seq`, and one or more matrices `...` representing function values defined on the mesh of `graph` at each time in `time_seq`, the `graph.plotter.3d()` function generates an interactive 3D visualization of these values over time.

```{r}
# Function to plot in 3D
graph.plotter.3d.old <- function(graph, time_seq, frame_val_to_display, ...) {
  U_list <- list(...)
  U_names <- sapply(substitute(list(...))[-1], deparse)

  # Spatial coordinates
  x <- plotting.order(graph$mesh$V[, 1], graph)
  y <- plotting.order(graph$mesh$V[, 2], graph)
  weights <- graph$mesh$weights

  # Apply plotting.order to each U
  U_list <- lapply(U_list, function(U) apply(U, 2, plotting.order, graph = graph))
  n_vars <- length(U_list)

  # Create plot_data frame with time and position replicated
  n_time <- ncol(U_list[[1]])
  base_data <- data.frame(
    x = rep(x, times = n_time),
    y = rep(y, times = n_time),
    the_graph = 0,
    frame = rep(time_seq, each = length(x))
  )

  # Add U columns to plot_data
  for (i in seq_along(U_list)) {
    base_data[[paste0("u", i)]] <- as.vector(U_list[[i]])
  }

  plot_data <- base_data

  # Generate vertical lines
  vertical_lines_list <- lapply(seq_along(U_list), function(i) {
    do.call(rbind, lapply(time_seq, function(t) {
      idx <- which(plot_data$frame == t)
      z_vals <- plot_data[[paste0("u", i)]][idx]
      data.frame(
        x = rep(plot_data$x[idx], each = 3),
        y = rep(plot_data$y[idx], each = 3),
        z = as.vector(t(cbind(0, z_vals, NA))),
        frame = rep(t, each = length(idx) * 3)
      )
    }))
  })

  # Set axis ranges
  z_range <- range(unlist(U_list))
  x_range <- range(x)
  y_range <- range(y)

  # Create plot
  p <- plot_ly(plot_data, frame = ~frame) %>%
    add_trace(x = ~x, y = ~y, z = ~the_graph, type = "scatter3d", mode = "lines",
              name = "", showlegend = FALSE,
              line = list(color = "black", width = 3))

  # Add traces for each variable
  colors <- rev(viridisLite::viridis(n_vars)) #RColorBrewer::brewer.pal(min(n_vars, 8), "Set1")
  for (i in seq_along(U_list)) {
    p <- add_trace(p,
      x = ~x, y = ~y, z = as.formula(paste0("~u", i)),
      type = "scatter3d", mode = "lines", name = U_names[i],
      line = list(color = colors[i], width = 3))
  }

  # Add vertical lines
  for (i in seq_along(vertical_lines_list)) {
    p <- add_trace(p,
      data = vertical_lines_list[[i]],
      x = ~x, y = ~y, z = ~z, frame = ~frame,
      type = "scatter3d", mode = "lines",
      line = list(color = "gray", width = 0.5),
      name = "Vertical lines",
      showlegend = FALSE)
  }
  frame_name <- deparse(substitute(frame_val_to_display))
  # Layout and animation controls
  p <- p %>%
    layout(
      scene = global.scene.setter(x_range, y_range, z_range),
      updatemenus = list(list(type = "buttons", showactive = FALSE,
                              buttons = list(
                                list(label = "Play", method = "animate",
                                     args = list(NULL, list(frame = list(duration = 2000 / length(time_seq), redraw = TRUE), fromcurrent = TRUE))),
                                list(label = "Pause", method = "animate",
                                     args = list(NULL, list(mode = "immediate", frame = list(duration = 0), redraw = FALSE)))
                              )
      )),
      title = paste0(frame_name,": ", formatC(frame_val_to_display[1], format = "f", digits = 4))
    ) %>%
    plotly_build()

  for (i in seq_along(p$x$frames)) {
    p$x$frames[[i]]$layout <- list(title = paste0(frame_name,": ", formatC(frame_val_to_display[i], format = "f", digits = 4)))
  }

  return(p)
}
```

```{r}
graph.plotter.3d <- function(graph, time_seq, frame_val_to_display, U_list) {
  U_names <- names(U_list) 
  # Spatial coordinates
  x <- plotting.order(graph$mesh$V[, 1], graph)
  y <- plotting.order(graph$mesh$V[, 2], graph)
  weights <- graph$mesh$weights

  # Apply plotting.order to each U
  U_list <- lapply(U_list, function(U) apply(U, 2, plotting.order, graph = graph))
  n_vars <- length(U_list)
  
  # Create plot_data frame with time and position replicated
  n_time <- ncol(U_list[[1]])
  base_data <- data.frame(
    x = rep(x, times = n_time),
    y = rep(y, times = n_time),
    the_graph = 0,
    frame = rep(time_seq, each = length(x))
  )

  # Add U columns to plot_data
  for (i in seq_along(U_list)) {
    base_data[[paste0("u", i)]] <- as.vector(U_list[[i]])
  }

  plot_data <- base_data

  # Generate vertical lines
  vertical_lines_list <- lapply(seq_along(U_list), function(i) {
    do.call(rbind, lapply(time_seq, function(t) {
      idx <- which(plot_data$frame == t)
      z_vals <- plot_data[[paste0("u", i)]][idx]
      data.frame(
        x = rep(plot_data$x[idx], each = 3),
        y = rep(plot_data$y[idx], each = 3),
        z = as.vector(t(cbind(0, z_vals, NA))),
        frame = rep(t, each = length(idx) * 3)
      )
    }))
  })

  # Set axis ranges
  z_range <- range(unlist(U_list))
  x_range <- range(x)
  y_range <- range(y)

  # Create plot
  p <- plot_ly(plot_data, frame = ~frame) %>%
    add_trace(x = ~x, y = ~y, z = ~the_graph, type = "scatter3d", mode = "lines",
              name = "", showlegend = FALSE,
              line = list(color = "black", width = 3))

  if (n_vars == 2) {
    colors <- RColorBrewer::brewer.pal(min(n_vars, 8), "Set1") 
    } else {
    colors <- rev(viridisLite::viridis(n_vars)) 
  }
  # RColorBrewer::brewer.pal(min(n_vars, 8), "Set1")
  for (i in seq_along(U_list)) {
    p <- add_trace(p,
      x = ~x, y = ~y, z = as.formula(paste0("~u", i)),
      type = "scatter3d", mode = "lines", name = U_names[i],
      line = list(color = colors[i], width = 3))
  }

  # Add vertical lines
  for (i in seq_along(vertical_lines_list)) {
    p <- add_trace(p,
      data = vertical_lines_list[[i]],
      x = ~x, y = ~y, z = ~z, frame = ~frame,
      type = "scatter3d", mode = "lines",
      line = list(color = "gray", width = 0.5),
      name = "Vertical lines",
      showlegend = FALSE)
  }
  frame_name <- deparse(substitute(frame_val_to_display))
  # Layout and animation controls
  p <- p %>%
    layout(
      scene = global.scene.setter(x_range, y_range, z_range),
      updatemenus = list(list(type = "buttons", showactive = FALSE,
                              buttons = list(
                                list(label = "Play", method = "animate",
                                     args = list(NULL, list(frame = list(duration = 2000 / length(time_seq), redraw = TRUE), fromcurrent = TRUE))),
                                list(label = "Pause", method = "animate",
                                     args = list(NULL, list(mode = "immediate", frame = list(duration = 0), redraw = FALSE)))
                              )
      )),
      title = paste0(frame_name,": ", formatC(frame_val_to_display[1], format = "f", digits = 4))
    ) %>%
    plotly_build()

  for (i in seq_along(p$x$frames)) {
    p$x$frames[[i]]$layout <- list(title = paste0(frame_name,": ", formatC(frame_val_to_display[i], format = "f", digits = 4)))
  }

  return(p)
}
```

### Function `error.at.each.time.plotter()`

Given a graph object `graph`, a matrix `U_true` of true values, a matrix `U_approx` of approximated values, a sequence of time points `time_seq`, and a time step `time_step`, function `error.at.each.time.plotter()` computes the error at each time step and generates a plot showing the error over time.

```{r}
# Function to plot the error at each time step
error.at.each.time.plotter <- function(graph, U_true, U_approx, time_seq, time_step) {
  weights <- graph$mesh$weights
  error_at_each_time <- t(weights) %*% (U_true - U_approx)^2
  error <- sqrt(as.double(t(weights) %*% (U_true - U_approx)^2 %*% rep(time_step, ncol(U_true))))
  p <- plot_ly() %>% 
  add_trace(
  x = ~time_seq, y = ~error_at_each_time, type = 'scatter', mode = 'lines+markers',
  line = list(color = 'blue', width = 2),
  marker = list(color = 'blue', size = 4),
  name = "",
  showlegend = TRUE
) %>% 
  layout(
  title = paste0("Error at Each Time Step (Total error = ", formatC(error, format = "f", digits = 9), ")"),
  xaxis = list(title = "t"),
  yaxis = list(title = "Error"),
  legend = list(x = 0.1, y = 0.9)
)
  return(p)
}
```

### Function `graph.plotter.3d.comparer()`

Given a graph object `graph`, matrices `U_true` and `U_approx` representing true and approximated values, and a sequence of time points `time_seq`, function `graph.plotter.3d.comparer()` generates a 3D plot comparing the true and approximated values over time, with color-coded traces for each time point.

```{r}
# Function to plot the 3D comparison of U_true and U_approx
graph.plotter.3d.comparer <- function(graph, U_true, U_approx, time_seq) {
  x <- graph$mesh$V[, 1]; y <- graph$mesh$V[, 2]
  x <- plotting.order(x, graph); y <- plotting.order(y, graph)

  U_true <- apply(U_true, 2, plotting.order, graph = graph)
  U_approx <- apply(U_approx, 2, plotting.order, graph = graph)
  n_times <- length(time_seq)
  
  x_range <- range(x); y_range <- range(y); z_range <- range(c(U_true, U_approx))
  
  # Normalize time_seq
  time_normalized <- (time_seq - min(time_seq)) / (max(time_seq) - min(time_seq))
  blues <- colorRampPalette(c("lightblue", "blue"))(n_times)
  reds <- colorRampPalette(c("mistyrose", "red"))(n_times)
  
  # Accurate colorscales
  colorscale_greens <- Map(function(t, col) list(t, col), time_normalized, blues)
  colorscale_reds <- Map(function(t, col) list(t, col), time_normalized, reds)
  
  p <- plot_ly()
  
  # Static black graph structure
  p <- p %>%
    add_trace(x = x, y = y, z = rep(0, length(x)),
              type = "scatter3d", mode = "lines",
              line = list(color = "black", width = 4),
              name = "Graph", showlegend = FALSE)
  
  # U_true traces (green)
  for (i in seq_len(n_times)) {
    z <- U_true[, i]
    p <- add_trace(
      p,
      type = "scatter3d",
      mode = "lines",
      x = x, y = y, z = z,
      line = list(color = blues[i], width = 4),
      showlegend = FALSE,
      scene = "scene"
    )
  }
  
  # U_approx traces (dashed red)
  for (i in seq_len(n_times)) {
    z <- U_approx[, i]
    p <- add_trace(
      p,
      type = "scatter3d",
      mode = "lines",
      x = x, y = y, z = z,
      line = list(color = reds[i], width = 4, dash = "dot"),
      showlegend = FALSE,
      scene = "scene"
    )
  }
  
  # Dummy green colorbar (True) – with ticks
  p <- add_trace(
    p,
    type = "heatmap",
    z = matrix(time_seq, nrow = 1),
    showscale = TRUE,
    colorscale = colorscale_greens,
    colorbar = list(
      title = list(font = list(size = 12, color = "black"), text = "Time", side = "top"),
      len = 0.9,
      thickness = 15,
      x = 1.02,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle",
      tickvals = NULL,   # hide tick values
      ticktext = NULL,
      ticks = ""         # also hides tick marks
    ),
    x = matrix(time_seq, nrow = 1),
    y = matrix(1, nrow = 1),
    hoverinfo = "skip",
    opacity = 0
  )

# Dummy red colorbar (Approx) – no ticks
  p <- add_trace(
    p,
    type = "heatmap",
    z = matrix(time_seq, nrow = 1),
    showscale = TRUE,
    colorscale = colorscale_reds,
    colorbar = list(
      title = list(font = list(size = 12, color = "black"), text = ".", side = "top"),
      len = 0.9,
      thickness = 15,
      x = 1.05,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle"
    ),
    x = matrix(time_seq, nrow = 1),
    y = matrix(1, nrow = 1),
    hoverinfo = "skip",
    opacity = 0
  )
  p <- p %>%
    add_trace(x = x, y = y, z = rep(0, length(x)),
              type = "scatter3d", mode = "lines",
              line = list(color = "black", width = 4),
              name = "Graph", showlegend = FALSE)
  p <- layout(p,
            scene = global.scene.setter(x_range, y_range, z_range),
            xaxis = list(visible = FALSE),
            yaxis = list(visible = FALSE),
            annotations = list(
  list(
    text = "Exact",
    x = 1.045,
    y = 0.5,
    xref = "paper",
    yref = "paper",
    showarrow = FALSE,
    font = list(size = 12, color = "black"),
    textangle = -90
  ),
  list(
    text = "Approx",
    x = 1.075,
    y = 0.5,
    xref = "paper",
    yref = "paper",
    showarrow = FALSE,
    font = list(size = 12, color = "black"),
    textangle = -90
  )
)

)

  
  return(p)
}
```

### Function `graph.plotter.3d.single()`

Given a graph object `graph`, a matrix `U_true` representing true values, and a sequence of time points `time_seq`, function `graph.plotter.3d.single()` generates a 3D plot of the true values over time, with color-coded traces for each time point.

```{r}
# Function to plot a single 3D line for 
graph.plotter.3d.single <- function(graph, U_true, time_seq) {
  x <- graph$mesh$V[, 1]; y <- graph$mesh$V[, 2]
  x <- plotting.order(x, graph); y <- plotting.order(y, graph)

  U_true <- apply(U_true, 2, plotting.order, graph = graph)
  n_times <- length(time_seq)
  
  x_range <- range(x); y_range <- range(y); z_range <- range(U_true)
  z_range[1] <- z_range[1] - 10^-6
  viridis_colors <- viridisLite::viridis(100)
  
  # Normalize time_seq
  time_normalized <- (time_seq - min(time_seq)) / (max(time_seq) - min(time_seq))
  #greens <- colorRampPalette(c("palegreen", "darkgreen"))(n_times)
  greens <- colorRampPalette(c(viridis_colors[1], viridis_colors[50],  viridis_colors[100]))(n_times)
  # Accurate colorscales
  colorscale_greens <- Map(function(t, col) list(t, col), time_normalized, greens)
  
  p <- plot_ly()
  
  # Add the 3D lines with fading green color
  for (i in seq_len(n_times)) {
    z <- U_true[, i]
    
    p <- add_trace(
      p,
      type = "scatter3d",
      mode = "lines",
      x = x,
      y = y,
      z = z,
      line = list(color = greens[i], width = 2),
      showlegend = FALSE,
      scene = "scene"
    )
  }
  p <- p %>%
    add_trace(x = x, y = y, z = rep(0, length(x)),
              type = "scatter3d", mode = "lines",
              line = list(color = "black", width = 5),
              name = "Graph", showlegend = FALSE)
  # Add dummy heatmap to show colorbar (not part of scene)
  p <- add_trace(
    p,
    type = "heatmap",
    z = matrix(time_seq, nrow = 1),
    showscale = TRUE,
    colorscale = colorscale_greens,
    colorbar = list(
    title = list(font = list(size = 12, color = "black"), text = "Time", side = "top"),
    len = 0.9,         # height (0 to 1)
    thickness = 15,     # width in pixels
    x = 1.02,           # shift it slightly right of the plot
    xanchor = "left",
    y = 0.5,
    yanchor = "middle"),
    x = matrix(time_seq, nrow = 1),
    y = matrix(1, nrow = 1),
    hoverinfo = "skip",
    opacity = 0
  )
  
  p <- layout(p,
              scene = global.scene.setter(x_range, y_range, z_range),
              xaxis = list(visible = FALSE),
              yaxis = list(visible = FALSE)
  )
  
  return(p)
}
```

### Function `error.convergence.plotter()`

```{r}
# Function to plot the error convergence
error.convergence.plotter <- function(x_axis_vector, 
                                      alpha_vector, 
                                      errors, 
                                      theoretical_rates, 
                                      observed_rates,
                                      line_equation_fun,
                                      fig_title,
                                      x_axis_label,
                                      apply_sqrt = FALSE) {
  
  x_vec <- if (apply_sqrt) sqrt(x_axis_vector) else x_axis_vector
  
  guiding_lines <- compute_guiding_lines(x_axis_vector = x_vec, 
                                         errors = errors, 
                                         theoretical_rates = theoretical_rates, 
                                         line_equation_fun = line_equation_fun)
  
  default_colors <- scales::hue_pal()(length(alpha_vector))
  
  plot_lines <- lapply(1:ncol(guiding_lines), function(i) {
    geom_line(
      data = data.frame(x = x_vec, y = guiding_lines[, i]),
      aes(x = x, y = y),
      color = default_colors[i],
      linetype = "dashed",
      show.legend = FALSE
    )
  })
  
  df <- as.data.frame(cbind(x_vec, errors))
  colnames(df) <- c("x_axis_vector", alpha_vector)
  df_melted <- melt(df, id.vars = "x_axis_vector", variable.name = "column", value.name = "value")
  
  custom_labels <- paste0(formatC(alpha_vector, format = "f", digits = 2), 
                          " | ", 
                          formatC(theoretical_rates, format = "f", digits = 4), 
                          " | ", 
                          formatC(observed_rates, format = "f", digits = 4))
  
  df_melted$column <- factor(df_melted$column, levels = alpha_vector, labels = custom_labels)

  p <- ggplot() +
    geom_line(data = df_melted, aes(x = x_axis_vector, y = value, color = column)) +
    geom_point(data = df_melted, aes(x = x_axis_vector, y = value, color = column)) +
    plot_lines +
    labs(
      title = fig_title,
      x = x_axis_label,
      y = expression(Error),
      color = "          α  | theo  | obs"
    ) +
    (if (apply_sqrt) {
      scale_x_continuous(breaks = x_vec, labels = round(x_axis_vector, 4))
    } else {
      scale_x_log10(breaks = x_axis_vector, labels = round(x_axis_vector, 4))
    }) +
    (if (apply_sqrt) {
      scale_y_continuous(trans = "log", labels = scales::scientific_format())
    } else {
      scale_y_log10(labels = scales::scientific_format())
    }) +
    theme_minimal() +
    theme(text = element_text(family = "Palatino"),
          legend.position = "bottom",
          legend.direction = "vertical",
          plot.margin = margin(0, 0, 0, 0),
          plot.title = element_text(hjust = 0.5, size = 18, face = "bold"))
  
  return(p)
}

```


```{r}
graph.plotter.3d.static <- function(graph, z_list) {
  x <- plotting.order(graph$mesh$V[, 1], graph)
  y <- plotting.order(graph$mesh$V[, 2], graph)
  U_names <- names(z_list)
  n_vars <- length(z_list)
  z_list <- lapply(z_list, function(z) plotting.order(z, graph))

  # Axis ranges
  z_range <- range(unlist(z_list))
  x_range <- range(x)
  y_range <- range(y)

  if (n_vars == 2) {
    colors <- RColorBrewer::brewer.pal(min(n_vars, 8), "Set1") 
    } else {
    colors <- rev(viridisLite::viridis(n_vars)) 
  }
  p <- plot_ly()

  for (i in seq_along(z_list)) {
    z <- z_list[[i]]

    # Main 3D curve
    p <- add_trace(
      p,
      x = x, y = y, z = z,
      type = "scatter3d", mode = "lines",
      line = list(color = colors[i], width = 3),
      name = U_names[i], showlegend = TRUE
    )

    # Efficient vertical lines: one trace with breaks (NA)
    x_vert <- rep(x, each = 3)
    y_vert <- rep(y, each = 3)
    z_vert <- unlist(lapply(z, function(zj) c(0, zj, NA)))

    p <- add_trace(
      p,
      x = x_vert, y = y_vert, z = z_vert,
      type = "scatter3d", mode = "lines",
      line = list(color = "gray", width = 0.5),
      showlegend = FALSE
    )
  }
  p <- p %>% add_trace(x = x, y = y, z = x*0, type = "scatter3d", mode = "lines",
              line = list(color = "black", width = 3),
              name = "thegraph", showlegend = FALSE) %>%
    layout(scene = global.scene.setter(x_range, y_range, z_range))
  return(p)
}

```


```{r}
graph.plotter.3d.two.meshes.time <- function(graph_finer, graph_coarser, 
                                             time_seq, frame_val_to_display,
                                             fs_finer = list(), fs_coarser = list()) {
  # Spatial coordinates (ordered for plotting)
  x_finer <- plotting.order(graph_finer$mesh$V[, 1], graph_finer)
  y_finer <- plotting.order(graph_finer$mesh$V[, 2], graph_finer)
  x_coarser <- plotting.order(graph_coarser$mesh$V[, 1], graph_coarser)
  y_coarser <- plotting.order(graph_coarser$mesh$V[, 2], graph_coarser)
  
  n_time <- if (length(fs_finer) > 0) ncol(fs_finer[[1]]) else ncol(fs_coarser[[1]])

  # Helper: make dataframe from one function
  make_df <- function(f_mat, graph, x, y, mesh_name) {
    z <- apply(f_mat, 2, plotting.order, graph = graph)
    data.frame(
      x = rep(x, times = n_time),
      y = rep(y, times = n_time),
      z = as.vector(z),
      frame = rep(time_seq, each = length(x)),
      mesh = mesh_name
    )
  }
  
  # Build data for finer functions
  data_finer_list <- lapply(names(fs_finer), function(nm) {
    make_df(fs_finer[[nm]], graph_finer, x_finer, y_finer, nm)
  })
  
  # Build data for coarser functions
  data_coarser_list <- lapply(names(fs_coarser), function(nm) {
    make_df(fs_coarser[[nm]], graph_coarser, x_coarser, y_coarser, nm)
  })
  
  # Combine
  all_data <- c(data_finer_list, data_coarser_list)
  
  # Baseline graph (on finer mesh for consistency)
  data_graph <- data.frame(
    x = rep(x_finer, times = n_time),
    y = rep(y_finer, times = n_time),
    z = 0,
    frame = rep(time_seq, each = length(x_finer)),
    mesh = "Graph"
  )
  
# --------- Vertical lines helper ----------
vertical_lines <- function(x, y, z, frame_vals, mesh_name) {
  do.call(rbind, lapply(seq_along(frame_vals), function(i) {
    idx <- ((i - 1) * length(x) + 1):(i * length(x))
    data.frame(
      x = rep(x, each = 3),
      y = rep(y, each = 3),
      z = as.vector(t(cbind(0, z[idx], NA))),
      frame = rep(frame_vals[i], each = length(x) * 3),
      mesh = mesh_name
    )
  }))
}

# --------- Compute vertical lines per mesh using max absolute value ---------
make_vertical_from_list <- function(data_list, x, y, mesh_name) {
  if (length(data_list) == 0) return(NULL)
  
  # Reshape each function's z back to matrix: (nodes × time)
  z_mats <- lapply(data_list, function(df) {
    matrix(df$z, nrow = length(x), ncol = length(time_seq))
  })
  
  # Stack into 3D array: (nodes × time × functions)
  arr <- array(unlist(z_mats), dim = c(length(x), length(time_seq), length(z_mats)))
  
  # For each node × time, select the entry with largest absolute value (keep sign)
  idx <- apply(arr, c(1, 2), function(v) which.max(abs(v)))
  z_signed_max <- mapply(function(i, j) arr[i, j, idx[i, j]],
                         rep(1:length(x), times = length(time_seq)),
                         rep(1:length(time_seq), each = length(x)))
  
  # Flatten back into long vector
  z_signed_max <- as.vector(z_signed_max)
  
  vertical_lines(x, y, z_signed_max, time_seq, mesh_name)
}


vertical_finer   <- make_vertical_from_list(data_finer_list,   x_finer,   y_finer,   "finer")
vertical_coarser <- make_vertical_from_list(data_coarser_list, x_coarser, y_coarser, "coarser")

  
  # Compute ranges
  all_z <- unlist(lapply(all_data, function(df) df$z))
  x_range <- range(c(x_finer, x_coarser))
  y_range <- range(c(y_finer, y_coarser))
  z_range <- range(all_z)
  
  # --------- Plotly object ----------
  p <- plot_ly(frame = ~frame)
  
  # Add traces for finer + coarser (looping automatically with names)
  for (df in all_data) {
    p <- p %>%
      add_trace(data = df,
                x = ~x, y = ~y, z = ~z,
                type = "scatter3d", mode = "lines",
                line = list(width = 3),
                name = unique(df$mesh))
  }
  
  # Add baseline
  p <- p %>%
    add_trace(data = data_graph,
              x = ~x, y = ~y, z = ~z,
              type = "scatter3d", mode = "lines",
              line = list(color = "black", width = 2),
              name = "Graph", showlegend = FALSE)
  
# Add verticals (one per mesh, envelope of all functions)
if (!is.null(vertical_finer)) {
  p <- p %>%
    add_trace(data = vertical_finer,
              x = ~x, y = ~y, z = ~z,
              type = "scatter3d", mode = "lines",
              line = list(color = "gray", width = 0.5),
              name = "Vertical finer", showlegend = FALSE)
}
if (!is.null(vertical_coarser)) {
  p <- p %>%
    add_trace(data = vertical_coarser,
              x = ~x, y = ~y, z = ~z,
              type = "scatter3d", mode = "lines",
              line = list(color = "gray", width = 0.5),
              name = "Vertical coarser", showlegend = FALSE)
}

  
  frame_name <- deparse(substitute(frame_val_to_display))
  
  p <- p %>%
    layout(
      scene = global.scene.setter(x_range, y_range, z_range),
      updatemenus = list(list(
        type = "buttons", showactive = FALSE,
        buttons = list(
          list(label = "Play", method = "animate",
               args = list(NULL, list(frame = list(duration = 2000 / length(time_seq), redraw = TRUE),
                                      fromcurrent = TRUE))),
          list(label = "Pause", method = "animate",
               args = list(NULL, list(mode = "immediate", 
                                      frame = list(duration = 0), redraw = FALSE)))
        )
      )),
      title = paste0(frame_name, ": ", formatC(frame_val_to_display[1], format = "f", digits = 4))
    ) %>%
    plotly_build()
  
  # Update frame titles
  for (i in seq_along(p$x$frames)) {
    p$x$frames[[i]]$layout <- list(
      title = paste0(frame_name, ": ", formatC(frame_val_to_display[i], format = "f", digits = 4))
    )
  }
  
  return(p)
}

```


```
{r}
myggsave <- function(plot, width = 9.22, height = 7.05) {
  
  dir_to_save <- here::here("data_files/tikzpic")
  obj_name <- deparse(substitute(plot))
  tex_name <- file.path(dir_to_save, paste0(obj_name, ".tex"))
  
  # Create directory if it doesn't exist
  if (!dir.exists(dir_to_save)) dir.create(dir_to_save, recursive = TRUE)

  # FORCE LaTeX FONT HERE, comment the following if smth breaks
  library(tikzDevice)
  options(
  tikzDocumentDeclaration = "\\documentclass[tikz,border=0pt]{standalone}",
  tikzLatexPackages = c(
    "\\usepackage{tikz}",
    "\\usetikzlibrary{calc}",
    "\\usepackage[T1]{fontenc}",
    "\\usepackage[sc]{mathpazo}"
  ))
  # Comment until here if it fails
  # the above create this
  # \documentclass[tikz,border=0pt]{standalone}\usepackage{tikz}
  # \usetikzlibrary{calc}
  # \usepackage[T1]{fontenc}
  # \usepackage[sc]{mathpazo}\begin{document}
  # Save TikZ plot
  tikzDevice::tikz(tex_name, standAlone = TRUE, width = width, height = height)
  print(plot)
  dev.off()
  
  # Compile to PDF
  system(paste0("pdflatex -output-directory=", dir_to_save, " ", tex_name), 
         ignore.stdout = TRUE,
         ignore.stderr = TRUE)
  
  # Remove auxiliary files
  aux_ext <- c(".aux", ".log", ".tex")
  for (ext in aux_ext) {
    f <- file.path(dir_to_save, paste0(obj_name, ext))
    if (file.exists(f)) file.remove(f)
  }
  
  # Remove any temporary raster images generated by tikzDevice
  ras_files <- list.files(
    dir_to_save,
    pattern = paste0(obj_name, "_ras[0-9]+\\.png$"),
    full.names = TRUE
  )
  if (length(ras_files) > 0) file.remove(ras_files)
  
  cat("PDF saved at: ", file.path(dir_to_save, paste0(obj_name, ".pdf")))
}
```

```{r}
myggsave <- function(plot,
                     width = 9.22,
                     height = 7.05,
                     use_mathpazo = TRUE) {
  
  dir_to_save <- here::here("data_files/tikzpic")
  obj_name <- deparse(substitute(plot))
  
  # --- Filename suffix ---
  suffix <- if (use_mathpazo) "_with_mathpazo" else ""
  
  tex_name <- file.path(dir_to_save, paste0(obj_name, suffix, ".tex"))
  pdf_name <- file.path(dir_to_save, paste0(obj_name, suffix, ".pdf"))
  png_name <- file.path(dir_to_save, paste0(obj_name, suffix, ".png"))
  
  # Create directory if needed
  if (!dir.exists(dir_to_save)) dir.create(dir_to_save, recursive = TRUE)

  # --- TikZ + LaTeX options ---
  tikz_packages <- c(
    "\\usepackage{tikz}",
    "\\usetikzlibrary{calc}",
    "\\usepackage[T1]{fontenc}"
  )
  
  if (use_mathpazo) {
    tikz_packages <- c(tikz_packages, "\\usepackage[sc]{mathpazo}")
  }

  options(
    tikzDocumentDeclaration = "\\documentclass[tikz,border=0pt]{standalone}",
    tikzLatexPackages = tikz_packages
  )

  # Save TikZ plot
  tikzDevice::tikz(tex_name, standAlone = TRUE, width = width, height = height)
  print(plot)
  dev.off()
  
  # Compile to PDF
  system(
    paste0("pdflatex -output-directory=", dir_to_save, " ", tex_name),
    ignore.stdout = TRUE,
    ignore.stderr = TRUE
  )
  
  # --- Convert PDF to PNG (ImageMagick, no trimming) ---
  cmd <- paste(
    "convert -density 400",
    shQuote(pdf_name),
    shQuote(png_name)
  )
  
  system(cmd, ignore.stdout = TRUE, ignore.stderr = TRUE)
  
  if (!file.exists(png_name)) {
    stop("PNG conversion failed. Likely ImageMagick PDF policy restriction.")
  }
  
  # --- Clean auxiliary files ---
  aux_ext <- c(".aux", ".log", ".tex")
  for (ext in aux_ext) {
    f <- file.path(dir_to_save, paste0(obj_name, suffix, ext))
    if (file.exists(f)) file.remove(f)
  }
  
  ras_files <- list.files(
    dir_to_save,
    pattern = paste0(obj_name, suffix, "_ras[0-9]+\\.png$"),
    full.names = TRUE
  )
  if (length(ras_files) > 0) file.remove(ras_files)
  
  # --- Copy PNG to presentation folder ---
  pres_dir <- path.expand("~/Desktop/leninPresentations/data_files")
  if (!dir.exists(pres_dir)) dir.create(pres_dir, recursive = TRUE)
  
  pres_png <- file.path(pres_dir, basename(png_name))
  file.copy(png_name, pres_png, overwrite = TRUE)
  
  # --- Output ---
  cat("PDF saved at:", pdf_name, "\n",
      "PNG saved at:", png_name, "\n",
      "PNG copied to:", pres_png, "\n")
}
```


## Check norm identity {#norm_identity}

The following code builds matrix $\boldsymbol{\mathfrak{B}}$ in \eqref{matrixB} (object `big_matrix` below) and compares its 2-norm to that of matrix $\mathbf{T}$ in \eqref{matrixT} (object `TT` below) times $\tau^2$.

```{r, eval = FALSE, purl = FALSE}
# check norm identity
T_final <- 2
time_step <- 0.001 
h <- 1
kappa <- 15
alpha <- 0.5 
m = 1
beta <- alpha/2

graph <- gets.graph.tadpole(h = h)
graph$compute_fem()
G <- graph$mesh$G
C <- graph$mesh$C
L <- kappa^2*C + G
I <- Matrix::Diagonal(nrow(C))

# Numerical solution
obj <- my.fractional.operators.frac(L, beta, C, scale.factor = kappa^2, m = m, time_step)
partial_fraction_terms <- obj$partial_fraction_terms
residues <- obj$residues
output <- I*0
for (i in 1:(m+1)) {output <- output + residues[i] * solve(partial_fraction_terms[[i]], I)}
R <- output

C_sqrt <- expm::sqrtm(C)       # matrix square root
Omega <- C_sqrt %*% R %*% C_sqrt
n <- nrow(Omega)
Omega2 <- Omega %*% Omega
Omega3 <- Omega2 %*% Omega
Omega4 <- Omega2 %*% Omega2
Omega5 <- Omega3 %*% Omega2
Omega6 <- Omega3 %*% Omega3
B11 <- matrix(0, nrow = n, ncol = n)
B12 <- Omega2 + Omega4 + Omega6
B13 <- Omega3 + Omega5
B14 <- Omega4

B21 <- matrix(0, nrow = n, ncol = n)
B22 <- Omega3 + Omega5
B23 <- Omega2 + Omega4
B24 <- Omega3

B31 <- matrix(0, nrow = n, ncol = n)
B32 <- Omega4
B33 <- Omega3
B34 <- Omega2

B41 <- matrix(0, nrow = n, ncol = n)
B42 <- matrix(0, nrow = n, ncol = n)
B43 <- matrix(0, nrow = n, ncol = n)
B44 <- matrix(0, nrow = n, ncol = n)

big_matrix <- time_step^2 * rbind(
  cbind(B11, B12, B13, B14),
  cbind(B21, B22, B23, B24),
  cbind(B31, B32, B33, B34),
  cbind(B41, B42, B43, B44)
)

omega <- 1/(1+time_step * kappa^(2*beta))

TT <- build_T_fast(omega, 3)

time_step^2 * norm(TT, type = "2")
norm(big_matrix, type = "2")
```

## Comparison between $\gamma$ and its upper bound $1/(\mu \kappa^{4 \beta})$ {#contraction_constant_comparison}

```{r, eval = FALSE, purl = FALSE}
# --- 3. Parameter grid ---
T_final <- 2
tau_vector <- c(0.001, 0.01, 0.1)
N_vector <- T_final / tau_vector
kappa_vector <- c(1, 4, 16)
beta_vector <- c(0.5, 0.7, 0.9)
mu_vector <- c(0.1, 1, 10)

results <- expand.grid(
  tau = tau_vector,
  kappa = kappa_vector,
  beta = beta_vector,
  mu = mu_vector
)

# --- 4. Parallel computation with caching (A depends on ω and N only) ---
cache <- new.env(hash = TRUE)

results$L_c <- unlist(pbmclapply(
  1:nrow(results),
  function(i) {
    tau <- results$tau[i]
    kappa <- results$kappa[i]
    beta <- results$beta[i]
    mu <- results$mu[i]
    
    omega <- 1 / (1 + tau * kappa^(2 * beta))
    N <- as.integer(T_final / tau)
    key <- paste0(round(omega, 10), "_", N)
    
    if (exists(key, envir = cache)) {
      lambda_max <- get(key, envir = cache)
    } else {
      A <- build_T_fast(omega, N)
      lambda_max <- max(eigen(A, symmetric = TRUE, only.values = TRUE)$values)
      assign(key, lambda_max, envir = cache)
    }
    
    (tau^2 * lambda_max) / mu
  },
  mc.cores = parallel::detectCores()
))

save(results, file = here::here("data_files/contraction_constant_results2.RData"))
```


```{r, eval = FALSE, purl = FALSE}
library(pbmcapply)
library(ggplot2)
library(scales)
library(dplyr)
library(patchwork)
T_final <- 2
load(here::here("data_files/contraction_constant_results2.RData"))
all_results <- results %>% 
  mutate(upperbound = 1/(mu*kappa^(4*beta))) %>%
  mutate(T_final = T_final) %>%
  mutate(N = T_final/tau) %>%
  mutate(L_cless1 = L_c < 1) %>%
  mutate(Tlessmukappa2beta = T_final < (mu * kappa^(2*beta))) %>%
  mutate(onelessmukappa4beta  = upperbound < 1) 


# Find combined y-limits
ymin <- min(all_results$L_c, all_results$upperbound, na.rm = TRUE)
ymax <- max(all_results$L_c, all_results$upperbound, na.rm = TRUE)

p <- ggplot(all_results, aes(x = N, y = L_c,
                             color = factor(kappa),
                             shape = factor(beta),
                             linetype = factor(mu))) +
  geom_point(size = 3) +
  geom_line(aes(group = interaction(kappa, beta, mu)), alpha = 1) +
  scale_x_log10() +
  scale_y_log10(
    limits = c(ymin, ymax),
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = label_scientific()
  ) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  labs(x = "$N\\quad(\\log_{10}\\mbox{ scale})$",
       y = "$\\gamma =  \\tau^2\\|\\mathbf{T}\\|_2/\\mu\\quad(\\log_{10}\\mbox{ scale})$",
       color = "$\\kappa$",
       shape = "$\\beta$",
       linetype = "$\\mu$") +
  theme_minimal(base_size = 14, base_family = "Palatino")

q <- ggplot(all_results, aes(x = N, y = upperbound,
                             color = factor(kappa),
                             shape = factor(beta),
                             linetype = factor(mu))) +
  geom_point(size = 3) +
  geom_line(aes(group = interaction(kappa, beta, mu)), alpha = 1) +
  scale_x_log10() +
  scale_y_log10(
    limits = c(ymin, ymax),
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = label_scientific(),
    position = "right"
  ) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  labs(x = "$N\\quad(\\log_{10}\\mbox{ scale})$",
       y = "$\\mathtt{upper bound}=1/(\\mu\\kappa^{4\\beta})\\quad(\\log_{10}\\mbox{ scale})$",# expression(1 / mu * kappa^{4 * beta}),
       color = "$\\kappa$",
       shape = "$\\beta$",
       linetype = "$\\mu$") +
  theme_minimal(base_size = 14, base_family = "Palatino")


fixedpointconvergence_combined2 <- (p | q) + 
  # plot_annotation(
  #   title = "$\\mbox{Contraction constant }\\gamma=\\tau^2\\|\\mathbf{T}\\|_2/\\mu\\mbox{ and its upper bound }\\mathtt{upper bound}=1/(\\mu\\kappa^{4\\beta})$",
  #   theme = theme(plot.title = element_text(size = 18, #face = "bold", 
  #                                           hjust = 0.5, 
  #                                           family = "Palatino"))
  # ) +
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

myggsave(fixedpointconvergence_combined2, width = 12, height = 7, use_mathpazo = FALSE)
```


```{r, eval = TRUE, purl = FALSE, fig.align='center', out.width="1200px", out.height="700px", fig.cap = captioner("Comparison of the contraction constant $\\gamma =  \\tau^2\\|\\mathbf{T}\\|_2/\\mu$ and its theoretical upper bound $1 / (\\mu \\kappa^{4 \\beta})$ as functions of the sample size $N$ for $T = 2$. Different colors, shapes, and line types correspond to variations in $\\kappa$, $\\beta$, and $\\mu$, respectively. Both plots have their x- and y-axes on a $\\log_{10}$ scale, and they share the same y-axis limits for direct comparability.")}
knitr::include_graphics(here::here("data_files/tikzpic/fixedpointconvergence_combined2.pdf"))
```



## References

```{r, purl = FALSE}
grateful::cite_packages(output = "paragraph", out.dir = ".")
```


