---
title: "Optimal control of fractional diffusion equations on metric graphs"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
  - \newcommand{\almosteverywhere}{\mathrm{a.e.}\;}
---

Go back to the [Contents](about.html) page.

<div style="color: #2c3e50; text-align: right;">
********  
<strong>Press Show to reveal the code chunks.</strong>  

********
</div>


```{r}
# Create a clipboard button on the rendered HTML page
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = FALSE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```

```{r}
library(MetricGraph)
library(ggplot2)
library(reshape2)
library(plotly)
library(tidyr)
```


```{r}
capture.output(
  knitr::purl(here::here("control_functionality.Rmd"), output = here::here("control_functionality.R")),
  file = here::here("purl_log.txt")
)
source(here::here("control_functionality.R"))
```

# Optimal control of fractional diffusion equations on metric graphs

## Problem statement {#optimal_control_problem}

Let $\Gamma = (\mathcal{V},\mathcal{E})$ be a metric graph. Let $u_d: \Gamma \times(0, T) \rightarrow \mathbb{R}$ be the desired state and  $\mu>0$ a regularization parameter. We define the cost functional

\begin{equation}
\label{eq:costfun}
\tag{1}
    J(u, z)=\frac{1}{2} \int_0^T\left(\left\|u-u_d\right\|_{L_2(\Gamma)}^2+\mu\|z\|_{L_2(\Gamma)}^2\right) dt
\end{equation}


Let $f:\Gamma\times (0,T)\rightarrow\ar$ and $u_0: \Gamma \rightarrow \mathbb{R}$ be fixed functions. We will call them right-hand side and initial datum, respectively. Let $\alpha\in(0,2]$ and $z: \Gamma \times(0, T) \rightarrow \mathbb{R}$ denote the control variable. We shall be concerned with the following PDE-constrained optimization problem: Find

\begin{equation}
\label{eq:min_pro}
\tag{2}
    \min\; J(u, z)
\end{equation}
subject to the fractional diffusion equation
\begin{equation}
\label{eq:maineq}
\tag{3}
\left\{
\begin{aligned}
    \partial_t u(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s,t) &= f(s,t)+z(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    u(s,0) &= u_0(s), && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
with $u(\cdot,t)$ satisfying the Kirchhoff vertex conditions
\begin{equation}
\label{eq:Kcond}
\tag{4}
   \mathcal{K} =  \left\{\phi\in C(\Gamma)\;\middle|\; \forall v\in \mathcal{V}:\; \sum_{e\in\mathcal{E}_v}\partial_e \phi(v)=0 \right\}
\end{equation}
and the control constraints 
\begin{align}
\label{control_constraints}
\tag{5}
    a(s,t)\leq z(s,t)\leq b(s,t)\;\almosteverywhere (s,t)\in\Gamma \times(0, T).
\end{align}


## Numerical scheme {#num_scheme_optim_control}

The optimal variables $(\bar{u}, \bar{p}, \bar{z})$ satisfy

\begin{equation}
\label{eq:maineqoptimal}
\tag{6}
\left\{
\begin{aligned}
    \partial_t \bar{u}(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{u}(s,t) &= f(s,t)+\bar{z}(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    \bar{u}(s,0) &= u_0(s), && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
and
\begin{equation}
\label{eq:adjointeq}
\tag{7}
\left\{
\begin{aligned}
    -\partial_t \bar{p}(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{p}(s,t) &= \bar{u}(s,t)-u_d(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    \bar{p}(s,T) &= 0, && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
with 
\begin{align}
\label{zz}
\tag{8}
        \bar{z}(s,t) = \max\left\{a(s,t),\min\left\{b(s,t),-\dfrac{1}{\mu}\bar{p}(s,t)\right\}\right\}.
\end{align}
By considering the change of variable $\bar{q}(s,t) = \bar{p}(s,T-t)$, the fractional adjoint problem \eqref{eq:adjointeq} becomes a forward-in-time problem where the transformed adjoint state $\bar{q}$ solves
\begin{equation}
\label{transformed_adjoint_state}
\tag{9}
\left\{
\begin{aligned}
    \partial_t \bar{q}(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{q}(s,t) &= \bar{v}(s,t)-v_d(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    \bar{q}(s,0) &= 0, && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
since $\partial_t\bar{q}(s,t) = -\partial_t\bar{p}(s,T-t)$ and $\bar{q}(s, 0) =\bar{p}(s,T-0)= \bar{p}(s,T)=0$. Here, $\bar{v}(s,t) = \bar{u}(s,T-t)$ and $v_d(s,t) = u_d(s,T-t)$. 


Given $\bar{u}$ and $u_d$, we can time-reverse them to obtain $\bar{v}$ and $v_d$ and then use the same numerical scheme we use for the forward problem \eqref{eq:maineqoptimal} to solve the adjoint problem \eqref{transformed_adjoint_state}. The control variable $\bar{z}$ is then computed using \eqref{zz}.



The numerical scheme for \eqref{eq:maineqoptimal} and \eqref{transformed_adjoint_state} are given by (see the [Functionality](functionality.html#num_scheme) page)

\begin{align}
\tag{10}
    \label{numericalscheme1}
    \begin{cases}
    &\mathbf{\bar{U}}_{k+1}  = \mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_k+\tau (\mathbf{F}_{k+1}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{k+1})),\quad k = 0,\dots, N-1,\\
    &\mathbf{\bar{U}}_{0}  = P_h u_0,
    \end{cases}
\end{align}
and
\begin{align}
\tag{11}
    \label{thenumericalscheme2}
    \begin{cases}
    &\mathbf{\bar{Q}}_{k+1}  = \mathbf{R}(\mathbf{C}\mathbf{\bar{Q}}_k+\tau ((\mathbf{C}\mathbf{\bar{U}}\mathbf{J}_{N+1})_{k+1}-\mathbf{D}_{k+1})),\quad k = 0,\dots, N-1,\\
    &\mathbf{\bar{Q}}_{0}  = \mathbf{0},
    \end{cases}
\end{align}
where $\mathbf{R} = \sum_{k=1}^{m+1} a_k\left(\mathbf{L}/\kappa^2-p_k\mathbf{C}\right)^{-1}$. Observe that \eqref{numericalscheme1}-\eqref{thenumericalscheme2} is a coupled problem.

Here

- $\mathbf{\bar{U}}$ has entries $\mathbf{U}_{j,k}  = \bar{u}(s_j,t_k)$
- $\mathbf{F}$ has entries $\mathbf{F}_{j,k} =(f^{k},\psi^j_h)_{L_2(\Gamma)}$.
- $\mathbf{\bar{\mathcal{Z}}}$ has entries $\mathbf{\bar{\mathcal{Z}}}_{j,k} = \bar{z}(s_j,t_k)$
- $\mathbf{D}$ has entries $\mathbf{D}_{j,k} =(u_d^{N-k},\psi^j_h)_{L_2(\Gamma)}$
- $\mathbf{\bar{P}} = \mathbf{\bar{Q}}\mathbf{J}_{N+1}$ and has entries $\mathbf{\bar{P}}_{j,k} = \bar{p}(s_j,t_k)$


If we change \eqref{thenumericalscheme2} to $\mathbf{\bar{P}}$, then we obtain 
\begin{align}
\tag{12}
    \label{thenumericalscheme3}
    \begin{cases}
    &\mathbf{\bar{P}}_{k}  = \mathbf{R}(\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau (\mathbf{C}\mathbf{\bar{U}}_{k}-\mathbf{D}_{k})),\quad k = N-1,\dots, 0,\\
    &\mathbf{\bar{P}}_{N}  = \mathbf{0},
    \end{cases}
\end{align}

## Numerical scheme derivation for the state equation

\begin{equation}
\label{system:fully_discrete_scheme}
\tag{3}
    \left\{
\begin{aligned}
    \langle\delta U_h^{k+1},\phi\rangle + \mathfrak{a}(
        U_h^{k+1},\phi) &= \langle f^{k+1}+Z_h^{k+1},\phi\rangle, \quad\forall\phi\in V_h,\quad k=0,\dots, N-1, \\
    U^0_h &= P_hu_0, 
\end{aligned}
\right.
\end{equation}
We have 
\begin{align}
\langle\dfrac{U_h^{k+1} - U_h^{k}}{\tau},\phi\rangle + \mathfrak{a}(
        U_h^{k+1},\phi) = \langle f^{k+1}+Z_h^{k+1},\phi\rangle
\end{align}
and
\begin{align}
\langle U_h^{k+1},\phi\rangle + \tau\mathfrak{a}(
        U_h^{k+1},\phi) = \langle U_h^{k},\phi\rangle + \tau\langle f^{k+1}+Z_h^{k+1},\phi\rangle
\end{align}
At each time step $t_k$, the finite element solution $U_h^k\in V_h$ to \eqref{system:fully_discrete_scheme} can be expressed as a linear combination of the basis functions  $\{\psi^i_h\}_{i=1}^{N_h}$ introduced in the [Preliminaries](preliminaries.html#fem-basis) page, namely, 
\begin{align}
\label{num_sol}
\tag{4}
    U_h^k(s) =  \sum_{i=1}^{N_h}u_i^k\psi^i_h(s), \;s\in\Gamma.
\end{align}
Replacing \eqref{num_sol} into \eqref{system:fully_discrete_scheme} yields the following linear system
\begin{align*}
    \sum_{j=1}^{N_h}u_j^{k+1}[(\psi_j,\psi_i)_{L_2(\Gamma)}+ \tau\mathfrak{a}(\psi_j,\psi_i)] = \sum_{j=1}^{N_h}u_j^{k}(\psi_j,\psi_i)_{L_2(\Gamma)}+\tau( f^{k+1},\psi_i)_{L_2(\Gamma)},
\end{align*}
for $i = 1,\dots, N_h$. In matrix notation,
\begin{align}
\label{diff_eq_discrete}
    (\mathbf{C}+\tau \mathbf{L}^{\alpha/2})\mathbf{U}^{k+1} = \mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1},
\end{align}
or by introducing the scaling parameter $\kappa^2>0$,
\begin{align}
    (\mathbf{C}+\tau (\kappa^2)^{\alpha/2}(\mathbf{L}/\kappa^2)^{\alpha/2})\mathbf{U}^{k+1} = \mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1},
\end{align}
where $\mathbf{C}$ has entries $\mathbf{C}_{ij} = (\psi_j,\psi_i)_{L_2(\Gamma)}$, $\mathbf{L}^{\alpha/2}$ has entries $\mathfrak{a}(\psi_j,\psi_i)$, $\mathbf{U}^k$ has entries $u_j^k$, and $\mathbf{F}^k$ has entries $( f^{k},\psi_i)_{L_2(\Gamma)}$. Applying $(\mathbf{L}/\kappa^2)^{-\alpha/2}$ to both sides yields
\begin{equation}
((\mathbf{L}/\kappa^2)^{-\alpha/2}\mathbf{C}+\tau (\kappa^2)^{\alpha/2}\mathbf{I})\mathbf{U}^{k+1} = (\mathbf{L}/\kappa^2)^{-\alpha/2}(\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}).
\end{equation}
Following @rSPDE2020, we approximate $(\mathbf{L}/\kappa^2)^{-\alpha/2}$ by $\mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top$ to arrive at
\begin{equation}
\label{eq:scheme2}
\tag{5}
(\mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top \mathbf{C}+\tau(\kappa^2)^{\alpha/2} \mathbf{I})\mathbf{U}^{k+1} = \mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top(\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}).
\end{equation}
where
\begin{equation}
\label{eq:PLPRbolin}
\tag{7}
\mathbf{P}_r = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right)\quad\text{and}\quad \mathbf{P}_\ell = \dfrac{\kappa^{2\beta}}{\texttt{factor}}\mathbf{C}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right),
\end{equation}
and $\texttt{factor} = \dfrac{c_m}{b_{m+1}}$, and
$\{r_{1i}\}_{i=1}^m$ and $\{r_{2j}\}_{j=1}^{m+1}$ are the roots of $q_1(x) =\sum_{i=0}^mc_ix^{i}$ and  $q_2(x)=\sum_{j=0}^{m+1}b_jx^{j}$, respectively. The coefficients  $\{c_i\}_{i=0}^m$ and  $\{b_j\}_{j=0}^{m+1}$ are determined as the best rational approximation $\hat{r}_m =q_1/q_2$ of the function $\hat{f}(x) := x^{\alpha/2-1}$ over the interval $J_h: = [\kappa^{2}\lambda_{N_h,h}^{-1}, \kappa^{2}\lambda_{1,h}^{-1}]$, where $\lambda_{1,h}, \lambda_{N_h,h}>0$ are the smallest and the largest eigenvalue of $L_h$, respectively. With this under consideration, we can rewrite \eqref{eq:scheme2} as
\begin{equation}
\tag{8}
(\mathbf{P}_r^\top \mathbf{C}+\tau \mathbf{P}_\ell^\top)\mathbf{U}^{k+1} = \mathbf{P}_r^\top(\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}),
\label{eq:scheme}
\end{equation}
where  
\begin{equation}
\mathbf{P}_r^\top = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\quad\text{and}\quad \mathbf{P}_\ell^\top = \dfrac{\kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot \mathbf{C}
\end{equation}
since $\mathbf{L}$ and $\mathbf{C}^{-1}$ are symmetric and the factors in the product commute. Replacing these two into \eqref{eq:scheme} yields
\begin{equation}
\left(\prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\right)\mathbf{C}\mathbf{U}^{k+1} = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot (\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}),
\end{equation}
that is,
\begin{equation}
\label{eq:final_scheme}
\tag{9}
\mathbf{U}^{k+1} = \mathbf{C}^{-1}\left(\prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\right)^{-1} \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot (\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}).
\end{equation}
Considering the partial fraction decomposition
\begin{equation}
\label{eq:partial_fraction}
\tag{10}
\dfrac{\prod_{i=1}^m (1-r_{1i}x)}{\prod_{i=1}^m (1-r_{1i}x)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}} \prod_{j=1}^{m+1} (1-r_{2j}x)}=\sum_{k=1}^{m+1} a_k(x-p_k)^{-1} + r,
\end{equation}
scheme \eqref{eq:final_scheme} can be expressed as
\begin{equation}
\label{eq:final_scheme2}
\tag{11}
\mathbf{U}^{k+1} = \mathbf{C}^{-1}\left(\sum_{k=1}^{m+1} a_k\left( \dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}-p_k\mathbf{I}\right)^{-1} + r\mathbf{I}\right) (\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1})
\end{equation}

In practice, since the rational function in \eqref{eq:partial_fraction} is proper, there is no remainder $r$. Moreover, since $\left( \dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}-p_k\mathbf{I}\right)^{-1}  = \mathbf{C}\left( \dfrac{\mathbf{L}}{\kappa^2}-p_k\mathbf{C}\right)^{-1}$, we have that \eqref{eq:final_scheme2} can be rewritten as

\begin{equation}
\label{eq:final_scheme3}
\tag{12}
\mathbf{U}^{k+1} = \left(\sum_{k=1}^{m+1} a_k\left( \dfrac{\mathbf{L}}{\kappa^2}-p_k\mathbf{C}\right)^{-1}\right) (\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}).
\end{equation}


## Numerical scheme derivation for the adjoint equation

\begin{equation}
\label{discrete_adjoint}
    \left\{
\begin{aligned}
    \langle\bar{\delta} P_h^{k},\phi\rangle + \mathfrak{a}(P_h^{k},\phi) &= \langle U_h^{k+1}-u_d^{k+1},\phi\rangle, \quad\forall\phi\in V_h,\quad k=N-1,\dots, 0, \\
    P^N_h &= 0
\end{aligned}
\right.
\end{equation}
We have 
\begin{align}
\langle\dfrac{P_h^{k} - P_h^{k+1}}{\tau},\phi\rangle + \mathfrak{a}(
        P_h^{k},\phi) = \langle U_h^{k+1} - u_d^{k+1},\phi\rangle
\end{align}
and
\begin{align}
\langle P_h^{k},\phi\rangle + \tau\mathfrak{a}(
        P_h^{k},\phi) = \langle P_h^{k+1},\phi\rangle + \tau\langle U_h^{k+1} - u_d^{k+1},\phi\rangle
\end{align}

At each time step $t_k$, the finite element solution $P_h^k\in V_h$ to \eqref{system:fully_discrete_scheme} can be expressed as a linear combination of the basis functions  $\{\psi^i_h\}_{i=1}^{N_h}$ introduced in the [Preliminaries](preliminaries.html#fem-basis) page, namely, 
\begin{align}
\label{num_sol2}
\tag{4}
    P_h^k(s) =  \sum_{i=1}^{N_h}p_i^k\psi^i_h(s), \;s\in\Gamma.
\end{align}
Replacing \eqref{num_sol} into \eqref{system:fully_discrete_scheme} yields the following linear system
\begin{align*}
    \sum_{j=1}^{N_h}p_j^{k}[(\psi_j,\psi_i)_{L_2(\Gamma)}+ \tau\mathfrak{a}(\psi_j,\psi_i)] = \sum_{j=1}^{N_h}p_j^{k+1}(\psi_j,\psi_i)_{L_2(\Gamma)}+\tau(U_h^{k+1} - u_d^{k+1},\psi_i)_{L_2(\Gamma)},
\end{align*}
for $i = 1,\dots, N_h$. In matrix notation,
\begin{align}
\label{diff_eq_discrete_adjoint}
    (\mathbf{C}+\tau \mathbf{L}^{\alpha/2})\mathbf{P}^{k} = \mathbf{C}\mathbf{P}^{k+1}+\tau \mathbf{U}^{k+1},
\end{align}
or by introducing the scaling parameter $\kappa^2>0$,
\begin{align}
    (\mathbf{C}+\tau (\kappa^2)^{\alpha/2}(\mathbf{L}/\kappa^2)^{\alpha/2})\mathbf{P}^{k} = \mathbf{C}\mathbf{P}^{k+1}+\tau \mathbf{U}^{k+1},
\end{align}
where $\mathbf{C}$ has entries $\mathbf{C}_{ij} = (\psi_j,\psi_i)_{L_2(\Gamma)}$, $\mathbf{L}^{\alpha/2}$ has entries $\mathfrak{a}(\psi_j,\psi_i)$, $\mathbf{U}^k$ has entries $u_j^k$, and $\mathbf{F}^k$ has entries $( f^{k},\psi_i)_{L_2(\Gamma)}$. Applying $(\mathbf{L}/\kappa^2)^{-\alpha/2}$ to both sides yields
\begin{equation}
((\mathbf{L}/\kappa^2)^{-\alpha/2}\mathbf{C}+\tau (\kappa^2)^{\alpha/2}\mathbf{I})\mathbf{P}^{k} = (\mathbf{L}/\kappa^2)^{-\alpha/2}(\mathbf{C}\mathbf{P}^{k+1}+\tau \mathbf{U}^{k+1}).
\end{equation}
Following @rSPDE2020, we approximate $(\mathbf{L}/\kappa^2)^{-\alpha/2}$ by $\mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top$ to arrive at
\begin{equation}
\label{eq:scheme2adjoint}
\tag{5}
(\mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top \mathbf{C}+\tau(\kappa^2)^{\alpha/2} \mathbf{I})\mathbf{P}^{k} = \mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top(\mathbf{C}\mathbf{P}^{k+1}+\tau \mathbf{U}^{k+1}).
\end{equation}
where
\begin{equation}
\label{eq:PLPRbolinadjoint}
\tag{7}
\mathbf{P}_r = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right)\quad\text{and}\quad \mathbf{P}_\ell = \dfrac{\kappa^{2\beta}}{\texttt{factor}}\mathbf{C}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right),
\end{equation}
and $\texttt{factor} = \dfrac{c_m}{b_{m+1}}$, and
$\{r_{1i}\}_{i=1}^m$ and $\{r_{2j}\}_{j=1}^{m+1}$ are the roots of $q_1(x) =\sum_{i=0}^mc_ix^{i}$ and  $q_2(x)=\sum_{j=0}^{m+1}b_jx^{j}$, respectively. The coefficients  $\{c_i\}_{i=0}^m$ and  $\{b_j\}_{j=0}^{m+1}$ are determined as the best rational approximation $\hat{r}_m =q_1/q_2$ of the function $\hat{f}(x) := x^{\alpha/2-1}$ over the interval $J_h: = [\kappa^{2}\lambda_{N_h,h}^{-1}, \kappa^{2}\lambda_{1,h}^{-1}]$, where $\lambda_{1,h}, \lambda_{N_h,h}>0$ are the smallest and the largest eigenvalue of $L_h$, respectively. With this under consideration, we can rewrite \eqref{eq:scheme2} as
\begin{equation}
\tag{8}
(\mathbf{P}_r^\top \mathbf{C}+\tau \mathbf{P}_\ell^\top)\mathbf{P}^{k} = \mathbf{P}_r^\top(\mathbf{C}\mathbf{P}^{k+1}+\tau \mathbf{U}^{k+1}),
\label{eq:schemeadjoint}
\end{equation}
where  
\begin{equation}
\mathbf{P}_r^\top = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\quad\text{and}\quad \mathbf{P}_\ell^\top = \dfrac{\kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot \mathbf{C}
\end{equation}
since $\mathbf{L}$ and $\mathbf{C}^{-1}$ are symmetric and the factors in the product commute. Replacing these two into \eqref{eq:scheme} yields
\begin{equation}
\left(\prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\right)\mathbf{C}\mathbf{P}^{k} = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot (\mathbf{C}\mathbf{P}^{k+1}+\tau \mathbf{U}^{k+1}),
\end{equation}
that is,
\begin{equation}
\label{eq:final_schemeadjoint}
\tag{9}
\mathbf{P}^{k} = \mathbf{C}^{-1}\left(\prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\right)^{-1} \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot (\mathbf{C}\mathbf{P}^{k+1}+\tau \mathbf{U}^{k+1}).
\end{equation}
Considering the partial fraction decomposition
\begin{equation}
\label{eq:partial_fractionadjoint}
\tag{10}
\dfrac{\prod_{i=1}^m (1-r_{1i}x)}{\prod_{i=1}^m (1-r_{1i}x)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}} \prod_{j=1}^{m+1} (1-r_{2j}x)}=\sum_{k=1}^{m+1} a_k(x-p_k)^{-1} + r,
\end{equation}
scheme \eqref{eq:final_scheme} can be expressed as
\begin{equation}
\label{eq:final_scheme3adjoint}
\tag{12}
\mathbf{P}^{k} = \left(\sum_{k=1}^{m+1} a_k\left( \dfrac{\mathbf{L}}{\kappa^2}-p_k\mathbf{C}\right)^{-1}\right) (\mathbf{C}\mathbf{P}^{k+1}+\tau \mathbf{U}^{k+1}).
\end{equation}

## Fixed point iteration

\begin{align}
\tag{10}
    \label{numstate}
    \begin{cases}
    &\mathbf{\bar{U}}_{k+1}  = \mathbf{R}\mathbf{C}\mathbf{\bar{U}}_k+\tau \mathbf{R}(\mathbf{F}_{k+1}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{k+1}),\quad k = 0,\dots, N-1,\\
    &\mathbf{\bar{U}}_{0}  = P_h u_0,
    \end{cases}
\end{align}
and
\begin{align}
\tag{12}
    \label{numadjoint}
    \begin{cases}
    &\mathbf{\bar{P}}_{k}  = \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau \mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{k+1}-\mathbf{D}_{k+1}),\quad k = N-1,\dots, 0,\\
    &\mathbf{\bar{P}}_{N}  = \mathbf{0},
    \end{cases}
\end{align}

Let $N=3$. Then
\begin{align}
\mathbf{\bar{U}}_{0} & = \mathbf{\bar{U}}_{0} \\
& \\
\mathbf{\bar{U}}_{1} & =  (\mathbf{R}\mathbf{C})^1\mathbf{\bar{U}}_0+\tau \mathbf{R}(\mathbf{F}_{1}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{1})\\
& \\
\mathbf{\bar{U}}_{2} & =  \mathbf{R}\mathbf{C}\mathbf{\bar{U}}_1+\tau \mathbf{R}(\mathbf{F}_{2}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{2})\\
& = \mathbf{R}\mathbf{C}((\mathbf{R}\mathbf{C})^1\mathbf{\bar{U}}_0+\tau \mathbf{R}(\mathbf{F}_{1}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{1}))+\tau \mathbf{R}(\mathbf{F}_{2}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{2})\\
& = (\mathbf{R}\mathbf{C})^2\mathbf{\bar{U}}_0+\tau (\mathbf{R}\mathbf{C})^1\mathbf{R}(\mathbf{F}_{1}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{1})+\tau \mathbf{R}(\mathbf{F}_{2}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{2})\\
& \\
\mathbf{\bar{U}}_{3} & =  \mathbf{R}\mathbf{C}\mathbf{\bar{U}}_2+\tau \mathbf{R}(\mathbf{F}_{3}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{3})\\
& =  \mathbf{R}\mathbf{C}((\mathbf{R}\mathbf{C})^2\mathbf{\bar{U}}_0+\tau (\mathbf{R}\mathbf{C})^1\mathbf{R}(\mathbf{F}_{1}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{1})+\tau \mathbf{R}(\mathbf{F}_{2}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{2}))+\tau \mathbf{R}(\mathbf{F}_{3}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{3})\\
& =  (\mathbf{R}\mathbf{C})^3\mathbf{\bar{U}}_0+\tau (\mathbf{R}\mathbf{C})^2\mathbf{R}(\mathbf{F}_{1}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{1})+\tau (\mathbf{R}\mathbf{C})^1\mathbf{R}(\mathbf{F}_{2}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{2})+\tau \mathbf{R}(\mathbf{F}_{3}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{3})
\end{align}
and
\begin{align}
\mathbf{\bar{P}}_{3} &= \mathbf{0}, \\
&\\
\mathbf{\bar{P}}_{2} &= \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{3} + \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{3}-\mathbf{D}_{3})\\
&= \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{3}-\mathbf{D}_{3}), \\
&\\
\mathbf{\bar{P}}_{1} &= \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{2} + \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{2}-\mathbf{D}_{2}) \\
&= \tau(\mathbf{R}\mathbf{C})^1\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{3}-\mathbf{D}_{3})
+ \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{2}-\mathbf{D}_{2}), \\
&\\
\mathbf{\bar{P}}_{0} &= \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{1} + \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{1}-\mathbf{D}_{1}) \\
&= \tau(\mathbf{R}\mathbf{C})^{2}\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{3}-\mathbf{D}_{3})
+ \tau(\mathbf{R}\mathbf{C})^1\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{2}-\mathbf{D}_{2})
+ \tau\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{1}-\mathbf{D}_{1}).
\end{align}



In block-matrix notation
\begin{align}
\begin{bmatrix}
\mathbf{\bar{U}}_{0} \\[1mm]
\mathbf{\bar{U}}_{1} \\[1mm]
\mathbf{\bar{U}}_{2} \\[1mm]
\mathbf{\bar{U}}_{3}
\end{bmatrix}
= 
\begin{bmatrix}
\mathbf{\bar{U}}_{0} \\[1mm]
(\mathbf{R}\mathbf{C})^1\mathbf{\bar{U}}_{0} \\[1mm]
(\mathbf{R}\mathbf{C})^2\mathbf{\bar{U}}_{0} \\[1mm]
(\mathbf{R}\mathbf{C})^3\mathbf{\bar{U}}_{0}
\end{bmatrix}
+\tau
\begin{bmatrix}
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{0}  \\[1mm]
\mathbf{0} & \mathbf{I} & \mathbf{0} & \mathbf{0}  \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^1 & \mathbf{I} & \mathbf{0}  \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^2 & (\mathbf{R}\mathbf{C})^1 & \mathbf{I} 
\end{bmatrix}
\begin{bmatrix}
\mathbf{R}(\mathbf{\bar{F}}_{0}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{0}) \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{1}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{1}) \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{2}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{2}) \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{3}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{3})
\end{bmatrix}
\end{align}
and
\begin{align}
\begin{bmatrix}
\mathbf{\bar{P}}_{0} \\[1mm]
\mathbf{\bar{P}}_{1} \\[1mm]
\mathbf{\bar{P}}_{2} \\[1mm]
\mathbf{\bar{P}}_{3}
\end{bmatrix}
=
\tau
\begin{bmatrix}
\mathbf{0} & \mathbf{I} & (\mathbf{R}\mathbf{C}) & (\mathbf{R}\mathbf{C})^{2} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{I} & (\mathbf{R}\mathbf{C}) \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{I} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{0}
\end{bmatrix}
\begin{bmatrix}
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{0}-\mathbf{D}_{0}) \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{1}-\mathbf{D}_{1}) \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{2}-\mathbf{D}_{2}) \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{3}-\mathbf{D}_{3})
\end{bmatrix}.
\end{align}



In general
\begin{align}
\begin{bmatrix}
\mathbf{\bar{U}}_{0} \\[1mm]
\mathbf{\bar{U}}_{1} \\[1mm]
\mathbf{\bar{U}}_{2} \\[1mm]
\vdots \\[1mm]
\mathbf{\bar{U}}_{N}
\end{bmatrix}
&=
\begin{bmatrix}
\mathbf{\bar{U}}_{0} \\[1mm]
(\mathbf{R}\mathbf{C})^1\mathbf{\bar{U}}_{0} \\[1mm]
(\mathbf{R}\mathbf{C})^2\mathbf{\bar{U}}_{0} \\[1mm]
\vdots \\[1mm]
(\mathbf{R}\mathbf{C})^N\mathbf{\bar{U}}_{0}
\end{bmatrix}

+\tau
\begin{bmatrix}
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & \mathbf{I} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^1 & \mathbf{I} & \cdots & \mathbf{0} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^{N-1} & (\mathbf{R}\mathbf{C})^{N-2} & \cdots & \mathbf{I}
\end{bmatrix}
\begin{bmatrix}
\mathbf{R}(\mathbf{\bar{F}}_{0}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{0}) \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{1}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{1}) \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{2}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{2}) \\[1mm]
\vdots \\[1mm]
\mathbf{R}(\mathbf{\bar{F}}_{N}+\mathbf{C}\mathbf{\bar{\mathcal{Z}}}_{N})
\end{bmatrix}.
\end{align}
and \begin{align}
\begin{bmatrix}
\mathbf{\bar{P}}_{0} \\[1mm]
\mathbf{\bar{P}}_{1} \\[1mm]
\vdots \\[1mm]
\mathbf{\bar{P}}_{N-1} \\[1mm]
\mathbf{\bar{P}}_{N}
\end{bmatrix}
=
\tau
\begin{bmatrix}
\mathbf{0} & \mathbf{I} & (\mathbf{R}\mathbf{C}) & \cdots & (\mathbf{R}\mathbf{C})^{N-1} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{I} & \cdots & (\mathbf{R}\mathbf{C})^{N-2} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{I} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0}
\end{bmatrix}
\begin{bmatrix}
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{0}-\mathbf{D}_{0}) \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{1}-\mathbf{D}_{1}) \\[1mm]
\vdots \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{N-1}-\mathbf{D}_{N-1}) \\[1mm]
\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{N}-\mathbf{D}_{N})
\end{bmatrix}.
\end{align}
By defining stacked vectors of length $(N+1)N_h$ 
\begin{align*}
 \mathbf{\bar{u}} = \text{vec}(\mathbf{\bar{U}}),\quad 
\mathbf{\bar{p}} = \text{vec}(\mathbf{\bar{P}}),\quad
\boldsymbol{\bar{z}} = \text{vec}(\boldsymbol{\bar{\mathcal{Z}}}),\quad \mathbf{f} = \text{vec}(\mathbf{F}),\quad
\mathbf{d} =\text{vec}(\mathbf{D}),
\end{align*}
\begin{align*}
    \mathbf{a} =\text{vec}(\mathbf{A}),\quad
\mathbf{b} = \text{vec}(\mathbf{B}),\quad\mathbf{h}= [\mathbf{\bar{U}}_{0}^\top ((\mathbf{R}\mathbf{C})^{1}\mathbf{\bar{U}}_{0})^\top ((\mathbf{R}\mathbf{C})^{2} \mathbf{\bar{U}}_{0})^\top\dots ((\mathbf{R}\mathbf{C})^{N}\mathbf{\bar{U}}_{0})^\top]^\top,
\end{align*}
and block matrices of dimension $(N+1)N_h\times (N+1)N_h$
\begin{align*}
\mathbf{M} = \mathbf{I}_{N+1}\otimes  \mathbf{C},\quad \mathbf{\hat{J}} = \mathbf{J}_{N+1}\otimes \mathbf{I}_{N_h},\quad \mathbf{\hat{R}} = \mathbf{I}_{N+1}\otimes  \mathbf{R} ,
\end{align*}
\begin{align}
\mathbf{S}_{\mathrm{F}}
= \tau
\begin{bmatrix}
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & \mathbf{I} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^1 & \mathbf{I} & \cdots & \mathbf{0} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^{N-1} & (\mathbf{R}\mathbf{C})^{N-2} & \cdots & \mathbf{I}
\end{bmatrix}\quad \text{and}\quad
\mathbf{S}_{\mathrm{B}}
= \tau
\begin{bmatrix}
\mathbf{0} & \mathbf{I} & (\mathbf{R}\mathbf{C}) & \cdots & (\mathbf{R}\mathbf{C})^{N-1} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{I} & \cdots & (\mathbf{R}\mathbf{C})^{N-2} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{I} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0}
\end{bmatrix}
\end{align}
then the vectorized form 
\begin{align}
\label{iterationvectorized}
    \begin{cases}
    \mathbf{\bar{u}} &= \mathbf{h} + \mathbf{S}_{\mathrm{F}}\mathbf{\hat{R}}(\mathbf{f}+\mathbf{M}\boldsymbol{\bar{z}})\\
    \mathbf{\bar{p}} & = \mathbf{S}_{\mathrm{B}}\mathbf{\hat{R}}(\mathbf{M}\mathbf{\bar{u}} - \mathbf{d})\\
    \boldsymbol{\bar{z}} & = \max\{\mathbf{a}, \min\{\mathbf{b}, -\mathbf{\bar{p}}/\mu\}\} \quad \text{(componentwise)}.
     \end{cases}
\end{align}
We need to estimate $\gamma = (1/\mu)\|\boldsymbol{\mathfrak{L}}\|_{\mathbf{C}_{N+1}}$ and $\boldsymbol{\mathfrak{L}} = \mathbf{S}_{\mathrm{B}}\mathbf{\hat{R}}\mathbf{M}\mathbf{S}_{\mathrm{F}}\mathbf{\hat{R}}\mathbf{M}$

Now 
\begin{align}
\mathbf{S}_{\mathrm{F}}\mathbf{\hat{R}}\mathbf{M}
= \tau
\begin{bmatrix}
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & \mathbf{R}\mathbf{C} & \mathbf{0} & \cdots & \mathbf{0} \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^2 & \mathbf{R}\mathbf{C} & \cdots & \mathbf{0} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^{N} & (\mathbf{R}\mathbf{C})^{N-1} & \cdots & \mathbf{R}\mathbf{C}
\end{bmatrix}\quad \text{and}\quad
\mathbf{S}_{\mathrm{B}}\mathbf{\hat{R}}\mathbf{M}
= \tau
\begin{bmatrix}
\mathbf{0} & \mathbf{R}\mathbf{C} & (\mathbf{R}\mathbf{C})^2 & \cdots & (\mathbf{R}\mathbf{C})^{N} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{R}\mathbf{C} & \cdots & (\mathbf{R}\mathbf{C})^{N-1} \\[1mm]
\vdots & \vdots & \vdots & \ddots & \vdots \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{R}\mathbf{C} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \cdots & \mathbf{0}
\end{bmatrix}
\end{align}

For example, with $N=3$
\begin{align}
\boldsymbol{\mathfrak{L}} = \mathbf{S}_{\mathrm{B}}\mathbf{\hat{R}}\mathbf{M}\mathbf{S}_{\mathrm{F}}\mathbf{\hat{R}}\mathbf{M} 
&= \tau^2
\begin{bmatrix}
\mathbf{0} & \mathbf{R}\mathbf{C} & (\mathbf{R}\mathbf{C})^2 & (\mathbf{R}\mathbf{C})^{3} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{R}\mathbf{C} & (\mathbf{R}\mathbf{C})^2 \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{R}\mathbf{C} \\[1mm]
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{0}
\end{bmatrix}
\begin{bmatrix}
\mathbf{0} & \mathbf{0} & \mathbf{0} & \mathbf{0}  \\[1mm]
\mathbf{0} & \mathbf{R}\mathbf{C} & \mathbf{0} & \mathbf{0}  \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^2 & \mathbf{R}\mathbf{C} & \mathbf{0}  \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^3 & (\mathbf{R}\mathbf{C})^2 & \mathbf{R}\mathbf{C}
\end{bmatrix}\\
&\\
& = \tau^2
\begin{bmatrix}
\mathbf{0} & (\mathbf{R}\mathbf{C})^2 + (\mathbf{R}\mathbf{C})^4 + (\mathbf{R}\mathbf{C})^6 
           & (\mathbf{R}\mathbf{C})^3 + (\mathbf{R}\mathbf{C})^5 
           & (\mathbf{R}\mathbf{C})^4 \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^3 + (\mathbf{R}\mathbf{C})^5 
           & (\mathbf{R}\mathbf{C})^2 + (\mathbf{R}\mathbf{C})^4 
           & (\mathbf{R}\mathbf{C})^3 \\[1mm]
\mathbf{0} & (\mathbf{R}\mathbf{C})^4 
           & (\mathbf{R}\mathbf{C})^3 
           & (\mathbf{R}\mathbf{C})^2 \\[1mm]
\mathbf{0} & \mathbf{0} 
           & \mathbf{0} 
           & \mathbf{0}
\end{bmatrix}
\end{align}
Let $\boldsymbol{\mathfrak{B}} = \mathbf{C}_{N+1}^{\frac{1}{2}}\boldsymbol{\mathfrak{L}}\mathbf{C}_{N+1}^{-\frac{1}{2}}$. Note for example that
\begin{align*}
 \mathbf{C}^{\frac{1}{2}}(\mathbf{R}\mathbf{C})^{3}\mathbf{C}^{-\frac{1}{2}}  =  \mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}\mathbf{R}\mathbf{C}\mathbf{R}\mathbf{C}\mathbf{C}^{-\frac{1}{2}} =  (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^3.
\end{align*}
Therefore
\begin{align}
\boldsymbol{\mathfrak{B}}  = \tau^2
\begin{bmatrix}
\mathbf{0} & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^2 + (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^4 + (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^6 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^3 + (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^5 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^4 \\[1mm]
\mathbf{0} & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^3 + (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^5 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^2 + (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^4 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^3 \\[1mm]
\mathbf{0} & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^4 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^3 
           & (\mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}})^2 \\[1mm]
\mathbf{0} & \mathbf{0} 
           & \mathbf{0} 
           & \mathbf{0}
\end{bmatrix}
\end{align}
By letting $\mathbf{\Omega} = \mathbf{C}^{\frac{1}{2}}\mathbf{R}\mathbf{C}^{\frac{1}{2}}$, we can write
\begin{align}
\boldsymbol{\mathfrak{B}}  = \tau^2
\begin{bmatrix}
\mathbf{0} & \mathbf{\Omega}^2 + \mathbf{\Omega}^4 + \mathbf{\Omega}^6 
           & \mathbf{\Omega}^3 + \mathbf{\Omega}^5 
           & \mathbf{\Omega}^4 \\[1mm]
\mathbf{0} & \mathbf{\Omega}^3 + \mathbf{\Omega}^5 
           & \mathbf{\Omega}^2 + \mathbf{\Omega}^4 
           & \mathbf{\Omega}^3 \\[1mm]
\mathbf{0} & \mathbf{\Omega}^4 
           & \mathbf{\Omega}^3 
           & \mathbf{\Omega}^2 \\[1mm]
\mathbf{0} & \mathbf{0} 
           & \mathbf{0} 
           & \mathbf{0}
\end{bmatrix}
\end{align}
Let $\boldsymbol{\mathfrak{A}} = \begin{bmatrix}
 \mathbf{\Omega}^2 + \mathbf{\Omega}^4 + \mathbf{\Omega}^6 
           & \mathbf{\Omega}^3 + \mathbf{\Omega}^5 
           & \mathbf{\Omega}^4 \\[1mm]
 \mathbf{\Omega}^3 + \mathbf{\Omega}^5 
           & \mathbf{\Omega}^2 + \mathbf{\Omega}^4 
           & \mathbf{\Omega}^3 \\[1mm]
 \mathbf{\Omega}^4 
           & \mathbf{\Omega}^3 
           & \mathbf{\Omega}^2 
\end{bmatrix}$

\begin{align}
\label{chain1}
    \|\boldsymbol{\mathfrak{L}}\|_{\mathbf{C}_{N+1}}  = \|\boldsymbol{\mathfrak{B}}\|_2= \|\boldsymbol{\mathfrak{B}} \mathbf{\hat{J}} \|_2= \tau^2  \| \begin{bmatrix}
 \boldsymbol{\mathfrak{A}}\mathbf{\hat{J}} & \mathbf{0} \\[1mm]
 \mathbf{0} & \mathbf{0} 
\end{bmatrix}\|_2  = \tau^2  \|\boldsymbol{\mathfrak{A}}\mathbf{\hat{J}} \|_2 = \tau^2  \|\boldsymbol{\mathfrak{A}}\|_2
\end{align}

We have that $\mathbf{Q}^\top\boldsymbol{\Omega}^k\mathbf{Q}  =  \boldsymbol{\Delta}^k$. Let $\mathbf{V} = \mathbf{I}_{N}\otimes  \mathbf{Q}$. By ,  $\mathbf{V}$ is orthogonal and
\begin{align*}
\mathbf{V}^\top \boldsymbol{\mathfrak{A}} \mathbf{V} = 
\begin{bmatrix}
 \mathbf{\Delta}^2 + \mathbf{\Delta}^4 + \mathbf{\Delta}^6 
           & \mathbf{\Delta}^3 + \mathbf{\Delta}^5 
           & \mathbf{\Delta}^4 \\[1mm]
 \mathbf{\Delta}^3 + \mathbf{\Delta}^5 
           & \mathbf{\Delta}^2 + \mathbf{\Delta}^4 
           & \mathbf{\Delta}^3 \\[1mm]
 \mathbf{\Delta}^4 
           & \mathbf{\Delta}^3 
           & \mathbf{\Delta}^2 
\end{bmatrix}
\end{align*}

$$
\begin{bmatrix}
\mu_1^2+\mu_1^4+\mu_1^6 & 0 & 0 & \mu_1^3+\mu_1^5 & 0 & 0 & \mu_1^4 & 0 & 0 \\[1mm]
0 & \mu_2^2+\mu_2^4+\mu_2^6 & 0 & 0 & \mu_2^3+\mu_2^5 & 0 & 0 & \mu_2^4 & 0 \\[1mm]
0 & 0 & \mu_3^2+\mu_3^4+\mu_3^6 & 0 & 0 & \mu_3^3+\mu_3^5 & 0 & 0 & \mu_3^4 \\[2mm]
\mu_1^3+\mu_1^5 & 0 & 0 & \mu_1^2+\mu_1^4 & 0 & 0 & \mu_1^3 & 0 & 0 \\[1mm]
0 & \mu_2^3+\mu_2^5 & 0 & 0 & \mu_2^2+\mu_2^4 & 0 & 0 & \mu_2^3 & 0 \\[1mm]
0 & 0 & \mu_3^3+\mu_3^5 & 0 & 0 & \mu_3^2+\mu_3^4 & 0 & 0 & \mu_3^3 \\[2mm]
\mu_1^4 & 0 & 0 & \mu_1^3 & 0 & 0 & \mu_1^2 & 0 & 0 \\[1mm]
0 & \mu_2^4 & 0 & 0 & \mu_2^3 & 0 & 0 & \mu_2^2 & 0 \\[1mm]
0 & 0 & \mu_3^4 & 0 & 0 & \mu_3^3 & 0 & 0 & \mu_3^2
\end{bmatrix}

$$
$$
P =
\begin{bmatrix}
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\[1mm]
0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\[1mm]
0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
\end{bmatrix}
$$

$$
P M P^T =
\begin{bmatrix}
\mu_1^2+\mu_1^4+\mu_1^6 & \mu_1^3+\mu_1^5 & \mu_1^4 & 0 & 0 & 0 & 0 & 0 & 0 \\[1mm]
\mu_1^3+\mu_1^5 & \mu_1^2+\mu_1^4 & \mu_1^3 & 0 & 0 & 0 & 0 & 0 & 0 \\[1mm]
\mu_1^4 & \mu_1^3 & \mu_1^2 & 0 & 0 & 0 & 0 & 0 & 0 \\[2mm]
0 & 0 & 0 & \mu_2^2+\mu_2^4+\mu_2^6 & \mu_2^3+\mu_2^5 & \mu_2^4 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & \mu_2^3+\mu_2^5 & \mu_2^2+\mu_2^4 & \mu_2^3 & 0 & 0 & 0 \\[1mm]
0 & 0 & 0 & \mu_2^4 & \mu_2^3 & \mu_2^2 & 0 & 0 & 0 \\[2mm]
0 & 0 & 0 & 0 & 0 & 0 & \mu_3^2+\mu_3^4+\mu_3^6 & \mu_3^3+\mu_3^5 & \mu_3^4 \\[1mm]
0 & 0 & 0 & 0 & 0 & 0 & \mu_3^3+\mu_3^5 & \mu_3^2+\mu_3^4 & \mu_3^3 \\[1mm]
0 & 0 & 0 & 0 & 0 & 0 & \mu_3^4 & \mu_3^3 & \mu_3^2
\end{bmatrix}
$$


## Construction of an exact solution to the optimal control problem {#exact_sol_for_optim_control_prob}

We follow the approach in @Glusa2021errorestimates to construct an exact solution to the optimal control problem \eqref{eq:costfun}-\eqref{control_constraints}. The solution to the elliptic problem
\begin{align}
    \label{eq:elliptic_problem}
    \tag{12}
   (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s)=f(s),\quad s\in\Gamma,
\end{align}
where $u(\cdot)$ satisfies the Kirchhoff vertex conditions \eqref{eq:Kcond}, is given by
\begin{align*}
    u(s) = \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}(f,e_j)_{L_2(\Gamma)}e_j(s),\quad s\in\Gamma.
\end{align*}
We choose coefficients $\{x_j\}_{j=1}^{\infty}$ and $\{y_j\}_{j=1}^{\infty}$ and set
\begin{align}
    \label{eq:rhsfunctions}
    \tag{13}
   f(s) = \sum_{j=1}^{\infty}x_je_j(s)\quad \text{ and } \quad g(s) = \sum_{j=1}^{\infty}y_je_j(s).
\end{align}
Hence the corresponding solutions are given by
\begin{align}
    \label{eq:sols}
    \tag{14}
   u(s) = \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)\quad \text{ and } \quad v(s) = \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s).
\end{align}
 We also define $\psi(t) = \cos(t)$ and $\phi(t) =\sin(T-t)$. Observe that $\psi(0)=1$ and $\phi(T)=0$. We also choose $\mu =0.1$, $a(s,t) = -0.5$ and $b(s,t)=0.5$. Now we set
\begin{align*}
    f(s,t) &= \psi'(t)u(s)+\psi(t)f(s) - \bar{z}(s,t) = -\sin(t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)+ \cos(t)\sum_{j=1}^{\infty}x_je_j(s)- \max\{-0.5,\min\{0.5, \sin(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)\}\}\\ 
    u_d(s,t) & = \psi(t)u(s)-\mu\phi'(t)v(s)+\mu\phi(t)g(s) = \cos(t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)+\mu\cos(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)+\mu\sin(T-t)\sum_{j=1}^{\infty}y_je_j(s)\\
    u_0(s)  & = u(s)= \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)
\end{align*}
The exact solution to the optimal control problem is given by
\begin{align}
    \bar{u}(s,t)&=\psi(t)u(s) = \cos(t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)\\
    \bar{p}(s,t)&=-\mu\phi(t)v(s)= - \mu\sin(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)\\
    \bar{z}(s,t)&=\max\{a(s,t),\min\{b(s,t), \phi(t)v(s)\}\}= \max\{-0.5,\min\{0.5, \sin(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)\}\} 
\end{align}

Observe that the choice of $f$ decouples the systems \eqref{numericalscheme1}-\eqref{thenumericalscheme2}.

### Checking that $\bar{u}$ and $\bar{p}$ satisfy \eqref{eq:maineqoptimal} and \eqref{eq:adjointeq}

At $t=0$, we get
\begin{align*}
    \bar{u}(s,0) = \psi(0)u(s) = u(s) = u_0(s),
\end{align*}
while for $(s,t) \in \Gamma \times (0, T)$, we obtain
\begin{align*}
    \partial_t \bar{u}(s,t) +(\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{u}(s,t) = \psi'(t)u(s) + \psi(t) (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s) = \psi'(t)u(s) + \psi(t)f(s) = f(s,t) + \bar{z}(s,t).
\end{align*}
Similarly, at $t=T$, we get
\begin{align*}
    \bar{p}(s,T) = -\mu\phi(T)v(s) = 0,
\end{align*}
while for $(s,t) \in \Gamma \times (0, T)$, we obtain
\begin{align*}
    -\partial_t \bar{p}(s,t) +(\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{p}(s,t) = \mu\phi'(t)v(s) -\mu \phi(t) (\kappa^2 - \Delta_\Gamma)^{\alpha/2} v(s) = \mu\phi'(t)v(s) -\mu \phi(t)g(s)
\end{align*}
and
\begin{align*}
    \bar{u}(s,t) - u_d(s,t) = \psi(t)u(s) - \psi(t)u(s) + \mu\phi'(t)v(s) - \mu\phi(t)g(s) = \mu\phi'(t)v(s) -\mu \phi(t)g(s).
\end{align*}
This shows that $\bar{u}$ and $\bar{p}$ satisfy \eqref{eq:maineqoptimal} and \eqref{eq:adjointeq}.

## Numerical implementation {#num_implementation_optimal_control_problem}

In this section, we implement the numerical scheme for the optimal control problem \eqref{eq:costfun}-\eqref{control_constraints} on the tadpole graph. We start by defining the parameters and constructing the exact solution. We them use \eqref{numericalscheme1}-\eqref{thenumericalscheme2} to solve the forward and adjoint problems. Finally, we compute the control variable $\bar{z}$ using \eqref{zz}.


```{r, message = TRUE}
# Parameters
T_final <- 2
time_step <- 0.02
h <- 0.02
kappa <- 4
alpha <- 1.8
m = 4
beta <- alpha/2
mu <- 0.1
a <- - 0.5*10
b <- 0.5*10
N_finite = 4 # choose even
adjusted_N_finite <- N_finite + N_finite/2 + 1
# Coefficients for f and g
coeff_elliptic_g <- rep(0, adjusted_N_finite)
coeff_elliptic_g[7] <- 100*1
coeff_elliptic_f <- rep(0, adjusted_N_finite)
coeff_elliptic_f[7] <- -10*1
# time and spatial mesh 
time_seq <- seq(0, T_final, length.out = ((T_final - 0) / time_step + 1))
graph <- gets.graph.tadpole(h = h)
graph$compute_fem()
G <- graph$mesh$G
C <- graph$mesh$C
L <- kappa^2*C + G
I <- Matrix::Diagonal(nrow(C))
# Compute the eigenvalues and eigenfunctions
eigen_params <- gets.eigen.params(N_finite = N_finite, 
                                  kappa = kappa, 
                                  alpha = alpha, 
                                  graph = graph)
EIGENVAL_MINUS_ALPHA <- eigen_params$EIGENVAL_MINUS_ALPHA
EIGENFUN <- eigen_params$EIGENFUN
# Construct the right hand side functions f and g for the elliptic problem
elliptic_f <- as.vector(EIGENFUN %*% coeff_elliptic_f)
elliptic_g <- as.vector(EIGENFUN %*% coeff_elliptic_g)
# Construct the corresponding elliptic solution u and v
elliptic_u <- as.vector(EIGENFUN %*% (coeff_elliptic_f * EIGENVAL_MINUS_ALPHA))
elliptic_v <- as.vector(EIGENFUN %*% (coeff_elliptic_g * EIGENVAL_MINUS_ALPHA))

A <- matrix(a, nrow = nrow(C), ncol = length(time_seq))
B <- matrix(b, nrow = nrow(C), ncol = length(time_seq))
# Construct the optimal variables
psi <- cos(time_seq)
psi_prime <- - sin(time_seq)
phi <- sin(T_final - time_seq)
phi_prime <- - cos(T_final - time_seq)

u_bar <- elliptic_u %*% t(psi)
p_bar <- - mu * (elliptic_v %*% t(phi))
z_bar <- pmax(A, pmin(B, - p_bar / mu))
# Construct the data for the problem
f <- elliptic_u %*% t(psi_prime) + 
  elliptic_f %*% t(psi) - 
  z_bar
u_d <- elliptic_u %*% t(psi) -
  mu * (elliptic_v %*% t(phi_prime)) +
  mu * (elliptic_g %*% t(phi))

# Construct the fractional operator, which is shared for the forward and adjoint problems
my_op_frac <- my.fractional.operators.frac(L, beta, C, scale.factor = kappa^2, m = m, time_step)
# Construct the integration mesh
overkill_graph <- gets.graph.tadpole(h = 0.0001)
overkill_graph$compute_fem()
overkill_weights <- overkill_graph$mesh$weights
overkill_EIGENFUN <- gets.eigen.params(N_finite = N_finite, 
                                       kappa = kappa, 
                                       alpha = alpha, 
                                       graph = overkill_graph)$EIGENFUN
# Construct the right hand side functions f and g for the elliptic problem on the integration mesh
overkill_elliptic_f <- as.vector(overkill_EIGENFUN %*% coeff_elliptic_f)
overkill_elliptic_g <- as.vector(overkill_EIGENFUN %*% coeff_elliptic_g)
# Construct the corresponding elliptic solution u and v on the integration mesh
overkill_elliptic_u <- as.vector(overkill_EIGENFUN %*% (coeff_elliptic_f * EIGENVAL_MINUS_ALPHA))
overkill_elliptic_v <- as.vector(overkill_EIGENFUN %*% (coeff_elliptic_g * EIGENVAL_MINUS_ALPHA))
# Construct the projection matrix
Psi <- graph$fem_basis(overkill_graph$get_mesh_locations())
R <- t(Psi) %*% overkill_graph$mesh$C
```



```{r}
overkill_A <- matrix(a, nrow = length(overkill_weights), ncol = length(time_seq))
overkill_B <- matrix(b, nrow = length(overkill_weights), ncol = length(time_seq))


z_bar_overkill <- pmax(overkill_A, pmin(overkill_B, overkill_elliptic_v %*% t(phi)))
f <- overkill_elliptic_u %*% t(psi_prime) +
  overkill_elliptic_f %*% t(psi) -
  z_bar_overkill
U_d <- overkill_elliptic_u %*% t(psi) -
  mu * (overkill_elliptic_v %*% t(phi_prime)) +
  mu * (overkill_elliptic_g %*% t(phi))
u_0 <- elliptic_u

F_proj <- R %*% f
V_d <- R %*% U_d
```

```{r, message = TRUE}
tol <- 1e-12
maxit <- 25
verbose <- TRUE
nested_spatial_mesh <- TRUE
res <- solve_coupled_system_multi_tol(
  my_op_frac = my_op_frac,
  time_step = time_step, 
  time_seq = time_seq, 
  u_0 = u_0, 
  F_proj = F_proj, 
  Z_ini = F_proj * 0,
  V_d = V_d, 
  u_d = u_d,
  Psi = Psi, 
  R = R,
  A = A, 
  B = B, 
  C = C,
  mu = mu,
  tol = tol, 
  maxit = maxit, 
  verbose = verbose,
  nested_spatial_mesh = nested_spatial_mesh,
  true_sol = list(
    u_bar = u_bar, 
    p_bar = p_bar, 
    z_bar = z_bar
  ))

U_bar <- res$U
P_bar <- res$P
Z_bar <- res$Z
res$converged
res$iterations
rel_history <- res$rel_history
abs_history <- res$abs_history
min_history <- res$min_history
tol_list <- res$tol_list
Z_list <- res$Z_list
U_list <- res$U_list
P_list <- res$P_list
```


```{r, message = TRUE}
plot_convergence_history(rel_history, tol_list, type = "relative")
plot_convergence_history(abs_history, tol_list, type = "absolute")
plot_convergence_history(min_history, tol_list, type = "minimum")
```


```{r}
start_idx <- which.min(abs(time_seq - 0))
end_idx <- which.min(abs(time_seq - T_final))
idx <- seq(start_idx, end_idx, by = 1)
```


## Visualization of the results {#visualization_results}

```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Exact (in red) and numerical (in blue) optimal state variable.")}
XX <- U_list[[length(U_list)]] - u_bar
sqrt(as.double(time_step * sum(XX * (C %*% XX))))
U_list$u_true <- u_bar
graph.plotter.3d(graph, time_seq[idx], time_seq[idx], U_list)
```


```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Exact (in red) and numerical (in blue) optimal adjoint variable.")}
XX <- P_list[[length(P_list)]] - p_bar
sqrt(as.double(time_step * sum(XX * (C %*% XX))))
P_list$p_true <- p_bar
graph.plotter.3d(graph, time_seq[idx], time_seq[idx], P_list)
```


```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Exact (in red) and numerical (in blue) optimal control variable.")}
XX <- Z_list[[length(Z_list)]] - z_bar
sqrt(as.double(time_step * sum(XX * (C %*% XX))))
Z_list$z_true <- z_bar
graph.plotter.3d(graph, time_seq[idx], time_seq[idx], Z_list)
```


# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```