---
title: "Optimal control of fractional diffusion equations on metric graphs"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
  - \newcommand{\almosteverywhere}{\mathrm{a.e.}\;}
---

Go back to the [Contents](about.html) page.

<div style="color: #2c3e50; text-align: right;">
********  
<strong>Press Show to reveal the code chunks.</strong>  

********
</div>


```{r}
# Create a clipboard button on the rendered HTML page
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = FALSE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```

```{r}
library(MetricGraph)
library(ggplot2)
library(reshape2)
library(plotly)
library(tidyr)
```


```{r}
capture.output(
  knitr::purl(here::here("control_functionality.Rmd"), output = here::here("control_functionality.R")),
  file = here::here("purl_log.txt")
)
source(here::here("control_functionality.R"))
```

# Optimal control of fractional diffusion equations on metric graphs

## Problem statement

Let $\Gamma = (\mathcal{V},\mathcal{E})$ be a metric graph. Let $u_d: \Gamma \times(0, T) \rightarrow \mathbb{R}$ be the desired state and  $\mu>0$ a regularization parameter. We define the cost functional

\begin{equation}
\label{eq:costfun}
\tag{1}
    J(u, z)=\frac{1}{2} \int_0^T\left(\left\|u-u_d\right\|_{L_2(\Gamma)}^2+\mu\|z\|_{L_2(\Gamma)}^2\right) dt
\end{equation}


Let $f:\Gamma\times (0,T)\rightarrow\ar$ and $u_0: \Gamma \rightarrow \mathbb{R}$ be fixed functions. We will call them right-hand side and initial datum, respectively. Let $\alpha\in(0,2]$ and $z: \Gamma \times(0, T) \rightarrow \mathbb{R}$ denote the control variable. We shall be concerned with the following PDE-constrained optimization problem: Find

\begin{equation}
\label{eq:min_pro}
\tag{2}
    \min\; J(u, z)
\end{equation}
subject to the fractional diffusion equation
\begin{equation}
\label{eq:maineq}
\tag{3}
\left\{
\begin{aligned}
    \partial_t u(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s,t) &= f(s,t)+z(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    u(s,0) &= u_0(s), && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
with $u(\cdot,t)$ satisfying the Kirchhoff vertex conditions
\begin{equation}
\label{eq:Kcond}
\tag{4}
   \mathcal{K} =  \left\{\phi\in C(\Gamma)\;\middle|\; \forall v\in \mathcal{V}:\; \sum_{e\in\mathcal{E}_v}\partial_e \phi(v)=0 \right\}
\end{equation}
and the control constraints 
\begin{align}
\label{control_constraints}
\tag{5}
    a(s,t)\leq z(s,t)\leq b(s,t)\;\text{a.e.} (s,t)\in\Gamma \times(0, T).
\end{align}


## Optimal solution

The optimal variables $(\bar{u}, \bar{p}, \bar{z})$ satisfy

\begin{equation}
\label{eq:maineqoptimal}
\tag{6}
\left\{
\begin{aligned}
    \partial_t \bar{u}(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{u}(s,t) &= f(s,t)+\bar{z}(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    \bar{u}(s,0) &= u_0(s), && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
and
\begin{equation}
\label{eq:adjointeq}
\tag{7}
\left\{
\begin{aligned}
    -\partial_t \bar{p}(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{p}(s,t) &= \bar{u}(s,t)-u_d(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    \bar{p}(s,T) &= 0, && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
with 
\begin{align}
\label{zz}
\tag{8}
        \bar{z}(s,t) = \max\left\{a(s,t),\min\left\{b(s,t),-\dfrac{1}{\mu}\bar{p}(s,t)\right\}\right\}.
\end{align}

## Numerical scheme {#num_scheme_optim_control}

Go to the [Control Functionality](control_functionality.html) page to see the details of the numerical scheme for solving the optimal control problem \eqref{eq:costfun}-\eqref{control_constraints}. Here is a short summary of the numerical scheme.

- **1. Initialization:** Initialize $\mathbf{\bar{Z}}\in \mathbb{R}^{N_{h}\times (N+1)}$.

- **2. State solve:** Given $\mathbf{\bar{Z}}$, compute $\mathbf{\bar{U}}\in \mathbb{R}^{N_{h}\times (N+1)}$ with the iteration

\begin{align}
\label{statesolve}
\tag{FS}
    \begin{cases}
    \mathbf{\bar{U}}_{k+1}  &= \mathbf{R}\mathbf{C}\mathbf{\bar{U}}_k+\tau \mathbf{R}(\mathbf{F}_{k+1}+\mathbf{C}\mathbf{\bar{Z}}_{k+1})\\
    & = (\mathbf{R}\mathbf{C})^{k+1}\mathbf{\bar{U}}_0
+ \tau \sum_{i=0}^{k} (\mathbf{R}\mathbf{C})^{k-i}\mathbf{R}(\mathbf{F}_{i+1}+\mathbf{C} \mathbf{\bar{Z}}_{i+1}), \quad k = 0,\dots, N-1,\\
    \mathbf{\bar{U}}_{0}  &= [u_0(s_1), \dots, u_0(s_{N_h})]^\top,
    \end{cases}
\end{align}

- **3. Adjoint solve:** Given $\mathbf{\bar{U}}$ from before, compute $\mathbf{\bar{P}}\in \mathbb{R}^{N_{h}\times (N+1)}$ with the iteration

\begin{align}
\label{adjointsolve}
\tag{BS}
    \begin{cases}
    \mathbf{\bar{P}}_{k}  & = \mathbf{R}\mathbf{C}\mathbf{\bar{P}}_{k+1}+\tau \mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{k+1}-\mathbf{D}_{k+1})\\
    &=  \tau \sum_{i=0}^{N-k-1} (\mathbf{R}\mathbf{C})^{i}\mathbf{R}(\mathbf{C}\mathbf{\bar{U}}_{k+i+1}-\mathbf{D}_{k+i+1}), \quad k = N-1,\dots, 0,\\
    \mathbf{\bar{P}}_{N} &  = \mathbf{0},
    \end{cases}
\end{align}

- **4. Control update:** Given $\mathbf{\bar{P}}$ from before, update 

\begin{align}
  \label{entrywisez}
  \mathbf{\bar{Z}} = \max\{\mathbf{A},\min\{\mathbf{B},-\mathbf{\bar{P}}/\mu\}\}\quad \text{(entry-wise)}.
\end{align}

- **5. Iteration:** Repeat steps 2--4 until convergence.

## Construction of an exact solution to the optimal control problem {#exact_sol_for_optim_control_prob}

We follow the approach in @Glusa2021errorestimates to construct an exact solution to the optimal control problem \eqref{eq:costfun}-\eqref{control_constraints}. The solution to the elliptic problem
\begin{align}
    \label{eq:elliptic_problem}
    \tag{12}
   (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s)=f(s),\quad s\in\Gamma,
\end{align}
where $u(\cdot)$ satisfies the Kirchhoff vertex conditions \eqref{eq:Kcond}, is given by
\begin{align*}
    u(s) = \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}(f,e_j)_{L_2(\Gamma)}e_j(s),\quad s\in\Gamma.
\end{align*}
We choose coefficients $\{x_j\}_{j=1}^{\infty}$ and $\{y_j\}_{j=1}^{\infty}$ and set
\begin{align}
    \label{eq:rhsfunctions}
    \tag{13}
   f(s) = \sum_{j=1}^{\infty}x_je_j(s)\quad \text{ and } \quad g(s) = \sum_{j=1}^{\infty}y_je_j(s).
\end{align}
Hence the corresponding solutions are given by
\begin{align}
    \label{eq:sols}
    \tag{14}
   u(s) = \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)\quad \text{ and } \quad v(s) = \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s).
\end{align}
 We also define $\psi(t) = \cos(t)$ and $\phi(t) =\sin(T-t)$. Observe that $\psi(0)=1$ and $\phi(T)=0$. We also choose $\mu =0.1$, $a(s,t) = -0.5$ and $b(s,t)=0.5$. Now we set
\begin{align*}
    f(s,t) &= \psi'(t)u(s)+\psi(t)f(s) - \bar{z}(s,t) = -\sin(t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)+ \cos(t)\sum_{j=1}^{\infty}x_je_j(s)- \max\{-0.5,\min\{0.5, \sin(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)\}\}\\ 
    u_d(s,t) & = \psi(t)u(s)-\mu\phi'(t)v(s)+\mu\phi(t)g(s) = \cos(t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)+\mu\cos(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)+\mu\sin(T-t)\sum_{j=1}^{\infty}y_je_j(s)\\
    u_0(s)  & = u(s)= \sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)
\end{align*}
The exact solution to the optimal control problem is given by
\begin{align}
    \bar{u}(s,t)&=\psi(t)u(s) = \cos(t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}x_je_j(s)\\
    \bar{p}(s,t)&=-\mu\phi(t)v(s)= - \mu\sin(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)\\
    \bar{z}(s,t)&=\max\{a(s,t),\min\{b(s,t), \phi(t)v(s)\}\}= \max\{-0.5,\min\{0.5, \sin(T-t)\sum_{j=1}^{\infty}\lambda_j^{-\alpha/2}y_je_j(s)\}\} 
\end{align}


### Checking that $\bar{u}$ and $\bar{p}$ satisfy \eqref{eq:maineqoptimal} and \eqref{eq:adjointeq}

At $t=0$, we get
\begin{align*}
    \bar{u}(s,0) = \psi(0)u(s) = u(s) = u_0(s),
\end{align*}
while for $(s,t) \in \Gamma \times (0, T)$, we obtain
\begin{align*}
    \partial_t \bar{u}(s,t) +(\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{u}(s,t) = \psi'(t)u(s) + \psi(t) (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s) = \psi'(t)u(s) + \psi(t)f(s) = f(s,t) + \bar{z}(s,t).
\end{align*}
Similarly, at $t=T$, we get
\begin{align*}
    \bar{p}(s,T) = -\mu\phi(T)v(s) = 0,
\end{align*}
while for $(s,t) \in \Gamma \times (0, T)$, we obtain
\begin{align*}
    -\partial_t \bar{p}(s,t) +(\kappa^2 - \Delta_\Gamma)^{\alpha/2} \bar{p}(s,t) = \mu\phi'(t)v(s) -\mu \phi(t) (\kappa^2 - \Delta_\Gamma)^{\alpha/2} v(s) = \mu\phi'(t)v(s) -\mu \phi(t)g(s)
\end{align*}
and
\begin{align*}
    \bar{u}(s,t) - u_d(s,t) = \psi(t)u(s) - \psi(t)u(s) + \mu\phi'(t)v(s) - \mu\phi(t)g(s) = \mu\phi'(t)v(s) -\mu \phi(t)g(s).
\end{align*}
This shows that $\bar{u}$ and $\bar{p}$ satisfy \eqref{eq:maineqoptimal} and \eqref{eq:adjointeq}.

## Numerical implementation {#num_implementation_optimal_control_problem}

In this section, we implement the numerical scheme for the optimal control problem \eqref{eq:costfun}-\eqref{control_constraints} on the tadpole graph. We start by defining the parameters and constructing the exact solution. We them use \eqref{numericalscheme1}-\eqref{thenumericalscheme2} to solve the forward and adjoint problems. Finally, we compute the control variable $\bar{z}$ using \eqref{zz}.


```{r, message = TRUE}
# Parameters
T_final <- 2
time_step <- 0.02
h <- 0.02
kappa <- 4
alpha <- 1.8
m = 4
beta <- alpha/2
mu <- 0.1
a <- - 0.5*10
b <- 0.5*10
N_finite = 4 # choose even
adjusted_N_finite <- N_finite + N_finite/2 + 1
# Coefficients for f and g
coeff_elliptic_g <- rep(0, adjusted_N_finite)
coeff_elliptic_g[7] <- 100*1
coeff_elliptic_f <- rep(0, adjusted_N_finite)
coeff_elliptic_f[7] <- -10*1
# time and spatial mesh 
time_seq <- seq(0, T_final, length.out = ((T_final - 0) / time_step + 1))
graph <- gets.graph.tadpole(h = h)
graph$compute_fem()
G <- graph$mesh$G
C <- graph$mesh$C
L <- kappa^2*C + G
I <- Matrix::Diagonal(nrow(C))
# Compute the eigenvalues and eigenfunctions
eigen_params <- gets.eigen.params(N_finite = N_finite, 
                                  kappa = kappa, 
                                  alpha = alpha, 
                                  graph = graph)
EIGENVAL_MINUS_ALPHA <- eigen_params$EIGENVAL_MINUS_ALPHA
EIGENFUN <- eigen_params$EIGENFUN
# Construct the right hand side functions f and g for the elliptic problem
elliptic_f <- as.vector(EIGENFUN %*% coeff_elliptic_f)
elliptic_g <- as.vector(EIGENFUN %*% coeff_elliptic_g)
# Construct the corresponding elliptic solution u and v
elliptic_u <- as.vector(EIGENFUN %*% (coeff_elliptic_f * EIGENVAL_MINUS_ALPHA))
elliptic_v <- as.vector(EIGENFUN %*% (coeff_elliptic_g * EIGENVAL_MINUS_ALPHA))

# Construct the optimal variables
psi <- cos(time_seq)
psi_prime <- - sin(time_seq)
phi <- sin(T_final - time_seq)
phi_prime <- - cos(T_final - time_seq)

u_bar <- elliptic_u %*% t(psi)
p_bar <- - mu * (elliptic_v %*% t(phi))
z_bar <- matrix(pmax(a, pmin(b, - p_bar / mu)), dim(p_bar))

# Construct the fractional operator, which is shared for the forward and adjoint problems
my_op_frac <- my.fractional.operators.frac(L, beta, C, scale.factor = kappa^2, m = m, time_step)
# Construct the integration mesh
overkill_graph <- gets.graph.tadpole(h = 0.0001)
overkill_graph$compute_fem()
overkill_weights <- overkill_graph$mesh$weights
overkill_EIGENFUN <- gets.eigen.params(N_finite = N_finite, 
                                       kappa = kappa, 
                                       alpha = alpha, 
                                       graph = overkill_graph)$EIGENFUN
# Construct the right hand side functions f and g for the elliptic problem on the integration mesh
overkill_elliptic_f <- as.vector(overkill_EIGENFUN %*% coeff_elliptic_f)
overkill_elliptic_g <- as.vector(overkill_EIGENFUN %*% coeff_elliptic_g)
# Construct the corresponding elliptic solution u and v on the integration mesh
overkill_elliptic_u <- as.vector(overkill_EIGENFUN %*% (coeff_elliptic_f * EIGENVAL_MINUS_ALPHA))
overkill_elliptic_v <- as.vector(overkill_EIGENFUN %*% (coeff_elliptic_g * EIGENVAL_MINUS_ALPHA))
# Construct the projection matrix
Psi <- graph$fem_basis(overkill_graph$get_mesh_locations())
R <- t(Psi) %*% overkill_graph$mesh$C
minus_p_bar_over_mu <- overkill_elliptic_v %*% t(phi)
z_bar_overkill <- matrix(pmax(a, pmin(b, minus_p_bar_over_mu)), dim(minus_p_bar_over_mu))
f <- overkill_elliptic_u %*% t(psi_prime) + overkill_elliptic_f %*% t(psi) - z_bar_overkill

U_d <- overkill_elliptic_u %*% t(psi) -
  mu * (overkill_elliptic_v %*% t(phi_prime)) +
  mu * (overkill_elliptic_g %*% t(phi))

V_d <- R %*% U_d
u_0 <- elliptic_u
F_proj <- R %*% f

u_d <- outer(elliptic_u, psi) -
      mu * outer(elliptic_v, phi_prime) +
      mu * outer(elliptic_g, phi)
```



```{r, message = TRUE}
tol <- 1e-12
maxit <- 25
verbose <- TRUE
nested_spatial_mesh <- TRUE
res <- solve_coupled_system_multi_tol(
  my_op_frac = my_op_frac,
  time_step = time_step, 
  time_seq = time_seq, 
  u_0 = u_0, 
  F_proj = F_proj, 
  Z_ini = F_proj * 0,
  V_d = V_d, 
  u_d = u_d,
  Psi = Psi, 
  R = R,
  a = a, 
  b = b, 
  C = C,
  mu = mu,
  tol = tol, 
  maxit = maxit, 
  verbose = verbose,
  nested_spatial_mesh = nested_spatial_mesh,
  true_sol = list(
    u_bar = u_bar, 
    p_bar = p_bar, 
    z_bar = z_bar
  ))

U_bar <- res$U
P_bar <- res$P
Z_bar <- res$Z
res$converged
res$iterations
rel_history <- res$rel_history
abs_history <- res$abs_history
min_history <- res$min_history
tol_list <- res$tol_list
Z_list <- res$Z_list
U_list <- res$U_list
P_list <- res$P_list
```


```{r, message = TRUE}
plot_convergence_history(rel_history, tol_list, type = "relative")
plot_convergence_history(abs_history, tol_list, type = "absolute")
plot_convergence_history(min_history, tol_list, type = "minimum")
```


```{r}
start_idx <- which.min(abs(time_seq - 0))
end_idx <- which.min(abs(time_seq - T_final))
idx <- seq(start_idx, end_idx, by = 1)
```


## Visualization of the results {#visualization_results}

```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Exact (in red) and numerical (in blue) optimal state variable.")}
XX <- U_list[[length(U_list)]] - u_bar
sqrt(as.double(time_step * sum(XX * (C %*% XX))))
# U_list$u_true <- u_bar
U_list <- c(list(u_true = u_bar), U_list)
graph.plotter.3d(graph, time_seq[idx], time_seq[idx], U_list)
```


```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Exact (in red) and numerical (in blue) optimal adjoint variable.")}
XX <- P_list[[length(P_list)]] - p_bar
sqrt(as.double(time_step * sum(XX * (C %*% XX))))
#P_list$p_true <- p_bar
P_list <- c(list(p_true = p_bar), P_list)
graph.plotter.3d(graph, time_seq[idx], time_seq[idx], P_list)
```


```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Exact (in red) and numerical (in blue) optimal control variable.")}
XX <- Z_list[[length(Z_list)]] - z_bar
sqrt(as.double(time_step * sum(XX * (C %*% XX))))
#Z_list$z_true <- z_bar
Z_list <- c(list(z_true = z_bar), Z_list)
graph.plotter.3d(graph, time_seq[idx], time_seq[idx], Z_list)
```


# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```