---
title: "Convergence in 𝘩"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

Go back to the [Contents](about.html) page.

<div style="color: #2c3e50; text-align: right;">
********  
<strong>Press Show to reveal the code chunks.</strong>  

********
</div>


Let us set some global options for all code chunks in this document.


```{r}
# Create a clipboard button
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = FALSE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```


::: {.custom-box}
To observe convergence in $h$, we determine $\tau$ numerically from the relation $\tau\propto h^{\alpha-\epsilon}|\ln(\tau)|$ and choose $m\propto\left\lceil\ln^2\!\bigl(h^{2-\epsilon}|\ln(\tau)|\tau^2\bigr)\big/\!\left(4\pi^2\left(1-\alpha/2\right)\right)\right\rceil$ so that the temporal and rational approximation errors are of the same order as the spatial error. This yields $E_h=E/|\ln(\tau)|\lesssim h^{\alpha-\epsilon}$, corresponding to a convergence rate of order $\alpha-\epsilon$ with respect to the mesh size $h$. Fitting the regression $\log_{10}E_h=s_h\log_{10}h+\log_{10}c_h$, we therefore expect $s_h\approx\alpha-\epsilon$.
:::


```{r, eval = TRUE}
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")
library(rSPDE)
library(MetricGraph)
library(grateful)

library(ggplot2)
library(reshape2)
library(plotly)
library(slackr)
source(here::here("keys.R"))
slackr_setup(token = token) # token comes from keys.R
```



```{r}
capture.output(
  knitr::purl(here::here("functionality.Rmd"), output = here::here("functionality.R")),
  file = here::here("purl_log.txt")
)
source(here::here("functionality.R"))
```




```{r}
# Parameters
T_final <- 2
kappa <- 4
N_finite = 4 # choose even
adjusted_N_finite <- N_finite + N_finite/2 + 1
# Coefficients for u_0 and f
coeff_U_0 <- 50*(1:adjusted_N_finite)^-1
coeff_U_0[-5] <- 0
coeff_FF <- rep(0, adjusted_N_finite)
coeff_FF[7] <- 10

AAA = 1
OMEGA = pi

# Time step and mesh size
POWERS <- seq(14, 9, by = -1)

time_steps <- 0.1 * 2^-POWERS
h_vector <- 10^seq(log10(0.0009), log10(0.01), length.out = 6) #time_steps^(1/2) 
alpha_vector <- seq(1, 1.8, by = 0.2)
epsilon <- 1e-10
```



```{r}
# Overkill parameters
overkill_time_step <- 5e-6 #0.1 * 2^-14
overkill_h <- 9e-4 #(0.1 * 2^-14)^(1/2)
# Finest time and space mesh
overkill_time_seq <- seq(0, T_final, length.out = ((T_final - 0) / overkill_time_step + 1))
overkill_graph <- gets.graph.tadpole(h = overkill_h)
# Compute the weights on the finest mesh
overkill_graph$compute_fem() # This is needed to compute the weights
overkill_C <- overkill_graph$mesh$C
m_values <- c()
```



```{r, eval = FALSE, class.source = "fold-show"}
# Create a matrix to store the errors
errors_projected <- matrix(NA, nrow = length(h_vector), ncol = length(alpha_vector))
for (j in 1:length(alpha_vector)) {
  alpha <- alpha_vector[j] # from 0.5 to 2
  beta <- alpha / 2

  # Compute the eigenvalues and eigenfunctions on the finest mesh
  auxX <- gets.eigen.params(N_finite = N_finite, kappa = kappa, alpha = alpha, graph = overkill_graph)
  EIGENVAL_ALPHA <- auxX$EIGENVAL_ALPHA # Eigenvalues (they are independent of the meshes)
  auxY <- auxX$EIGENFUN # Eigenfunctions on the finest mesh
  
  # Compute the true solution on the finest mesh
  auxX <- auxY %*% 
    outer(1:length(coeff_U_0), 
          1:length(overkill_time_seq), 
          function(i, j) (coeff_U_0[i] + coeff_FF[i] * G_sin(t = overkill_time_seq[j], A = AAA, lambda_j_alpha_half = EIGENVAL_ALPHA[i], omega = OMEGA)) * exp(-EIGENVAL_ALPHA[i] * overkill_time_seq[j]))
  
  for (i in 1:length(h_vector)) {
    h <- h_vector[i]
    time_step <- tau_from_h_exact(h, alpha - epsilon)#tau_from_h(h, alpha, C_c)#h^(2*alpha) #1e-5 #tau_from_h_exact(h, alpha) #h^alpha
    m <- min(4, ceiling((log(h^(2 - epsilon) * time_step^2 * abs(log(time_step))))^2 / (4 * pi^2 * (1 - alpha / 2)))) #min(8, m_from_h(h, alpha, C_c, time_step)) 
    m_values <- c(m_values, m)
    graph <- gets.graph.tadpole(h = h)
    graph$compute_fem()
    L <- graph$mesh$G
    C <- graph$mesh$C
    L <- kappa^2*C + L
    U_0 <- gets.eigen.params(N_finite = N_finite, kappa = kappa, alpha = alpha, graph = graph)$EIGENFUN
    U_0 <- U_0 %*% coeff_U_0 # Compute U_0 on the current mesh
    Psi <- graph$fem_basis(overkill_graph$get_mesh_locations())
  
    time_seq <- seq(0, T_final, length.out = ((T_final - 0) / time_step + 1))
    my_op_frac <- my.fractional.operators.frac(L, beta, C, scale.factor = kappa^2, m = m, time_step)
    # Compute matrix F with columns F^k
    U_approx <- t(Psi) %*% overkill_C %*% auxY %*%  
      outer(1:length(coeff_FF), 
            1:length(time_seq), 
        function(i, j) coeff_FF[i] * g_sin(r = time_seq[j], A = AAA, omega = OMEGA))

    U_approx <- Psi %*% solve_fractional_evolution(my_op_frac, time_step, time_seq, val_at_0 = U_0, RHST = U_approx)
    U_approx <- auxX - construct_piecewise_projection(U_approx, time_seq, overkill_time_seq)
    errors_projected[i,j] <- sqrt(as.double(overkill_time_step * sum(U_approx * (overkill_C %*% U_approx)))) / abs(log(time_step))
    slackr_msg(text = paste0("m =", m, ", alpha =", alpha, ", h =", h, ", time_step =", time_step), channel = "#research")
  }
}
save(errors_projected, file = here::here("data_files/errors_projected_h_with_logtau_and_tauminustwo.RData"))
```


```{r}
# Load the errors_projected data
load(here::here("data_files/errors_projected_h_with_logtau_and_tauminustwo.RData"))
```


```{r}
observed_rates <- numeric(length(alpha_vector))
for (u in 1:length(alpha_vector)) {
  observed_rates[u] <- coef(lm(log10(errors_projected[, u]) ~ log10(h_vector)))[2]
}

theoretical_rates <- alpha_vector  - epsilon


p_h_tadpole_with_logtau_and_tauminustwo <- error.convergence.plotter(x_axis_vector = h_vector, 
                                 alpha_vector, 
                                 errors_projected, 
                                 theoretical_rates, 
                                 observed_rates,
                                 line_equation_fun = loglog_line_equation,
                                 fig_title = expression("Convergence in " * italic(h)),
                                 x_axis_label = expression(italic(h)))

p_h_nicer_tadpole_with_logtau_and_tauminustwo <- errorConvergencePlotterLatex(x_axis_vector = h_vector, 
                                 alpha_vector, 
                                 errors_projected, 
                                 theoretical_rates, 
                                 observed_rates,
                                 line_equation_fun = loglog_line_equation,
                                 fig_title = "$\\mbox{Convergence in }h$",
                                 x_axis_label = "$h\\quad(\\log_{10}\\mbox{ scale})$",
                                 y_axis_label = "$\\|u-U_{h,m}^{\\tau}\\|_{L_2(\\Gamma\\times[0,T])}\\quad(\\log_{10}\\mbox{ scale})$",
                                 color_label = "$\\quad\\quad\\;\\;\\alpha\\;|\\;\\mathtt{theo}\\;|\\;\\;\\mathtt{obs}\\;\\;|\\%\\mathtt{re}$")

ggsave(here::here("data_files/conv_rates_h_tadpole_with_logtau_and_tauminustwo.png"), width = 4, height = 5, plot = p_h_tadpole_with_logtau_and_tauminustwo, dpi = 300)
myggsave(p_h_nicer_tadpole_with_logtau_and_tauminustwo, width = 4, height = 5, use_mathpazo = TRUE)
save(p_h_tadpole_with_logtau_and_tauminustwo, file = here::here("data_files/p_h_tadpole_with_logtau_and_tauminustwo.RData"))
save(p_h_nicer_tadpole_with_logtau_and_tauminustwo, file = here::here("data_files/p_h_nicer_tadpole_with_logtau_and_tauminustwo.RData"))
```


```{r, eval = TRUE, fig.align='center', out.width="400px", out.height="500px", fig.cap = captioner("Comparison between theoretical and observed convergence behavior for the $L_2(\\Gamma\\times(0,T))$-error with respect to $h$ on a $\\text{log}_{10}$–$\\text{log}_{10}$ scale. Dashed lines indicate the theoretical rates, and solid lines represent the observed error curves. The legend below each plot shows the value of $\\alpha$ along with the corresponding theoretical ('theo'), and observed ('obs') rates for each case.")}
knitr::include_graphics(here::here("data_files/tikzpic/p_h_nicer_tadpole_with_logtau_and_tauminustwo_with_mathpazo.pdf"))
```



```{r, eval = TRUE, echo = FALSE}
initial_comment <- "Here’s the latest plot update for the convergence in h!"
# option 1
slackr_upload(
  filename = here::here("data_files/conv_rates_h_tadpole_with_logtau_and_tauminustwo.png"),        # path to your image
  initial_comment = initial_comment,
  channels = "#research"
)
get_numeric_scalars <- function() {
  vars <- ls(envir = .GlobalEnv)
  scalars <- sapply(vars, function(v) {
    val <- get(v, envir = .GlobalEnv)
    is.numeric(val) && length(val) == 1
  })
  mget(vars[scalars], envir = .GlobalEnv)
}

# Collect the scalars
num_scalars <- get_numeric_scalars()

# Format scalars
scalar_msg <- paste(
  sprintf("%s = %s", names(num_scalars), unlist(num_scalars)),
  collapse = "\n"
)

# Format tau_vector
h_msg <- paste(sprintf("%.12f", h_vector), collapse = ", ")

# Send one Slack message
slackr::slackr_msg(
  text = paste(
    "Numeric scalars in this run:\n",
    scalar_msg,
    "\n\n",
    "h_vector =",
    h_msg
  ),
  channel = "#research"
)
```

# References

```{r, eval = TRUE}
cite_packages(output = "paragraph", out.dir = ".")
```
