---
title: "Convergence in τ"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    # toc: true
    # toc_float:
    #   collapsed: true
    #   smooth_scroll: true
    number_sections: false
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---


<div style="color: #2c3e50; text-align: right;">
********  
<strong>Press Show to reveal the code chunks.</strong>  

********
</div>


Let us set some global options for all code chunks in this document.


```{r}
# Create a clipboard button
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```


::: {.custom-box}
To observe convergence in $\tau$, we fix $\tau$ and balance the spatial and rational-approx error components by setting $h \sim \tau^{1/\alpha}$ and then $m \sim \lceil25\log_e^2(h)/(16\pi^2(1-\alpha/2))\rceil$, respectively. Under this scaling, the total convergence rate with respect to the time step $\tau$ is of order $1$. That is, $\|u-U_{h,m}^\tau\|_{L_2((0,T);L_2(\Gamma))} \leq C_\tau\tau$. This can be verified by estimating the slope $S_\tau$ in the regression $\log_{10} E = S_\tau\log_{10} \tau + \log_{10} C_\tau$, where we expect $S_\tau\sim1$. 
:::


```{r}
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")
library(rSPDE)
library(MetricGraph)
library(grateful)

library(ggplot2)
library(reshape2)
library(plotly)
```


```{r}
capture.output(
  knitr::purl(here::here("functionality.Rmd"), output = here::here("functionality.R")),
  file = here::here("purl_log.txt")
)
source(here::here("functionality.R"))
```




```{r}
# Parameters
T_final <- 2
kappa <- 15
N_finite = 4 # choose even
adjusted_N_finite <- N_finite + N_finite/2 + 1
# Coefficients for u_0 and f
coeff_U_0 <- 50*(1:adjusted_N_finite)^-1
coeff_U_0[-5] <- 0
coeff_FF <- rep(0, adjusted_N_finite)
coeff_FF[7] <- 10

AAA = 1
OMEGA = pi

# Time step and mesh size
POWERS <- seq(6.125, 5.625, by = -0.1)
tau_vector <- 0.1 * 2^-POWERS

# Overkill parameters
overkill_time_step <- 0.1 * 2^-14
overkill_h <- (0.1 * 2^-14)^(1/2)

# Finest time and space mesh
overkill_time_seq <- seq(0, T_final, length.out = ((T_final - 0) / overkill_time_step + 1))
overkill_graph <- gets.graph.tadpole(h = overkill_h)

# Compute the weights on the finest mesh
overkill_graph$compute_fem() # This is needed to compute the weights
overkill_weights <- overkill_graph$mesh$weights

m_values <- c()
alpha_vector <- seq(1, 1.8, by = 0.2)
```


```{r, eval = FALSE, class.source = "fold-show"}
# Create a matrix to store the errors
errors_projected <- matrix(NA, nrow = length(tau_vector), ncol = length(alpha_vector))
for (j in 1:length(alpha_vector)) {
  alpha <- alpha_vector[j] # from 0.5 to 2
  beta <- alpha / 2

  # Compute the eigenvalues and eigenfunctions on the finest mesh
  overkill_eigen_params <- gets.eigen.params(N_finite = N_finite, kappa = kappa, alpha = alpha, graph = overkill_graph)
  EIGENVAL_ALPHA <- overkill_eigen_params$EIGENVAL_ALPHA # Eigenvalues (they are independent of the meshes)
  overkill_EIGENFUN <- overkill_eigen_params$EIGENFUN # Eigenfunctions on the finest mesh
  
  # Compute the true solution on the finest mesh
  overkill_U_true <- overkill_EIGENFUN %*% 
    outer(1:length(coeff_U_0), 
          1:length(overkill_time_seq), 
          function(i, j) (coeff_U_0[i] + coeff_FF[i] * G_sin(t = overkill_time_seq[j], A = AAA, lambda_j_alpha_half = EIGENVAL_ALPHA[i], omega = OMEGA)) * exp(-EIGENVAL_ALPHA[i] * overkill_time_seq[j]))
  
  for (i in 1:length(tau_vector)) {
    time_step <- tau_vector[i]
    h <- time_step^(1/alpha)
    m <- min(4, ceiling(25 * (log(h))^2 / (16 * pi^2 * (1 - alpha / 2))))
    m_values <- c(m_values, m)
    graph <- gets.graph.tadpole(h = h)
    graph$compute_fem()
    G <- graph$mesh$G
    C <- graph$mesh$C
    L <- kappa^2*C + G
    eigen_params <- gets.eigen.params(N_finite = N_finite, kappa = kappa, alpha = alpha, graph = graph)
    EIGENFUN <- eigen_params$EIGENFUN
    U_0 <- EIGENFUN %*% coeff_U_0 # Compute U_0 on the current mesh
    A <- graph$fem_basis(overkill_graph$get_mesh_locations())
  
    time_seq <- seq(0, T_final, length.out = ((T_final - 0) / time_step + 1))
    my_op_frac <- my.fractional.operators.frac(L, beta, C, scale.factor = kappa^2, m = m, time_step)
    INT_BASIS_EIGEN <- t(overkill_EIGENFUN) %*% overkill_graph$mesh$C %*% A
    # Compute matrix F with columns F^k
    FF_approx <- t(INT_BASIS_EIGEN) %*% 
      outer(1:length(coeff_FF), 
            1:length(time_seq), 
        function(i, j) coeff_FF[i] * g_sin(r = time_seq[j], A = AAA, omega = OMEGA))
    
    U_approx <- matrix(NA, nrow = nrow(C), ncol = length(time_seq))
    U_approx[, 1] <- U_0
    for (k in 1:(length(time_seq) - 1)) {
      U_approx[, k + 1] <- as.matrix(my.solver.frac(my_op_frac, my_op_frac$C %*% U_approx[, k] + time_step * FF_approx[, k + 1]))
    }
    
    projected_U_approx <- A %*% U_approx
    projected_U_piecewise <- construct_piecewise_projection(projected_U_approx, time_seq, overkill_time_seq)
    errors_projected[i,j] <- sqrt(as.double(t(overkill_weights) %*% (overkill_U_true - projected_U_piecewise)^2 %*% rep(overkill_time_step, length(overkill_time_seq))))
  }
}
save(errors_projected, file = here::here("data_files/errors_projected_tau.RData"))
```

```{r}
# Load the errors_projected data
load(here::here("data_files/errors_projected_tau.RData"))
```



```{r, fig.align='center', fig.dim= c(4,5), fig.cap = captioner("Comparison of theoretical and observed convergence behavior for the $L_2((0,T);L_2(\\Gamma))$-error with respect to $\\tau$ on a $\\text{log}_{10}$–$\\text{log}_{10}$ scale. Dashed lines indicate the theoretical rates, and solid lines represent the observed error curves. The legend below each plot shows the value of $\\alpha$ along with the corresponding theoretical ('theo'), and observed ('obs') rates for each case.")}
observed_rates <- numeric(length(alpha_vector))
for (u in 1:length(alpha_vector)) {
  observed_rates[u] <- coef(lm(log10(errors_projected[, u]) ~ log10(tau_vector)))[2]
}

theoretical_rates <- rep(1, length(alpha_vector))

guiding_lines <- compute_guiding_lines(x_axis_vector = tau_vector, 
                                       errors = errors_projected, 
                                       theoretical_rates = theoretical_rates, 
                                       line_equation_fun = loglog_line_equation)
default_colors <- scales::hue_pal()(length(alpha_vector))
plot_lines <- lapply(1:ncol(guiding_lines), function(i) {
  geom_line(
    data = data.frame(x = tau_vector, y = guiding_lines[, i]),
    aes(x = x, y = y),
    color = default_colors[i],
    linetype = "dashed",
    show.legend = FALSE
  )
})

df <- as.data.frame(cbind(tau_vector, errors_projected))
colnames(df) <- c("tau_vector", alpha_vector)
df_melted <- melt(df, id.vars = "tau_vector", variable.name = "column", value.name = "value")
alpha_formatted <- formatC(alpha_vector, format = "f", digits = 2)

custom_labels <- paste0(alpha_formatted, 
                        " | ", 
                        formatC(theoretical_rates, format = "f", digits = 4), 
                        " | ", 
                        formatC(observed_rates, format = "f", digits = 4))

df_melted$column <- factor(df_melted$column, levels = alpha_vector, labels = custom_labels)

p <- ggplot() +
  geom_line(data = df_melted, aes(x = tau_vector, y = value, color = column)) +
  geom_point(data = df_melted, aes(x = tau_vector, y = value, color = column)) +
  plot_lines +
  labs(
    title = expression("Convergence in " * italic(tau)),
    x = expression(italic(tau)),
    y = expression(Error),
    color = "          α  | theo  | obs"
  ) +
  scale_x_log10(breaks = tau_vector, labels = round(tau_vector, 4)) +
  scale_y_log10(labels = function(x) formatC(x, format = "f", digits = 4)) +
  theme_minimal() +
  theme(text = element_text(family = "Palatino"),
        legend.position = "bottom",
        legend.direction = "vertical",
        plot.margin = margin(0, 0, 0, 0),
        plot.title = element_text(hjust = 0.5, size = 18, face = "bold"))
p
```


```{r}
ggsave(here::here("data_files/conv_rates_tau.png"), width = 4, height = 5, plot = p, dpi = 300)
```


# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
