---
title: "Experiment"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

Go back to the [Contents](about.html) page.

<div style="color: #2c3e50; text-align: right;">
********  
<strong>Press Show to reveal the code chunks.</strong>  

********
</div>


Let us set some global options for all code chunks in this document.


```{r}
# Create a clipboard button
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = FALSE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```



```{r}
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")
library(rSPDE)
library(MetricGraph)
library(grateful)

library(ggplot2)
library(reshape2)
library(plotly)
```



```{r}
capture.output(
  knitr::purl(here::here("functionality.Rmd"), output = here::here("functionality.R")),
  file = here::here("purl_log.txt")
)
source(here::here("functionality.R"))
```


## The equation

We analyze and numerically approximate solutions to fractional diffusion equations on metric graphs of the form
\begin{equation}
\label{eq:maineq}
\tag{1}
\left\{
\begin{aligned}
    \partial_t u(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s,t) &= f(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    u(s,0) &= u_0(s), && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
with $u(\cdot,t)$ satisfying the Kirchhoff vertex conditions
\begin{equation}
\label{eq:Kcond}
\tag{2}
   \mathcal{K} =  \left\{\phi\in C(\Gamma)\;\middle|\; \forall v\in \mathcal{V}:\; \textstyle\sum_{e\in\mathcal{E}_v}\partial_e \phi(v)=0 \right\}.
\end{equation}
Here $\Gamma = (\mathcal{V},\mathcal{E})$ is a metric graph, $\kappa>0$, $\alpha\in(0,2]$ determines the smoothness of $u(\cdot,t)$, $\Delta_{\Gamma}$ is the so-called Kirchhoff--Laplacian, and $f:\Gamma\times (0,T)\to\mathbb{R}$ and $u_0: \Gamma \to \mathbb{R}$ are fixed functions, called right-hand side and initial condition, respectively.

## Eigenfunction-based construction of an exact solution {#eigensolconst}

The exact solution to \eqref{eq:maineq} can be expressed as
\begin{equation}
\label{eq:sol_reprentation}
\tag{3}
        u(s,t) = \displaystyle\sum_{j\in\mathbb{N}}e^{-\lambda^{\alpha/2}_jt}\left(u_0, e_j\right)_{L_2(\Gamma)}e_j(s) + \int_0^t \displaystyle\sum_{j\in\mathbb{N}}e^{-\lambda^{\alpha/2}_j(t-r)}\left(f(\cdot, r), e_j\right)_{L_2(\Gamma)}e_j(s)dr.
\end{equation}

To construct an exact solution, we expand both the initial condition and the right-hand side in terms of the eigenfunctions. We choose coefficients $\{x_j\}_{j=1}^{\infty}$ and  $\{y_j\}_{j=1}^{\infty}$ and a scalar function $g(t)$ and set
\begin{align}
\tag{4}
    u_0(s) = \sum_{j=1}^{\infty} x_j e_j(s)\quad \text{ and }\quad
    f(s,t) = g(t) \sum_{j=1}^{\infty} y_j e_j(s).
\end{align}
 Substituting these expressions into the solution formula \eqref{eq:sol_reprentation}, we obtain
\begin{align}
\label{sollll}
\tag{5}
    u(s,t) = \sum_{j=1}^{\infty}(x_j+y_j G_j(t))e^{-\lambda^{\alpha/2}_jt}e_j(s),\quad G_j(t)= \int_0^t e^{\lambda^{\alpha/2}_jr}g(r)dr,
\end{align}
where the integral can be evaluated analytically for some choices of $g(t)$, as shown in the  [Functionality](functionality.html#exact_solution) page. The following list shows how the above expressions are implemented in $\texttt{R}$.
\begin{aligned}
\text{expression} \iff&\quad \text{In } \texttt{R}\\
\{x_j\}_{j=1}^{\infty} \iff&\quad  \texttt{coeff_U_0}\\
\{y_j\}_{j=1}^{\infty} \iff&\quad  \texttt{coeff_FF}\\
[e_1 e_2 \dots e_{N_f}] \iff&\quad  \texttt{EIGENFUN}\\
u_0(s) \iff&\quad  \texttt{U_0 <- EIGENFUN %*% coeff_U_0}\\
f(s,t) \iff&\quad  \texttt{FF_true <- EIGENFUN %*%}\\
&\quad\texttt{ outer(1:length(coeff_FF),}\\
&\qquad\qquad\texttt{ 1:length(time_seq),}\\
&\qquad\qquad\texttt{ function(i, j) coeff_FF[i] * g_sin(r = time_seq[j], A = AAA, omega = OMEGA))}\\
u(s,t) \iff&\quad  \texttt{U_true <- EIGENFUN %*%}\\
&\quad\texttt{ outer(1:length(coeff_U_0),}\\
&\qquad\qquad\texttt{ 1:length(time_seq),}\\
&\qquad\qquad\texttt{ function(i, j) (coeff_U_0[i] + coeff_FF[i] * G_sin(t = time_seq[j], A = AAA, lambda_j_alpha_half = EIGENVAL_ALPHA[i], omega = OMEGA)) *}\\
&\qquad\qquad\qquad\qquad\qquad\qquad\quad\texttt{exp(-EIGENVAL_ALPHA[i] * time_seq[j]))}
\end{aligned}

Let us check that \eqref{sollll} is a solution of \eqref{eq:maineq}. At $t=0$, we get
\begin{align*}
    u(s,0) = \sum_{j=1}^{\infty}(x_j+y_j G_j(0))e^{-\lambda^{\alpha/2}_j0}e_j(s) = \sum_{j=1}^{\infty}x_je_j(s) =u_0(s),
\end{align*}
while for $(s,t) \in \Gamma \times (0, T)$, we obtain
\begin{align*}
    \partial_t u(s,t) = \sum_{j=1}^{\infty}y_j G'_j(t)e^{-\lambda^{\alpha/2}_jt}e_j(s) - \sum_{j=1}^{\infty}(x_j+y_j G_j(t))e^{-\lambda^{\alpha/2}_jt}\lambda^{\alpha/2}_je_j(s)
\end{align*}
and
\begin{align*}
    (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s,t) &= \sum_{j=1}^{\infty}(x_j+y_j G_j(t))e^{-\lambda^{\alpha/2}_jt}(\kappa^2 - \Delta_\Gamma)^{\alpha/2}e_j(s)\\
    &= \sum_{j=1}^{\infty}(x_j+y_j G_j(t))e^{-\lambda^{\alpha/2}_jt}\lambda^{\alpha/2}_je_j(s).
\end{align*}
Adding these two and using the fact that $G'_j(t) = e^{\lambda^{\alpha/2}_jt}g(t)$ yield
\begin{align*}
    \partial_t u(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s,t)&= g(t)\sum_{j=1}^{\infty}y_je^{\lambda^{\alpha/2}_jt} e^{-\lambda^{\alpha/2}_jt}e_j(s) = f(s,t).
\end{align*}
This shows that \eqref{sollll} is indeed a solution of \eqref{eq:maineq}.

From the [Functionality](functionality.html#num_scheme) page, we know that the solution to \eqref{eq:maineq} can be approximated by a numerical scheme of the form
\begin{equation}
\label{eq:final_scheme2}
\tag{6}
\mathbf{U}^{k+1} = \mathbf{R}(\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}),\quad \mathbf{R} = \sum_{k=1}^{m+1} a_k\left( \dfrac{\mathbf{L}}{\kappa^2}-p_k\mathbf{C}\right)^{-1}
\end{equation}
Let $\mathbf{F}$ be the matrix with columns $\mathbf{F}^k$. This matrix has entries $\mathbf{F}_{ik}$ given by
\begin{align}
\label{eq:innerprod}
\tag{7}
    (f^{k},\psi^i_h)_{L_2(\Gamma)} = (f(\cdot,t_k),\psi^i_h)_{L_2(\Gamma)} = \sum_{j=1}^{N_f} y_j g(t_k) (e_j,\psi^i_h)_{L_2(\Gamma)}.
\end{align}
To approximate the inner products $(e_j,\psi^i_h)_{L_2(\Gamma)}$, we use the quadrature formula $\boldsymbol{\psi}_i^\top\mathbf{C}^{\star}\mathbf{e}_j$, where $\mathbf{e}_j=[e_j(s_1)\dots e_j(s_{N_{h_{\star}}})]^\top$ and $\boldsymbol{\psi}_i=[\psi^i_h(s_1)\dots \psi^i_h(s_{N_{h_{\star}}})]^\top$  are vectors of function evaluations on a fine spatial mesh with mesh size $h_{\star}$ and nodes $\{s_\ell\}_{\ell=1}^{N_{h_{\star}}}$. Matrix $\mathbf{C}^{\star}$ contains the corresponding quadrature weights, with entries $\mathbf{C}^{\star}_{nm} = (\psi_{h_\star}^n,\psi_{h_\star}^m)_{L_2(\Gamma)}$ for $n,m = 1,\dots, N_{h_\star}$. We emphasize that two spatial meshes are involved in this construction. The basis functions $\{\psi^i_h\}_{i=1}^{N_h}$ are defined on a coarse spatial mesh with mesh size $h$, while the quadrature is carried out over a finer mesh with associated basis functions $\{\psi^\ell_{h_\star}\}_{\ell=1}^{N_{h_\star}}$ and mesh size $h_\star$. If $\mathbf{N}$ denotes the matrix with entries $\mathbf{N}_{jk} = y_j g(t_k)$, then $\mathbf{F}$ can be approximated as $[\boldsymbol{\psi}_1\dots \boldsymbol{\psi}_{N_h}]^\top \mathbf{C}^{\star} [\mathbf{e}_1\dots \mathbf{e}_{N_f}] \mathbf{N}$. In $\texttt{R}$,  $\mathbf{F}$ and $\mathbf{U}^{k+1}$ are implemented as
\begin{aligned}
\text{expression} \iff&\quad \text{In } \texttt{R}\\
[\boldsymbol{\psi}_1\dots \boldsymbol{\psi}_{N_h}]^\top \iff&\quad \texttt{t(graph\$fem_basis(overkill_graph\$get_mesh_locations()))}\\
\mathbf{C}^{\star} \iff&\quad \texttt{overkill_graph\$mesh\$C}\\
[\mathbf{e}_1\dots \mathbf{e}_{N_f}] \iff&\quad \texttt{overkill_EIGENFUN}\\
\mathbf{N} \iff&\quad \texttt{outer(1:length(coeff_FF),}\\
&\qquad\quad\;\;\texttt{ 1:length(time_seq),}\\
&\qquad\quad\;\;\texttt{ function(i, j) coeff_FF[i] * g_sin(r = time_seq[j], A = AAA, omega = OMEGA))}\\
\mathbf{F}\iff&\quad  \texttt{FF_innerprod <- t(graph\$fem_basis(overkill_graph\$get_mesh_locations())) %*%}\\
&\quad\texttt{overkill_graph\$mesh\$C %*%}\\
&\quad\texttt{overkill_EIGENFUN %*%}\\
&\quad\texttt{outer(1:length(coeff_FF),}\\
&\qquad\qquad\texttt{1:length(time_seq),}\\
&\qquad\qquad\texttt{function(i, j) coeff_FF[i] * g_sin(r = time_seq[j], A = AAA, omega = OMEGA))}\\
\mathbf{U}^{k+1}\iff&\quad \texttt{U_approx[, k + 1] <- as.matrix(my.solver.frac(my_op_frac, my_op_frac\$C %*% U_approx[, k] + time_step * FF_innerprod[, k + 1]))}
\end{aligned}

```{r}
T_final <- 2
time_step <- 0.001
h <- 0.001
kappa <- 15
alpha <- 0.5
m = 4
beta <- alpha/2
N_finite = 4 # choose even
adjusted_N_finite <- N_finite + N_finite/2 + 1
# Coefficients for u_0 and f
coeff_U_0 <- 50*(1:adjusted_N_finite)^-1
coeff_U_0[-5] <- 0
coeff_FF <- rep(0, adjusted_N_finite)
coeff_FF[7] <- 10


time_seq <- seq(0, T_final, by = time_step)
graph <- gets.graph.tadpole(h = h)
graph$compute_fem()
G <- graph$mesh$G
C <- graph$mesh$C
L <- kappa^2*C + G
I <- Matrix::Diagonal(nrow(C))

eigen_params <- gets.eigen.params(N_finite = N_finite, kappa = kappa, alpha = alpha, graph = graph)
EIGENVAL_ALPHA <- eigen_params$EIGENVAL_ALPHA
EIGENFUN <- eigen_params$EIGENFUN

AAA = 1
OMEGA = pi

U_0 <- EIGENFUN %*% coeff_U_0

U_true <- EIGENFUN %*% 
  outer(1:length(coeff_U_0), 
        1:length(time_seq), 
        function(i, j) (coeff_U_0[i] + coeff_FF[i] * G_sin(t = time_seq[j], A = AAA, lambda_j_alpha_half = EIGENVAL_ALPHA[i], omega = OMEGA) ) * exp(-EIGENVAL_ALPHA[i] * time_seq[j]))

overkill_graph <- gets.graph.tadpole(h = 0.0001)
overkill_graph$compute_fem()
overkill_EIGENFUN <- gets.eigen.params(N_finite = N_finite, kappa = kappa, alpha = alpha, graph = overkill_graph)$EIGENFUN


FF_innerprod <- t(graph$fem_basis(overkill_graph$get_mesh_locations())) %*%
  overkill_graph$mesh$C %*%
  overkill_EIGENFUN %*%
  outer(1:length(coeff_FF), 
        1:length(time_seq), 
        function(i, j) coeff_FF[i] * g_sin(r = time_seq[j], A = AAA, omega = OMEGA))

FF_true <- EIGENFUN %*% outer(1:length(coeff_FF), 
                         1:length(time_seq), 
                         function(i, j) coeff_FF[i] * g_sin(r = time_seq[j], A = AAA, omega = OMEGA))

# Numerical solution
my_op_frac <- my.fractional.operators.frac(L, beta, C, scale.factor = kappa^2, m = m, time_step)


U_approx <- solve_fractional_evolution(my_op_frac, time_step, time_seq, val_at_0 = U_0, RHST = FF_innerprod)
U_approx2 <- U_approx
XX <- U_true - U_approx

error <- sqrt(as.double(time_step * sum(XX * (C %*% XX))))
save(error, file = here::here("data_files/error_exp1.RData"))
```

The following is the error reported in the paper for the considered experiment.

```{r, eval = TRUE}
load(here::here("data_files/error_exp1.RData"))
print(paste0("Total error = ", formatC(error, format = "f", digits = 9)))
```



Because of GitHub storage limits, we project the solution (and all other visualizations) to a coarser mesh as follows.

```{r}
aux_graph <- gets.graph.tadpole(h = 0.01)
A_proj <- graph$fem_basis(aux_graph$get_mesh_locations())
FF_true <- A_proj %*% FF_true
U_true <- A_proj %*% U_true
U_approx <- A_proj %*% U_approx

aux_graph$compute_fem()
weights <- aux_graph$mesh$weights
error_at_each_time <- t(weights) %*% (U_true - U_approx)^2

start_idx <- which.min(abs(time_seq - 0))
end_idx <- which.min(abs(time_seq - 2))
idx <- seq(start_idx, end_idx, by = 40)
```


# Another comparison

```{r}
error_for_project2 <- graph.plotter.3d(aux_graph, time_seq[idx], error_at_each_time[idx], U_list = list(U_true = U_true[,idx], U_approx = U_approx[,idx]))
save_dual_for_presentation(error_for_project2)
```

```{r, eval = TRUE, fig.height = 6, out.width = "100%", fig.cap = captioner("Comparison of the exact and approximate solutions.")}
load(here::here("data_files/error_for_project2.RData"))
error_for_project2
```


# RHS function

```{r}
start_idx <- which.min(abs(time_seq - 0.5))
end_idx <- which.min(abs(time_seq - 1.5))
idx <- seq(start_idx, end_idx, by = 40)


idx_for1 <- which.min(abs(time_seq - 0.5))

exp1_fig <- graph.plotter.3d.single(aux_graph, FF_true[, idx], time_seq[idx])
save(exp1_fig, file = here::here("data_files/exp1_fig.RData"))
```


```{r, eval = TRUE, fig.height = 6, out.width = "100%", fig.cap = captioner("Time evolution of the right-hand side function $f$.")}
load(here::here("data_files/exp1_fig.RData"))
exp1_fig
```

# Error

```{r}
diff <- U_true - U_approx
abs_error <- abs(diff)[, idx]
exp1_error <- graph.plotter.3d.single(aux_graph, abs_error, time_seq[idx])
save(exp1_error, file = here::here("data_files/exp1_error.RData"))
```


```{r, eval = TRUE, fig.height = 6, out.width = "100%", fig.cap = captioner("Time evolution of the absolute difference between the exact and approximate solution.")}
load(here::here("data_files/exp1_error.RData"))
exp1_error
```


# Error at 1 frame


```{r}
error_tadpole <- graph.plotter.3d.onecol(graph, abs(XX[,idx_for1]))
save(error_tadpole, file = here::here("data_files/exp1_error_tadpole.RData"))
```


```{r, eval = TRUE, fig.height = 6, out.width = "100%", fig.cap = captioner("Error at some time point.")}
load(here::here("data_files/exp1_error_tadpole.RData"))
error_tadpole
```


# Solution at 1 frame

```{r}
f_tadpole <- graph.plotter.3d.onecol(graph, U_approx2[,idx_for1])
save(f_tadpole, file = here::here("data_files/exp1_f_tadpole.RData"))
```


```{r, eval = TRUE, fig.height = 6, out.width = "100%", fig.cap = captioner("Error at some time point.")}
load(here::here("data_files/exp1_f_tadpole.RData"))
f_tadpole
```


# Comparer

```{r}
idx2 <- seq(start_idx, end_idx, by = 40)
exp1_com <- graph.plotter.3d.comparer(aux_graph, U_true[,idx2], U_approx[,idx2], time_seq[idx2])
save(exp1_com, file = here::here("data_files/exp1_com.RData"))
```


```{r, eval = TRUE, fig.height = 6, out.width = "100%", fig.cap = captioner("Comparison of the exact and approximate solution at selected time points.")}
load(here::here("data_files/exp1_com.RData"))
exp1_com
```


## References

```{r, eval = TRUE}
grateful::cite_packages(output = "paragraph", out.dir = ".")
```

