---
title: "Functionality"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: show # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

Go back to the [Contents](about.html) page.

<div style="color: #2c3e50; text-align: right;">
********  
<strong>Press Show to reveal the code chunks.</strong>  

********
</div>


```{r, purl = FALSE, echo = FALSE}
# Create a clipboard button on the rendered HTML page
source(here::here("clipboard.R")); clipboard
```


```{r, purl = FALSE, class.source = "fold-hide"}
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```



```{r}
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")
library(rSPDE)
library(MetricGraph)
library(grateful)

library(ggplot2)
library(reshape2)
library(plotly)
```


## Fractional diffusion equations on metric graphs

We analyze and numerically approximate solutions to fractional diffusion equations on metric graphs of the form
\begin{equation}
\label{eq:maineq}
\tag{1}
\left\{
\begin{aligned}
    \partial_t u(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s,t) &= f(s,t), && \quad (s,t) \in \Gamma \times (0, T), \\
    u(s,0) &= u_0(s), && \quad s \in \Gamma,
\end{aligned}
\right.
\end{equation}
with $u(\cdot,t)$ satisfying the Kirchhoff vertex conditions
\begin{equation}
\label{eq:Kcond}
\tag{2}
   \mathcal{K} =  \left\{\phi\in C(\Gamma)\;\middle|\; \forall v\in \mathcal{V}:\; \sum_{e\in\mathcal{E}_v}\partial_e \phi(v)=0 \right\}.
\end{equation}
Here $\Gamma = (\mathcal{V},\mathcal{E})$ is a metric graph, $\kappa>0$, $\alpha\in(0,2]$ determines the smoothness of $u(\cdot,t)$, $\Delta_{\Gamma}$ is the so-called Kirchhoff--Laplacian, and $f:\Gamma\times (0,T)\to\mathbb{R}$ and $u_0: \Gamma \to \mathbb{R}$ are fixed functions, called right-hand side and initial condition, respectively.

## Numerical Scheme {#num_scheme}

Let $\alpha\in(0,2]$ and $U_h^\tau$ denote the sequence of approximations of the solution to the weak form of problem \eqref{eq:maineq} at each time step on a mesh indexed by $h$. The scheme computes $U_h^{\tau}\subset V_h$, which solves
\begin{equation}
    \left\{
\begin{aligned}
    \langle\delta U_h^{k+1},\phi\rangle + \mathfrak{a}(
        U_h^{k+1},\phi) &= \langle f^{k+1},\phi\rangle, \quad\forall\phi\in V_h,\quad k=0,\dots, N-1, \\
    U^0_h &= P_hu_0, 
\end{aligned}
\right.
\end{equation}
where $f^{k+1} = \displaystyle\dfrac{1}{\tau}\int_{t_k}^{t_{k+1}}f(t)dt$. The above expression can be equivalently written as
\begin{equation}
    \left\{
\begin{aligned}
    \langle\dfrac{U_h^{k+1} - U_h^{k}}{\tau},\phi\rangle + \mathfrak{a}(
        U_h^{k+1},\phi) &= \langle f^{k+1},\phi\rangle, \quad\forall\phi\in V_h,\quad k=0,\dots, N-1, \\
    U^0_h &= P_hu_0, 
\end{aligned}
\right.
\end{equation}
or
\begin{equation}
\label{system:fully_discrete_scheme}
\tag{3}
    \left\{
\begin{aligned}
    \langle U_h^{k+1},\phi\rangle + \tau\mathfrak{a}(
        U_h^{k+1},\phi) &= \langle U_h^{k},\phi\rangle + \tau\langle f^{k+1},\phi\rangle, \quad\forall\phi\in V_h,\quad k=0,\dots, N-1, \\
    U^0_h &= P_hu_0.
\end{aligned}
\right.
\end{equation}

At each time step $t_k$, the finite element solution $U_h^k\in V_h$ to \eqref{system:fully_discrete_scheme} can be expressed as a linear combination of the basis functions  $\{\psi^i_h\}_{i=1}^{N_h}$ introduced in the [Preliminaries](preliminaries.html#fem-basis) page, namely, 
\begin{align}
\label{num_sol}
\tag{4}
    U_h^k(s) =  \sum_{i=1}^{N_h}u_i^k\psi^i_h(s), \;s\in\Gamma.
\end{align}

Replacing \eqref{num_sol} into \eqref{system:fully_discrete_scheme} yields the following system
\begin{align*}
    \sum_{j=1}^{N_h}u_j^{k+1}[(\psi_h^j,\psi_h^i)_{L_2(\Gamma)}+ \tau\mathfrak{a}(\psi_h^j,\psi_h^i)] = \sum_{j=1}^{N_h}u_j^{k}(\psi_h^j,\psi_h^i)_{L_2(\Gamma)}+\tau( f^{k+1},\psi_h^i)_{L_2(\Gamma)},\quad i = 1,\dots, N_h.
\end{align*}
In matrix notation,
\begin{align}
\label{diff_eq_discrete}
    (\mathbf{C}+\tau \mathbf{L}^{\alpha/2})\mathbf{U}^{k+1} = \mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1},
\end{align}
or by introducing the scaling parameter $\kappa^2>0$,
\begin{align}
    (\mathbf{C}+\tau (\kappa^2)^{\alpha/2}(\mathbf{L}/\kappa^2)^{\alpha/2})\mathbf{U}^{k+1} = \mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1},
\end{align}
where $\mathbf{C}$ has entries $\mathbf{C}_{i,j} = (\psi_h^i,\psi_h^j)_{L_2(\Gamma)}$, $\mathbf{L}^{\alpha/2}$ has entries $\mathfrak{a}(\psi_h^i,\psi_h^j)$, $\mathbf{U}^k$ has components $u_j^k$, and $\mathbf{F}^k$ has components $( f^{k},\psi_h^i)_{L_2(\Gamma)}$. Applying $(\mathbf{L}/\kappa^2)^{-\alpha/2}$ to both sides yields
\begin{equation}
((\mathbf{L}/\kappa^2)^{-\alpha/2}\mathbf{C}+\tau (\kappa^2)^{\alpha/2}\mathbf{I})\mathbf{U}^{k+1} = (\mathbf{L}/\kappa^2)^{-\alpha/2}(\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}).
\end{equation}
Following @rSPDE2020, we approximate $(\mathbf{L}/\kappa^2)^{-\alpha/2}$ by $\mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top$ to arrive at
\begin{equation}
\label{eq:scheme2}
\tag{5}
(\mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top \mathbf{C}+\tau(\kappa^2)^{\alpha/2} \mathbf{I})\mathbf{U}^{k+1} = \mathbf{P}_\ell^{-\top}\mathbf{P}_r^\top(\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}).
\end{equation}
where
\begin{equation}
\label{eq:PLPR}
\tag{6}
\mathbf{P}_r = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right)\quad\text{and}\quad \mathbf{P}_\ell = \dfrac{1}{\texttt{factor}}\mathbf{C}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right),
\end{equation}
and $\texttt{factor} = \dfrac{c_m}{b_{m+1}}$, and
$\{r_{1i}\}_{i=1}^m$ and $\{r_{2j}\}_{j=1}^{m+1}$ are the roots of $q_1(x) =\sum_{i=0}^mc_ix^{i}$ and  $q_2(x)=\sum_{j=0}^{m+1}b_jx^{j}$, respectively. The coefficients  $\{c_i\}_{i=0}^m$ and  $\{b_j\}_{j=0}^{m+1}$ are determined as the best rational approximation $q_1/q_2$ of the function $x^{\alpha/2-1}$ over the interval $J_h: = [\kappa^{2}\lambda_{N_h,h}^{-1}, \kappa^{2}\lambda_{1,h}^{-1}]$, where $\lambda_{1,h}, \lambda_{N_h,h}>0$ are the smallest and the largest eigenvalue of $L_h$, respectively.


For the sake of clarity, we note that the numerical implementation of @rSPDE2020 actually defines $\mathbf{P}_r$ and $\mathbf{P}_\ell$ as
\begin{equation}
\label{eq:PLPRbolin}
\tag{7}
\mathbf{P}_r = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right)\quad\text{and}\quad \mathbf{P}_\ell = \dfrac{\kappa^{2\beta}}{\texttt{factor}}\mathbf{C}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{C}^{-1}\mathbf{L}}{\kappa^2}\right),
\end{equation}
where $\beta = \alpha/2$ and the scaling factor $(\kappa^2)^{\alpha/2}$ or $\kappa^{2\beta}$ is already incorporated in $\mathbf{P}_\ell$, a convention we adopt in the following. With this under consideration, we can rewrite \eqref{eq:scheme2} as
\begin{equation}
\tag{8}
(\mathbf{P}_r^\top \mathbf{C}+\tau \mathbf{P}_\ell^\top)\mathbf{U}^{k+1} = \mathbf{P}_r^\top(\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}),
\label{eq:scheme}
\end{equation}
where  
\begin{equation}
\mathbf{P}_r^\top = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\quad\text{and}\quad \mathbf{P}_\ell^\top = \dfrac{\kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot \mathbf{C}
\end{equation}
since $\mathbf{L}$ and $\mathbf{C}^{-1}$ are symmetric and the factors in the product commute. Replacing these two into \eqref{eq:scheme} yields
\begin{equation}
\left(\prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\right)\mathbf{C}\mathbf{U}^{k+1} = \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot (\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}),
\end{equation}
that is,
\begin{equation}
\label{eq:final_scheme}
\tag{9}
\mathbf{U}^{k+1} = \mathbf{C}^{-1}\left(\prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}}\prod_{j=1}^{m+1} \left(\mathbf{I}-r_{2j}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\right)^{-1} \prod_{i=1}^m \left(\mathbf{I}-r_{1i}\dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}\right)\cdot (\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}).
\end{equation}
Considering the partial fraction decomposition
\begin{equation}
\label{eq:partial_fraction}
\tag{10}
\dfrac{\prod_{i=1}^m (1-r_{1i}x)}{\prod_{i=1}^m (1-r_{1i}x)+\dfrac{\tau \kappa^{2\beta}}{\texttt{factor}} \prod_{j=1}^{m+1} (1-r_{2j}x)}=\sum_{k=1}^{m+1} a_k(x-p_k)^{-1} + r,
\end{equation}
scheme \eqref{eq:final_scheme} can be expressed as
\begin{equation}
\label{eq:final_scheme2}
\tag{11}
\mathbf{U}^{k+1} = \mathbf{C}^{-1}\left(\sum_{k=1}^{m+1} a_k\left( \dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}-p_k\mathbf{I}\right)^{-1} + r\mathbf{I}\right) (\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1})
\end{equation}

In practice, since the rational function in \eqref{eq:partial_fraction} is proper, there is no remainder $r$. Moreover, since $\left( \dfrac{\mathbf{L}\mathbf{C}^{-1}}{\kappa^2}-p_k\mathbf{I}\right)^{-1}  = \mathbf{C}\left( \dfrac{\mathbf{L}}{\kappa^2}-p_k\mathbf{C}\right)^{-1}$, we have that \eqref{eq:final_scheme2} can be rewritten as

\begin{equation}
\label{eq:final_scheme3}
\tag{12}
\mathbf{U}^{k+1} = \mathbf{R}(\mathbf{C}\mathbf{U}^k+\tau \mathbf{F}^{k+1}),\quad \mathbf{R} = \sum_{k=1}^{m+1} a_k\left( \dfrac{\mathbf{L}}{\kappa^2}-p_k\mathbf{C}\right)^{-1}.
\end{equation}

## Numerical implementation {#num_implementation}

### Function `my.get.roots()`

For each rational order $m$ (1,2,3,4) and smoothness parameter $\beta$ (= $\alpha/2$ with $\alpha$ between 0.5 and 2), function `my.get.roots()` (adapted from the `rSPDE` package) returns $\texttt{factor} = \dfrac{c_m}{b_{m+1}}$, and the roots $\{r_{1i}\}_{i=1}^m$ and $\{r_{2j}\}_{j=1}^{m+1}$.

```{r}
# Function to compute the roots and factor for the rational approximation
my.get.roots <- function(m, # rational order, m = 1, 2, 3, or 4
                         beta # smoothness parameter, beta = alpha/2 with alpha between 0.5 and 2
                         ) {
  m1table <- rSPDE:::m1table
  m2table <- rSPDE:::m2table
  m3table <- rSPDE:::m3table
  m4table <- rSPDE:::m4table
  mt <- get(paste0("m", m, "table"))
  rb <- rep(0, m + 1)
  rc <- rep(0, m)
  if(m == 1) {
    rc = approx(mt$beta, mt[[paste0("rc")]], beta)$y
  } else {
    rc = sapply(1:m, function(i) {
      approx(mt$beta, mt[[paste0("rc.", i)]], beta)$y
    })
  }
  rb = sapply(1:(m+1), function(i) {
    approx(mt$beta, mt[[paste0("rb.", i)]], xout = beta)$y
  })
  factor = approx(mt$beta, mt$factor, xout = beta)$y
  return(list(pl_roots = rb, # roots \{r_{2j}\}_{j=1}^{m+1}
              pr_roots = rc, # roots \{r_{1i}\}_{i=1}^m
              factor = factor # this is c_m/b_{m+1}
              ))
}
```

### Function `poly.from.roots()`

Function `poly.from.roots()` computes the coefficients of a polynomial from its roots.

```{r}
# Function to compute polynomial coefficients from roots
poly.from.roots <- function(roots) {
  coef <- 1
  for (r in roots) {coef <- convolve(coef, c(1, -r), type = "open")}
  return(coef) # returned in increasing order like a+bx+cx^2+...
}
```

### Function `compute.partial.fraction.param()`

Given `factor`$=\texttt{factor} = \dfrac{c_m}{b_{m+1}}$, `pr_roots`$=\{r_{1i}\}_{i=1}^m$, `pl_roots`$=\{r_{2j}\}_{j=1}^{m+1}$, `time_step`$=\tau$, and `scaling`$=\kappa^{2\beta}$, function `compute.partial.fraction.param()` computes the parameters for the partial fraction decomposition \eqref{eq:partial_fraction}.

```{r}
# Function to compute the parameters for the partial fraction decomposition
compute.partial.fraction.param <- function(factor, # c_m/b_{m+1}
                                           pr_roots, # roots \{r_{1i}\}_{i=1}^m
                                           pl_roots, # roots \{r_{2j}\}_{j=1}^{m+1}
                                           time_step, # \tau
                                           scaling # \kappa^{2\beta}
                                           ) {
  pr_coef <- c(0, poly.from.roots(pr_roots))
  pl_coef <- poly.from.roots(pl_roots)
  factor_pr_coef <- pr_coef
  pr_plus_pl_coef <- factor_pr_coef + ((scaling * time_step)/factor) * pl_coef
  res <- gsignal::residue(factor_pr_coef, pr_plus_pl_coef)
  return(list(r = res$r, # residues \{a_k\}_{k=1}^{m+1}
              p = res$p, # poles \{p_k\}_{k=1}^{m+1}
              k = res$k # remainder r
              ))
}
```

### Function `my.fractional.operators.frac()`

Given the Laplacian matrix `L`, the smoothness parameter `beta`, the mass matrix `C` (not lumped), the scaling factor `scale.factor`$=\kappa^2$, the rational order `m`, and the time step `time_step`$=\tau$, function `my.fractional.operators.frac()` computes the fractional operator and returns a list containing the necessary matrices and parameters for the fractional diffusion equation.

```{r}
# Function to compute the fractional operator
my.fractional.operators.frac <- function(L, # Laplacian matrix
                                         beta, # smoothness parameter beta
                                         C, # mass matrix (not lumped)
                                         scale.factor, # scaling parameter = kappa^2
                                         m = 1, # rational order, m = 1, 2, 3, or 4
                                         time_step # time step = tau
                                         ) {
  I <- Matrix::Diagonal(dim(C)[1])
  L <- L / scale.factor
  if(beta == 1){
    L <- L * scale.factor^beta
    return(list(C = C, # mass matrix
                L = L, # Laplacian matrix scaled
                m = m, # rational order
                beta = beta, # smoothness parameter
                LHS = C + time_step * L # left-hand side of the linear system
                ))
  } else {
    scaling <- scale.factor^beta
    roots <- my.get.roots(m, beta)
    poles_rs_k <- compute.partial.fraction.param(roots$factor, roots$pr_roots, roots$pl_roots, time_step, scaling)

    partial_fraction_terms <- list()
    for (i in 1:(m+1)) {
      # Here is where the terms in the sum in eq 12 are computed
      partial_fraction_terms[[i]] <- (L - poles_rs_k$p[i] * C)#/poles_rs_k$r[i]
      }
    return(list(C = C, # mass matrix
                L = L, # Laplacian matrix scaled
                m = m, # rational order
                beta = beta, # smoothness parameter
                partial_fraction_terms = partial_fraction_terms, # partial fraction terms
                residues = poles_rs_k$r # residues \{a_k\}_{k=1}^{m+1}
                ))
  }
}
```

### Function `my.solver.frac()`

Given the object returned by `my.fractional.operators.frac()` and a vector `v`, function `my.solver.frac()` solves the linear \eqref{eq:final_scheme2} for the vector `v`. If `beta = 1`, it solves the system directly; otherwise, it uses the partial fraction decomposition.

```{r}
# Function to solve the iteration
my.solver.frac <- function(obj, # object returned by my.fractional.operators.frac()
                           v # vector to be solved for
                           ){
  beta <- obj$beta
  m <- obj$m
  if (beta == 1){
    return(solve(obj$LHS, v) # solve the linear system directly for beta = 1
           )
  } else {
    partial_fraction_terms <- obj$partial_fraction_terms
    residues <- obj$residues
    output <- v*0
    for (i in 1:(m+1)) {output <- output + residues[i] * solve(partial_fraction_terms[[i]], v)}
    return(output # solve the linear system using the partial fraction decomposition
           )
  }
}


# my.solver.frac <- function(obj, v) {
#   beta <- obj$beta
#   m <- obj$m
#   
#   if (beta == 1) {
#     return(solve(obj$LHS, v))
#   } 
#   
#   partial_fraction_terms <- obj$partial_fraction_terms
#   residues <- obj$residues
#   
#   # --- PARALLEL PART ---
#   results <- parallel::mclapply(
#     1:(m+1),
#     function(i) residues[i] * solve(partial_fraction_terms[[i]], v),
#     mc.cores = parallel::detectCores()  
#   )
#   
#   return(Reduce(`+`, results))
# }

```


### Function `solve_fractional_evolution()`

Given the fractional operator object `my_op_frac`, a time step `time_step`, a sequence of time points `time_seq`, an initial value `val_at_0`, and the right-hand side matrix `RHST`, function `solve_fractional_evolution()` computes the solution to the fractional diffusion equation at each time step using scheme \eqref{eq:final_scheme2}.


```{r}
solve_fractional_evolution <- function(my_op_frac, time_step, time_seq, val_at_0, RHST) {
  CC <- my_op_frac$C
  SOL <- matrix(NA, nrow = nrow(CC), ncol = length(time_seq))
  SOL[, 1] <- val_at_0
  for (k in 1:(length(time_seq) - 1)) {
    rhs <- CC %*% SOL[, k] + time_step * RHST[, k + 1]
    SOL[, k + 1] <- as.matrix(my.solver.frac(my_op_frac, rhs))
  }
  return(SOL)
}
```


## Auxiliary functions {#auxiliary_functions}

### Function `gets.graph.tadpole()`

Given a mesh size `h`, function `gets.graph.tadpole()` builds a tadpole graph and creates a mesh.


```{r}
# Function to build a tadpole graph and create a mesh
gets.graph.tadpole <- function(h){
  edge1 <- rbind(c(0,0),c(1,0))
  theta <- seq(from=-pi,to=pi,length.out = 10000)
  edge2 <- cbind(1+1/pi+cos(theta)/pi,sin(theta)/pi)
  edges <- list(edge1, edge2)
  graph <- metric_graph$new(edges = edges, verbose = 0)
  graph$set_manual_edge_lengths(edge_lengths = c(1,2))
  graph$build_mesh(h = h)
  return(graph)
}
```


### Function `tadpole.eig()`

Given a mode number `k` and a tadpole graph `graph`, function `tadpole.eig()` computes the eigenpairs of the tadpole graph.


```{r}
# Function to compute the eigenfunctions of the tadpole graph
tadpole.eig <- function(k,graph){
x1 <- c(0,graph$get_edge_lengths()[1]*graph$mesh$PtE[graph$mesh$PtE[,1]==1,2]) 
x2 <- c(0,graph$get_edge_lengths()[2]*graph$mesh$PtE[graph$mesh$PtE[,1]==2,2]) 

if(k==0){ 
  f.e1 <- rep(1,length(x1)) 
  f.e2 <- rep(1,length(x2)) 
  f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
  f = list(phi=f1/sqrt(3)) 
  
} else {
  f.e1 <- -2*sin(pi*k*1/2)*cos(pi*k*x1/2) 
  f.e2 <- sin(pi*k*x2/2)                  
  
  f1 = c(f.e1[1],f.e2[1],f.e1[-1], f.e2[-1]) 
  
  if((k %% 2)==1){ 
    f = list(phi=f1/sqrt(3)) 
  } else { 
    f.e1 <- (-1)^{k/2}*cos(pi*k*x1/2)
    f.e2 <- cos(pi*k*x2/2)
    f2 = c(f.e1[1],f.e2[1],f.e1[-1],f.e2[-1]) 
    f <- list(phi=f1,psi=f2/sqrt(3/2))
  }
}
return(f)
}
```

### Function `gets.eigen.params()`

Given a finite number of modes `N_finite`, a scaling parameter `kappa`, a smoothness parameter `alpha`, and a tadpole graph `graph`, function `gets.eigen.params()` computes `EIGENVAL_ALPHA` (a vector with entries $\lambda_j^{\alpha/2}$), `EIGENVAL_MINUS_ALPHA` (a vector with entries $\lambda_j^{-\alpha/2}$), and `EIGENFUN` (a matrix with columns $e_j$ on the mesh of `graph`).


```{r}
# Function to compute the eigenpairs of the tadpole graph
gets.eigen.params <- function(N_finite = 4, kappa = 1, alpha = 0.5, graph){
  EIGENVAL <- NULL
  EIGENVAL_ALPHA <- NULL
  EIGENVAL_MINUS_ALPHA <- NULL
  EIGENFUN <- NULL
  INDEX <- NULL
  for (j in 0:N_finite) {
    lambda_j <- kappa^2 + (j*pi/2)^2
    lambda_j_alpha_half <- lambda_j^(alpha/2)
    lambda_j_minus_alpha_half <- lambda_j^(-alpha/2)
    e_j <- tadpole.eig(j,graph)$phi
    EIGENVAL <- c(EIGENVAL, lambda_j)
    EIGENVAL_ALPHA <- c(EIGENVAL_ALPHA, lambda_j_alpha_half)  
    EIGENVAL_MINUS_ALPHA <- c(EIGENVAL_MINUS_ALPHA, lambda_j_minus_alpha_half)
    EIGENFUN <- cbind(EIGENFUN, e_j)
    INDEX <- c(INDEX, j)
    if (j>0 && (j %% 2 == 0)) {
      lambda_j <- kappa^2 + (j*pi/2)^2
      lambda_j_alpha_half <- lambda_j^(alpha/2)
      lambda_j_minus_alpha_half <- lambda_j^(-alpha/2)
      e_j <- tadpole.eig(j,graph)$psi
      EIGENVAL <- c(EIGENVAL, lambda_j)
      EIGENVAL_ALPHA <- c(EIGENVAL_ALPHA, lambda_j_alpha_half)    
      EIGENVAL_MINUS_ALPHA <- c(EIGENVAL_MINUS_ALPHA, lambda_j_minus_alpha_half)
      EIGENFUN <- cbind(EIGENFUN, e_j)
      INDEX <- c(INDEX, j+0.1)
      }
    }
  return(list(EIGENVAL = EIGENVAL,
              EIGENVAL_ALPHA = EIGENVAL_ALPHA, 
              EIGENVAL_MINUS_ALPHA = EIGENVAL_MINUS_ALPHA,
              EIGENFUN = EIGENFUN,
              INDEX = INDEX))
}
```


### Function `construct_piecewise_projection()` {#construct_piecewise_projection}

Given a matrix `projected_U_approx` with approximated values at discrete time points, a sequence of time points `time_seq`, and an extended sequence of time points `overkill_time_seq`, function `construct_piecewise_projection()` constructs a piecewise constant projection of the approximated values over the extended time sequence.

```{r}
# Function to construct a piecewise constant projection of approximated values
construct_piecewise_projection <- function(projected_U_approx, time_seq, overkill_time_seq) {
  projected_U_piecewise <- matrix(NA, nrow = nrow(projected_U_approx), ncol = length(overkill_time_seq))
  
  # Assign value at t = 0
  projected_U_piecewise[, which(overkill_time_seq == 0)] <- projected_U_approx[, 1]
  
  # Assign values for intervals (t_{k-1}, t_k]
  for (k in 2:length(time_seq)) {
    idxs <- which(overkill_time_seq > time_seq[k - 1] & overkill_time_seq <= time_seq[k])
    projected_U_piecewise[, idxs] <- projected_U_approx[, k]
  }
  
  return(projected_U_piecewise)
}
```

### Function `compute_guiding_lines()`

Given a vector `x_axis_vector`, a matrix `errors` with error values, a vector `theoretical_rates` with theoretical convergence rates, and a function `line_equation_fun` (either `loglog_line_equation` or `exp_line_equation`), function `compute_guiding_lines()` computes guiding lines for convergence plots.

```{r}
loglog_line_equation <- function(x1, y1, slope) {
  b <- log10(y1 / (x1 ^ slope))
  
  function(x) {
    (x ^ slope) * (10 ^ b)
  }
}
exp_line_equation <- function(x1, y1, slope) {
  lnC <- log(y1) - slope * x1
  
  function(x) {
    exp(lnC + slope * x)
  }
}
compute_guiding_lines <- function(x_axis_vector, errors, theoretical_rates, line_equation_fun) {
  guiding_lines <- matrix(NA, nrow = length(x_axis_vector), ncol = length(theoretical_rates))
  
  for (j in seq_along(theoretical_rates)) {
    guiding_lines_aux <- matrix(NA, nrow = length(x_axis_vector), ncol = length(x_axis_vector))
    
    for (k in seq_along(x_axis_vector)) {
      point_x1 <- x_axis_vector[k]
      point_y1 <- errors[k, j]
      slope <- theoretical_rates[j]
      
      line <- line_equation_fun(x1 = point_x1, y1 = point_y1, slope = slope)
      guiding_lines_aux[, k] <- line(x_axis_vector)
    }
    
    guiding_lines[, j] <- rowMeans(guiding_lines_aux)
  }
  
  return(guiding_lines)
}
```



### Functions for computing an exact solution to the fractional diffusion equation {#exact_solution}

Below we present closed-form expressions for $\displaystyle G_j(t)= \int_0^t e^{\lambda^{\alpha/2}_jr}g(r)dr$ corresponding to some choices of $g(r)$.

\begin{aligned}
g(r) &= Ae^{-\lambda^{\alpha/2}_j r} &\implies\quad G_j(t) &= At, \quad A\in\mathbb{R} \\[1.5ex]
g(r) &= Ae^{\mu r} &\implies\quad G_j(t) &= A \frac{e^{(\lambda^{\alpha/2}_j+\mu)t} - 1}{\lambda^{\alpha/2}_j + \mu}, \quad -\lambda^{\alpha/2}_j \ne \mu \in \mathbb{R} \\[1.5ex]
g(r) &= Ar^n &\implies\quad G_j(t) &= A\frac{(-1)^{n+1} n!}{(\lambda_j^{\alpha/2})^{n+1}} \left(1 - e^{\lambda_j^{\alpha/2} t} \sum_{k=0}^n \frac{(-\lambda_j^{\alpha/2} t)^k}{k!} \right), \quad n=0,1,\dots \\[1.5ex]
g(r) &= A\sin(\omega r) &\implies\quad G_j(t) &= A \frac{e^{\lambda_j^{\alpha/2} t} \left( \lambda_j^{\alpha/2} \sin(\omega t) - \omega \cos(\omega t) \right) + \omega}{(\lambda_j^{\alpha/2})^2 + \omega^2}, \quad \omega \in \mathbb{R} \\[1.5ex]
g(r) &= A\cos(\theta r) &\implies\quad G_j(t) &= A \frac{e^{\lambda_j^{\alpha/2} t} \left( \lambda_j^{\alpha/2} \cos(\theta t) + \theta \sin(\theta t) \right) - \lambda_j^{\alpha/2}}{(\lambda_j^{\alpha/2})^2 + \theta^2}, \quad \theta \in \mathbb{R}
\end{aligned}


```{r}
# Functions to compute the exact solution to the fractional diffusion equation
g_linear <- function(r, A, lambda_j_alpha_half) {
  return(A * exp(-lambda_j_alpha_half * r))
  }
G_linear <- function(t, A) {
  return(A * t)
  }
g_exp <- function(r, A, mu) {
  return(A * exp(mu * r))
  }
G_exp <- function(t, A, lambda_j_alpha_half, mu) {
  exponent <- lambda_j_alpha_half + mu
  return(A * (exp(exponent * t) - 1) / exponent)
  }
g_poly <- function(r, A, n) {
  return(A * r^n)
}
G_poly <- function(t, A, lambda_j_alpha_half, n) {
  t <- as.vector(t)
  k_vals <- 0:n
  sum_term <- sapply(t, function(tt) {
    sum(((-lambda_j_alpha_half * tt)^k_vals) / factorial(k_vals))
  })
  coeff <- ((-1)^(n + 1)) * factorial(n) / (lambda_j_alpha_half^(n + 1))
  return(A * coeff * (1 - exp(lambda_j_alpha_half * t) * sum_term))
}
g_sin <- function(r, A, omega) {
  return(A * sin(omega * r))
}
G_sin <- function(t, A, lambda_j_alpha_half, omega) {
  denom <- lambda_j_alpha_half^2 + omega^2
  numerator <- exp(lambda_j_alpha_half * t) * (lambda_j_alpha_half * sin(omega * t) - omega * cos(omega * t)) + omega
  return(A * numerator / denom)
}
g_cos <- function(r, A, theta) {
  return(A * cos(theta * r)) 
}
G_cos <- function(t, A, lambda_j_alpha_half, theta) {
  denom <- lambda_j_alpha_half^2 + theta^2
  numerator <- exp(lambda_j_alpha_half * t) * (lambda_j_alpha_half * cos(theta * t) + theta * sin(theta * t)) - lambda_j_alpha_half
  return(A * numerator / denom)
}
```

### Function `reversecolumns()`

Given a matrix `mat`, function `reversecolumns()` reverses the order of its columns.


```{r}
reversecolumns <- function(mat) {
  return(mat[, rev(seq_len(ncol(mat)))])
}
```


## For Euclidean domains (rectangle) {#euclidean_domains}

```{r}
gets.mesh.and.FEM.on.rectangle <- function(a, b, n) {
  mesh <- fmesher::fm_rcdt_2d(
  lattice = fmesher::fm_lattice_2d(
  x = seq(0, a, length.out = n + 1),
  y = seq(0, b, length.out = n + 1)
), extend = FALSE)
  FEM <- fmesher::fm_fem(mesh, order = 1)
  return(list(mesh = mesh, 
              Cl = FEM$c0,
              C = FEM$c1,
              G = FEM$g1))
}
```



```{r}
compute_true_eigen_rectangle <- function(a = 1,
                                         b = 1,
                                         loc,
                                         kappa = 1, 
                                         alpha = 1,
                                         m_max = 3, 
                                         n_max = 3) {
  loc_x <- loc[,1]
  loc_y <- loc[,2]
  N <- (m_max+1)*(n_max+1)
  # Prepare lists
  EIGENVAL <- numeric(N)
  m_vector <- numeric(N)
  n_vector <- numeric(N)
  EIGENFUN <- matrix(0, nrow = nrow(loc), ncol = N)
  
  i <- 0 
  # Loop over mode indices
  for (m in 0:m_max) {
    for (n in 0:n_max) {
      lambda_mn <- kappa^2 + pi^2 * ((m^2 / a^2) + (n^2 / b^2))
      EIGENVAL[i + 1] <- lambda_mn
      # Evaluate eigenfunction on mesh grid
      phi_mn <-  cos(m*pi*loc_x/a) * cos(n*pi*loc_y/b)
      EIGENFUN[, i + 1] <- phi_mn
      m_vector[i + 1] <- m
      n_vector[i + 1] <- n
      i <- i + 1
    }
  }
  # Sort eigenvalues and corresponding eigenfunctions
  idx <- order(EIGENVAL)
  EIGENVAL <- EIGENVAL[idx]
  EIGENVAL_ALPHA <- EIGENVAL^(alpha/2)
  EIGENFUN <- EIGENFUN[, idx]
  m_vector <- m_vector[idx]
  n_vector <- n_vector[idx]
  
  return(list(EIGENVAL = EIGENVAL,
              EIGENVAL_ALPHA = EIGENVAL_ALPHA,
              EIGENFUN = EIGENFUN,
              m_vector = m_vector,
              n_vector = n_vector))
}
```


## For manifolds (sphere) {#manifolds}

```{r}
# Function to get globe from h using spline
globe_from_h <- function(h) {
  readed_data <- readRDS(
    file = here::here("data_files/h_min_max_vs_globe.RDS")
  )
  
  globe_vector <- readed_data$globe_vector
  h_min_vector <- readed_data$h_min_vector
  
  ord <- order(h_min_vector)
  h_sorted <- h_min_vector[ord]
  globe_sorted <- globe_vector[ord]
  
  # Create a smooth spline function: globe as function of h_max
  spline_func <- splinefun(x = h_sorted, 
                           y = globe_sorted, 
                           method = "monoH.FC")

  return(round(spline_func(h)))
}
```


```{r}
calculate_laplace_beltrami_eigenvalues <- function(kappa = 0, L_max = 5, rot.inv = FALSE) {
  
  eigenvalues <- numeric(0)
  
  for (l in 0:L_max) {
    lambda_l <- kappa^2 + l * (l + 1)
    
    if (rot.inv) {
      # Only one eigenvalue per l (rotational invariance)
      eigenvalues <- c(eigenvalues, lambda_l)
    } else {
      # Full multiplicity: (m = -l,...,l)
      multiplicity <- 2 * l + 1
      eigenvalues <- c(eigenvalues, rep(lambda_l, multiplicity))
    }
  }
  
  return(eigenvalues)
}


compute_true_eigen_sphere <- function(mesh, 
                                      kappa,
                                      alpha,
                                      L_max,
                                      rot.inv){
  EIGENFUN <- fmesher::fm_raw_basis(
    mesh = mesh, 
    type = "sph.harm",
    n = L_max, 
    rot.inv = rot.inv)
  EIGENVAL <- calculate_laplace_beltrami_eigenvalues(
    kappa = kappa, 
    L_max = L_max, 
    rot.inv = rot.inv)
  idx <- order(EIGENVAL)
  EIGENVAL <- EIGENVAL[idx]
  EIGENVAL_ALPHA <- EIGENVAL^(alpha/2)
  EIGENFUN <- EIGENFUN[, idx]
  return(list(EIGENVAL = EIGENVAL,
              EIGENVAL_ALPHA = EIGENVAL_ALPHA,
              EIGENFUN = EIGENFUN))
}
```


## Plotting functions {#plotting_functions}

### Function `plotting.order()`

Given a vector `v` and a graph object `graph`, function `plotting.order()` orders the mesh values for plotting.

```{r}
# Function to order the vertices for plotting
plotting.order <- function(v, graph){
  edge_number <- graph$mesh$VtE[, 1]
  pos <- sum(edge_number == 1)+1
  return(c(v[1], v[3:pos], v[2], v[(pos+1):length(v)], v[2]))
}
```

### Function `global.scene.setter()`

Given ranges for the `x`, `y`, and `z` axes, and an optional aspect ratio for the `z` axis, function `global.scene.setter()` sets the scene for 3D plots so that all plots have the same aspect ratio and camera position.

```{r}
# Function to set the scene for 3D plots
global.scene.setter <- function(x_range, y_range, z_range, z_aspectratio = 4) {
  
  return(list(xaxis = list(title = "x", range = x_range),
              yaxis = list(title = "y", range = y_range),
              zaxis = list(title = "z", range = z_range),
              aspectratio = list(x = 2*(1+2/pi), 
                                 y = 2*(2/pi), 
                                 z = z_aspectratio*(2/pi)),
              camera = list(eye = list(x = (1+2/pi)/2, 
                                       y = 4, 
                                       z = 2),
                            center = list(x = (1+2/pi)/2, 
                                          y = 0, 
                                          z = 0))))
}
```


```{r}
global.scene.setter.rec.and.sphere <- function(x_range, y_range, z_range) {
  
  return(list(xaxis = list(title = "x", range = x_range),
              yaxis = list(title = "y", range = y_range),
              zaxis = list(title = "z", range = z_range),
              aspectratio = list(x = 1, 
                                 y = 1, 
                                 z = 1),
              camera = list(eye = list(x = 2, 
                                       y = 2, 
                                       z = 2),
                            center = list(x = 0, 
                                          y = 0, 
                                          z = 0))))
}


global.scene.setter.rec.and.sphere.for.hat <- function(x_range, y_range, z_range) {
  
  return(list(xaxis = list(title = "x", range = x_range),
              yaxis = list(title = "y", range = y_range),
              zaxis = list(title = "z", range = z_range),
              aspectratio = list(x = 1, 
                                 y = 1, 
                                 z = 0.4),
              camera = list(eye = list(x = 2, 
                                       y = 2, 
                                       z = 2),
                            center = list(x = 0, 
                                          y = 0, 
                                          z = 0))))
}
```

### Function `graph.plotter.3d()`

Given a graph object `graph`, a sequence of time points `time_seq`, and one or more matrices `...` representing function values defined on the mesh of `graph` at each time in `time_seq`, the `graph.plotter.3d()` function generates an interactive 3D visualization of these values over time.

```{r}
# Function to plot in 3D
graph.plotter.3d.old <- function(graph, time_seq, frame_val_to_display, ...) {
  U_list <- list(...)
  U_names <- sapply(substitute(list(...))[-1], deparse)

  # Spatial coordinates
  x <- plotting.order(graph$mesh$V[, 1], graph)
  y <- plotting.order(graph$mesh$V[, 2], graph)
  weights <- graph$mesh$weights

  # Apply plotting.order to each U
  U_list <- lapply(U_list, function(U) apply(U, 2, plotting.order, graph = graph))
  n_vars <- length(U_list)

  # Create plot_data frame with time and position replicated
  n_time <- ncol(U_list[[1]])
  base_data <- data.frame(
    x = rep(x, times = n_time),
    y = rep(y, times = n_time),
    the_graph = 0,
    frame = rep(time_seq, each = length(x))
  )

  # Add U columns to plot_data
  for (i in seq_along(U_list)) {
    base_data[[paste0("u", i)]] <- as.vector(U_list[[i]])
  }

  plot_data <- base_data

  # Generate vertical lines
  vertical_lines_list <- lapply(seq_along(U_list), function(i) {
    do.call(rbind, lapply(time_seq, function(t) {
      idx <- which(plot_data$frame == t)
      z_vals <- plot_data[[paste0("u", i)]][idx]
      data.frame(
        x = rep(plot_data$x[idx], each = 3),
        y = rep(plot_data$y[idx], each = 3),
        z = as.vector(t(cbind(0, z_vals, NA))),
        frame = rep(t, each = length(idx) * 3)
      )
    }))
  })

  # Set axis ranges
  z_range <- range(unlist(U_list))
  x_range <- range(x)
  y_range <- range(y)

  # Create plot
  p <- plot_ly(plot_data, frame = ~frame) %>%
    add_trace(x = ~x, y = ~y, z = ~the_graph, type = "scatter3d", mode = "lines",
              name = "", showlegend = FALSE,
              line = list(color = "black", width = 3))

  # Add traces for each variable
  colors <- rev(viridisLite::viridis(n_vars)) #RColorBrewer::brewer.pal(min(n_vars, 8), "Set1")
  for (i in seq_along(U_list)) {
    p <- add_trace(p,
      x = ~x, y = ~y, z = as.formula(paste0("~u", i)),
      type = "scatter3d", mode = "lines", name = U_names[i],
      line = list(color = colors[i], width = 3))
  }

  # Add vertical lines
  for (i in seq_along(vertical_lines_list)) {
    p <- add_trace(p,
      data = vertical_lines_list[[i]],
      x = ~x, y = ~y, z = ~z, frame = ~frame,
      type = "scatter3d", mode = "lines",
      line = list(color = "gray", width = 0.5),
      name = "Vertical lines",
      showlegend = FALSE)
  }
  frame_name <- deparse(substitute(frame_val_to_display))
  # Layout and animation controls
  p <- p %>%
    layout(
      scene = global.scene.setter(x_range, y_range, z_range),
      updatemenus = list(list(type = "buttons", showactive = FALSE,
                              buttons = list(
                                list(label = "Play", method = "animate",
                                     args = list(NULL, list(frame = list(duration = 2000 / length(time_seq), redraw = TRUE), fromcurrent = TRUE))),
                                list(label = "Pause", method = "animate",
                                     args = list(NULL, list(mode = "immediate", frame = list(duration = 0), redraw = FALSE)))
                              )
      )),
      title = paste0(frame_name,": ", formatC(frame_val_to_display[1], format = "f", digits = 4))
    ) %>%
    plotly_build()

  for (i in seq_along(p$x$frames)) {
    p$x$frames[[i]]$layout <- list(title = paste0(frame_name,": ", formatC(frame_val_to_display[i], format = "f", digits = 4)))
  }

  return(p)
}
```


### Function `graph.plotter.3d()`

Given a graph object `graph`, a sequence of time points `time_seq`, a vector `frame_val_to_display` with values to display at each frame, and a list `U_list` representing function values defined on the mesh of `graph` at each time in `time_seq`, the `graph.plotter.3d()` function generates an interactive 3D visualization of these functions over time.

```{r}
graph.plotter.3d <- function(graph, time_seq, frame_val_to_display, U_list) {
  U_names <- names(U_list) 
  # Spatial coordinates
  x <- plotting.order(graph$mesh$V[, 1], graph)
  y <- plotting.order(graph$mesh$V[, 2], graph)
  weights <- graph$mesh$weights

  # Apply plotting.order to each U
  U_list <- lapply(U_list, function(U) apply(U, 2, plotting.order, graph = graph))
  n_vars <- length(U_list)
  
  # Create plot_data frame with time and position replicated
  n_time <- ncol(U_list[[1]])
  base_data <- data.frame(
    x = rep(x, times = n_time),
    y = rep(y, times = n_time),
    the_graph = 0,
    frame = rep(time_seq, each = length(x))
  )

  # Add U columns to plot_data
  for (i in seq_along(U_list)) {
    base_data[[paste0("u", i)]] <- as.vector(U_list[[i]])
  }

  plot_data <- base_data

  # Generate vertical lines
  vertical_lines_list <- lapply(seq_along(U_list), function(i) {
    do.call(rbind, lapply(time_seq, function(t) {
      idx <- which(plot_data$frame == t)
      z_vals <- plot_data[[paste0("u", i)]][idx]
      data.frame(
        x = rep(plot_data$x[idx], each = 3),
        y = rep(plot_data$y[idx], each = 3),
        z = as.vector(t(cbind(0, z_vals, NA))),
        frame = rep(t, each = length(idx) * 3)
      )
    }))
  })

  # Set axis ranges
  z_range <- range(unlist(U_list))
  x_range <- range(x)
  y_range <- range(y)

  # Create plot
  p <- plot_ly(plot_data, frame = ~frame) %>%
    add_trace(x = ~x, y = ~y, z = ~the_graph, type = "scatter3d", mode = "lines",
              name = "", showlegend = FALSE,
              line = list(color = "black", width = 3))

  if (n_vars == 2) {
    colors <- RColorBrewer::brewer.pal(min(n_vars, 8), "Set1") 
    } else {
    colors <- rev(viridisLite::viridis(n_vars)) 
  }
  # RColorBrewer::brewer.pal(min(n_vars, 8), "Set1")
  for (i in seq_along(U_list)) {
    p <- add_trace(p,
      x = ~x, y = ~y, z = as.formula(paste0("~u", i)),
      type = "scatter3d", mode = "lines", name = U_names[i],
      line = list(color = colors[i], width = 3))
  }

  # Add vertical lines
  for (i in seq_along(vertical_lines_list)) {
    p <- add_trace(p,
      data = vertical_lines_list[[i]],
      x = ~x, y = ~y, z = ~z, frame = ~frame,
      type = "scatter3d", mode = "lines",
      line = list(color = "gray", width = 0.5),
      name = "Vertical lines",
      showlegend = FALSE)
  }
  frame_name <- deparse(substitute(frame_val_to_display))
  # Layout and animation controls
  p <- p %>%
    layout(
      scene = global.scene.setter(x_range, y_range, z_range),
      updatemenus = list(list(type = "buttons", showactive = FALSE,
                              buttons = list(
                                list(label = "Play", method = "animate",
                                     args = list(NULL, list(frame = list(duration = 2000 / length(time_seq), redraw = TRUE), fromcurrent = TRUE))),
                                list(label = "Pause", method = "animate",
                                     args = list(NULL, list(mode = "immediate", frame = list(duration = 0), redraw = FALSE)))
                              )
      )),
      title = paste0(frame_name,": ", formatC(frame_val_to_display[1], format = "f", digits = 4))
    ) %>%
    plotly_build()

  for (i in seq_along(p$x$frames)) {
    p$x$frames[[i]]$layout <- list(title = paste0(frame_name,": ", formatC(frame_val_to_display[i], format = "f", digits = 4)))
  }

  return(p)
}
```

### Function `error.at.each.time.plotter()`

Given a graph object `graph`, a matrix `U_true` of true values, a matrix `U_approx` of approximated values, a sequence of time points `time_seq`, and a time step `time_step`, function `error.at.each.time.plotter()` computes the error at each time step and generates a plot showing the error over time.

```{r}
# Function to plot the error at each time step
error.at.each.time.plotter <- function(graph, U_true, U_approx, time_seq, time_step) {
  weights <- graph$mesh$weights
  error_at_each_time <- t(weights) %*% (U_true - U_approx)^2
  error <- sqrt(as.double(t(weights) %*% (U_true - U_approx)^2 %*% rep(time_step, ncol(U_true))))
  p <- plot_ly() %>% 
  add_trace(
  x = ~time_seq, y = ~error_at_each_time, type = 'scatter', mode = 'lines+markers',
  line = list(color = 'blue', width = 2),
  marker = list(color = 'blue', size = 4),
  name = "",
  showlegend = TRUE
) %>% 
  layout(
  title = paste0("Error at Each Time Step (Total error = ", formatC(error, format = "f", digits = 9), ")"),
  xaxis = list(title = "t"),
  yaxis = list(title = "Error"),
  legend = list(x = 0.1, y = 0.9)
)
  return(p)
}
```

### Function `graph.plotter.3d.comparer()`

Given a graph object `graph`, matrices `U_true` and `U_approx` representing true and approximated values, and a sequence of time points `time_seq`, function `graph.plotter.3d.comparer()` generates a 3D plot comparing the true and approximated values over time, with color-coded traces for each time point.

```{r}
# Function to plot the 3D comparison of U_true and U_approx
graph.plotter.3d.comparer <- function(graph, U_true, U_approx, time_seq) {
  x <- graph$mesh$V[, 1]; y <- graph$mesh$V[, 2]
  x <- plotting.order(x, graph); y <- plotting.order(y, graph)

  U_true <- apply(U_true, 2, plotting.order, graph = graph)
  U_approx <- apply(U_approx, 2, plotting.order, graph = graph)
  n_times <- length(time_seq)
  
  x_range <- range(x); y_range <- range(y); z_range <- range(c(U_true, U_approx))
  
  # Normalize time_seq
  time_normalized <- (time_seq - min(time_seq)) / (max(time_seq) - min(time_seq))
  blues <- colorRampPalette(c("lightblue", "blue"))(n_times)
  reds <- colorRampPalette(c("mistyrose", "red"))(n_times)
  
  # Accurate colorscales
  colorscale_greens <- Map(function(t, col) list(t, col), time_normalized, blues)
  colorscale_reds <- Map(function(t, col) list(t, col), time_normalized, reds)
  
  p <- plot_ly()
  
  # Static black graph structure
  p <- p %>%
    add_trace(x = x, y = y, z = rep(0, length(x)),
              type = "scatter3d", mode = "lines",
              line = list(color = "black", width = 4),
              name = "Graph", showlegend = FALSE)
  
  # U_true traces (green)
  for (i in seq_len(n_times)) {
    z <- U_true[, i]
    p <- add_trace(
      p,
      type = "scatter3d",
      mode = "lines",
      x = x, y = y, z = z,
      line = list(color = blues[i], width = 4),
      showlegend = FALSE,
      scene = "scene"
    )
  }
  
  # U_approx traces (dashed red)
  for (i in seq_len(n_times)) {
    z <- U_approx[, i]
    p <- add_trace(
      p,
      type = "scatter3d",
      mode = "lines",
      x = x, y = y, z = z,
      line = list(color = reds[i], width = 4, dash = "dot"),
      showlegend = FALSE,
      scene = "scene"
    )
  }
  
  # Dummy green colorbar (True) – with ticks
  p <- add_trace(
    p,
    type = "heatmap",
    z = matrix(time_seq, nrow = 1),
    showscale = TRUE,
    colorscale = colorscale_greens,
    colorbar = list(
      title = list(font = list(size = 12, color = "black"), text = "Time", side = "top"),
      len = 0.9,
      thickness = 15,
      x = 1.02,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle",
      tickvals = NULL,   # hide tick values
      ticktext = NULL,
      ticks = ""         # also hides tick marks
    ),
    x = matrix(time_seq, nrow = 1),
    y = matrix(1, nrow = 1),
    hoverinfo = "skip",
    opacity = 0
  )

# Dummy red colorbar (Approx) – no ticks
  p <- add_trace(
    p,
    type = "heatmap",
    z = matrix(time_seq, nrow = 1),
    showscale = TRUE,
    colorscale = colorscale_reds,
    colorbar = list(
      title = list(font = list(size = 12, color = "black"), text = ".", side = "top"),
      len = 0.9,
      thickness = 15,
      x = 1.05,
      xanchor = "left",
      y = 0.5,
      yanchor = "middle"
    ),
    x = matrix(time_seq, nrow = 1),
    y = matrix(1, nrow = 1),
    hoverinfo = "skip",
    opacity = 0
  )
  p <- p %>%
    add_trace(x = x, y = y, z = rep(0, length(x)),
              type = "scatter3d", mode = "lines",
              line = list(color = "black", width = 4),
              name = "Graph", showlegend = FALSE)
  p <- layout(p,
            scene = global.scene.setter(x_range, y_range, z_range),
            xaxis = list(visible = FALSE),
            yaxis = list(visible = FALSE),
            annotations = list(
  list(
    text = "Exact",
    x = 1.045,
    y = 0.5,
    xref = "paper",
    yref = "paper",
    showarrow = FALSE,
    font = list(size = 12, color = "black"),
    textangle = -90
  ),
  list(
    text = "Approx",
    x = 1.075,
    y = 0.5,
    xref = "paper",
    yref = "paper",
    showarrow = FALSE,
    font = list(size = 12, color = "black"),
    textangle = -90
  )
)

)

  
  return(p)
}
```

### Function `graph.plotter.3d.single()`

Given a graph object `graph`, a matrix `U_true` representing true values, and a sequence of time points `time_seq`, function `graph.plotter.3d.single()` generates a 3D plot of the true values over time, with color-coded traces for each time point.

```{r}
# Function to plot a single 3D line for 
graph.plotter.3d.single <- function(graph, U_true, time_seq) {
  x <- graph$mesh$V[, 1]; y <- graph$mesh$V[, 2]
  x <- plotting.order(x, graph); y <- plotting.order(y, graph)

  U_true <- apply(U_true, 2, plotting.order, graph = graph)
  n_times <- length(time_seq)
  
  x_range <- range(x); y_range <- range(y); z_range <- range(U_true)
  z_range[1] <- z_range[1] - 10^-6
  viridis_colors <- viridisLite::viridis(100)
  
  # Normalize time_seq
  time_normalized <- (time_seq - min(time_seq)) / (max(time_seq) - min(time_seq))
  #greens <- colorRampPalette(c("palegreen", "darkgreen"))(n_times)
  greens <- colorRampPalette(c(viridis_colors[1], viridis_colors[50],  viridis_colors[100]))(n_times)
  # Accurate colorscales
  colorscale_greens <- Map(function(t, col) list(t, col), time_normalized, greens)
  
  p <- plot_ly()
  
  # Add the 3D lines with fading green color
  for (i in seq_len(n_times)) {
    z <- U_true[, i]
    
    p <- add_trace(
      p,
      type = "scatter3d",
      mode = "lines",
      x = x,
      y = y,
      z = z,
      line = list(color = greens[i], width = 2),
      showlegend = FALSE,
      scene = "scene"
    )
  }
  p <- p %>%
    add_trace(x = x, y = y, z = rep(0, length(x)),
              type = "scatter3d", mode = "lines",
              line = list(color = "black", width = 5),
              name = "Graph", showlegend = FALSE)
  # Add dummy heatmap to show colorbar (not part of scene)
  p <- add_trace(
    p,
    type = "heatmap",
    z = matrix(time_seq, nrow = 1),
    showscale = TRUE,
    colorscale = colorscale_greens,
    colorbar = list(
    title = list(font = list(size = 12, color = "black"), text = "Time", side = "top"),
    len = 0.9,         # height (0 to 1)
    thickness = 15,     # width in pixels
    x = 1.02,           # shift it slightly right of the plot
    xanchor = "left",
    y = 0.5,
    yanchor = "middle"),
    x = matrix(time_seq, nrow = 1),
    y = matrix(1, nrow = 1),
    hoverinfo = "skip",
    opacity = 0
  )
  
  p <- layout(p,
              scene = global.scene.setter(x_range, y_range, z_range),
              xaxis = list(visible = FALSE),
              yaxis = list(visible = FALSE)
  )
  
  return(p)
}
```

### Function `error.convergence.plotter()`

Given a vector `x_axis_vector`, a vector `alpha_vector` of parameter values, a matrix `errors` with error values, vectors `theoretical_rates` and `observed_rates` with convergence rates, a function `line_equation_fun` (either `loglog_line_equation` or `exp_line_equation`), and plot titles and labels, function `error.convergence.plotter()` generates a convergence plot showing errors and guiding lines.

```{r}
# Function to plot the error convergence
error.convergence.plotter <- function(x_axis_vector, 
                                      alpha_vector, 
                                      errors, 
                                      theoretical_rates, 
                                      observed_rates,
                                      line_equation_fun,
                                      fig_title,
                                      x_axis_label,
                                      apply_sqrt = FALSE) {
  
  x_vec <- if (apply_sqrt) sqrt(x_axis_vector) else x_axis_vector
  
  guiding_lines <- compute_guiding_lines(x_axis_vector = x_vec, 
                                         errors = errors, 
                                         theoretical_rates = theoretical_rates, 
                                         line_equation_fun = line_equation_fun)
  
  default_colors <- scales::hue_pal()(length(alpha_vector))
  
  plot_lines <- lapply(1:ncol(guiding_lines), function(i) {
    geom_line(
      data = data.frame(x = x_vec, y = guiding_lines[, i]),
      aes(x = x, y = y),
      color = default_colors[i],
      linetype = "dashed",
      show.legend = FALSE
    )
  })
  
  df <- as.data.frame(cbind(x_vec, errors))
  colnames(df) <- c("x_axis_vector", alpha_vector)
  df_melted <- melt(df, id.vars = "x_axis_vector", variable.name = "column", value.name = "value")
  
  custom_labels <- paste0(formatC(alpha_vector, format = "f", digits = 2), 
                          " | ", 
                          formatC(theoretical_rates, format = "f", digits = 4), 
                          " | ", 
                          formatC(observed_rates, format = "f", digits = 4))
  
  df_melted$column <- factor(df_melted$column, levels = alpha_vector, labels = custom_labels)

  p <- ggplot() +
    geom_line(data = df_melted, aes(x = x_axis_vector, y = value, color = column)) +
    geom_point(data = df_melted, aes(x = x_axis_vector, y = value, color = column)) +
    plot_lines +
    labs(
      title = fig_title,
      x = x_axis_label,
      y = expression(Error),
      color = "          α  | theo  | obs"
    ) +
    (if (apply_sqrt) {
      scale_x_continuous(breaks = x_vec, labels = round(x_axis_vector, 4))
    } else {
      scale_x_log10(breaks = x_axis_vector, labels = round(x_axis_vector, 4))
    }) +
    (if (apply_sqrt) {
      scale_y_continuous(trans = "log", labels = scales::scientific_format())
    } else {
      scale_y_log10(labels = scales::scientific_format())
    }) +
    theme_minimal() +
    theme(text = element_text(family = "Palatino"),
          legend.position = "bottom",
          legend.direction = "vertical",
          plot.margin = margin(0, 0, 0, 0),
          plot.title = element_text(hjust = 0.5, size = 18, face = "bold"))
  
  return(p)
}

errorConvergencePlotterLatex <- function(x_axis_vector, 
                                      alpha_vector, 
                                      errors, 
                                      theoretical_rates, 
                                      observed_rates,
                                      line_equation_fun,
                                      fig_title,
                                      x_axis_label,
                                      y_axis_label,
                                      color_label,
                                      apply_sqrt = FALSE) {

  relative_per_error <- 100 * abs(theoretical_rates - observed_rates) / abs(theoretical_rates)
  
  x_vec <- if (apply_sqrt) sqrt(x_axis_vector) else x_axis_vector
  
  guiding_lines <- compute_guiding_lines(x_axis_vector = x_vec, 
                                         errors = errors, 
                                         theoretical_rates = theoretical_rates, 
                                         line_equation_fun = line_equation_fun)
  
  default_colors <- scales::hue_pal()(length(alpha_vector))
  
  plot_lines <- lapply(1:ncol(guiding_lines), function(i) {
    geom_line(
      data = data.frame(x = x_vec, y = guiding_lines[, i]),
      aes(x = x, y = y),
      color = default_colors[i],
      linetype = "dashed",
      linewidth = 0.6,
      show.legend = FALSE
    )
  })
  
  df <- as.data.frame(cbind(x_vec, errors))
  colnames(df) <- c("x_axis_vector", alpha_vector)
  df_melted <- melt(df, id.vars = "x_axis_vector", variable.name = "column", value.name = "value")
  
  custom_labels <- sprintf("$%s \\; | \\; %s \\; | \\; %s \\; | \\; %s$",
                           formatC(alpha_vector, format = "f", digits = 2),
                           formatC(theoretical_rates, format = "f", digits = 4),
                           formatC(observed_rates, format = "f", digits = 4),
                           formatC(relative_per_error, format = "f", digits = 2))
  
  df_melted$column <- factor(df_melted$column, levels = alpha_vector, labels = custom_labels)

  p <- ggplot() +
    geom_line(data = df_melted, aes(x = x_axis_vector, y = value, color = column), linewidth = 1.6) +
    geom_point(data = df_melted, aes(x = x_axis_vector, y = value, color = column), size = 2) +
    plot_lines +
    labs(
      title = fig_title,
      x = x_axis_label,
      y = y_axis_label,
      color = color_label
    ) +
    (if (apply_sqrt) {
      scale_x_continuous(breaks = x_vec, labels = round(x_axis_vector, 4))
    } else {
      scale_x_log10(breaks = x_axis_vector, labels = round(x_axis_vector, 4))
    }) +
    (if (apply_sqrt) {
      #scale_y_continuous(trans = "log", labels = scales::scientific_format())
      scale_y_continuous(trans = scales::log_trans(base = exp(1)), labels = scales::scientific_format())
    } else {
      scale_y_log10(labels = scales::scientific_format())
    }) +
    theme_minimal() +
    theme(text = element_text(family = "Palatino"),
          legend.position = "bottom",
          legend.direction = "vertical",
          #plot.margin = margin(0, 0, 0, 0),
          plot.title = element_text(hjust = 0.5, size = 18, face = "plain"))
  
  return(p)
}
```


### Function `graph.plotter.3d.static()`

Given a graph object `graph` and a list `z_list` of function values defined on the mesh of `graph`, function `graph.plotter.3d.static()` generates a static 3D plot of these values.

```{r}
graph.plotter.3d.static <- function(graph, z_list) {
  x <- plotting.order(graph$mesh$V[, 1], graph)
  y <- plotting.order(graph$mesh$V[, 2], graph)
  U_names <- names(z_list)
  n_vars <- length(z_list)
  z_list <- lapply(z_list, function(z) plotting.order(z, graph))

  # Axis ranges
  z_range <- range(unlist(z_list))
  x_range <- range(x)
  y_range <- range(y)

  if (n_vars == 2) {
    colors <- RColorBrewer::brewer.pal(min(n_vars, 8), "Set1") 
    } else {
    colors <- rev(viridisLite::viridis(n_vars)) 
  }
  p <- plot_ly()

  for (i in seq_along(z_list)) {
    z <- z_list[[i]]

    # Main 3D curve
    p <- add_trace(
      p,
      x = x, y = y, z = z,
      type = "scatter3d", mode = "lines",
      line = list(color = colors[i], width = 3),
      name = U_names[i], showlegend = TRUE
    )

    # Efficient vertical lines: one trace with breaks (NA)
    x_vert <- rep(x, each = 3)
    y_vert <- rep(y, each = 3)
    z_vert <- unlist(lapply(z, function(zj) c(0, zj, NA)))

    p <- add_trace(
      p,
      x = x_vert, y = y_vert, z = z_vert,
      type = "scatter3d", mode = "lines",
      line = list(color = "gray", width = 0.5),
      showlegend = FALSE
    )
  }
  p <- p %>% add_trace(x = x, y = y, z = x*0, type = "scatter3d", mode = "lines",
              line = list(color = "black", width = 3),
              name = "thegraph", showlegend = FALSE) %>%
    layout(scene = global.scene.setter(x_range, y_range, z_range))
  return(p)
}

```


### Function `graph.plotter.3d.two.meshes.time()`

Given two graph objects `graph_finer` and `graph_coarser`, sequences of time points `time_seq` and `frame_val_to_display`, and lists `fs_finer` and `fs_coarser` of function values defined on the meshes of `graph_finer` and `graph_coarser` respectively, function `graph.plotter.3d.two.meshes.time()` generates an interactive 3D visualization comparing these functions over time.


```{r}
graph.plotter.3d.two.meshes.time <- function(graph_finer, graph_coarser, 
                                             time_seq, frame_val_to_display,
                                             fs_finer = list(), fs_coarser = list()) {
  # Spatial coordinates (ordered for plotting)
  x_finer <- plotting.order(graph_finer$mesh$V[, 1], graph_finer)
  y_finer <- plotting.order(graph_finer$mesh$V[, 2], graph_finer)
  x_coarser <- plotting.order(graph_coarser$mesh$V[, 1], graph_coarser)
  y_coarser <- plotting.order(graph_coarser$mesh$V[, 2], graph_coarser)
  
  n_time <- if (length(fs_finer) > 0) ncol(fs_finer[[1]]) else ncol(fs_coarser[[1]])

  # Helper: make dataframe from one function
  make_df <- function(f_mat, graph, x, y, mesh_name) {
    z <- apply(f_mat, 2, plotting.order, graph = graph)
    data.frame(
      x = rep(x, times = n_time),
      y = rep(y, times = n_time),
      z = as.vector(z),
      frame = rep(time_seq, each = length(x)),
      mesh = mesh_name
    )
  }
  
  # Build data for finer functions
  data_finer_list <- lapply(names(fs_finer), function(nm) {
    make_df(fs_finer[[nm]], graph_finer, x_finer, y_finer, nm)
  })
  
  # Build data for coarser functions
  data_coarser_list <- lapply(names(fs_coarser), function(nm) {
    make_df(fs_coarser[[nm]], graph_coarser, x_coarser, y_coarser, nm)
  })
  
  # Combine
  all_data <- c(data_finer_list, data_coarser_list)
  
  # Baseline graph (on finer mesh for consistency)
  data_graph <- data.frame(
    x = rep(x_finer, times = n_time),
    y = rep(y_finer, times = n_time),
    z = 0,
    frame = rep(time_seq, each = length(x_finer)),
    mesh = "Graph"
  )
  
# --------- Vertical lines helper ----------
vertical_lines <- function(x, y, z, frame_vals, mesh_name) {
  do.call(rbind, lapply(seq_along(frame_vals), function(i) {
    idx <- ((i - 1) * length(x) + 1):(i * length(x))
    data.frame(
      x = rep(x, each = 3),
      y = rep(y, each = 3),
      z = as.vector(t(cbind(0, z[idx], NA))),
      frame = rep(frame_vals[i], each = length(x) * 3),
      mesh = mesh_name
    )
  }))
}

# --------- Compute vertical lines per mesh using max absolute value ---------
make_vertical_from_list <- function(data_list, x, y, mesh_name) {
  if (length(data_list) == 0) return(NULL)
  
  # Reshape each function's z back to matrix: (nodes × time)
  z_mats <- lapply(data_list, function(df) {
    matrix(df$z, nrow = length(x), ncol = length(time_seq))
  })
  
  # Stack into 3D array: (nodes × time × functions)
  arr <- array(unlist(z_mats), dim = c(length(x), length(time_seq), length(z_mats)))
  
  # For each node × time, select the entry with largest absolute value (keep sign)
  idx <- apply(arr, c(1, 2), function(v) which.max(abs(v)))
  z_signed_max <- mapply(function(i, j) arr[i, j, idx[i, j]],
                         rep(1:length(x), times = length(time_seq)),
                         rep(1:length(time_seq), each = length(x)))
  
  # Flatten back into long vector
  z_signed_max <- as.vector(z_signed_max)
  
  vertical_lines(x, y, z_signed_max, time_seq, mesh_name)
}


vertical_finer   <- make_vertical_from_list(data_finer_list,   x_finer,   y_finer,   "finer")
vertical_coarser <- make_vertical_from_list(data_coarser_list, x_coarser, y_coarser, "coarser")

  
  # Compute ranges
  all_z <- unlist(lapply(all_data, function(df) df$z))
  x_range <- range(c(x_finer, x_coarser))
  y_range <- range(c(y_finer, y_coarser))
  z_range <- range(all_z)
  
  # --------- Plotly object ----------
  p <- plot_ly(frame = ~frame)
  
  # Add traces for finer + coarser (looping automatically with names)
  for (df in all_data) {
    p <- p %>%
      add_trace(data = df,
                x = ~x, y = ~y, z = ~z,
                type = "scatter3d", mode = "lines",
                line = list(width = 3),
                name = unique(df$mesh))
  }
  
  # Add baseline
  p <- p %>%
    add_trace(data = data_graph,
              x = ~x, y = ~y, z = ~z,
              type = "scatter3d", mode = "lines",
              line = list(color = "black", width = 2),
              name = "Graph", showlegend = FALSE)
  
# Add verticals (one per mesh, envelope of all functions)
if (!is.null(vertical_finer)) {
  p <- p %>%
    add_trace(data = vertical_finer,
              x = ~x, y = ~y, z = ~z,
              type = "scatter3d", mode = "lines",
              line = list(color = "gray", width = 0.5),
              name = "Vertical finer", showlegend = FALSE)
}
if (!is.null(vertical_coarser)) {
  p <- p %>%
    add_trace(data = vertical_coarser,
              x = ~x, y = ~y, z = ~z,
              type = "scatter3d", mode = "lines",
              line = list(color = "gray", width = 0.5),
              name = "Vertical coarser", showlegend = FALSE)
}

  
  frame_name <- deparse(substitute(frame_val_to_display))
  
  p <- p %>%
    layout(
      scene = global.scene.setter(x_range, y_range, z_range),
      updatemenus = list(list(
        type = "buttons", showactive = FALSE,
        buttons = list(
          list(label = "Play", method = "animate",
               args = list(NULL, list(frame = list(duration = 2000 / length(time_seq), redraw = TRUE),
                                      fromcurrent = TRUE))),
          list(label = "Pause", method = "animate",
               args = list(NULL, list(mode = "immediate", 
                                      frame = list(duration = 0), redraw = FALSE)))
        )
      )),
      title = paste0(frame_name, ": ", formatC(frame_val_to_display[1], format = "f", digits = 4))
    ) %>%
    plotly_build()
  
  # Update frame titles
  for (i in seq_along(p$x$frames)) {
    p$x$frames[[i]]$layout <- list(
      title = paste0(frame_name, ": ", formatC(frame_val_to_display[i], format = "f", digits = 4))
    )
  }
  
  return(p)
}

```

### Function `plot_3d_rectangle_scatter()`

```{r}
plot_3d_rectangle_scatter <- function(loc, eigvals, eigfuncs, fixed_colorscale = TRUE) {

  x <- loc[,1]
  y <- loc[,2]

  colorscale <- "Viridis"

  # Compute global color limits if fixed
  if (fixed_colorscale) {
    cmin <- min(eigfuncs)
    cmax <- max(eigfuncs)
  } else {
    cmin <- NULL
    cmax <- NULL
  }

  # Frames ----------------------------------------------------------------
  frames <- lapply(seq_len(ncol(eigfuncs)), function(i) {
    list(
      name = as.character(i),
      data = list(list(
        x = x,
        y = y,
        z = eigfuncs[,i],
        type = "scatter3d",
        mode = "markers",
        marker = list(
          size = 5,
          color = eigfuncs[,i],
          colorscale = colorscale,
          showscale = TRUE,
          cmin = cmin,
          cmax = cmax
        )
      ))
    )
  })

  # Initial Plot ----------------------------------------------------------
  p <- plot_ly(
    x = x,
    y = y,
    z = eigfuncs[,1],
    type = "scatter3d",
    mode = "markers",
    marker = list(
      size = 5,
      color = eigfuncs[,1],
      colorscale = colorscale,
      showscale = TRUE,
      cmin = cmin,
      cmax = cmax
    ),
    frame = "1"
  )

  p$x$frames <- frames

  z_range <- range(eigfuncs)
  x_range <- range(x)
  y_range <- range(y)

  frame_name <- deparse(substitute(eigvals))

  # Layout + slider -------------------------------------------------------
  p <- p %>% layout(
    title = paste0(frame_name, ": ", eigvals[1]),
    scene = global.scene.setter.rec.and.sphere(x_range, y_range, z_range),
    sliders = list(
      list(
        active = 0,
        currentvalue = list(prefix = "Frame: "),
        pad = list(t = 50),
        steps = lapply(seq_len(ncol(eigfuncs)), function(i) {
          list(
            label = as.character(i),
            method = "animate",
            args = list(
              list(as.character(i)),
              list(frame = list(duration = 300, redraw = TRUE),
                   mode = "immediate")
            )
          )
        })
      )
    ),
    updatemenus = list(
      list(
        type = "buttons",
        showactive = FALSE,
        y = 1,
        x = 1.15,
        xanchor = "right",
        yanchor = "top",
        buttons = list(
          list(label = "Play",
               method = "animate",
               args = list(
                 NULL,
                 list(frame = list(duration = 300, redraw = TRUE),
                      fromcurrent = TRUE)
               )),
          list(label = "Pause",
               method = "animate",
               args = list(NULL, list(frame = list(duration = 0))))
        )
      )
    )
  ) %>% plotly_build()

  # Update title in each frame --------------------------------------------
  for (i in seq_len(ncol(eigfuncs))) {
    p$x$frames[[i]]$layout <- list(
      title = paste0(frame_name, ": ", eigvals[i])
    )
  }

  return(p)
}

plot_3d_rectangle_onecol_vir <- function(mesh, fvals) {
  
  colorscale = "Viridis"
  opacity = 1
  
  
  x <- mesh$loc[, 1]
  y <- mesh$loc[, 2]
  tri <- mesh$graph$tv
  
  # fvals is a vector
  
  x_range <- range(x)
  y_range <- range(y)
  z_range <- range(fvals)
  
  # Static mesh3d plot
  p <- plot_ly(
    x = x, y = y, z = fvals,
    i = tri[,1] - 1,
    j = tri[,2] - 1,
    k = tri[,3] - 1,
    type = "mesh3d",
    intensity = fvals,
    colorscale = colorscale,
    opacity = opacity,
    flatshading = TRUE,
    text = paste0(
      "x: ", sprintf("%.3f", x), "<br>",
      "y: ", sprintf("%.3f", y), "<br>",
      "f: ", sprintf("%.5f", fvals)
    ),
    hoverinfo = "text"
  ) %>% layout(
    scene = global.scene.setter.rec.and.sphere(x_range, y_range, z_range)
  )
  
  return(p)
}

plot_3d_rectangle_onecol <- function(mesh, fvals) {
  
  opacity <- 1
  surface_color <- "blue"
  
  x <- mesh$loc[, 1]
  y <- mesh$loc[, 2]
  tri <- mesh$graph$tv
  
  x_range <- range(x)
  y_range <- range(y)
  z_range <- range(fvals)
  
  # Static mesh3d plot
  p <- plot_ly(
    x = x, y = y, z = fvals,
    i = tri[,1] - 1,
    j = tri[,2] - 1,
    k = tri[,3] - 1,
    type = "mesh3d",
    color = surface_color,     # Fixed blue color
    opacity = opacity,
    flatshading = TRUE,
    text = paste0(
      "x: ", sprintf("%.3f", x), "<br>",
      "y: ", sprintf("%.3f", y), "<br>",
      "f: ", sprintf("%.5f", fvals)
    ),
    hoverinfo = "text"
  ) %>% layout(
    scene = global.scene.setter.rec.and.sphere(x_range, y_range, z_range)
  )
  
  return(p)
}

plot_3d_square_mesh <- function(mesh) {
  
  x <- mesh$loc[,1]
  y <- mesh$loc[,2]
  # Flat square (or use mesh$loc[,3] if available)
  if (ncol(mesh$loc) >= 3) {
    z <- mesh$loc[,3]
  } else {
    z <- rep(0, length(x))
  }
  
  x_range <- range(x)
  y_range <- range(y)
  z_range <- c(0,1)
  
  tri <- mesh$graph$tv
  
  # ---- Uniform gray color for faces ----
  gray_rgb <- rep(list(rgb(180, 180, 180, maxColorValue = 255)), length(x))
  white_rgb <- rep(list(rgb(255, 255, 255, alpha = 255, maxColorValue = 255)), length(x))

  
  df3 <- data.frame(x = x, 
                  y = y, 
                  z = z)
  
  # ---- Plot mesh faces ----
  p <- plot_ly() %>% add_trace(
    x = x,
    y = y,
    z = z,
    i = tri[,1] - 1,
    j = tri[,2] - 1,
    k = tri[,3] - 1,
    type = "mesh3d",
    vertexcolor = white_rgb,
    flatshading = TRUE,
    opacity = 1,
    showscale = FALSE
  )
  
  # ---- Extract unique edges ----
  edges <- rbind(
    tri[,1:2],
    tri[,2:3],
    tri[,c(3,1)]
  )
  edges <- t(apply(edges, 1, sort))
  edges <- unique(edges)
  
  # ---- Efficient edges as one trace using NA breaks ----
  x_edges <- as.vector(t(cbind(x[edges[,1]], x[edges[,2]], NA)))
  y_edges <- as.vector(t(cbind(y[edges[,1]], y[edges[,2]], NA)))
  z_edges <- as.vector(t(cbind(z[edges[,1]], z[edges[,2]], NA)))
  
  p <- add_trace(
    p,
    x = x_edges, y = y_edges, z = z_edges,
    type = "scatter3d",
    mode = "lines",
    line = list(color = "black", width = 3),
    showlegend = FALSE,
    hoverinfo = "none"
  ) %>% 
    add_trace(data = df3, x = ~x, y = ~y, z = ~z, mode = "markers", type = "scatter3d", 
            marker = list(size = 4, color = "gray", symbol = 104)) %>%
    layout(
    scene = global.scene.setter.rec.and.sphere(x_range, y_range, z_range)
  )
  
  return(p)
}


plot_3d_square_mesh_with_hat <- function(mesh,
                                hat_nodes = NULL,
                                hat_height = 1,
                                hat_color = NULL,
                                hat_alpha = 0.6) {
  
  x <- mesh$loc[,1]
  y <- mesh$loc[,2]
  n <- length(x)
  
  z0 <- rep(0, n)
  tri <- mesh$graph$tv
  
  x_range <- range(x)
  y_range <- range(y)
  z_range <- c(0, hat_height)
  
  # ---- Extract unique edges once ----
  edges <- rbind(tri[,1:2], tri[,2:3], tri[,c(3,1)])
  edges <- t(apply(edges, 1, sort))
  edges <- unique(edges)
  
  # Base mesh color
  white_rgb <- rep(list(rgb(255, 255, 255, alpha = 255, maxColorValue = 255)), n)
  
  df3 <- data.frame(x = x, 
                  y = y, 
                  z = rep(0, length(x)))
  
  # ---- Base mesh ----
  p <- plot_ly() %>% add_trace(
    x = x, y = y, z = z0,
    i = tri[,1] - 1, j = tri[,2] - 1, k = tri[,3] - 1,
    type = "mesh3d",
    vertexcolor = white_rgb,
    flatshading = TRUE,
    opacity = 1,
    name = "Mesh"
  )
  
  # ---- Base mesh edges ----
  x_edges <- as.vector(t(cbind(x[edges[,1]], x[edges[,2]], NA)))
  y_edges <- as.vector(t(cbind(y[edges[,1]], y[edges[,2]], NA)))
  z_edges <- as.vector(t(cbind(z0[edges[,1]], z0[edges[,2]], NA)))
  
  # ---- Hat functions ----
  if (!is.null(hat_nodes)) {
    
    if (is.null(hat_color)) {
      hat_color <- grDevices::rainbow(length(hat_nodes))
    }
    if (length(hat_color) != length(hat_nodes)) {
      stop("length(hat_color) must equal length(hat_nodes)")
    }
    
    for (k in seq_along(hat_nodes)) {
      i0 <- hat_nodes[k]
      
      phi <- rep(0, n)
      phi[i0] <- 1
      z_hat <- hat_height * phi
      
      hat_rgb <- rep(list(adjustcolor(hat_color[k], alpha.f = hat_alpha)), n)
      
      # ---- Hat surface ----
      p <- add_trace(p,
        x = x, y = y, z = z_hat,
        i = tri[,1] - 1, j = tri[,2] - 1, k = tri[,3] - 1,
        type = "mesh3d",
        vertexcolor = hat_rgb,
        flatshading = TRUE,
        opacity = 1,
        name = paste0("phi_", i0),
        showlegend = FALSE
      )
      
      # ---- Hat edges (wireframe) ----
      x_hat_edges <- as.vector(t(cbind(x[edges[,1]], x[edges[,2]], NA)))
      y_hat_edges <- as.vector(t(cbind(y[edges[,1]], y[edges[,2]], NA)))
      z_hat_edges <- as.vector(t(cbind(z_hat[edges[,1]], z_hat[edges[,2]], NA)))
      
      p <- add_trace(p,
        x = x_hat_edges, y = y_hat_edges, z = z_hat_edges,
        type = "scatter3d", mode = "lines",
        line = list(color = hat_color[k], width = 2),
        showlegend = FALSE
      )
      
      # # Mark hat center node
      # p <- add_trace(p,
      #   x = x[i0], y = y[i0], z = z_hat[i0],
      #   type = "scatter3d", mode = "markers",
      #   marker = list(size = 8, color = hat_color[k]),
      #   name = paste0("node_", i0)
      # )
    }
  }
  
  p <- add_trace(p,
    x = x_edges, y = y_edges, z = z_edges,
    type = "scatter3d", mode = "lines",
    line = list(color = "black", width = 4),
    showlegend = FALSE
  )
  
  p <- p %>% 
    add_trace(data = df3, x = ~x, y = ~y, z = ~z, mode = "markers", type = "scatter3d", 
            marker = list(size = 4, color = "gray", symbol = 104), showlegend = FALSE) %>% 
    layout(
    scene = global.scene.setter.rec.and.sphere.for.hat(x_range, y_range, z_range)
  )
  
  return(p)
}


```



### Function `plot_3d_sphere_scatter()`

```{r}
plot_3d_sphere_scatter <- function(mesh, 
                                          eigvals, 
                                          eigfuncs,
                                          fixed_colorscale = TRUE) {
  
  colorscale = "Viridis"
  
  x <- mesh$loc[,1]
  y <- mesh$loc[,2]
  z <- mesh$loc[,3]
  
  # Compute global color limits if fixed
  if (fixed_colorscale) {
    cmin <- min(eigfuncs)
    cmax <- max(eigfuncs)
  } else {
    cmin <- NULL
    cmax <- NULL
  }
  
  # Create frames
  frames <- lapply(seq_len(ncol(eigfuncs)), function(i) {
    
    fvals <- eigfuncs[,i]
    
    list(
      name = as.character(i),
      data = list(list(
        x = x,
        y = y,
        z = z,
        type = "scatter3d",
        mode = "markers",
        marker = list(
          size = 5,
          color = fvals,
          colorscale = colorscale,
          showscale = TRUE,
          cmin = cmin,
          cmax = cmax
        ),
        text = paste0(
          "x: ", sprintf("%.3f", x), "<br>",
          "y: ", sprintf("%.3f", y), "<br>",
          "z: ", sprintf("%.3f", z), "<br>",
          "f: ", sprintf("%.5f", fvals)
        ),
        hoverinfo = "text"
      ))
    )
  })
  
  # Initial plot (frame 1)
  fvals0 <- eigfuncs[,1]
  
  p <- plot_ly(
    x = x, y = y, z = z,
    type = "scatter3d",
    mode = "markers",
    marker = list(
      size = 5,
      color = fvals0,
      colorscale = colorscale,
      showscale = TRUE,
      cmin = cmin,
      cmax = cmax
    ),
    text = paste0(
      "x: ", sprintf("%.3f", x), "<br>",
      "y: ", sprintf("%.3f", y), "<br>",
      "z: ", sprintf("%.3f", z), "<br>",
      "f: ", sprintf("%.5f", fvals0)
    ),
    hoverinfo = "text",
    frame = "1"
  )
  
  frame_name <- deparse(substitute(eigvals))
  
  p$x$frames <- frames
  
  x_range <- range(x)
  y_range <- range(y)
  z_range <- range(z)
  
  # Slider + play/pause
  p <- p %>% layout(
    title = paste0(frame_name, ": ", eigvals[1]),
    scene = global.scene.setter.rec.and.sphere(x_range, y_range, z_range),
    sliders = list(
      list(
        active = 0,
        currentvalue = list(prefix = "Mode: "),
        pad = list(t = 50),
        steps = lapply(seq_len(ncol(eigfuncs)), function(i) {
          list(
            label = as.character(i),
            method = "animate",
            args = list(list(as.character(i)),
                        list(mode = "immediate",
                             frame = list(duration = 300, redraw = TRUE),
                             transition = list(duration = 0)))
          )
        })
      )
    ),
    updatemenus = list(
      list(
        type = "buttons",
        showactive = FALSE,
        y = 1,
        x = 1.15,
        xanchor = "right",
        yanchor = "top",
        buttons = list(
          list(
            label = "Play",
            method = "animate",
            args = list(NULL,
                        list(frame = list(duration = 300, redraw = TRUE),
                             fromcurrent = TRUE,
                             mode = "immediate"))
          ),
          list(
            label = "Pause",
            method = "animate",
            args = list(NULL,
                        list(frame = list(duration = 0, redraw = FALSE),
                             mode = "immediate"))
          )
        )
      )
    )
  ) %>% plotly_build()
  
  # Update title per frame
  for (i in seq_len(ncol(eigfuncs))) {
    p$x$frames[[i]]$layout <- list(
      title = paste0(frame_name, ": ", eigvals[i])
    )
  }
  
  return(p)
}

plot_3d_slider_sphere <- function(mesh, eigvals, eigfuncs, fixed_colorscale = TRUE) {
  
  colorscale = "Viridis"
  opacity = 1
  
  x <- mesh$loc[, 1]
  y <- mesh$loc[, 2]
  z <- mesh$loc[, 3]
  tri <- mesh$graph$tv
  
  # Global intensity limits if fixed
  if (fixed_colorscale) {
    cmin <- min(eigfuncs)
    cmax <- max(eigfuncs)
  } else {
    cmin <- NULL
    cmax <- NULL
  }
  
  # Create frames
  frames <- lapply(seq_len(ncol(eigfuncs)), function(i) {
    fvals <- eigfuncs[,i]
    list(
      name = as.character(i),
      data = list(list(
        x = x,
        y = y,
        z = z,
        i = tri[,1] - 1,
        j = tri[,2] - 1,
        k = tri[,3] - 1,
        type = "mesh3d",
        intensity = fvals,
        colorscale = colorscale,
        opacity = opacity,
        flatshading = TRUE,
        cmin = cmin,
        cmax = cmax,
        text = paste0(
          "x: ", sprintf("%.3f", x), "<br>",
          "y: ", sprintf("%.3f", y), "<br>",
          "z: ", sprintf("%.3f", z), "<br>",
          "f: ", sprintf("%.5f", fvals)
        ),
        hoverinfo = "text"
      ))
    )
  })
  
  # Initial plot
  fvals0 <- eigfuncs[,1]
  
  p <- plot_ly(
    x = x, y = y, z = z,
    i = tri[,1] - 1,
    j = tri[,2] - 1,
    k = tri[,3] - 1,
    type = "mesh3d",
    intensity = fvals0,
    colorscale = colorscale,
    opacity = opacity,
    flatshading = TRUE,
    cmin = cmin,
    cmax = cmax,
    text = paste0(
      "x: ", sprintf("%.3f", x), "<br>",
      "y: ", sprintf("%.3f", y), "<br>",
      "z: ", sprintf("%.3f", z), "<br>",
      "f: ", sprintf("%.5f", fvals0)
    ),
    colorbar = list(
      thickness = 15,
      len = 0.5,      # fraction of the height of the plot
      y = 0.5,        # center vertically (0 = bottom, 1 = top)
      yanchor = "middle",
      title = ""
    ),
    hoverinfo = "text",
    frame = "1"
  )
  
  frame_name <- deparse(substitute(eigvals))
  
  p$x$frames <- frames
  
  # Layout + slider + buttons
  p <- p %>% layout(
    title = paste0(frame_name, ": ", eigvals[1]),
    scene = global.scene.setter.rec.and.sphere(range(x), range(y), range(z)),
    sliders = list(
      list(
        active = 0,
        currentvalue = list(prefix = "Frame: "),
        pad = list(t = 50),
        steps = lapply(seq_len(ncol(eigfuncs)), function(i) {
          list(
            label = as.character(i),
            method = "animate",
            args = list(list(as.character(i)),
                        list(mode = "immediate",
                             frame = list(duration = 300, redraw = TRUE),
                             transition = list(duration = 0)))
          )
        })
      )
    ),
    updatemenus = list(
      list(
        type = "buttons",
        showactive = FALSE,
        y = 1,
        x = 1.15,
        xanchor = "right",
        yanchor = "top",
        buttons = list(
          list(label = "Play",
               method = "animate",
               args = list(NULL, list(frame = list(duration = 300, redraw = TRUE),
                                      fromcurrent = TRUE, mode = "immediate"))),
          list(label = "Pause",
               method = "animate",
               args = list(NULL, list(frame = list(duration = 0, redraw = FALSE),
                                      mode = "immediate")))
        )
      )
    )
  ) %>% plotly_build()
  
  # Add titles for frames
  for (i in seq_len(ncol(eigfuncs))) {
    p$x$frames[[i]]$layout <- list(
      title = paste0(frame_name, ": ", eigvals[i])
    )
  }
  
  return(p)
}


plot_3d_sphere_scatter_onecol <- function(mesh, eigfuncs) {
  
  colorscale = "Viridis"
  
  x <- mesh$loc[,1]
  y <- mesh$loc[,2]
  z <- mesh$loc[,3]
  
  # eigfuncs is a vector
  fvals <- eigfuncs
  
  x_range <- range(x)
  y_range <- range(y)
  z_range <- range(z)
  
  # Build static 3D scatter plot
  p <- plot_ly(
    x = x, y = y, z = z,
    type = "scatter3d",
    mode = "markers",
    marker = list(
      size = 5,
      color = fvals,
      colorscale = colorscale,
      showscale = TRUE
    ),
    text = paste0(
      "x: ", sprintf("%.3f", x), "<br>",
      "y: ", sprintf("%.3f", y), "<br>",
      "z: ", sprintf("%.3f", z), "<br>",
      "f: ", sprintf("%.5f", fvals)
    ),
    hoverinfo = "text"
  ) %>% layout(
    scene = global.scene.setter.rec.and.sphere(x_range, y_range, z_range)
  )
  
  return(p)
}



# plot_3d_sphere_surface_onecol <- function(mesh, fvals) {
#   
#   colorscale = "Viridis"
#   opacity = 1
#   
#   
#   x <- mesh$loc[, 1]
#   y <- mesh$loc[, 2]
#   z <- mesh$loc[, 3]
#   tri <- mesh$graph$tv
#   
#   # fvals is a vector
#   
#   x_range <- range(x)
#   y_range <- range(y)
#   z_range <- range(z)
#   
#   # Static mesh3d plot
#   p <- plot_ly(
#     x = x, y = y, z = z,
#     i = tri[,1] - 1,
#     j = tri[,2] - 1,
#     k = tri[,3] - 1,
#     type = "mesh3d",
#     intensity = fvals,
#     colorscale = colorscale,
#     opacity = opacity,
#     flatshading = TRUE,
#     text = paste0(
#       "x: ", sprintf("%.3f", x), "<br>",
#       "y: ", sprintf("%.3f", y), "<br>",
#       "z: ", sprintf("%.3f", z), "<br>",
#       "f: ", sprintf("%.5f", fvals)
#     ),
#     hoverinfo = "text"
#   ) %>% layout(
#     scene = global.scene.setter.rec.and.sphere(x_range, y_range, z_range)
#   )
#   
#   return(p)
# }


plot_3d_sphere_surface_onecol <- function(mesh, fvals) {
  
  colorscale = "Viridis"
  opacity = 1
  
  x <- mesh$loc[, 1]
  y <- mesh$loc[, 2]
  z <- mesh$loc[, 3]
  tri <- mesh$graph$tv
  
  x_range <- range(x)
  y_range <- range(y)
  z_range <- range(z)
  
  # Static mesh3d plot
  p <- plot_ly(
    x = x, y = y, z = z,
    i = tri[,1] - 1,
    j = tri[,2] - 1,
    k = tri[,3] - 1,
    type = "mesh3d",
    intensity = fvals,
    colorscale = colorscale,
    opacity = opacity,
    flatshading = TRUE,
    text = paste0(
      "x: ", sprintf("%.3f", x), "<br>",
      "y: ", sprintf("%.3f", y), "<br>",
      "z: ", sprintf("%.3f", z), "<br>",
      "f: ", sprintf("%.5f", fvals)
    ),
    hoverinfo = "text",
    colorbar = list(
      thickness = 15,
      len = 0.5,      # fraction of the height of the plot
      y = 0.5,        # center vertically (0 = bottom, 1 = top)
      yanchor = "middle",
      title = ""
    )
  ) %>% layout(
    scene = global.scene.setter.rec.and.sphere(x_range, y_range, z_range)
  )
  
  return(p)
}


plot_3d_mesh_edges_faces_old <- function(mesh) {
  
  x <- mesh$loc[,1]
  y <- mesh$loc[,2]
  z <- mesh$loc[,3]
  tri <- mesh$graph$tv
  
  # ---- Uniform gray color for faces (vertexcolor) ----
  gray_rgb <- rep(list(rgb(180, 180, 180, maxColorValue = 255)), length(x))
  
  p <- plot_ly() %>% add_trace(
    x = x,
    y = y,
    z = z,
    i = tri[,1] - 1,
    j = tri[,2] - 1,
    k = tri[,3] - 1,
    type = "mesh3d",
    vertexcolor = gray_rgb,   # <- the correct way
    flatshading = TRUE,
    opacity = 1,
    showscale = FALSE
  )
  
  # ---- Extract unique edges ----
  edges <- rbind(
    tri[,1:2],
    tri[,2:3],
    tri[,c(3,1)]
  )
  
  edges <- t(apply(edges, 1, sort))
  edges <- unique(edges)
  
  # ---- Add edges as blue lines ----
  for (e in 1:nrow(edges)) {
    i1 <- edges[e,1]
    i2 <- edges[e,2]
    
    p <- p %>% add_trace(
      x = c(x[i1], x[i2]),
      y = c(y[i1], y[i2]),
      z = c(z[i1], z[i2]),
      type = "scatter3d",
      mode = "lines",
      line = list(color = "blue", width = 3),
      hoverinfo = "none",
      showlegend = FALSE
    )
  }
  
  return(p)
}

plot_3d_mesh_edges_faces <- function(mesh) {
  
  x <- mesh$loc[,1]
  y <- mesh$loc[,2]
  z <- mesh$loc[,3]
  tri <- mesh$graph$tv
  
  x_range <- range(x)
  y_range <- range(y)
  z_range <- range(z)
  
  # ---- Uniform gray color for faces ----
  gray_rgb <- rep(list(rgb(180, 180, 180, maxColorValue = 255)), length(x))
  white_rgb <- rep(list(rgb(255, 255, 255, alpha = 255, maxColorValue = 255)), length(x))
  
  df3 <- data.frame(x = x, 
                  y = y, 
                  z = z)
  
  # ---- Plot mesh faces ----
  p <- plot_ly() %>% add_trace(
    x = x,
    y = y,
    z = z,
    i = tri[,1] - 1,
    j = tri[,2] - 1,
    k = tri[,3] - 1,
    type = "mesh3d",
    vertexcolor = white_rgb,
    flatshading = TRUE,
    opacity = 1,
    showscale = FALSE
  )
  
  # ---- Extract unique edges ----
  edges <- rbind(
    tri[,1:2],
    tri[,2:3],
    tri[,c(3,1)]
  )
  edges <- t(apply(edges, 1, sort))
  edges <- unique(edges)
  
  # ---- Efficient edges as a single trace using NA breaks ----
  x_edges <- as.vector(t(cbind(x[edges[,1]], x[edges[,2]], NA)))
  y_edges <- as.vector(t(cbind(y[edges[,1]], y[edges[,2]], NA)))
  z_edges <- as.vector(t(cbind(z[edges[,1]], z[edges[,2]], NA)))
  
  p <- add_trace(
    p,
    x = x_edges, y = y_edges, z = z_edges,
    type = "scatter3d",
    mode = "lines",
    line = list(color = "black", width = 3),
    showlegend = FALSE,
    hoverinfo = "none"
  ) %>% 
    add_trace(data = df3, x = ~x, y = ~y, z = ~z, mode = "markers", type = "scatter3d", 
            marker = list(size = 4, color = "gray", symbol = 104)) %>%
    layout(
    scene = global.scene.setter.rec.and.sphere(x_range, y_range, z_range)
  )
  
  return(p)
}

plot_3d_mesh_edges_faces_with_hat <- function(mesh,
                                              hat_nodes = NULL,
                                              hat_height = 0.1,
                                              hat_color = NULL,
                                              hat_alpha = 0.6) {
  
  x <- mesh$loc[,1]
  y <- mesh$loc[,2]
  z <- mesh$loc[,3]
  n <- length(x)
  tri <- mesh$graph$tv
  
  x_range <- range(x)*2
  y_range <- range(y)*2
  z_range <- range(z)*2
  
  # ---- Extract unique edges once ----
  edges <- rbind(tri[,1:2], tri[,2:3], tri[,c(3,1)])
  edges <- t(apply(edges, 1, sort))
  edges <- unique(edges)
  
  # Base mesh colors
  white_rgb <- rep(list(rgb(255, 255, 255, alpha = 255, maxColorValue = 255)), n)
  
  df3 <- data.frame(x = x, y = y, z = z)
  
  # ---- Base spherical mesh ----
  p <- plot_ly() %>% add_trace(
    x = x, y = y, z = z,
    i = tri[,1] - 1, j = tri[,2] - 1, k = tri[,3] - 1,
    type = "mesh3d",
    vertexcolor = white_rgb,
    flatshading = TRUE,
    opacity = 1,
    showscale = FALSE,
    name = "Sphere mesh"
  )
  
  # ---- Base mesh edges (single efficient trace) ----
  x_edges <- as.vector(t(cbind(x[edges[,1]], x[edges[,2]], NA)))
  y_edges <- as.vector(t(cbind(y[edges[,1]], y[edges[,2]], NA)))
  z_edges <- as.vector(t(cbind(z[edges[,1]], z[edges[,2]], NA)))
  
 
  
  # ---- Hat functions on sphere ----
  if (!is.null(hat_nodes)) {
    
    # Validate indices
    if (any(hat_nodes < 1 | hat_nodes > n)) {
      stop("hat_nodes contains invalid node indices")
    }
    
    # Color recycling
    if (is.null(hat_color)) {
      hat_color <- grDevices::rainbow(length(hat_nodes))
    }
    hat_color <- rep(hat_color, length.out = length(hat_nodes))
    
    for (k in seq_along(hat_nodes)) {
      i0 <- hat_nodes[k]
      
      # Nodal hat basis
      phi <- rep(0, n)
      phi[i0] <- 1
      
      # Lift radially outward (important for sphere!)
      r <- sqrt(x^2 + y^2 + z^2)
      nx <- x / r
      ny <- y / r
      nz <- z / r
      
      x_hat <- x + hat_height * phi * nx
      y_hat <- y + hat_height * phi * ny
      z_hat <- z + hat_height * phi * nz
      
      # Hat surface color
      hat_rgb <- rep(list(adjustcolor(hat_color[k], alpha.f = hat_alpha)), n)
      
      # ---- Hat surface ----
      p <- add_trace(p,
        x = x_hat, y = y_hat, z = z_hat,
        i = tri[,1] - 1, j = tri[,2] - 1, k = tri[,3] - 1,
        type = "mesh3d",
        vertexcolor = hat_rgb,
        flatshading = TRUE,
        opacity = 1,
        name = paste0("phi_", i0),
        showlegend = FALSE
      )
      
      # ---- Hat wireframe edges ----
      x_hat_edges <- as.vector(t(cbind(x_hat[edges[,1]], x_hat[edges[,2]], NA)))
      y_hat_edges <- as.vector(t(cbind(y_hat[edges[,1]], y_hat[edges[,2]], NA)))
      z_hat_edges <- as.vector(t(cbind(z_hat[edges[,1]], z_hat[edges[,2]], NA)))
      
      p <- add_trace(p,
        x = x_hat_edges, y = y_hat_edges, z = z_hat_edges,
        type = "scatter3d", mode = "lines",
        line = list(color = hat_color[k], width = 2),
        showlegend = FALSE,
        hoverinfo = "none"
      )
    }
  }
  
   p <- add_trace(p,
    x = x_edges, y = y_edges, z = z_edges,
    type = "scatter3d", mode = "lines",
    line = list(color = "black", width = 3),
    showlegend = FALSE,
    hoverinfo = "none"
  )
   
  # ---- Nodes ----
  p <- add_trace(p,
    data = df3,
    x = ~x, y = ~y, z = ~z,
    mode = "markers",
    type = "scatter3d",
    marker = list(size = 4, color = "gray", symbol = 104),
    showlegend = FALSE
  )
  
  # ---- Scene ----
  p <- p %>% layout(
    scene = global.scene.setter.rec.and.sphere(x_range, y_range, z_range)
  )
  
  return(p)
}

```


```{r}
graph.plotter.3d.onecol <- function(graph, vec) {
  
  # Coordinates on the mesh
  x <- plotting.order(graph$mesh$V[, 1], graph)
  y <- plotting.order(graph$mesh$V[, 2], graph)
  z <- plotting.order(vec, graph)  # vector to plot
  
  # Axis ranges
  x_range <- range(x)
  y_range <- range(y)
  z_range <- range(z)
  z_range[1] <- z_range[1] - 10^-4
  
  # Vertical lines (vectorized)
  n <- length(x)
  x_vert <- rep(x, each = 3)
  y_vert <- rep(y, each = 3)
  z_vert <- as.vector(t(cbind(0, z, NA)))
  
  # Create plot
  p <- plot_ly() %>%
    
    # Main 3D curve over the graph
    add_trace(
      x = x, y = y, z = z,
      type = "scatter3d", mode = "lines",
      line = list(color = "blue", width = 3),
      showlegend = FALSE
    ) %>%
    # Graph base
    add_trace(
      x = x, y = y, z = z*0,
      type = "scatter3d", mode = "lines",
      line = list(color = "black", width = 3),
      showlegend = FALSE
    ) %>%
    
    # Vertical lines from base to curve
    add_trace(
      x = x_vert, y = y_vert, z = z_vert,
      type = "scatter3d", mode = "lines",
      line = list(color = "gray", width = 0.5),
      showlegend = FALSE
    ) %>%
    
    layout(
      scene = list(xaxis = list(title = "x", range = x_range),
              yaxis = list(title = "y", range = y_range),
              zaxis = list(
      title = "z",
      range = z_range,
      exponentformat = "e",
      tickformat = NULL#".4f"     # prevents SI prefixes like μ, m, k
    ),
              aspectratio = list(x = 2*(1+2/pi), 
                                 y = 2*(2/pi), 
                                 z = 1*(2/pi)),
              camera = list(eye = list(x = 5, 
                                       y = 3, 
                                       z = 3.5),
                            center = list(x = (1+2/pi)/2, 
                                          y = 0, 
                                          z = 0)))
    )
  
  return(p)
}

```


```{r}
myggsave <- function(plot,
                     width = 9.22,
                     height = 7.05,
                     use_mathpazo = TRUE) {
  
  dir_to_save <- here::here("data_files/tikzpic")
  obj_name <- deparse(substitute(plot))
  
  # --- Filename suffix ---
  suffix <- if (use_mathpazo) "_with_mathpazo" else ""
  
  tex_name <- file.path(dir_to_save, paste0(obj_name, suffix, ".tex"))
  pdf_name <- file.path(dir_to_save, paste0(obj_name, suffix, ".pdf"))
  png_name <- file.path(dir_to_save, paste0(obj_name, suffix, ".png"))
  
  # Create directory if needed
  if (!dir.exists(dir_to_save)) dir.create(dir_to_save, recursive = TRUE)

  # --- TikZ + LaTeX options ---
  tikz_packages <- c(
    "\\usepackage{tikz}",
    "\\usetikzlibrary{calc}",
    "\\usepackage[T1]{fontenc}"
  )
  
  if (use_mathpazo) {
    tikz_packages <- c(tikz_packages, "\\usepackage[sc]{mathpazo}")
  }

  options(
    tikzDocumentDeclaration = "\\documentclass[tikz,border=0pt]{standalone}",
    tikzLatexPackages = tikz_packages
  )

  # Save TikZ plot
  tikzDevice::tikz(tex_name, standAlone = TRUE, width = width, height = height)
  print(plot)
  dev.off()
  
  # Compile to PDF
  system(
    paste0("pdflatex -output-directory=", dir_to_save, " ", tex_name),
    ignore.stdout = TRUE,
    ignore.stderr = TRUE
  )
  
  # --- Convert PDF to PNG (ImageMagick, no trimming) ---
  cmd <- paste(
    "convert -density 400",
    shQuote(pdf_name),
    shQuote(png_name)
  )
  
  system(cmd, ignore.stdout = TRUE, ignore.stderr = TRUE)
  
  if (!file.exists(png_name)) {
    stop("PNG conversion failed. Likely ImageMagick PDF policy restriction.")
  }
  
  # --- Clean auxiliary files ---
  aux_ext <- c(".aux", ".log", ".tex")
  for (ext in aux_ext) {
    f <- file.path(dir_to_save, paste0(obj_name, suffix, ext))
    if (file.exists(f)) file.remove(f)
  }
  
  ras_files <- list.files(
    dir_to_save,
    pattern = paste0(obj_name, suffix, "_ras[0-9]+\\.png$"),
    full.names = TRUE
  )
  if (length(ras_files) > 0) file.remove(ras_files)
  
  # --- Copy PNG to presentation folder ---
  pres_dir <- path.expand("~/Desktop/leninPresentations/data_files")
  if (!dir.exists(pres_dir)) dir.create(pres_dir, recursive = TRUE)
  
  pres_png <- file.path(pres_dir, basename(png_name))
  file.copy(png_name, pres_png, overwrite = TRUE)
  
  # --- Output ---
  cat("PDF saved at:", pdf_name, "\n",
      "PNG saved at:", png_name, "\n",
      "PNG copied to:", pres_png, "\n")
}
```


```{r}
#function to save
save_dual_for_presentation <- function(obj) {
  obj_name <- deparse(substitute(obj))
  filename <- paste0(obj_name, ".RData")
  
  path1 <- here::here("data_files", filename)
  path2 <- file.path("~", "Desktop", "leninPresentations", "data_files", filename)
  
  e <- new.env()
  assign(obj_name, obj, envir = e)
  
  save(list = obj_name, file = path1, envir = e)
  save(list = obj_name, file = path2, envir = e)
  
  cat("Saved file in:\n",
      " -", path1, "\n",
      " -", path2, "\n")
}
```



```{r}
reticulate::use_python(
  "/home/rierasl/miniconda3/envs/phdenv/bin/python",
  required = TRUE
)

reticulate::py_config()
```


```{r}
save_plotly_figure_fixed <- function(fig,
                                     dpi = 600,
                                     scale = 2,
                                     viewer_change = 1) {
  
  folder <- here::here("data_files/plotlypdf")
  
  if (!dir.exists(folder)) {
    dir.create(folder, recursive = TRUE)
  }
  
  fig_name <- deparse(substitute(fig))
  
  file_name_pdf <- file.path(folder, paste0(fig_name, ".pdf"))
  file_name_png <- file.path(folder, paste0(fig_name, ".png"))
  
  ## ---- CAMERA (correct Plotly location) ----
  cam <- fig$x$layoutAttrs[[1]]$scene$camera
  
  if (!is.null(cam)) {
    
    eye <- cam$eye
    center <- cam$center
    
    if (is.null(eye)) eye <- list(x = 1, y = 1, z = 1)
    if (is.null(center)) center <- list(x = 0, y = 0, z = 0)
    
    # vector center -> eye
    vx <- eye$x - center$x
    vy <- eye$y - center$y
    vz <- eye$z - center$z
    
    # norm (distance from center)
    norm_v <- sqrt(vx^2 + vy^2 + vz^2)
    
    if (norm_v > 0) {
      
      # normalize direction
      vx <- vx / norm_v
      vy <- vy / norm_v
      vz <- vz / norm_v
      
      # apply zoom scaling on radius
      norm_v <- norm_v * viewer_change
      
      fig$x$layoutAttrs[[1]]$scene$camera$eye <- list(
        x = center$x + vx * norm_v,
        y = center$y + vy * norm_v,
        z = center$z + vz * norm_v
      )
    }
  }
  
  
  # Save PDF
  plotly::save_image(
    fig,
    file_name_pdf,
    width = NULL,
    height = NULL,
    scale = scale#,engine = "kaleido"
  )
  
  # Convert PDF to PNG
  cmd <- paste(
    "convert -density", dpi,
    "-background white -trim +repage",
    "-bordercolor white -border 10",
    shQuote(file_name_pdf),
    shQuote(file_name_png)
  )
  
  # use below in case
  # label Loading[MathJax]/extensions/MathMenu.js appears
  # cmd <- paste(
  # "convert -density", dpi,
  # "-background white -trim +repage",
  # "-gravity southwest -chop 0x200", #200 it is the heigth of the cropped area
  # "-trim +repage",
  # "-bordercolor white -border 10",
  # shQuote(file_name_pdf),
  # shQuote(file_name_png)
  # )
  
  system(cmd, ignore.stdout = TRUE, ignore.stderr = TRUE)
  
  # Copy to presentation folder
  pres_dir <- path.expand("~/Desktop/leninPresentations/data_files")
  if (!dir.exists(pres_dir)) dir.create(pres_dir, recursive = TRUE)
  
  pres_png <- file.path(pres_dir, basename(file_name_png))
  pres_pdf <- file.path(pres_dir, basename(file_name_pdf))
  
  file.copy(file_name_png, pres_png, overwrite = TRUE)
  file.copy(file_name_pdf, pres_pdf, overwrite = TRUE)
  
  cat(
    "Files saved:\n",
    "  PDF:", file_name_pdf, "\n",
    "  PNG:", file_name_png, "\n",
    "  PDF copied to:", pres_pdf, "\n",
    "  PNG copied to:", pres_png, "\n",
    "  viewer change:", viewer_change, "\n"
  )
}

```


## References

```{r, purl = FALSE}
grateful::cite_packages(output = "paragraph", out.dir = ".")
```


