---
title: "Preliminaries"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

Go back to the [Contents](about.html) page.

<div style="color: #2c3e50; text-align: right;">
********  
<strong>Press Show to reveal the code chunks.</strong>  

********
</div>


```{r}
# Create a clipboard button on the rendered HTML page
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```

```{r}
library(MetricGraph)
library(ggplot2)
library(reshape2)
library(plotly)
```


```{r}
capture.output(
  knitr::purl(here::here("functionality.Rmd"), output = here::here("functionality.R")),
  file = here::here("purl_log.txt")
)
source(here::here("functionality.R"))
```


## Finite element basis functions on a metric graph {#fem-basis}

Let each edge $e\in\Ecal$ be subdivided into $n_{e}\geq 2$ regular segments of length $h_{e}$, and be delimited by the nodes $0 = x_0^{e},x_1^{e},\dots,x_{n_{e}-1}^{e}, x_{n_{e}}^{e} = \ell_{e}$. For each $j = 1,\dots,n_{e}-1$, we consider the following standard hat basis functions 
\begin{equation*}
    \varphi_j^{e}(x)=\begin{cases}
        1-\dfrac{|x_j^{e}-x|}{h_{e}},&\text{ if }x_{j-1}^{e}\leq x\leq x_{j+1}^{e},\\
        0,&\text{ otherwise}.
    \end{cases}
\end{equation*}
For each $e\in\Ecal$, the set of hat functions $\llav{\varphi_1^{e},\dots,\varphi_{n_{e}-1}^{e}}$ is a basis for the space
\begin{equation*}
    V_{h_{e}} = \llav{w\in H_0^1(e)\;\Big|\;\forall j = 0,1,\dots,n_{e}-1:w|_{[x_j^{e}, x_{j+1}^{e}]}\in\mathbb{P}^1},
\end{equation*}
where $\mathbb{P}^1$ is the space of linear functions on $[0,\ell_{e}]$. For each vertex $v\in\Vcal$, we define
\begin{equation*}
    \Ncal_v = \llav{\bigcup_{e\in\llav{e\in\Ecal_v: v = x_0^e}}[v,x_1^e]}\bigcup\llav{\bigcup_{e\in\llav{e\in\Ecal_v: v = x^e_{n_e}}}[x^e_{n_e-1},v]},
\end{equation*}
which is a star-shaped set with center at $v$ and rays made of the segments contiguous to $v$. On $\mathcal{N}_v$, we define the hat functions as
\begin{equation*}
    \phi_v(x)=\begin{cases}
        1-\dfrac{|x_v^{e}-x|}{h_{e}},&\text{ if }x\in\mathcal{N}_v\cap e \text{ and }e\in\Ecal_v,\\
        0,&\text{ otherwise},
    \end{cases}
\end{equation*}
where $x_v^e$ is either $x_0^e$ or $x_{n_e}^e$ depending on the edge direction and its parameterization. See @arioli2018finite for more details. Figure 1 shows an illustration of the basis function system $\{\varphi_j^e, \phi_v\}$ on the tadpole graph (in black). Standard hat functions associated with internal edge nodes are shown in blue, while special vertex-centered functions are highlighted in red.


For an additional illustration, go to the [Basis](basis.html#basisf) page.

```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Illustration of the system of basis functions $\\{\\varphi_j^e, \\phi_v\\}$ on the tadpole graph. Notice that the sets $\\Ncal_{v}$ are depicted in green and their corresponding basis functions are shown in red.")}
graph <- gets.graph.tadpole(h = 1/4)
graph_to_get_loc <- gets.graph.tadpole(h = 1/40)
loc <- graph_to_get_loc$get_mesh_locations()


A <- as.matrix(graph$fem_basis(loc))
A <- apply(A, 2, function(x) plotting.order(x, graph_to_get_loc))
A <- rbind(A, rep(NA, ncol(A))) # Add a row of NAs for the plotting

x <- graph_to_get_loc$mesh$V[, 1]
y <- graph_to_get_loc$mesh$V[, 2]
x <- c(plotting.order(x, graph_to_get_loc), NA)
y <- c(plotting.order(y, graph_to_get_loc), NA)

x_range <- range(x, na.rm = TRUE)
y_range <- range(y, na.rm = TRUE)
z_range <- c(0,1)

# Start plot
p <- plot_ly() |> 
  add_trace(x = rep(x, times = graph$nV), 
            y = rep(y, times = graph$nV), 
            z = as.vector(A[, 1:graph$nV]), 
            type = "scatter3d", 
            mode = "lines",
            line = list(color = "red", width = 3), 
            showlegend = FALSE) |>
  add_trace(x = rep(x, times = ncol(A) - graph$nV), 
            y = rep(y, times = ncol(A) - graph$nV), 
            z = as.vector(A[, (graph$nV+1):ncol(A)]), 
            type = "scatter3d",
            mode = "lines", 
            line = list(color = "blue", width = 3), 
            showlegend = FALSE) |>
  add_trace(x = rep(x, each = 3), 
            y = rep(y, each = 3), 
            z = unlist(lapply(apply(A, 1, max, na.rm = TRUE), function(zj) c(0, zj, NA))),
            type = "scatter3d", 
            mode = "lines",
            line = list(color = "gray", width = 0.5),
            showlegend = FALSE) |>
  add_trace(x = x, 
            y = y, 
            z = x*0, 
            type = "scatter3d",
            mode = "lines",  
            line = list(color = "black", width = 3),
            showlegend = FALSE) |>
  add_trace(x = rep(x, times = graph$nV), 
            y = rep(y, times = graph$nV), 
            z = c(replace(rep(NA, nrow(A)), 1:11, 0), 
                  replace(rep(NA, nrow(A)), c(31:51, 111:121), 0)), 
            type = "scatter3d", 
            mode = "lines",
            line = list(color = "green", width = 3), 
            showlegend = FALSE) |>
  layout(scene = global.scene.setter(x_range, y_range, z_range, z_aspectratio = 1))
p
```


Having introduced the system of basis functions $\{\varphi_j^e, \phi_v\}$, which will be referred to as $\{\psi_j\}_{j=1}^{N_h}$, we can now define the finite element space $V_h\subset H^1(\Gamma)$ as $V_h = \pare{\bigoplus_{e\in\Ecal} V_{h_e}}\bigoplus V_v$, where $V_v = \text{span}\pare{\{\phi_v:v\in\Vcal\}}$ and $\dim\pare{V_h}$ is given by $N_h = |\Vcal| + \sum_{e\in\Ecal}n_e$.

## Eigenfunctions and eigenvalues on the tadpole graph {#eigenfunctions}

Let $\Gamma_T = (\Vcal,\Ecal)$ characterize the tadpole graph with $\Vcal = \{v_1,v_2\}$ and $\Ecal = \{e_1,e_2\}$. The left edge $e_1$ has length 1 and the circular edge $e_2$ has length 2. A point on $e_1$ is parameterized via $s=\left(e_1, t\right)$ for $t \in[0,1]$ and a point on $e_2$ via $s=\left(e_2, t\right)$ for $t\in[0,2]$. One can verify that $-\Delta_\Gamma$ has eigenvalues $0,\left\{(i \pi / 2)^2\right\}_{i \in \mathbb{N}}$ and $\left\{(i \pi / 2)^2\right\}_{2 i \in \mathbb{N}}$ with corresponding eigenfunctions $\phi_0$, $\left\{\phi_i\right\}_{i \in \mathbb{N}}$, and $\left\{\psi_i\right\}_{2 i \in \mathbb{N}}$ given by $\phi_0(s)=1 / \sqrt{3}$ and 
\begin{equation*}
    \phi_i(s)=C_{\phi, i}\begin{cases}
        -2 \sin \left(\dfrac{i\pi}{2}\right) \cos \left(\dfrac{i \pi s}{2}\right), & s \in e_1, \\
\sin \left(\dfrac{i \pi s}{2}\right), & s \in e_2,
    \end{cases},
\quad 
    \psi_i(s)=\dfrac{\sqrt{3}}{\sqrt{2}} \begin{cases}
    (-1)^{i / 2} \cos \left(\dfrac{i \pi s}{2}\right), & s \in e_1, \\
\cos \left(\dfrac{i \pi s}{2}\right), & s \in e_2,
\end{cases},
\end{equation*}
where $C_{\phi, i}=1$ if $i$ is even and $C_{\phi, i}=1 / \sqrt{3}$ otherwise. Moreover, these functions form an orthonormal basis for $L_2(\Gamma_T)$.


```{r}
graph <- gets.graph.tadpole(h = 0.02)
eigenpairs <- gets.eigen.params(N_finite = 10, kappa = 1, alpha = 0.5, graph)
EIGENFUN <- eigenpairs$EIGENFUN
EIGENVAL <- eigenpairs$EIGENVAL
INDEX <- eigenpairs$INDEX
```


```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Illustration of the eigenfunctions on the tadpole graph.")}
graph.plotter.3d(graph, INDEX, EIGENVAL, EIGENFUN)
```


## Projection onto a fine space-time mesh {#piecewise_projection}

### Temporal piecewise projection

The piecewise constant projection $U_{\text{piecewise}}(t^*_\ell)$ of the approximated values $U_{\text{approx}}(t_k)$, defined on a coarse time grid $\{t_k\}_{k=0}^{N}$, onto a finer grid $\{t^*_\ell\}_{\ell=0}^{M}$, is given by

\begin{equation}
U_{\text{piecewise}}(t^*_\ell) =
\begin{cases}
U_{\text{approx}}(t_0), & \text{if } t^*_\ell = 0 \\\\
U_{\text{approx}}(t_k), & \text{if } t^*_\ell \in (t_{k-1}, t_k], \quad \text{for } k = 1, \dots, N.
\end{cases}
\end{equation}

This defines a function that is constant on each interval $(t_{k-1}, t_k]$, and takes the value of the approximation at the **right endpoint** $t_k$ of the interval.

See the [Fuctionality](functionality.html#construct_piecewise_projection) page for the implementation of the function `construct_piecewise_projection()` that performs this operation.

### Spatial projection

A function $U_{\text{coarse}}(s)$ defined on a coarse mesh with $N_{h}$ nodes can be projected onto a fine mesh with $N_{h_{\text{ok}}}$ nodes by doing $U_{\text{fine}}(s) = \boldsymbol{\Psi} U_{\text{coarse}}(s)$, where $\boldsymbol{\Psi}$ is a matrix with entries $\boldsymbol{\Psi}_{ij}=\psi^j_h(s_i)$ for $j=1\dots, N_h$ and $i = 1,\dots N_{h_{\text{ok}}}$.

```{r}
# Parameters
T_final <- 2
kappa <- 15
N_finite = 4 # choose even
adjusted_N_finite <- N_finite + N_finite/2 + 1
# Coefficients for u_0 and f
coeff_U_0 <- 50*(1:adjusted_N_finite)^-1
coeff_U_0[-5] <- 0
coeff_FF <- rep(0, adjusted_N_finite)
coeff_FF[7] <- 10

AAA = 1
OMEGA = pi

# Overkill parameters
overkill_time_step <- 0.02
overkill_h <- 0.02

# Finest time and space mesh
overkill_time_seq <- seq(0, T_final, length.out = ((T_final - 0) / overkill_time_step + 1))
overkill_graph <- gets.graph.tadpole(h = overkill_h)

# Compute the weights on the finest mesh
overkill_graph$compute_fem() # This is needed to compute the weights
overkill_weights <- overkill_graph$mesh$weights


alpha <- 0.5 # from 0.5 to 2
beta <- alpha / 2

# Compute the eigenvalues and eigenfunctions on the finest mesh
overkill_eigen_params <- gets.eigen.params(N_finite = N_finite, kappa = kappa, alpha = alpha, graph = overkill_graph)
EIGENVAL_ALPHA <- overkill_eigen_params$EIGENVAL_ALPHA # Eigenvalues (they are independent of the meshes)
overkill_EIGENFUN <- overkill_eigen_params$EIGENFUN # Eigenfunctions on the finest mesh

# Compute the true solution on the finest mesh
overkill_U_true <- overkill_EIGENFUN %*% 
  outer(1:length(coeff_U_0), 
        1:length(overkill_time_seq), 
        function(i, j) (coeff_U_0[i] + coeff_FF[i] * G_sin(t = overkill_time_seq[j], A = AAA, lambda_j_alpha_half = EIGENVAL_ALPHA[i], omega = OMEGA)) * exp(-EIGENVAL_ALPHA[i] * overkill_time_seq[j]))


h <- 0.2
time_step <- 0.2
m <- 1
graph <- gets.graph.tadpole(h = h)
graph$compute_fem()
G <- graph$mesh$G
C <- graph$mesh$C
L <- kappa^2*C + G
eigen_params <- gets.eigen.params(N_finite = N_finite, kappa = kappa, alpha = alpha, graph = graph)
EIGENFUN <- eigen_params$EIGENFUN
U_0 <- EIGENFUN %*% coeff_U_0 # Compute U_0 on the current mesh
A <- graph$fem_basis(overkill_graph$get_mesh_locations())

time_seq <- seq(0, T_final, length.out = ((T_final - 0) / time_step + 1))
my_op_frac <- my.fractional.operators.frac(L, beta, C, scale.factor = kappa^2, m = m, time_step)
INT_BASIS_EIGEN <- t(overkill_EIGENFUN) %*% overkill_graph$mesh$C %*% A
# Compute matrix F with columns F^k
FF_approx <- t(INT_BASIS_EIGEN) %*% 
  outer(1:length(coeff_FF), 
        1:length(time_seq), 
    function(i, j) coeff_FF[i] * g_sin(r = time_seq[j], A = AAA, omega = OMEGA))

U_approx <- matrix(NA, nrow = nrow(C), ncol = length(time_seq))
U_approx[, 1] <- U_0
for (k in 1:(length(time_seq) - 1)) {
  U_approx[, k + 1] <- as.matrix(my.solver.frac(my_op_frac, my_op_frac$C %*% U_approx[, k] + time_step * FF_approx[, k + 1]))
}

projected_U_approx <- A %*% U_approx
projected_U_piecewise <- construct_piecewise_projection(projected_U_approx, time_seq, overkill_time_seq)
```


```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("Illustration of the projection of the approximated solution $U_{\\text{approx}}$ (called `projected_U_piecewise`) onto a fine space-time mesh. `overkill_U_true` is the true solution $U_{\\text{true}}$.")}
graph.plotter.3d(overkill_graph, overkill_time_seq, overkill_time_seq, overkill_U_true, projected_U_piecewise)
```

```{r, fig.height = 6, out.width = "100%", fig.cap = captioner("`U_approx` is the coarse approximated solution on the coarse mesh, `projected_U_approx` is the coarse solution on the fine mesh, and `overkill_U_true` is the true solution on the fine mesh. We use the coarse time sequence.")}
start_idx <- which.min(abs(overkill_time_seq - 0))
end_idx <- which.min(abs(overkill_time_seq - T_final))
idx <- seq(start_idx, end_idx, by = 10)



graph.plotter.3d.two.meshes.time(overkill_graph, graph, time_seq, time_seq, fs_finer = list(projected_U_approx = projected_U_approx, overkill_U_true = overkill_U_true[,idx]), fs_coarser = list(U_approx = U_approx))
```

## References

```{r}
grateful::cite_packages(output = "paragraph", out.dir = ".")
```


