---
title: "Numerical analysis and optimal control of fractional diffusion equations on metric graphs"
subtitle: ""
author: "Lenin Rafael Riera Segura \n Joint work with David Bolin and Alexandre B. Simas"
institute: "KAUST"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: '16:9'
      slideNumberFormat: '%current%'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      #countdown: 60000
      navigation:
        scroll: true # disable slide transitions by scrolling
    css: 
      - default
      - xaringan-themer.css
      - lenin.css
mathjax: "https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.5/MathJax.js?config=TeX-MML-AM_CHTML"
---

```{r, load_refs, include=FALSE, cache=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("./references.bib", check = FALSE)
```

```{r setup, echo = FALSE}
options(htmltools.dir.version = FALSE)
library(xaringanthemer)
style_mono_accent(base_color = "#0D2244")
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
knitr::opts_chunk$set(
  message = FALSE,    
  warning = FALSE,    
  echo = FALSE,        
  include = TRUE,     
  eval = TRUE,       
  cache = FALSE,       
  fig.align = "center",
  out.width = "90%",
  dpi = 300,
  error = TRUE,
  collapse = TRUE
)
```


class: center, middle

<div style="position:absolute; top:40px; right:40px;">
  <img src="kaustlogo.svg" width="120">
</div>

# Part I


#### State equation

---

class: left, top

<div style="position:absolute; top:40px; right:40px;">
  <img src="kaustlogo.svg" width="120">
</div>

## Problem setting

*********

We analyze and numerically approximate solutions to fractional diffusion equations on metric graphs of the form

.content-box-gray[
$$
\begin{cases}
  \partial_t u(s,t) + (\kappa^2 - \Delta_\Gamma)^{\alpha/2} u(s,t) & = f(s,t), \quad (s,t) \in \Gamma \times (0, T), \\
  u(s,0) & = u_0(s), \quad s \in \Gamma,
\end{cases}
\tag{1}
$$
]

with $u(\cdot,t)$ satisfying the Kirchhoff vertex conditions

\begin{equation}
\label{eq:Kcondfirstpart}
\tag{2}
   \mathcal{K} =  \left\{\phi\in C(\Gamma)\;\middle|\; \forall v\in \mathcal{V}:\;  \sum_{e\in\mathcal{E}_v}\partial_e \phi(v)=0 \right\}.
\end{equation}

Here $\Gamma = (\mathcal{V},\mathcal{E})$ is a compact metric graph, $\kappa>0$, $\alpha\in(0,2]$ determines the smoothness of $u(\cdot,t)$, $\Delta_{\Gamma}$ is the so-called Kirchhoff--Laplacian.

---

class: left, top

<div style="position:absolute; top:40px; right:40px;">
  <img src="kaustlogo.svg" width="120">
</div>

## Problem setting

*********

The fractional operator $L^{\alpha/2} : D(L^{\alpha/2}) \longmapsto L_2(\Gamma)$ is defined via

\begin{equation}
\label{lfractional}
         \phi \longmapsto L^{\alpha/2}\phi = \sum_{j\in\mathbb{N}}\lambda_j^{\alpha/2}(\phi, e_j)_{L_2(\Gamma)}e_j,
\end{equation}

and $-L^{\alpha/2}$ generates

\begin{align}
\label{spectral_semigroup}
    T(t)\phi = \sum_{j\in\mathbb{N}}e^{-\lambda^{\alpha/2}_jt}\left(\phi, e_j\right)_{L_2(\Gamma)}e_j,
\end{align}

---

class: left, top

<div style="position:absolute; top:40px; right:40px;">
  <img src="kaustlogo.svg" width="120">
</div>

## Exsistence and uniqueness

*********

We interpret $(1)$ as an abstract Cauchy problem in $L_2(\Gamma)$, given by

.content-box-gray[
$$
\begin{cases}
    \partial_tu(t)+L^{\alpha/2}u(t)=f(t),\quad t\in (0, T),\\  
    u(0)=u_0\in D(L^{\alpha/2}).
\end{cases}
\tag{3}
$$
]


<div class="content-box-myblue">
Theorem 1
</div>

.content-box-gray[
If $f\in L_2((0,T);L_2(\Gamma))$  and $u_0\in \dot{H}^{\alpha}(\Gamma)$, then $(3)$ has a unique mild solution given by the variation of parameters formula
\begin{equation}
    \label{eq:mildsolution}
    u(t) = T(t)u_0 + \int_0^t T(t-r)f(r)dr,\quad t\geq 0.
\end{equation}
]

---

layout: false

These slides are based on the work of `r Cite(myBib, "xaringanpackage")`.
## References

```{r refs, echo=FALSE, results="asis"}
PrintBibliography(myBib, .opts = list(bib.style = "alphabetic"))
```


