---
title: "Numerical Analysis and Optimal Control of fractional diffusion equations on metric graphs"
subtitle: ""
author: "Lenin Rafael Riera Segura \n Joint work with David Bolin and Alexandre B. Simas"
institute: "KAUST"
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      ratio: '16:9'
      slideNumberFormat: '%current%'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      #countdown: 60000
      navigation:
        scroll: true # disable slide transitions by scrolling
    css: 
      - default
      - xaringan-themer.css
      - lenin.css
mathjax: "https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.5/MathJax.js?config=TeX-MML-AM_CHTML"
---
```{r, load_refs, include=FALSE, cache=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("./references.bib", check = FALSE)
```

```{r setup, echo = FALSE}
options(htmltools.dir.version = FALSE)
library(xaringanthemer)
style_mono_accent(base_color = "#005EB8")
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa-solid fa-clipboard\" style=\"color: #00008B\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
knitr::opts_chunk$set(
  message = FALSE,    
  warning = FALSE,    
  echo = FALSE,        
  include = TRUE,     
  eval = TRUE,       
  cache = FALSE,       
  fig.align = "center",
  out.width = "90%",
  dpi = 300,
  error = TRUE,
  collapse = TRUE
)
```

class: left, top

## Introduction | Neural Networks | 1

Neural Networks (NNs) are universal function approximators, meaning that given a set of input-output pairs $(x,y)$, where $y= f(x)$, a NN $N$ can approximate $f$ with arbitrary precision.  

---

## Introduction | Neural Networks | 2

.pull-left35[
```{r plot-first, out.width="100%", echo = FALSE}
library(neuralnet)
x <- seq(-1, 1, by = 0.05)
y <- x^2 + rnorm(length(x), mean = 0, sd = 0.01)
data <- data.frame(x, y)
model <- neuralnet(y ~ x, data = data, hidden = c(5)) #<<
new_data <- data.frame(x = seq(-1, 1, by = 0.01))
predictions <- compute(model, new_data)$net.result
par(family = "Palatino", cex = 1.5)
plot(x, x^2, type = "l", lwd = 2, col = "blue", 
     ylim = c(0, 1.2), xlab = bquote(x), ylab = bquote(y), 
     main = bquote("Aproximation of"~f(x) == x^2))
lines(new_data$x, predictions, col = "red", lwd = 2, lty = "dotted")
legend("topright", legend = c("f", "N"), col = c("blue", "red"), lwd = 2)
```
]
--
.pull-right65[
```{r ref.label = 'plot-first', eval = FALSE, echo = TRUE}
```
]

This is a 1 hidden layer NN with 5 neurons and sigmoid activation function.

---

## Introduction | Neural Networks | 3


---

layout: false

These slides are based on the work of `r Cite(myBib, "xaringanpackage")`.
## References

```{r refs, echo=FALSE, results="asis"}
PrintBibliography(myBib, .opts = list(bib.style = "alphabetic"))
```


