---
title: "Convergence in 𝑚, rectangle case"
date: "Last modified: `r format(Sys.time(), '%d-%m-%Y.')`"
output:
  html_document:
    mathjax: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    highlight: pygments
    theme: flatly
    code_folding: hide # class.source = "fold-hide" to hide code and add a button to show it
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    fig_caption: true
    code_download: true
    css: visual.css
always_allow_html: true
bibliography: 
  - references.bib
  - grateful-refs.bib
header-includes:
  - \newcommand{\ar}{\mathbb{R}}
  - \newcommand{\llav}[1]{\left\{#1\right\}}
  - \newcommand{\pare}[1]{\left(#1\right)}
  - \newcommand{\Ncal}{\mathcal{N}}
  - \newcommand{\Vcal}{\mathcal{V}}
  - \newcommand{\Ecal}{\mathcal{E}}
  - \newcommand{\Wcal}{\mathcal{W}}
---

Go back to the [Contents](about.html) page.

<div style="color: #2c3e50; text-align: right;">
********  
<strong>Press Show to reveal the code chunks.</strong>  

********
</div>


Let us set some global options for all code chunks in this document.


```{r}
# Create a clipboard button
source(here::here("clipboard.R")); clipboard
# Set seed for reproducibility
set.seed(1982) 
# Set global options for all code chunks
knitr::opts_chunk$set(
  # Disable messages printed by R code chunks
  message = FALSE,    
  # Disable warnings printed by R code chunks
  warning = FALSE,    
  # Show R code within code chunks in output
  echo = TRUE,        
  # Include both R code and its results in output
  include = TRUE,     
  # Evaluate R code chunks
  eval = TRUE,       
  # Enable caching of R code chunks for faster rendering
  cache = FALSE,      
  # Align figures in the center of the output
  fig.align = "center",
  # Enable retina display for high-resolution figures
  retina = 2,
  # Show errors in the output instead of stopping rendering
  error = TRUE,
  # Do not collapse code and output into a single block
  collapse = FALSE
)
# Start the figure counter
fig_count <- 0
# Define the captioner function
captioner <- function(caption) {
  fig_count <<- fig_count + 1
  paste0("Figure ", fig_count, ": ", caption)
}
```


::: {.custom-box}
To observe convergence in $m$, we calibrate $h$ and $\tau$ to $m$ as  $h \sim e^{-\pi\sqrt{(1-\alpha/2)m}}$ and then $\tau \sim h^{\alpha}$, respectively. With this scaling, the convergence with respect to the rational order $m$ is exponential. That is, $\|u-U_{h,m}^\tau\|_{L_2((0,T);L_2(\Gamma))} \leq C_me^{-\alpha\pi\sqrt{(1-\alpha/2)m}}$. This can be verified by estimating the slope $S_m$ in the regression $\log_e E = S_m\sqrt{m} +\log_e C_m$, where we expect $S_m\sim -\alpha\pi\sqrt{(1-\alpha/2)}$.
:::


```{r}
# remotes::install_github("davidbolin/rspde", ref = "devel")
# remotes::install_github("davidbolin/metricgraph", ref = "devel")
library(rSPDE)
library(MetricGraph)
library(fmesher)
library(grateful)
library(Matrix)

library(ggplot2)
library(reshape2)
library(plotly)
library(slackr)
source("keys.R")
slackr_setup(token = token) # token comes from keys.R
```



```{r}
capture.output(
  knitr::purl(here::here("functionality.Rmd"), output = here::here("functionality.R")),
  file = here::here("purl_log.txt")
)
source(here::here("functionality.R"))
```



```{r}
# Parameters
T_final <- 2
kappa <- 4
divider <- 1

a <- b <- 1

m_max <- 2
n_max <- 2
N_eig <- (m_max + 1) * (n_max + 1)
coeff_U_0 <- rep(0, N_eig)
coeff_U_0[9] <- 1
coeff_FF <- rep(0, N_eig)
coeff_FF[2] <- 10

AAA = 1
OMEGA = pi

# Time step and mesh size
m_vector <- c(1, 2, 3, 4)
alpha_vector <- seq(1.6, 1.8, by = 0.2)
```

```{r, eval = FALSE, echo = FALSE}
# Enforce nestedness
for (j in 1:length(alpha_vector)) {
  alpha <- alpha_vector[j]
  for (i in 1:length(m_vector)) {
    m <- m_vector[i]
    h <- exp(- pi * sqrt((1 - alpha / 2) * m))/divider 
    n <- ceiling(-1+sqrt(a*b)/h)
    time_step <- (h^alpha)
    print(n)
    print(time_step)
  }}

```


```{r}
# Overkill parameters
overkill_time_step <- 0.01
overkill_h <- sqrt(a*b)* 0.05
n_overkill <- ceiling(-1+sqrt(a*b)/overkill_h)
# Finest time and space mesh
overkill_time_seq <- seq(0, T_final, length.out = ((T_final - 0) / overkill_time_step + 1))

mesh_and_FEM_overkill <- gets.mesh.and.FEM.on.rectangle(a, b, n_overkill)
overkill_mesh_loc <- mesh_and_FEM_overkill$mesh$loc
overkill_C <- mesh_and_FEM_overkill$C
```


```{r, eval = FALSE, class.source = "fold-show"}
# Create a matrix to store the errors
errors_projected <- matrix(NA, nrow = length(m_vector), ncol = length(alpha_vector))
for (j in 1:length(alpha_vector)) {
  alpha <- alpha_vector[j] # from 0.5 to 2
  beta <- alpha / 2

  # Compute the eigenvalues and eigenfunctions on the finest mesh
  overkill_eigen_params <- compute_true_eigen_rectangle(
  a = a, 
  b = b, 
  loc = overkill_mesh_loc,
  kappa = kappa,
  alpha = alpha,
  m_max = m_max, 
  n_max = n_max
)
  EIGENVAL_ALPHA <- overkill_eigen_params$EIGENVAL_ALPHA # Eigenvalues (they are independent of the meshes)
  overkill_EIGENFUN <- overkill_eigen_params$EIGENFUN # Eigenfunctions on the finest mesh
  
  # Compute the true solution on the finest mesh
  overkill_U_true <- overkill_EIGENFUN %*% 
    outer(1:length(coeff_U_0), 
          1:length(overkill_time_seq), 
          function(i, j) (coeff_U_0[i] + coeff_FF[i] * G_sin(t = overkill_time_seq[j], A = AAA, lambda_j_alpha_half = EIGENVAL_ALPHA[i], omega = OMEGA)) * exp(-EIGENVAL_ALPHA[i] * overkill_time_seq[j]))
  
  for (i in 1:length(m_vector)) {
    m <- m_vector[i]
    h <- sqrt(a*b)*exp(- pi * sqrt((1 - alpha / 2) * m))/divider 
    time_step <- (h^alpha)
    n <- ceiling(-1+sqrt(a*b)/h)
    mesh_and_FEM <- gets.mesh.and.FEM.on.rectangle(a, b, n)
    mesh <- mesh_and_FEM$mesh
    G <- mesh_and_FEM$G
    C <- mesh_and_FEM$C
    L <- kappa^2*C + G
    eigen_params <- compute_true_eigen_rectangle(
    a = a, 
    b = b, 
    loc = mesh$loc,
    kappa = kappa,
    alpha = alpha,
    m_max = m_max, 
    n_max = n_max)
    EIGENFUN <- eigen_params$EIGENFUN
    U_0 <- EIGENFUN %*% coeff_U_0 # Compute U_0 on the current mesh
    Psi <- fm_basis(mesh, overkill_mesh_loc)
  
    time_seq <- seq(0, T_final, length.out = ((T_final - 0) / time_step + 1))
    my_op_frac <- my.fractional.operators.frac(L, beta, C, scale.factor = kappa^2, m = m, time_step)
    INT_BASIS_EIGEN <- t(overkill_EIGENFUN) %*% overkill_C %*% Psi
    # Compute matrix F with columns F^k
    FF_approx <- t(INT_BASIS_EIGEN) %*% 
      outer(1:length(coeff_FF), 
            1:length(time_seq), 
        function(i, j) coeff_FF[i] * g_sin(r = time_seq[j], A = AAA, omega = OMEGA))
    
    U_approx <- solve_fractional_evolution(my_op_frac, time_step, time_seq, val_at_0 = U_0, RHST = FF_approx)
    
    projected_U_approx <- Psi %*% U_approx
    projected_U_piecewise <- construct_piecewise_projection(projected_U_approx, time_seq, overkill_time_seq)
    XX <- overkill_U_true - projected_U_piecewise
    error_ij <-  sqrt(as.double(overkill_time_step * sum(XX * (overkill_C %*% XX)))) 
    errors_projected[i,j] <- error_ij
    slackr_msg(text = paste0("m =", m, ", alpha =", alpha, ", h =", h, ", time_step =", time_step, ", error =", error_ij), channel = "#research")
  }
}
save(errors_projected, file = here::here("data_files/errors_projected_m_square.RData"))
```


```{r}
load(here::here("data_files/errors_projected_m_square.RData"))
```


```{r, fig.align='center', fig.dim= c(4,5), fig.cap = captioner("Comparison between theoretical and observed convergence behavior for the $L_2(\\Gamma\\times(0,T))$-error with respect to $m$ on a semi-$\\text{log}_{e}$ scale, with $m$ plotted on a square-root scale. Dashed lines indicate the theoretical rates, and solid lines represent the observed error curves. The legend below each plot shows the value of $\\alpha$ along with the corresponding theoretical ('theo'), and observed ('obs') rates for each case.")}
observed_rates <- numeric(length(alpha_vector))
for (u in 1:length(alpha_vector)) {
  observed_rates[u] <- coef(lm(log(errors_projected[, u]) ~ sqrt(m_vector)))[2]
}

theoretical_rates <- - pi * alpha_vector * sqrt((1 - alpha_vector / 2))

p_m <- error.convergence.plotter(x_axis_vector = m_vector, 
                                 alpha_vector, 
                                 errors_projected, 
                                 theoretical_rates, 
                                 observed_rates,
                                 line_equation_fun = exp_line_equation,
                                 fig_title = expression("Convergence in " * italic(m)),
                                 x_axis_label = expression(italic(m)),
                                 apply_sqrt = TRUE)
p_m
```


```{r}
ggsave(here::here("data_files/conv_rates_m_square.png"), width = 4, height = 5, plot = p_m, dpi = 300)
save(p_m, file = here::here("data_files/p_m_square.RData"))
```


```{r, eval = TRUE, echo = FALSE}
initial_comment <- paste0("Here’s the latest plot update for the convergence in m for the optimal control problem! with kappa = ", kappa, " and divider = ", divider)
# option 1
slackr_upload(
  filename = "data_files/conv_rates_m_square.png",        # path to your image
  initial_comment = initial_comment,
  channels = "#research"
)
```

# References

```{r}
cite_packages(output = "paragraph", out.dir = ".")
```
